### 2021-03-03 10-Conclusions.tex
LOW / Init Latex article

including  lessons  learned,  impulses  for future work, etc...

### 2021-03-03 5-Analysis.tex
LOW / Init Latex article

any disambiguations of the  case  description and assumptions made, any potentially added require-ments

### 2021-03-03 6-ModelPresentation.tex
LOW / Init Latex article

detailed  presentation  of  a  model,including justifications for design decisions

### 2021-03-03 7-RequirementsSatisfaction.tex
LOW / Init Latex article

demonstration of how the solution satisfies the challenge requirements

### 2021-03-03 9-Relatedworks.tex
LOW / Init Latex article

Positioning  and  contrasting  the  presented solution with related work

### 2021-03-03 1-Main.tex

%\documentclass[preprint]{acmart}
\documentclass[preprint]{elsarticle}
\journal{Science of Computer Programming}

\usepackage{todonotes}

\usepackage{listings}
\usepackage{xcolor}
\usepackage{hyperref}
\usepackage{tikz-uml}

\definecolor{codegreen}{rgb}{0,0.6,0}
\definecolor{codegray}{rgb}{0.5,0.5,0.5}
\definecolor{codepurple}{rgb}{0.58,0,0.82}
\definecolor{backcolour}{rgb}{0.95,0.95,0.92}

\lstdefinestyle{mystyle}{
    backgroundcolor=\color{backcolour},   
    commentstyle=\color{codegreen},
    keywordstyle=\color{magenta},
    numberstyle=\tiny\color{codegray},
    stringstyle=\color{codepurple},
    basicstyle=\ttfamily\footnotesize,
    breakatwhitespace=false,         
    breaklines=true,                 
    captionpos=b,                    
    keepspaces=true,                 
    numbers=left,                    
    numbersep=5pt,                  
    showspaces=false,                
    showstringspaces=false,
    showtabs=false,                  
    tabsize=2https://www.overleaf.com/project/5efee0e8ad6bc10001d47854
}

\lstset{style=mystyle}

\DeclareTextFontCommand{\mytexttt}{\ttfamily\hyphenchar\font=45\relax}

%\acmPrice{15.00}
%\acmDOI{10.1145/1122445.1122456}
%\acmISBN{978-1-4503-9999-9/18/06}

\begin{document}
\begin{frontmatter}

\title{MULTI process challenge : the FML solution}

%% Group authors per affiliation:
%\author{Sylvain Guérin, Joel Champeau}
%\address{ENSTA Bretagne}

%\author{Jean-Christophe Bach, Salvador Martinez, Antoine Beugnard}
%\address{IMT Atlantique}

%% or include affiliations in footnotes:
\author[mymainaddress]{Sylvain Guérin\corref{mycorrespondingauthor}}
\cortext[mycorrespondingauthor]{Corresponding author}
%\ead{firstname.lastname@ensta-bretagne.fr}

\author[mymainaddress]{Joel Champeau}
%\ead{firstname.lastname@ensta-bretagne.fr}

\author[mysecondaryaddress]{Jean-Christophe Bach}
%\ead{firstname.lastname@imt-atlantique.fr}

\author[mysecondaryaddress]{Salvador Mart\'inez}
%\ead{firstname.lastname@imt-atlantique.fr}

\author[mysecondaryaddress]{Antoine Beugnard}
%\ead{firstname.lastname@imt-atlantique.fr}

\address[mymainaddress]{ENSTA Bretagne, Lab-STICC, UMR 6285, Brest, France}%vérifier si c'est l'affiliation officielle
\address[mysecondaryaddress]{IMT Atlantique, Lab-STICC, UMR 6285, F-29238 Brest, France}

\input{2-Abstract}

\begin{keyword}
Multilevel modeling \sep Challenge
\end{keyword}

\end{frontmatter}

%%
%% This command processes the author and affiliation and title
%% information and builds the first part of the formatted document.

\section{Introduction}
\label{sec:introduction}
\input{3-Introduction.tex}

\section{Technology}
\label{sec:technology}
\input{4-Technology.tex}

\section{Analysis}
\label{sec:analysis}
\input{5-Analysis.tex}

\section{Model presentation}
\label{sec:model}
\input{6-ModelPresentation.tex}

\section{Satisfaction of Requirements}
\label{sec:requirements}
\input{7-RequirementsSatisfaction.tex}

\section{Assessment of the Modeling Solution}
\label{sec:discussion}
\input{8-Discussion.tex}

\section{Related work}
\label{sec:relatedwork}
\input{9-RelatedWork.tex}

\input{10-Conclusions.tex}

%%
%% The next two lines define the bibliography style to be used, and
%% the bibliography file.
\bibliographystyle{ACM-Reference-Format}
\bibliography{biblio}

\end{document}
\endinput
%%

### 2021-03-03 4-Technology.tex

precise  description  of  the  technology  / approach that is used

### 2021-03-03 8-Discussion.tex

discussing choices  made, pointing out potential compromises/deficiencies

### 2021-03-03 1-Main.tex

@@
 \documentclass[preprint]{elsarticle}
-\journal{Science of Computer Programming}
+\journal{EMISA Journal}
 
@@
 \definecolor{backcolour}{rgb}{0.95,0.95,0.92}
-
-\lstdefinestyle{mystyle}{
-    backgroundcolor=\color{backcolour},   
-    commentstyle=\color{codegreen},
-    keywordstyle=\color{magenta},
-    numberstyle=\tiny\color{codegray},
-    stringstyle=\color{codepurple},
-    basicstyle=\ttfamily\footnotesize,
-    breakatwhitespace=false,         
-    breaklines=true,                 
-    captionpos=b,                    
-    keepspaces=true,                 
-    numbers=left,                    
-    numbersep=5pt,                  
-    showspaces=false,                
-    showstringspaces=false,
-    showtabs=false,                  
-    tabsize=2https://www.overleaf.com/project/5efee0e8ad6bc10001d47854
-}
-
-\lstset{style=mystyle}
-
-\DeclareTextFontCommand{\mytexttt}{\ttfamily\hyphenchar\font=45\relax}
 
@@
 \label{sec:relatedwork}
-\input{9-RelatedWork.tex}
+\input{9-Relatedworks.tex}
 
+\section{Conclusions}
 \input{10-Conclusions.tex}

### 2021-03-12 1-Main.tex

@@
 %\documentclass[preprint]{acmart}
-\documentclass[preprint]{elsarticle}
-\journal{EMISA Journal}
+\documentclass[american]{emisa}
 
 \usepackage{todonotes}
+
+\newcommand{\noteSM}[1]{\todo[color=red!40, author=\textbf{Salvador}, inline, caption={}]{#1}}
+\newcommand{\noteJC}[1]{\todo[color=blue!40, author=\textbf{JC}, inline, caption={}]{#1}}
+\newcommand{\noteSylvain}[1]{\todo[color=green!40, author=\textbf{Sylvain}, inline, caption={}]{#1}}
+\newcommand{\noteAntoine}[1]{\todo[color=yellow!40, author=\textbf{Antoine}, inline, caption={}]{#1}}
 
@@
 \usepackage{tikz-uml}
+
+\usepackage{amsmath,amssymb,mathtools}
+\usepackage{algorithmic,algorithm}
 
@@
 \begin{document}
-\begin{frontmatter}
-
+\begin{article}{
 \title{MULTI process challenge : the FML solution}
@@
 %% or include affiliations in footnotes:
-\author[mymainaddress]{Sylvain Guérin\corref{mycorrespondingauthor}}
-\cortext[mycorrespondingauthor]{Corresponding author}
+\author*{Sylvain Guérin}{firstname.lastname@ensta-bretagne.fr}
+\address{ENSTA Bretagne, Lab-STICC, UMR 6285, Brest, France}
 %\ead{firstname.lastname@ensta-bretagne.fr}
 
-\author[mymainaddress]{Joel Champeau}
-%\ead{firstname.lastname@ensta-bretagne.fr}
+\author{Joel Champeau}
+\address[a]{}
 
-\author[mysecondaryaddress]{Jean-Christophe Bach}
+\author{Jean-Christophe Bach}
+\address{IMT Atlantique, Lab-STICC, UMR 6285, F-29238 Brest, France}
 %\ead{firstname.lastname@imt-atlantique.fr}
 
-\author[mysecondaryaddress]{Salvador Mart\'inez}
-%\ead{firstname.lastname@imt-atlantique.fr}
+\author{Salvador Mart\'inez}
+\address[b]{}
 
-\author[mysecondaryaddress]{Antoine Beugnard}
-%\ead{firstname.lastname@imt-atlantique.fr}
+\author{Antoine Beugnard}
+\address[b]{}
 
-\address[mymainaddress]{ENSTA Bretagne, Lab-STICC, UMR 6285, Brest, France}%vérifier si c'est l'affiliation officielle
-\address[mysecondaryaddress]{IMT Atlantique, Lab-STICC, UMR 6285, F-29238 Brest, France}
-
-\input{2-Abstract}
-
-\begin{keyword}
-Multilevel modeling \sep Challenge
-\end{keyword}
-
-\end{frontmatter}
-
+\abstract{\input{2-Abstract}}
+\keywords{Multilevel Modeling Challenge, Federation Modeling Language, OpenFlexo}
+\bibliography{}
+}
 %%
@@
 \input{3-Introduction.tex}
+
+\noteSM{teste}
+\noteJC{test}
+\noteSylvain{Test}
+\noteAntoine{Test}
 
@@
 
+\end{article}
 \end{document}

### 2021-04-27 2-Abstract.tex
Updates from Overleaf

\begin{abstract}
TODO, deadline = 17/05/2021
\end{abstract}

### 2021-04-29 3-Introduction.tex

TODO

\cite{multiProcessChallenge2019}

\emph{Submissions responding to the challenge should describe
a multi-level model conforming to the case description,
including justifications for non-trivial design decisions. In
order to foster comparability between solutions, respondents
are asked to make sure that concepts of the case description
are explicitly represented by one or more model elements.}\\

\emph{criteria:
\begin{enumerate}
    \item Does the response address the established domain as described in the challenge description and demonstrate the use of modeling features?
    \item Does it evaluate/discuss its proposed modeling solution against the criteria described in the challenge description?
    \item Does it discuss the merits and limitations of the employed modeling technique?
    Authors are invited to suggest further requirements to demonstrate particular benefits of the approaches they adopt.
\end{enumerate}
}

structure : plus ou moins imposée par le challenge (structure actuelle = sections au moins attendues)

### 2021-04-29 5-Analysis.tex

@@
-any disambiguations of the  case  description and assumptions made, any potentially added require-ments
+any disambiguations of the  case  description and assumptions made, any potentially added requirements
+
+quelles hypothèses supplémentaires ?
+
++ architecture de la modélisation (ce qui se trouve dans le ppt), le détail se retrouvera dans la section suivante
+
+Joel commence

### 2021-05-04 4-Technology.tex

@@
-precise  description  of  the  technology  / approach that is used
+% Inspiré de la section 10 du papier formose
+
+To meet the \mpc, we have decided to use the general
+approach of model federation~\cite{Golra2016-federation}. \emph{Model
+  federation} is a way models are assembled together with some kind of
+low-coupling links. It has been studied to answer the OMG' RFP
+Semantic Modeling for Information Federation (SIMF)~\cite{simf}. In
+contrast to approaches that compose meta-models into a single large
+meta-model welcoming all needed entities, model federation build links
+among models and meta-models (even through levels) to make ``things''
+work together. As an illustration of this feature, we developed a
+free-modeling editor -- freeing oneself from the bonds of
+model/meta-model conformity -- that is presented
+in~\cite{models2016-freemodel}. Another very interesting feature of
+this approach is the strong decoupling among tools that remain usable
+after federations are made.
+
+We decided to use this approach since it offers the possibility to
+link and to navigate among levels. Before we describe the overall
+architecture in next section, here are some key concepts implemented
+in the Openflexo~\cite{openflexo_link} framework.
+
+% figure
+% Dans la suite les références à la figure sont commentés
+
+This framework relies on the following architecture. A federation
+gathers a set of conceptual models, named \emph{virtual models} and a
+set of \emph{federated models}. Each federated model pertains to a
+\emph{technological space} and uses the language of its specific
+paradigm while a virtual model is built using the Flexo Modeling
+Language (FML). Each federated model can be viewed as an autonomous
+component that may evolve with its own tooling. The virtual models
+serve as control components.
+% Figure~\ref{fig:space-ex} presents a
+% simplified version of the \textsf{Goal-xls} federation example that
+% groups these different models. The upper half of the figure
+% illustrates the design of a federation with a model coming from the
+% \smk technological space (the goal diagram), one from the Excel
+% technological space (the file) and one virtual model, named
+% \emph{ReqXls}, reifying the dependencies between the goal model and
+% the reporting file.
+
+A \emph{technology adapter} (TA) is a reusable library that defines
+connections between the FML execution engine and a particular
+technological space. The model federation framework provides ways to
+define TA.
+%The \emph{ReqXls} virtual model relies on two technology
+%adapters (Excel TA and \smk TA) for manipulating the models in their
+%respective technological spaces.
+
+FML is a language designed to define virtual models. A virtual model
+is composed of a set of \emph{concepts}, while itself being a concept.
+Hence, virtual models are structuring units while concepts are the
+core entities. A concept has a set of \emph{roles} and
+\emph{behaviors}. A parallel to object-oriented approach can be useful
+to understand FML\footnote{Beware, even though useful for
+  comprehension, this correspondence is not
+  reliable, as some aspects of FML do not map to object oriented
+  concepts}. A concept corresponds to a class, its roles to the
+attributes of the class and its behaviors to the methods of the class.
+% In our example, a concept \texttt{Requirement} is presented with
+% three roles \texttt{row}, \texttt{goal} and \texttt{refines}. The
+% first two are represented as attributes whereas the third as an
+% association.
+These roles have types defining the kind of value the role will point
+at runtime.
+% Our example illustrates three forms of such types \ie another
+% concept for \textsf{refines}, a type defined in a TA for
+% \textsf{Row} or a type defined in an external model for
+% \textsf{Goal}.
+Whenever a type external to the federation space (from a TA or an
+external model) is used, one needs to use a \emph{model slot}. A model
+slot is a mediation entity, associated with a TA, in charge of giving
+access to external elements of the corresponding technological space.
+
+We should stress that building a virtual model is purely conceptual,
+model slots / TA are in charge of the technicalities. Additionally,
+when a virtual model or a concept point to an external model, it comes
+with an explicit interpretation of the external data. The virtual
+model can hide information, transform elements and does not need to
+cover the whole external metamodel.
+% In our example, it ignores all other elements of the \smk metamodel.
+We did not required TA in our proposal. The whole solution uses the
+internal (FML) language.
+
+FML is designed to define not only the structure of virtual models but
+also to define the collection of actions an engineer can perform on
+them. These actions are called \emph{business behaviors} and may be
+compared to methods of object-oriented approach.
+% The rename operation already cited for the \textsf{Goal-xls}
+% federation is an example of such a behavior. One could also define
+% operations to add or remove requirements, to refine relationships,
+% \emph{etc.} The renaming operation is defined by a \texttt{rename}
+% behavior in the concept \texttt{Requirement}. It uses the
+% \texttt{setCell} action of the Excel TA and the \texttt{rnGoal} of
+% the \smk TA.
+When the FML execution engine runs a federation, it creates virtual
+model instances containing concept instances. Some concept instances
+are connected to external elements through model slot instances.
+% The lower part of the figure~\ref{fig:space-ex} illustrates this
+% runtime phase.
+
+FML also supports another kind of behavior which can be described as
+interactors. Such behavior is reactive and triggers when any of the
+federated models evolves. Two scenarios are then possible: (1) an
+action in the federation is propagated to the federated models or (2)
+a modification is detected in a federated model and triggers a
+behavioral interactor of the federation.
+% Hence the double arrows in the links of Figure~\ref{fig:space-ex}.
+This feature is useful when models have to be synchronised. We did not
+exploit this possibility in the challenge.
+
+The tool support for model federation framework,
+Openflexo\footnote{\url{https://github.com/openflexo-team}}, is
+developed as an open source initiative. This tool offers a FML
+execution engine with an interactive virtual model design environment.
+It has been used in several case studies including model mapping,
+multi-paradigm process modeling and enterprise architecting. It has
+also been used to build a tool, the freemodeling editor that has been
+used in industrial projects~\cite{models2016}. As of today, this tool
+offers some mature technology adapters (\emph{docx} and \emph{excel}
+for documents, EMF and OWL for modeling languages, JDBC for databases,
+REST and XMLRPC for external services and one for diagraming tools)
+and other more rudimentary (pdf, HTTP, XML, etc.).
+
+Finally, tools have their own model. We have taken advantage of
+Openflexo features to build in parallel with the models and virtual
+models required by the challenge a drawing tool that makes our
+solution (partially) executable.

### 2021-05-05 1-Main.tex
Updates from Overleaf

@@
 \newcommand{\noteSM}[1]{\todo[color=red!40, author=\textbf{Salvador}, inline, caption={}]{#1}}
-\newcommand{\noteJC}[1]{\todo[color=blue!40, author=\textbf{JC}, inline, caption={}]{#1}}
+\newcommand{\noteJC}[1]{\todo[color=pink!40, author=\textbf{JC}, inline, caption={}]{#1}}
 \newcommand{\noteSylvain}[1]{\todo[color=green!40, author=\textbf{Sylvain}, inline, caption={}]{#1}}
 \newcommand{\noteAntoine}[1]{\todo[color=yellow!40, author=\textbf{Antoine}, inline, caption={}]{#1}}
+\newcommand{\noteFabien}[1]{\todo[color=orange!40, author=\textbf{Fabiem}, inline, caption={}]{#1}}
+
+\newcommand{\mpc}{MULTI process challenge\xspace}
 
@@
 
-\usepackage{amsmath,amssymb,mathtools}
+%\usepackage{amsmath,mathtools}%amssymb,
 \usepackage{algorithmic,algorithm}
@@
 \begin{article}{
-\title{MULTI process challenge : the FML solution}
+\title{\mpc: the FML solution}
 
@@
 
-%\author{Jean-Christophe Bach, Salvador Martinez, Antoine Beugnard}
+%\author{Jean-Christophe Bach, Salvador Martinez, Antoine Beugnard, Fabien Dagnat}
 %\address{IMT Atlantique}
@@
 
+\author{Fabien Dagnat}
+\address[b]{}
+
 \abstract{\input{2-Abstract}}
@@
 
-\noteSM{teste}
+\noteSM{test}
 \noteJC{test}
@@
 \noteAntoine{Test}
+\noteFabien{test}
 
@@
 
-%%
 %% The next two lines define the bibliography style to be used, and
 %% the bibliography file.
-\bibliographystyle{ACM-Reference-Format}
+%\bibliographystyle{ACM-Reference-Format}
 \bibliography{biblio}
+
+% à utiliser d'après le style (avec biber), voir comment ça fonctionne sous overleaf, sinon CLI et tant pis/mieux
+%\printbibliography
 

### 2021-05-05 6-ModelPresentation.tex
Updates from Overleaf

@@
-detailed  presentation  of  a  model,including justifications for design decisions
+% detailed  presentation of a model, including justifications for design decisions
+
+Figure \ref{fig:MultilevelArchitecture} shows the overview of the multivel architecture designed to capture the two use cases described in Process Challenge.
+
+% Expliquer que le use case XSure ne nécessitait pas l'extension de metamodèle requise pour le usecase Acme Software Developement Process
+
+\subsection{Base metamodel for the Process Challenge}
+
+\begin{figure*}
+ \centering
+    % \includegraphics[width=1.0 \columnwidth]{Figures/Metamodel.pdf}
+    \includegraphics[width=1.0 \textwidth]{Figures/Metamodel.pdf}
+    \caption{Process management base metamodel}
+    \label{fig:BaseMetamodel}
+\end{figure*}
+
+\subsection{The XSure insurance domain}
+
+The use case presented in this section is straightforward implementable while instantiating base metamodel presented in figure \ref{fig:BaseMetamodel}.
+
+% Passer rapidement sur ce cas pour illustrer le metamodèle décrit plus haut
+
+% We will illustrate the approach...
+
+\subsection{The Acme software development process}
+
+\subsection{Openflexo tooling}
+
+% Expliquer l'outillage construit et comment il marche

### 2021-05-05 8-Discussion.tex

@@
-discussing choices  made, pointing out potential compromises/deficiencies
+discussing choices  made (ex : héritage vs attr), pointing out potential compromises/deficiencies
+jc, fabien ; 
+pas d'explicitation de niveaux de « concrétude » : conséquences ? facile de tracer des liens entre les niveaux ; difficile d'avoir le niveau ; 
+outil adaptable, niveau d'abstraction, on fournit une solution + l'outil pour exécuter le processus ; réutilisabilité des modèles ; 
+\todo{on pourrait aussi répondre ici aux 3 questions posées explicitement (copiées dansl'intro pour le moment)}
+
+difficulctés/facilités
+points saillants :  outillage "gratuit" ;
+
+Although our tool was not designed with a focus on multi-level modelling, we were able to propose a fully working solution.
+
+Our approach is level-blind, therefore it is not easy to quickly retrieve the abstraction level of a concept.

### 2021-05-07 6-ModelPresentation.tex

@@
 
-Figure \ref{fig:MultilevelArchitecture} shows the overview of the multivel architecture designed to capture the two use cases described in Process Challenge.
+Figure \ref{fig:MultilevelArchitecture} shows the overview of the multivel architecture designed to capture the two use cases described in the Process Challenge.
 

### 2021-05-07 1-Main.tex
Change title
Updates from Overleaf

@@
-%\documentclass[preprint]{acmart}
 \documentclass[american]{emisa}
@@
 \newcommand{\noteAntoine}[1]{\todo[color=yellow!40, author=\textbf{Antoine}, inline, caption={}]{#1}}
-\newcommand{\noteFabien}[1]{\todo[color=orange!40, author=\textbf{Fabiem}, inline, caption={}]{#1}}
+\newcommand{\noteFabien}[1]{\todo[color=orange!40, author=\textbf{Fabien}, inline, caption={}]{#1}}
+\newcommand{\noteJoel}[1]{\todo[color=blue!20, author=\textbf{Joel}, inline, caption={}]{#1}}
 
@@
 
-%\acmPrice{15.00}
-%\acmDOI{10.1145/1122445.1122456}
-%\acmISBN{978-1-4503-9999-9/18/06}
-
 \begin{document}
 \begin{article}{
-\title{\mpc: the FML solution}
+%\title{\mpc: the FML solution}
+\title{Multi-level modelling with Openflexo/FML - A contribution to the \mpc}
 
@@
 \keywords{Multilevel Modeling Challenge, Federation Modeling Language, OpenFlexo}
-\bibliography{}
+\bibliography{biblio.bib}
+
 }
@@
 \noteJC{test}
-\noteSylvain{Test}
-\noteAntoine{Test}
+\noteSylvain{test}
+\noteAntoine{test}
 \noteFabien{test}
+\noteJoel{test}
 
@@
 
-%% The next two lines define the bibliography style to be used, and
-%% the bibliography file.
-%\bibliographystyle{ACM-Reference-Format}
-\bibliography{biblio}
-
-% à utiliser d'après le style (avec biber), voir comment ça fonctionne sous overleaf, sinon CLI et tant pis/mieux
-%\printbibliography
+\printbibliography
 
@@
 \end{document}
-\endinput
 %%

### 2021-05-07 6-ModelPresentation.tex
Updates from Overleaf

@@
 
+We will illustrate the base metamodel presented in this section with XSure insurance domain use case, whose a partial description was provided in the challenge description.
+Note that the capture of the requirements P1 to P19 in the context of XSure insurance domain are straightforward implementable while instantiating base metamodel presented in figure \ref{fig:BaseMetamodel}.
+
 The use case presented in this section is straightforward implementable while instantiating base metamodel presented in figure \ref{fig:BaseMetamodel}.

### 2021-05-07 6-ModelPresentation.tex
synchro de overleaf

--- 2021-05-07 6-ModelPresentation.tex
@@
 
+We will illustrate the base metamodel presented in figure \ref{fig:BaseMetamodel}  with XSure insurance domain use case, whose a partial description was provided in the challenge description.
+Note that the capture of the requirements P1 to P19 in the context of XSure insurance domain are straightforward implementable while instantiating base metamodel.
+
 The use case presented in this section is straightforward implementable while instantiating base metamodel presented in figure \ref{fig:BaseMetamodel}.

### 2021-05-07 6-ModelPresentation.tex

--- 2021-05-07 6-ModelPresentation.tex
@@
 
-\subsection{The XSure insurance domain}
+\todo{Augmenter la taille des fontes}
 
-The use case presented in this section is straightforward implementable while instantiating base metamodel presented in figure \ref{fig:BaseMetamodel}.
+In this section, We will present and illustrate the base metamodel presented in figure \ref{fig:BaseMetamodel} with XSure insurance domain use case, whose a partial description was provided in the challenge description.
+Note that the capture of the requirements P1 to P19 in the context of XSure insurance domain are straightforward implementable while instantiating \textit{XSure model}, as an instance of base metamodel (the left side of figure \ref{fig:MultilevelArchitecture}).
+
+Figure \ref{fig:BaseMetamodel} represents this base metamodel with a UML-like formalism, well adapted to represent object-oriented FML concepts and their instances.
+Our proposition relies on ontologic instantiation as presented in figure \ref{fig:LinguisticAndOntologicInstantiation}, with a common root concept \textit{ModelingElement}.
+Two types of ontological instantiation are needed to meet the requirements of the challenge.
+One provides that some instances conform to only one type (\textit{Process} and \textit{Task} do respectively conform to one \textit{ProcessType} and \textit{TaskType}) while some entities define their conformity to several types (\textit{Actor} and \textit{artifact} do respectively conform to several \textit{ProcessType} and \textit{TaskType})
 

### 2021-05-07 6-ModelPresentation.tex
synchro

@@
 
-In this section, We will present and illustrate the base metamodel presented in figure \ref{fig:BaseMetamodel} with XSure insurance domain use case, whose a partial description was provided in the challenge description.
+In this section, we present and illustrate the base metamodel presented in figure \ref{fig:BaseMetamodel} with XSure insurance domain use case, whose a partial description was provided in the challenge description.
 Note that the capture of the requirements P1 to P19 in the context of XSure insurance domain are straightforward implementable while instantiating \textit{XSure model}, as an instance of base metamodel (the left side of figure \ref{fig:MultilevelArchitecture}).
@@
 Two types of ontological instantiation are needed to meet the requirements of the challenge.
-One provides that some instances conform to only one type (\textit{Process} and \textit{Task} do respectively conform to one \textit{ProcessType} and \textit{TaskType}) while some entities define their conformity to several types (\textit{Actor} and \textit{artifact} do respectively conform to several \textit{ProcessType} and \textit{TaskType})
+One provides that some instances conform to only one type (\textit{Process} and \textit{Task} do respectively conform to one \textit{ProcessType} and \textit{TaskType}) while some entities define their conformity to several types (\textit{Actor} and \textit{Artifact} do respectively conform to several \textit{ActorType} and \textit{ArtifactType}).
+This instantiation is reified with the relations \textit{type} and \textit{types} between \textit{Type}, \textit{Instance} and \textit{MultiInstance} concepts.
 

### 2021-05-07 8-Discussion.tex

@@
 outil adaptable, niveau d'abstraction, on fournit une solution + l'outil pour exécuter le processus ; réutilisabilité des modèles ; 
-\todo{on pourrait aussi répondre ici aux 3 questions posées explicitement (copiées dansl'intro pour le moment)}
+\todo{on pourrait aussi répondre ici aux 3 questions posées explicitement (copiées dansl'intro pour le moment), a priori non, le chef trouve que c'est chiant}
 
-difficulctés/facilités
+ajouter quelques métriquse "gros grains"
+
+difficultés/facilités : contraintes faites à la main par rapport aux langages où on peut spécifier le niveau d'extension lorsque l'on trace un lien multi niveaux -> outil plus souple, mais moins de comportements automatisés ; + ouvert ; moins de support (faudrait implémenter ce qui manque) ; requiert de la méthode, ajouter des comportemets, etc.
 points saillants :  outillage "gratuit" ;

### 2021-05-16 8-Discussion.tex
Updates from Overleaf

@@
 outil adaptable, niveau d'abstraction, on fournit une solution + l'outil pour exécuter le processus ; réutilisabilité des modèles ; 
-\todo{on pourrait aussi répondre ici aux 3 questions posées explicitement (copiées dansl'intro pour le moment), a priori non, le chef trouve que c'est chiant}
+\todo[inline]{on pourrait aussi répondre ici aux 3 questions posées explicitement (copiées dansl'intro pour le moment), a priori non, le chef trouve que c'est chiant}
 

### 2021-05-16 9-Relatedworks.tex
Updates from Overleaf

@@
-Positioning  and  contrasting  the  presented solution with related work
+%Positioning  and  contrasting  the  presented solution with related work
+
+%Salvador commence
+
+A first description of the process modeling challenge as a multi-level modeling problem was proposed in \cite{lara2018refactoring} in the context of a catalogue of refactoring for multi-level models.
+A solution is provided at a conceptual level using deep instantiation and linguistic extensions.
+
+\cite{multiecore2019} use MultiEcore \citep{multecore2016} to provide a solution to the challenge.
+MultEcore uses cascading and potency.
+They do not need the synthetic typing relations we introduce.
+However they need to do frequent model transformations to transform instance models into instantiable models.
+With respect to them, we use less models as we do not include a generic software engineering process model nor a generic assurance process model.
+Finally, they use multi-level couple model transformations for the definition and enforcement of constraints.
+
+More similar to us, \cite{deeptelos2019} use DeepTelos\citep{deeptelos2016}, an extension of the Telos language \citep{telos1990} in order to solve the process challenge.
+DeepTelos is based on the use of the PowerType \citep{atkinson2001essence} pattern.
+As with OpenFlexo and FML, they do not use explicit (numbered) levels nor potency.
+
+%Their solution to the challenge includes 5 (ontological) levels and 8 models, including Ecore at the top.
+%They include two supplementary hierarchies (linguistic extensions) that are not bound to any of the %aforementioned 5 levels.
+Finally, they use multi-level couple model transformations for the definition and %enforcement of constraints.
+Their approach is implemented on top of EMF and achieves integration with its %strict modeling paradigm by extending the two-level cascading mechanism to all the hierarchy.
+
+\cite{dmla2019} Use the Dynamic Multi-Layer Algebra (DMLA)\citep{dmla2017} in order to solve the process challenge.
+DMLA is a level-blind modeling framework which is fully customizable (e.g., different types of instantiaton may be implemented).
+Their proposed solution separates the process challenge in two separate domains, namely, the tasks definition domain and the process definition domain.
+Wrappers are used in order to reuse tasks in processes.
+Their solution does not use inheritance, as it is not supported by the framework.
+It does not use explicitly separated models either.
+
+%DMLA does not directly supports inheritance nor bi-directional navigation which prevents an easy solution to %some of the challenge requirements (e.g., p9 and s13).
+
+\noteSM{Say something about constraints in FML}

### 2021-05-17 2-Abstract.tex
Updates from Overleaf
add a comment to Salvador's comment

@@
-\begin{abstract}
-TODO, deadline = 17/05/2021
-\end{abstract}
+Model federation is a multi model \emph{management} approach based on the use of virtual models and loosely coupled links.
+The models in a federation remain autonomous and represented in their original technological spaces whereas virtual models and links serve as control components used to present different views to the users and  maintain synchronization.
+In this paper we tackle the \emph{process modeling challenge}, which consists in providing a solution to the problem of specifying and enacting processes.
+Solutions must fulfill a number of requirements for a process representation defined at an abstract process-definition level and at various more concrete domain-specific levels, resulting in a multi-level hierarchy of related models.
+We present a solution based on model federation and discuss the advantages and limitations of using this approach for multi-level modeling.
+Our solution is fully implemented with the OpenFlexo framework and its internal Flexo Modeling Language (FML).
+It fulfills all the challenge requirements.
+
+\noteSM{J'ai fait un premier essai d'abstract... un peu trop vendeur peut-etre?}
+\noteJC{bof, pas de manière éhontée : il faut qu'un abstract soit un peu
+  vendeur et les exigences sont bien remplies. La fin est très positive, mais
+afactuelle selon moi.}

### 2021-05-17 4-Technology.tex

@@
 also been used to build a tool, the freemodeling editor that has been
-used in industrial projects~\cite{models2016}. As of today, this tool
+used in industrial projects~\cite{models2016-freemodel}. As of today, this tool
 offers some mature technology adapters (\emph{docx} and \emph{excel}
@@
 Finally, tools have their own model. We have taken advantage of
-Openflexo features to build in parallel with the models and virtual
-models required by the challenge a drawing tool that makes our
+Openflexo features to build, in parallel with the models and virtual
+models required by the challenge, a drawing tool that makes our
 solution (partially) executable.

### 2021-05-17 1-Main.tex
format and cleaning
Updates from Overleaf
OpenFlexo -> Openflexo

@@
 
+\DeclareTextFontCommand{\mytexttt}{\ttfamily\hyphenchar\font=45\relax}
+
+\lstdefinestyle{mystyle}{
+    backgroundcolor=\color{backcolour},   
+    commentstyle=\color{codegreen},
+    keywordstyle=\color{magenta},
+    numberstyle=\tiny\color{codegray},
+    stringstyle=\color{codepurple},
+    basicstyle=\ttfamily\footnotesize,
+    breakatwhitespace=false,         
+    breaklines=true,                 
+    captionpos=b,                    
+    keepspaces=true,                 
+    numbers=left,                    
+    numbersep=5pt,                  
+    showspaces=false,                
+    showstringspaces=false,
+    showtabs=false,                  
+    tabsize=2
+}
+
+\lstset{style=mystyle}
+
 \begin{document}
 \begin{article}{
-%\title{\mpc: the FML solution}
+    % \title{\mpc: the FML solution}
     \title{Multi-level modelling with Openflexo/FML - A contribution to the \mpc}
-
-%% Group authors per affiliation:
-%\author{Sylvain Guérin, Joel Champeau}
-%\address{ENSTA Bretagne}
-
-%\author{Jean-Christophe Bach, Salvador Martinez, Antoine Beugnard, Fabien Dagnat}
-%\address{IMT Atlantique}
 
@@
     \address{ENSTA Bretagne, Lab-STICC, UMR 6285, Brest, France}
-%\ead{firstname.lastname@ensta-bretagne.fr}
+    % \ead{firstname.lastname@ensta-bretagne.fr}
 
@@
     \address{IMT Atlantique, Lab-STICC, UMR 6285, F-29238 Brest, France}
-%\ead{firstname.lastname@imt-atlantique.fr}
+    % \ead{firstname.lastname@imt-atlantique.fr}
 
@@
     \bibliography{biblio.bib}
-
 }
-%%
-%% This command processes the author and affiliation and title
-%% information and builds the first part of the formatted document.
 
@@
 \end{article}
+
 \end{document}

### 2021-05-17 6-ModelPresentation.tex
sync

@@
 
-In this section, we present and illustrate the base metamodel presented in figure \ref{fig:BaseMetamodel} with XSure insurance domain use case, whose a partial description was provided in the challenge description.
-Note that the capture of the requirements P1 to P19 in the context of XSure insurance domain are straightforward implementable while instantiating \textit{XSure model}, as an instance of base metamodel (the left side of figure \ref{fig:MultilevelArchitecture}).
+In this section, we present and illustrate the base metamodel presented in figure \ref{fig:BaseMetamodel} with \textit{XSure} insurance domain use case, whose a partial description was provided in the challenge description.
+Note that the capture of the requirements \textbf{P1} to \textbf{P19} in the context of XSure insurance domain are straightforward implementable while instantiating \textit{XSure model}, as an instance of base metamodel (the left side of figure \ref{fig:MultilevelArchitecture}).
 
 Figure \ref{fig:BaseMetamodel} represents this base metamodel with a UML-like formalism, well adapted to represent object-oriented FML concepts and their instances.
+Base conceptual element is called a \textit{FlexoConcept} in FML language, and is similar to the concept of class in object-oriented modeling.
+We use the UML representation of class to represent a \textit{FlexoConcept}, while cardinalities are shown using the UML-like classical rules.
+
 Our proposition relies on ontologic instantiation as presented in figure \ref{fig:LinguisticAndOntologicInstantiation}, with a common root concept \textit{ModelingElement}.
@@
 
-% Passer rapidement sur ce cas pour illustrer le metamodèle décrit plus haut
+\subsubsection{Process type definition}
 
-% We will illustrate the approach...
+We first present the process definition part of base metamodel, located left of figure \ref{fig:BaseMetamodel}.
+
+% Illustrer avec des instances de XSure
+
+\textit{ProcessType} is defined as a specialization of Type \textit{FlexoConcept}, and references a collection of \textit{TaskType} through the composition relation \textit{taskTypes} with (0..*) cardinality (\textbf{P1}).
+A \textit{TaskType} is embedded in a \textit{ProcessType} and inherits from it context.
+To illustrate this in \textit{XSure} insurance domain use case, \textit{XSure model} defines \textit{Claim Handling}, instance of \textit{ProcessType}, and \textit{Receive Claim}, \textit{Asses Claim} and \textit{Pay premium}, instances of \textit{TaskType}.
+
+\textit{ProcessType} also references a collection of gateways, reified with the \textit{Gateway} concept hierarchy.
+\textit{Gateway} is defined as a specialization of \textit{Type} and is specialized with \textit{Sequencing}, \textit{AndSplit}, \textit{AndJoin}, \textit{OrSplit} and \textit{OrJoin} concepts (\textbf{P2}).
+Depending of its type, a \textit{Gateway} defines one or more inputs and one or more outputs, depending of its underlying execution semantics.
+A \textit{ProcessType} additionally exposes unique initial \textit{TaskType} and a collection of final \textit{TaskType} with both attributes \textit{initialTaskType} (single cardinality) and \textit{finalTaskType} (cardinality 0..*) (\textbf{P3}).
+TaskType exposes a creator attribute as an association relation to Actor concept (\textbf{P4}), which is in the same conceptual level.
+
+\textit{ActorType} is defined as a sub-concept of \textit{Type} and the \textit{allowedActorTypes} relation to \textit{ActorType} which is defined in \textit{TaskType} (with 0..* cardinality) captures \textbf{P5} requirement.
+Requirement \textbf{P6} is symmetrically satisfied with \textit{allowedActors} relation to \textit{Actor} also defined in \textit{TaskType} (with same 0..* cardinality).
+Same modeling pattern applies for \textit{ArtefactType} defined as a sub-concept of \textit{Type}, and both relations \textit{usedArtifactTypes} and \textit{producedArtifactTypes} defined in \textit{TaskType} (\textbf{P7}).
+\textit{TaskType} additionnaly exposes an \textit{expectedDuration} attribute (expressed in number of days), satisfying \textbf{P8}.
+
+Actor concept defines a boolean attribute called \textit{isSenior}, while \textit{TaskType} defines an additional \textit{isCritical} boolean attribute, indicating that some instance of \textit{TaskType} are flagged as critical and must be performed by senior actors.
+To fullfill \textbf{P9} requirement, a supplementary constraint is required for \textit{TaskType} and is captured through this invariant expressed in FML language:
+
+\begin{lstlisting}[breaklines=true, language=java, basicstyle=\ttfamily\scriptsize, mathescape=true]
+forEach (actor : allowedActors) {
+    assert !isCritical | actor.isSenior
+}
+\end{lstlisting}
+
+This invariant should be completed with additional constraints defined in \textit{Task} \textit{FlexoConcept}, which apply to performing actors actually assigned to enacted tasks.
+
+\subsubsection{Process enactment}
+
+We now present the process enactment part of base metamodel, located right of figure \ref{fig:BaseMetamodel}.
+All concepts defined in this subsection are either specialization of \textit{Instance} concept (if they are associated with exactly one \textit{Type}) or \textit{MultiInstance} concept (for those which have several types).
+
+\textit{Process} represents an enacted \textit{ProcessType}, as defined in previous subsection (\textbf{P10}).
+FML defines behavioural features (operations in object-oriented modelling), called \texttt{FlexoBehaviour}.
+\textit{ProcessType} defines following behaviour \texttt{newProcess(String)}, taking a \texttt{String} argument (the name of the process to enact):
+
+\begin{lstlisting}[breaklines=true, language=XML, basicstyle=\ttfamily\scriptsize, mathescape=true]
+public Process newProcess(String name) {    
+  Process newProcess = new Process(parameters.name,this);      
+  for (taskType : taskTypes) {      
+    Task newTask = taskType.newTask(newProcess.name+"-"+taskType.name),newProcess);        
+  }      
+  return newProcess;    
+}    
+\end{lstlisting}
+
+This scheme allows to rely on FML dynamic binding mechanism to delegate to types the responsability for instances creation.
+
+% P10: Process <> ProcessType
+% P11: Tasks, toutes les TaskType sont instanciées en Task
+% P12: begin et end date dans Task
+% P13: instantiation des Artefacts et association des actors
+% P14: instantiation des Artefacts
+% P15: Actor/ActorType
+% P16: Artefact/ArtefactType
+% P17: cf P5&P6
+% P18: héritage des ActorTypes
+% P19: lastUpdated dans ModelingElement
 

### 2021-05-19 6-ModelPresentation.tex

@@
 
-\textit{ProcessType} is defined as a specialization of Type \textit{FlexoConcept}, and references a collection of \textit{TaskType} through the composition relation \textit{taskTypes} with (0..*) cardinality (\textbf{P1}).
+\textit{ProcessType} is defined as a specialization of \textit{Type} \texttt{FlexoConcept}, and references a collection of \textit{TaskType} through the composition relation \textit{taskTypes} with (0..*) cardinality (\textbf{P1}).
 A \textit{TaskType} is embedded in a \textit{ProcessType} and inherits from it context.
@@
 A \textit{ProcessType} additionally exposes unique initial \textit{TaskType} and a collection of final \textit{TaskType} with both attributes \textit{initialTaskType} (single cardinality) and \textit{finalTaskType} (cardinality 0..*) (\textbf{P3}).
-TaskType exposes a creator attribute as an association relation to Actor concept (\textbf{P4}), which is in the same conceptual level.
+TaskType exposes a creator attribute as an association relation to \textit{Actor} concept (\textbf{P4}), which is in the same conceptual level.
+
+Following listing shows an excerpt of FML code modeling some core concepts of process modeling base metamodel.
+
+\begin{lstlisting}[breaklines=true, language=java, basicstyle=\ttfamily\scriptsize, mathescape=true]
+model MetaModel {
+
+  concept ModelingElement {
+    ...
+  }
+
+  concept Type extends ModelingElement {
+    ...
+  }
+
+  concept ProcessType extends Type {
+    TaskType[0..*] taskTypes;
+    TaskType initialTaskType;
+    TaskType[0..*] finalTaskTypes;
+    Gateway[0..*] gateways;
+
+    concept TaskType extends Type {
+      Actor creator;
+      Actor[0..*] allowedActors;
+      ActorType[0..*] allowedActorTypes;
+      ...
+    }
+
+    abstract concept Gateway extends Type {
+      abstract void executeGateway(Process process);
+      ...
+    }
+
+    concept Sequencing extends Gateway {
+      TaskType in;
+      TaskType out;
+    }
+
+    // Other core concepts
+    ...
+  }
+}    
+\end{lstlisting}
 
@@
 
-\begin{lstlisting}[breaklines=true, language=XML, basicstyle=\ttfamily\scriptsize, mathescape=true]
+\begin{lstlisting}[breaklines=true, language=java, basicstyle=\ttfamily\scriptsize, mathescape=true]
 public Process newProcess(String name) {    
-  Process newProcess = new Process(parameters.name,this);      
+  Process newProcess = new Process(name,this);  
   for (taskType : taskTypes) {      
@@
 This scheme allows to rely on FML dynamic binding mechanism to delegate to types the responsability for instances creation.
+An instance of \textit{Process} references a unique \textit{ProcessType} trough the specialized \textit{/type} relation.
+Each instance of \textit{TaskType} is ontologically instantiated with a \textit{Task} (\textbf{P11}), using the same pattern where \textit{TaskType} has the responsability to manage this ontological instantiation.
+A \textit{Task} references its unique \textit{TaskType}, and defines a \textit{begin date} and an \textit{end date} as basic attributes of the concept (\textbf{P12}).
 
-% P10: Process <> ProcessType
-% P11: Tasks, toutes les TaskType sont instanciées en Task
-% P12: begin et end date dans Task
-% P13: instantiation des Artefacts et association des actors
-% P14: instantiation des Artefacts
-% P15: Actor/ActorType
-% P16: Artefact/ArtefactType
-% P17: cf P5&P6
-% P18: héritage des ActorTypes
-% P19: lastUpdated dans ModelingElement
+Same pattern applies for artifacts used and produced with relations \textit{usedArtifacts}, \textit{producedArtifacts} and \textit{performingActors} defined in \textit{Task} concept (\textbf{P13}).
+An instance of \textit{Artifact}, defined as a specialization of \textit{MultiInstance} concept, references a set of \textit{ArtifactType} through the specialized \textit{/types} relation (\textbf{P14} and \textbf{P16}).
+Likewise, concept \textit{Actor} is defined as a specialization of \textit{MultiInstance} concept and references a set of \textit{ActorType} through the specialized \textit{/types} relation (\textbf{P15}).
+
+Authorization for an actor to perform a task (\textbf{P17}) is captured either through the relations \textit{allowedActors} and \textit{allowedActorTypes} defined in \textit{TaskType}.
+This mechanism is completed by following behaviours \textit{isAuthorizedActor(Actor)}, \textit{isValidActor(Actor)} and \textit{isValidActorType(ActorType)} defined in \textit{TaskType} :
+
+\begin{lstlisting}[breaklines=true, language=java, basicstyle=\ttfamily\scriptsize, mathescape=true]
+concept TaskType extends Type {
+ ...
+ // Check that an Actor is authorized to perform a task, according to allowed Actor and ActorTypes
+ boolean isAuthorizedActor(Actor actor) {      
+  for (actType : allowedActorTypes) {        
+   if (actor.hasActorType(actType))         
+    return this.isValidActorType(actType);          
+  }        
+  for (act : allowedActors) {        
+   if (actor == act)         
+    return this.isValidActor(actor);          
+   return false;      
+ }
+
+ // Check that an Actor may perform this TaskType (override when required)
+ boolean isValidActor(Actor actor) {
+  return true;
+ }
+
+ // Check that an ActorType may perform this TaskType (override when required)
+ boolean isValidActorType(ActorType actorType) {
+  return true;
+ }
+ ...
+}
+\end{lstlisting}
+
+\textit{Task} concept basically delegates this authorization to its related \textit{TaskType}, as shown in following FML excerpt:
+
+\begin{lstlisting}[breaklines=true, language=java, basicstyle=\ttfamily\scriptsize, mathescape=true]
+concept Task extends Instance {
+ ...
+ boolean isAuthorizedActor(Actor actor) {      
+  return type.isAuthorizedActor(actor);      
+ }  
+ ...
+}
+\end{lstlisting}
+
+Enforcing those constraints is finally performed by the definition of this invariant in \textit{Task} concept:
+
+\begin{lstlisting}[breaklines=true, language=java, basicstyle=\ttfamily\scriptsize, mathescape=true]
+forEach (actor : performingActors) {
+    assert isAuthorizedActor(actor);
+}
+\end{lstlisting}
+
+Default behaviour states that all actors and actor types are valid for all task types.
+This modelling scheme offers many extension points, by the redefinition of one or some behaviours in inherited concepts (although none were required in the context of XSure insurance use case).
+
+Actor types specialization is captured by the \textit{parentActorTypes} relation defined in \textit{ActorType} (\textbf{P18}).
+This is completed by both the definition of \textit{hasActorType(ActorType)} behaviour in \textit{Actor} concept and recursive behaviour \textit{isOrSpecializes(ActorType)} in \textit{ActorType}:
+
+\begin{lstlisting}[breaklines=true, language=java, basicstyle=\ttfamily\scriptsize, mathescape=true]
+concept ActorType extends Type {
+ ActorType[0..*] parentActorTypes;
+ ...
+ boolean isOrSpecializes(ActorType actorType) {    
+  if (actorType = this)   
+   return true;      
+  for (p : parentActorTypes) {      
+   if (p.isOrSpecializes(actorType)
+    return true;        
+  return false;    
+ }
+ ...
+}
+
+concept Actor extends MultiInstance {
+ ...
+ boolean hasActorType(ActorType actType) {      
+  for (type : types) {
+   if (type.isOrSpecializes(actType)) {
+    return true;
+   }
+  }     
+ }  
+ ...
+}
+\end{lstlisting}
+
+All concepts inherits from \textit{ModelingElement}, which defines a \textit{lastUpdated} attribute with \textit{Date} type, and thus satisfies \textbf{P19} requirement.
 
 \subsection{The Acme software development process}
+
+The challenge describes in a second part a Software engineering process for fictional Acme company.
+Base metamodel as described in previous section is too generic to capture all domain-specific aspects of this use case.
+We choose to complete the architectual hierarchy with a specific metamodel, specific to Acme software development process, as shown in figure \ref{fig:MultilevelArchitecture}.
+\textit{Acme metamodel} inherits and specializes base metamodel, and \textit{Acme model} is defined as an instance of \textit{Acme metamodel}.
 

### 2021-05-19 6-ModelPresentation.tex
Updates from Overleaf

@@
 \end{figure*}
-
-\todo{Augmenter la taille des fontes}
 
@@
 Base metamodel as described in previous section is too generic to capture all domain-specific aspects of this use case.
-We choose to complete the architectual hierarchy with a specific metamodel, specific to Acme software development process, as shown in figure \ref{fig:MultilevelArchitecture}.
+We choose to complete the architectual hierarchy with a specific metamodel, specific to Acme software development process, as shown in figure \ref{fig:AcmeArchitecture}.
 \textit{Acme metamodel} inherits and specializes base metamodel, and \textit{Acme model} is defined as an instance of \textit{Acme metamodel}.
+
+\begin{figure}
+ \centering
+    \includegraphics[width=1.0 \textwidth]{Figures/AcmeArchitecture.pdf}
+     \caption{Acme software development process architecture}
+    \label{fig:AcmeArchitecture}
+\end{figure}
 

### 2021-05-20 6-ModelPresentation.tex
Updates from Overleaf

@@
 
-The challenge describes in a second part a Software engineering process for fictional Acme company.
+The challenge describes in a second part a Software engineering process for a fictional Acme company.
 Base metamodel as described in previous section is too generic to capture all domain-specific aspects of this use case.
@@
 \textit{Acme metamodel} inherits and specializes base metamodel, and \textit{Acme model} is defined as an instance of \textit{Acme metamodel}.
+Any instance of Acme model is either an instance of a concept defined in base metamodel, or a concept defined in specialized Acme metamodel, which may or not inherit from a concept defined in base metamodel.
+% FML follows a classical object-oriented semantics. ???
 
@@
  \centering
-    \includegraphics[width=1.0 \textwidth]{Figures/AcmeArchitecture.pdf}
+    \includegraphics[width=1.0 \columnwidth]{Figures/AcmeArchitecture.pdf}
      \caption{Acme software development process architecture}
@@
 
+\begin{figure*}
+ \centering
+    % \includegraphics[width=1.0 \columnwidth]{Figures/Metamodel.pdf}
+    \includegraphics[width=1.0 \textwidth]{Figures/AcmeMetamodel.pdf}
+     \caption{Process management base metamodel}
+    \label{fig:BaseMetamodel}
+\end{figure*}
+
 \subsection{Openflexo tooling}

### 2021-05-20 5-Analysis.tex
Updates from Overleaf

@@
+During the requirement analysis of the challenge, the general model federation approach remains the foundation of our reflections and modeling intentions.
+
+First, the federation approach is mainly based on modeling relationships among several models, independently of abstraction levels and model architectures.
+The next step in our approach, we try to take into account the reusability of the relationships by identifying the meaning, or semantics, associated with each relationship, in a goal of concept reification with their behavior.
+The last step is to organize or structure these concepts to improve reusability and extensibility, to create VirtualModels in our Openflexo framework.
+
+Like any modeling approaches, these different steps could also be achieved in any order.
+But our goal during the challenge requirement analysis remains to lead to define a FML Virtual Models architecture as a federation of virtualmodels.
+
+~\\
+~\\
+
 any disambiguations of the  case  description and assumptions made, any potentially added requirements
@@
 
-Joel commence
+%Joel commence
+
+% Sylvain
+% Joel, je te propose les deux figures suivantes:
+
+\begin{figure}
+    \centering
+    \includegraphics[width=1.0 \columnwidth]{Figures/Instantiation.pdf}
+    \caption{Linguistic and ontologic instantiation}
+    \label{fig:LinguisticAndOntologicInstantiation}
+\end{figure}
+
+\begin{figure}
+    \centering
+    \includegraphics[width=0.7 \columnwidth]{Figures/MultilevelArchitecture.pdf}
+    \caption{Multilevel architecture for the Process Challenge}
+    \label{fig:MultilevelArchitecture}
+\end{figure}

### 2021-05-21 5-Analysis.tex
Updates from Overleaf

@@
 But our goal during the challenge requirement analysis remains to lead to define a FML Virtual Models architecture as a federation of virtualmodels.
+
+Also as our understanding, the use case analysis leads to identify two main modeling axis (as presented in figure \ref{fig:LinguisticAndOntologicInstantiation}):
+\begin{itemize}
+    \item the horizontal axis is characterized as the an ontological instantiation axis, in the sense that the domain type definition (ProcessType as example) is referenced by an instance definition (Process, a.e).
+In our base  VirtualModel, each domain type definition is referenced by its instance definition, as developed in the section \ref{sec:ProcessEnactmen}
+
+\item the vertical axis is viewed as the linguistic instantiation, relative to the use of the FML language.
+The VirtualModel, defined as the core definition model, generates a model founded on the instantiation mechanism of the FML Language.
+This mechanism is the classical Object/Instance mechanism of the object languages but without any constraint on the referenced VirtualModel.
+The set of resulted instances comes from several VirtualModels
+from any abstraction level as detailed in section \ref{sec:AcmeSoftwareDevelopmentProcess}
+
+\end{itemize}
+
+% il reste à décrire l'architecture générale des VirtualModels
 

### 2021-05-21 2-Abstract.tex
OpenFlexo -> Openflexo

@@
 We present a solution based on model federation and discuss the advantages and limitations of using this approach for multi-level modeling.
-Our solution is fully implemented with the OpenFlexo framework and its internal Flexo Modeling Language (FML).
+Our solution is fully implemented with the Openflexo framework and its internal Flexo Modeling Language (FML).
 It fulfills all the challenge requirements.
@@
 afactuelle selon moi.}
+\noteSylvain{Ce qui m'embête un peu avec cet abstract c'est qu'on met en avant la fédération de modèles alors que sur ce cas précis on en fait pas vraiment (sauf pour la partie outillage avec les éditeurs diagrammatiques).
+Concernant la fédération de modèle proprement dite, je crois qu'il faut qu'on montre qu'on a besoin d'une couche conceptuelle (qu'on relie à l'espace technique via les connecteurs techniques, les model slots et les rôles).
+C'est ce besoin qu'adresse le language FML.
+On montre ici plutôt que le language FML - utilisé comme language conceptuel - est assez intéressant, assez expressif pour capturer les exigences du challenge d'une part, et va jusqu'à permettre l'exécution des processus.
+Et cerise sur le gâteau: on peut construire un outillage graphique pour editer et exécuter ce use case.}
+\noteSylvain{En revanche, on peut arguer que pouvoir au même niveau manipuler modèle et métamodèle nous simplifie grandement les choses pour adresser des problématiques multi-niveaux > un bout de la conclusion ?}.

### 2021-05-21 8-Discussion.tex

--- 2021-05-05 8-Discussion.tex
@@
-discussing choices  made (ex : héritage vs attr), pointing out potential compromises/deficiencies
-jc, fabien ; 
-pas d'explicitation de niveaux de « concrétude » : conséquences ? facile de tracer des liens entre les niveaux ; difficile d'avoir le niveau ; 
-outil adaptable, niveau d'abstraction, on fournit une solution + l'outil pour exécuter le processus ; réutilisabilité des modèles ; 
-\todo{on pourrait aussi répondre ici aux 3 questions posées explicitement (copiées dansl'intro pour le moment)}
+\noteJC{notes en vrac
+  \begin{itemize}
+    \item qui ? jc, fabien
+    \item l'appel contraint plus ou monis la structure et le contenu de cette
+      section
+  \begin{itemize}
+    \item discussing choices  made (ex : héritage vs attr), pointing out potential compromises/deficiencies ;
+    \item \emph{Challenge respondents must discuss their multilevel
+model solution with regard to the following aspects, each
+of which should be treated in a specific sub-section of the
+“Assessment” section of the article}
+  \end{itemize}
+    \item pas d'explicitation de niveaux de « concrétude » : conséquences ? facile de tracer des liens entre les niveaux ; difficile d'avoir le niveau ;
+    \item outil adaptable
+    \item niveau d'abstraction
+    \item réutilisabilité des modèles ; 
+    \item ajouter quelques métriques "gros grains"
+    \item difficultés/facilités : contraintes faites à la main par rapport aux
+      langages où on peut spécifier le niveau d'extension lorsque l'on trace un
+      lien multi niveaux -> outil plus souple, mais moins de comportements
+      automatisés
+    \item + ouvert
+    \item moins de support (faudrait implémenter ce qui manque)
+    \item requiert de la méthode, ajouter des comportements, etc.
+    \item point saillant :  outillage "gratuit" = on fournit une solution + l'outil pour exécuter le processus
+\end{itemize}}
 
-difficulctés/facilités
-points saillants :  outillage "gratuit" ;
+\todo[inline]{on pourrait aussi répondre ici aux 3 questions posées
+explicitement (copiées dansl'intro pour le moment), a priori non, le chef
+trouve que c'est chiant}
 
-Although our tool was not designed with a focus on multi-level modelling, we were able to propose a fully working solution.
+%
+%%niveaux de ``concrétude'' -> level-blind solution
+%
+%Although our tool was not designed with a focus on multi-level modelling, we were able to propose a fully working solution. 
+%
+%Our approach is level-blind, therefore it is not easy to quickly retrieve the abstraction level of a concept.
+%
+%%métriques gros grains ?
+%
+%
+%%choices, compromises/deficiences
+%During the modeling phase, we had to choose 
+%(for instance requirement xx 
+%
+%%+/-
+%
+%%points saillants
 
-Our approach is level-blind, therefore it is not easy to quickly retrieve the abstraction level of a concept.
+  %plan plus ou moins imposé par l'appel
+\todo[inline]{mandatory discussion aspects (on verra plus tard si on fait des
+sous-sections ou pas) ↓}
+
+  \subsection{Basic modeling constructs}
+  %Explain the basic modeling constructs used in the solution.
+
+  \subsection{Levels}
+  %Levels (or other model content organization schemes employed)
+
+  %Explain the nature of “levels” in the model, how model elements are arranged
+  %on these levels and which relationships (such as “instance-of”) may feature
+  %between elements at different levels. The nature of levels should be
+  %captured by explicitly stating the level segregation and the level cohesion
+  %principles used [5]. Avoid vague language such as “higher level concepts are
+  %more abstract” if the inter-level relationship is more specific. If the
+  %inter-level relationship is deliberately allowed to be vague, state this
+  %explicitly.
+
+  The Openflexo approach is level-blind: ``levels'' have no specific
+  nature and there are no numbered levels.
+
+  \subsection{Number of levels}
+  %Elaborate whether the submitted solution could have had more or fewer levels
+  %and explain how any potentially existing degrees of freedom were resolved.
+
+  Due to the fact our approach is level-blind, our solution could have had more
+  or fewer levels depending on the variations of the use case. %As levels are
+  %not specifically encoded,
+
+  \subsection{Cross-level relashionships}
+
+  %Discuss if and how associations and links can connect model elements at
+  %different levels. State well-formedness constraints, if any apply.
+
+  Thanks to the level-blindness nature of the Openflexo approach, cross-level
+  relashionships are not an issue. Model elements of different levels can be
+  linked each other in a transparent way. %If the user has defined
+  %well-formedness constraints, then they are
+
+  \subsection{Cross-level constraints}
+
+%Discuss if and how constraints can span multiple levels, especially with
+%regard to cross-level relationships.
+
+  As for cross-level relationships, cross-level constraints do not have a
+  specific nature. They are like any constraint a user can define on a model
+  element. Therefore, if a constraint is mandatory when establishing a
+  relationship between elements from different levels, the user has create it
+  manually. This can be a limit of our approach: the flexibility of our tooling
+  at the cost of a limited number fo automated behaviors.
+
+  \subsection{Integrity mechanism}
+
+  %Discuss how the integrity of level contents is preserved when changes to
+  %level contents occur.
+
+  In order to preserve the integrity of level contents when changes to level
+  contents occur, a validation has to be triggered. The tool reports all
+  inconsistencies that the user is free to fix.
+
+  \subsection{Deep characterization}
+
+  %Discuss if and how higher levels influence elements at lower levels with a
+  %level distance of two or more. Such an influence may be desired to ensure
+  %properties of lower level elements regardless of the design choices that
+  %modelers make at intermediate levels, including future extensions to
+  %intermediate levels.
+
+  \noteJC{pas trop clair pour moi pour le moment}
+
+  \subsection{Generality}
+
+  %Discuss the generality of the solution. Is (part of) it applicable to other
+  %domains? Does it embody invariant principles of the domain(s) it covers with
+  %minimal redundancy?
+
+  Due to the nature of model federation, the models are highly reusable. Our
+  solution separates the domain-specific elements from general purpose ones,
+  making easily possible to apply it to other domains. 
+  \noteJC{à développer + ajouter avantage de l'outil gratuit ici ?}
+
+  \subsection{Extensibility}
+
+  %Elaborate how the solution would respond to further requirements, such as
+  %further special tasks that must be taken care of by special actors. Identify
+  %expected extension points in the solution, e.g., subtyping opportunities. If
+  %level insertion is a possibility in your chosen approach, explain how it
+  %would be performed.
+
+  A strength of the approach of model federation we have adopted resides in its
+  flexibility. One can extend a solution easily:
+
+  \noteJC{comment étendre : dépend un peu des extensions, ne prendre que celle
+  donnée en exemple ? ; level insertion possible : oui, vu que pas de notion de
+level}
+
+\todo[inline]{recommended discussion aspects ↓}
+\subsection{formalisms and tools ?}
+
+%Indicate whether there are formalisms to establish the semantics of the MLM
+%technique and/or tools that support the present solution
+
+\noteJC{on a du tooling, un peu moins de formalisme}
+
+\subsection{model verification}
+
+%Discuss model verification (e.g., consistency analyses) or other quality
+%assessment mechanisms supported by the MLM technique employed.

### 2021-05-21 9-Relatedworks.tex
OpenFlexo -> Openflexo

@@
 
-\cite{multiecore2019} use MultiEcore \citep{multecore2016} to provide a solution to the challenge.
+\cite{multiecore2019} used MultiEcore \citep{multecore2016} to provide a solution to the challenge.
 MultEcore uses cascading and potency.
@@
 
-More similar to us, \cite{deeptelos2019} use DeepTelos\citep{deeptelos2016}, an extension of the Telos language \citep{telos1990} in order to solve the process challenge.
+More similar to us, \cite{deeptelos2019} used DeepTelos\citep{deeptelos2016}, an extension of the Telos language \citep{telos1990} in order to solve the process challenge.
 DeepTelos is based on the use of the PowerType \citep{atkinson2001essence} pattern.
-As with OpenFlexo and FML, they do not use explicit (numbered) levels nor potency.
+As with Openflexo and FML, they do not use explicit (numbered) levels nor potency.
 
@@
 
-\cite{dmla2019} Use the Dynamic Multi-Layer Algebra (DMLA)\citep{dmla2017} in order to solve the process challenge.
+\cite{dmla2019} used the Dynamic Multi-Layer Algebra (DMLA)\citep{dmla2017} in order to solve the process challenge.
 DMLA is a level-blind modeling framework which is fully customizable (e.g., different types of instantiaton may be implemented).

### 2021-05-21 6-ModelPresentation.tex
Updates from Overleaf

@@
 \subsubsection{Process enactment}
-
+\label{sec:ProcessEnactment}
 We now present the process enactment part of base metamodel, located right of figure \ref{fig:BaseMetamodel}.
@@
 \subsection{The Acme software development process}
+\label{sec:AcmeSoftwareDevelopmentProcess}
 
@@
 
+Figure \ref{fig:AcmeFullArchitecture} shows the full architecture capturing Acme software development process, with an instance of enacted Software Engineering Process, and highlights all conceptual levels.
+
 \begin{figure*}
  \centering
-    % \includegraphics[width=1.0 \columnwidth]{Figures/Metamodel.pdf}
-    \includegraphics[width=1.0 \textwidth]{Figures/AcmeMetamodel.pdf}
-     \caption{Process management base metamodel}
-    \label{fig:BaseMetamodel}
+     \includegraphics[width=1.0 \textwidth]{Figures/AcmeFullArchitecture.pdf}
+     \caption{Acme software development process architecture}
+    \label{fig:AcmeFullArchitecture}
 \end{figure*}

### 2021-05-25 8-Discussion.tex
more stuff in assessment section

@@
   element. Therefore, if a constraint is mandatory when establishing a
-  relationship between elements from different levels, the user has create it
-  manually. This can be a limit of our approach: the flexibility of our tooling
-  at the cost of a limited number fo automated behaviors.
+  relationship between elements from different levels, the user has to create
+  it manually. This can be a limit of our approach: the flexibility of our
+  tooling at the cost of a limited number of automated behaviors.
 
@@
   contents occur, a validation has to be triggered. The tool reports all
-  inconsistencies that the user is free to fix.
+  inconsistencies that the user is free to fix. Note that in our approach, the
+  metamodels are built in a ad hoc way  together with the models
+  (co-construction), threfore the user validates on a ad hoc basis.
 
@@
 
-  Due to the nature of model federation, the models are highly reusable. Our
-  solution separates the domain-specific elements from general purpose ones,
-  making easily possible to apply it to other domains. 
-  \noteJC{à développer + ajouter avantage de l'outil gratuit ici ?}
+  Due to the nature of model federation, the models are highly reusable.
+  Although we build ad hoc metamodels in our approach, our solution separates
+  the domain-specific elements from general purpose ones, making easily
+  possible to apply it to other domains.
 
@@
   A strength of the approach of model federation we have adopted resides in its
-  flexibility. One can extend a solution easily:
+  flexibility. As we build metamodels and models together instead of fitting
+  into a metamodel, our solution is more flexible. Therefore one can extend a
+  solution easily. The challenge itself can be seen as a validation of the
+  extensibility ability of our solution: it was decomposed into two steps
+  (generic process Xsure, then ACME process) which is kind of simulation of the
+  evolution of the  specification. We observed that our approach has been
+  resilient to change.
 
-  \noteJC{comment étendre : dépend un peu des extensions, ne prendre que celle
-  donnée en exemple ? ; level insertion possible : oui, vu que pas de notion de
-level}
+  Due to the level-blind aspect of our approach, we could insert a new level,
+  for instance. It would consist in creating new concepts we would link to
+  other ones. Then, we would add any necessary constraints on those concepts
+  and on the relashiontips linking them to the other existing concepts.
+
+  \noteJC{comment étendre : prendre l'exemple donné ?; level insertion possible
+  : oui, vu que pas de notion de level}
 
@@
 
+Currently, there are no formalism to establish the semantics of the MLM
+technique in our solution. \textcolor{red}{To develop…}
+
+Our solution is fully tooled by a ``free'' ad hoc tool. We use the Openflexo
+infrastructure to design models and metamodels together. In addition, our
+infrastructure also produces a dedicated tool which allows processes execution.
+In our approach,
+
 \subsection{model verification}

### 2021-05-25 2-Abstract.tex
Updates from Overleaf

@@
 Et cerise sur le gâteau: on peut construire un outillage graphique pour editer et exécuter ce use case.}
-\noteSylvain{En revanche, on peut arguer que pouvoir au même niveau manipuler modèle et métamodèle nous simplifie grandement les choses pour adresser des problématiques multi-niveaux > un bout de la conclusion ?}.
+\noteSylvain{En revanche, on peut arguer que pouvoir au même niveau manipuler modèle et métamodèle nous simplifie grandement les choses pour adresser des problématiques multi-niveaux > un bout de la conclusion}.
+\noteSM{En fait l'idee de l'abstract c'est juste de nomer les approche utilisé. Les détails sont decrites dans la section 2}

### 2021-05-25 7-RequirementsSatisfaction.tex
Updates from Overleaf

@@
-demonstration of how the solution satisfies the challenge requirements
+Now we discuss how the solution satisfies the challenge requirements.
+
+\noteSylvain{
+Je voulais dire ici qu'on avait construit la solution pas à pas en suivant les requirements les uns après les autres, et qu'on avait à chaque fois veillé à traduire et respecter les requirements.
+Ca me semble très lourd de tout rejustifier ici.
+
+On peut cependant expliciter les différentes techniques employées:
+\begin{itemize}
+    \item Requirements satisfaits par conformance syntaxique: obligation de respecter le modèle structurel
+    \item Requirements satisfaits par ajout de contraintes: on peut instantier quelque chose de syntaxiquement correct, mais qui viole certaines contraintes (sur l'outillage les croix rouges)
+    \item Requirements satisfaits par l'outillage: ajout de comportements de manipulation du modèle qui "forcent" à respecter les contraintes d'intégrité resultant des requirements.
+\end{itemize}
+
+Peut-être peut-on faire un tableau qui rescence comment on a satisfait chaque contrainte ? Est ce que c'est intéressant ?
+}
+
+Sylvain

### 2021-05-25 3-Introduction.tex

@@
 TODO
+MDE has traditionally adopted a two-level,top-down meta-modelling approach, where meta-models reside in a certain meta-level, and models are created one meta-level below by using types from the metamodel (cite MOF and EMF).
 
-\cite{multiProcessChallenge2019}
+This strict approach shows its limitations when modeling complex domains which require more than one level of specialization, e.g., to adapt the model to application sub domains.
+Indeed, the two-level approach fails to acknowledge the existence of two different forms of classification (ontological vs linguistic) and the duality class-object, leading to the introduction of accidental complexity when dealing with the aforementioned complex domains.
 
-\emph{Submissions responding to the challenge should describe
-a multi-level model conforming to the case description,
-including justifications for non-trivial design decisions. In
-order to foster comparability between solutions, respondents
-are asked to make sure that concepts of the case description
-are explicitly represented by one or more model elements.}\\
+In order to tackle this problem multi-level modeling have been introduced.
+Contrary to the strict approach, multi-level modeling advocates the use of a flexible number of levels as well as a flexiblzation of the instantiation concept.
 
-\emph{criteria:
-\begin{enumerate}
-    \item Does the response address the established domain as described in the challenge description and demonstrate the use of modeling features?
-    \item Does it evaluate/discuss its proposed modeling solution against the criteria described in the challenge description?
-    \item Does it discuss the merits and limitations of the employed modeling technique?
-    Authors are invited to suggest further requirements to demonstrate particular benefits of the approaches they adopt.
-\end{enumerate}
-}
+This approach is gaining interest lately within the modeling community as shown by the MULTI level-modeling workshop and the daghtul seminar and number of tools.
+Taking advantage of this vibrant state of affairs, and in order to foster discussion and enable comparison between competing approaches, a multi-level modeling challenge has been created.
+This paper is a response to the latest multi-level modeling challenge which consist of modeling processes.
+We do so by using model federation and the OpenFlexo framework.
 
-structure : plus ou moins imposée par le challenge (structure actuelle = sections au moins attendues)
+Model federation is a modeling approach in which...
+
+Our solution, which is fully implemented and executable, meets all the requirements.
+
+The rest of the paper is organized as follows.
+
+% \cite{multiProcessChallenge2019}
+
+% \emph{Submissions responding to the challenge should describe
+% a multi-level model conforming to the case description,
+% including justifications for non-trivial design decisions. In
+% order to foster comparability between solutions, respondents
+% are asked to make sure that concepts of the case description
+% are explicitly represented by one or more model elements.}\\
+
+% \emph{criteria:
+% \begin{enumerate}
+%     \item Does the response address the established domain as described in the challenge description and demonstrate the use of modeling features?
+%     \item Does it evaluate/discuss its proposed modeling solution against the criteria described in the challenge description?
+%     \item Does it discuss the merits and limitations of the employed modeling technique?
+%     Authors are invited to suggest further requirements to demonstrate particular benefits of the approaches they adopt.
+% \end{enumerate}
+% }
+
+% structure : plus ou moins imposée par le challenge (structure actuelle = sections au moins attendues)

### 2021-05-25 6-ModelPresentation.tex

@@
 This invariant should be completed with additional constraints defined in \textit{Task} \textit{FlexoConcept}, which apply to performing actors actually assigned to enacted tasks.
+
+\todo{Gérer également la fin de P9, non traité pour le moment: artifact they produced must be associated with a validation task}
 
@@
 
-Figure \ref{fig:AcmeFullArchitecture} shows the full architecture capturing Acme software development process, with an instance of enacted Software Engineering Process, and highlights all conceptual levels.
-
 \begin{figure*}
@@
 
+\todo{Augmenter la taille des fontes de la figure \ref{fig:AcmeFullArchitecture}}
+
+Figure \ref{fig:AcmeFullArchitecture} shows the full architecture capturing Acme software development process, with an instance of enacted \textit{Software Engineering Process}, and highlights all conceptual levels.
+Acme metamodel enriches base metamodel by offering specializing concepts: \textit{SETaskType} extending \textit{TaskType}, a more specialized concept \textit{CodingTaskType} extending \textit{SETaskType}, \textit{SEArtifactType} extending \textit{ArtifactType}, a more specialized concept \textit{CodeArtifactType} extending \textit{SEArtifactType}.
+Acme metamodel also defines task type \textit{Coding} as a specialized concept of \textit{Task}, and provides concept \textit{Developer} extending \textit{ActorType}.
+This metamodel is completed with \textit{ProgrammingLanguage} concept, defined as an enumeration (\textit{Java}, \textit{C}, \textit{COBOL}).
+All instances required to capture challenge use case are still defined in final Acme model itself instance of Acme metamodel.
+Instances are represented with rounded boxes and linguistic instantiation are represented with dashed connectors.
+
+Figure \ref{fig:AcmeSoftwareDevelopmentProcess} shows a screen-capture from the tooling developed in the context of the challenge, and presents the \textit{Software Development Process} for Acme company.
+
+\begin{figure}
+ \centering
+    \includegraphics[width=1.0 \columnwidth]{Figures/SoftwareEngineeringProcessCroped.pdf}
+     \caption{Acme software development process}
+    \label{fig:AcmeSoftwareDevelopmentProcess}
+\end{figure}
+
+\todo{Supprimer ombres et dégradés de la figure \ref{fig:AcmeSoftwareDevelopmentProcess}, et faire en sorte qu'elle soit plus haute et moins large}
+
+\textit{"Requirements analysis"} is defined as an instance of \textit{SETaskType}, \textit{Analyst} as an instance of \textit{ActorType} and \textit{Requirement specifications} as an instance of \textit{SEArtifactType}.
+\textit{"Requirements analysis"} instance defines property values \textit{producedArtifactTypes=\{Requirement specifications\}} and \textit{allowedActorTypes=\{Analyst\}} (\textbf{S1}).
+In the same way, \textit{"Test case design"} is defined as an instance of \textit{SETaskType}, marked as critical with property value \textit{isCritical=\{true\}}, and bound to \textit{Analyst} with \textit{allowedActorTypes} relation.
+\textit{"Test case design"} produces \textit{"Test cases"} defined as an instance of \textit{SETaskType}.
+The latter is used in "Test design review" (\textit{producedArtifactTypes} relation) (\textbf{S1} and \textbf{S13}, satisfied with \textbf{P9}).
+
+The capture of \textbf{S3} requirement is performed while defining \textit{Coding} as an instance of concept \textit{CodingTaskType} extending \textit{SETaskType} and defining a relation \textit{languages} to \textit{ProgrammingLanguage} with 1..* cardinality.
+\textit{newTask(String)} behaviour overrides generic implementation while instantiating a \textit{Coding} concept, specializing \textit{SEArtifact}, as shown in following excerpt of FML code:
+
+\begin{lstlisting}[breaklines=true, language=java, basicstyle=\ttfamily\scriptsize, mathescape=true]
+concept CodingTaskType extends SETaskType 
+{
+  ProgrammingLanguage[1..*] languages;
+  ...
+  public Process newTask(String name) {    
+    return new Coding(name,this); 
+  }
+}    
+\end{lstlisting}
+
+\textit{Code} concept extends \textit{SEArtifact}, and defines a relation \textit{language} to \textit{ProgrammingLanguage} with single cardinality (\textbf{S4}).
+\textbf{S5} is more ambigous as task type \textit{Coding} defines one or more programming languages but produces code which is expressed in one language only.
+Semantics is unclear relatively to instantiation policy for artifacts.
+We assume that a Code artifact must be produced for each programming language defined in a \textit{CodingTaskType}.
+If \textit{COBOL} is referenced in languages defined for a given Coding instance, then we must instantiate a \textit{Code} instance with \textit{COBOL} language (\textbf{S5} and \textbf{S6}).
+
+\begin{lstlisting}[breaklines=true, language=java, basicstyle=\ttfamily\scriptsize, mathescape=true]
+concept CodingTaskType extends SETaskType 
+{
+  ...
+  public Artifact[] instantiateArtifacts() {
+    Artifact[] returned = {}; 
+    for (langage : languages) {
+      returned.add(new Code(langage); 
+    }
+    return returned;
+  }
+  ...
+}    
+\end{lstlisting}
+
 \subsection{Openflexo tooling}

### 2021-05-25 3-Introduction.tex
Updates from Overleaf

@@
-TODO
-MDE has traditionally adopted a two-level,top-down meta-modelling approach, where meta-models reside in a certain meta-level, and models are created one meta-level below by using types from the metamodel (cite MOF and EMF).
+Model-driven engineering (MDE) has traditionally adopted a two-level, top-down approach, where meta-models reside in a certain meta-level, and models are created one meta-level below by using types from the metamodel (e.g., EMF - \cite{emf} and MOF - \cite{omg2013mof}).
 

### 2021-05-25 6-ModelPresentation.tex
Updates from Overleaf

@@
 Semantics is unclear relatively to instantiation policy for artifacts.
-We assume that a Code artifact must be produced for each programming language defined in a \textit{CodingTaskType}.
+We assume that task execution implies that for each used artifact type defined in related task type, at least one artifact
+
+Code artifact must be produced for each programming language defined in a \textit{CodingTaskType}.
 If \textit{COBOL} is referenced in languages defined for a given Coding instance, then we must instantiate a \textit{Code} instance with \textit{COBOL} language (\textbf{S5} and \textbf{S6}).
@@
 
+\begin{figure*}
+ \centering
+     \includegraphics[width=1.0 \textwidth]{Figures/ToolingArchitecture.pdf}
+     \caption{Tooling architecture}
+    \label{fig:ToolingArchitecture}
+\end{figure*}
+
 % Expliquer l'outillage construit et comment il marche

### 2021-05-25 8-Discussion.tex
Updates from Overleaf

@@
 \todo[inline]{on pourrait aussi répondre ici aux 3 questions posées
-explicitement (copiées dansl'intro pour le moment), a priori non, le chef
+explicitement (copiées dans l'intro pour le moment), a priori non, le chef
 trouve que c'est chiant}

### 2021-05-25 8-Discussion.tex
update

--- 2021-05-25 8-Discussion.tex
@@
 \todo[inline]{on pourrait aussi répondre ici aux 3 questions posées
-explicitement (copiées dansl'intro pour le moment), a priori non, le chef
-trouve que c'est chiant}
+explicitement (copiées dans l'intro pour le moment), a priori non, le chef
+« trouve que c'est chiant »}
 
@@
 Our solution is fully tooled by a ``free'' ad hoc tool. We use the Openflexo
-infrastructure to design models and metamodels together. In addition, our
-infrastructure also produces a dedicated tool which allows processes execution.
-In our approach,
+infrastructure to build models and metamodels together. In addition to the
+solution, it also provides a dedicated tool which allows processes execution. 
+\noteJC{dire qqch sur le fait que produire une solution avec son outil associé
+  est une force, genre ``Providing a solution and its associated tool can be
+…'' mais pour le moment je ne vois pas comment l'enchainer}
 
-\subsection{model verification}
+\subsection{Model verification}
 
@@
 %assessment mechanisms supported by the MLM technique employed.
+
+By default, our tooling does not include a lot of automated verifications.
+However, a user can add verification mechanisms by writing new behaviors
+attached to concepts. Presently, the verifications rely on the users when they
+build the metamodels and the models.

### 2021-05-25 3-Introduction.tex

--- 2021-05-25 3-Introduction.tex
@@
-TODO
-MDE has traditionally adopted a two-level,top-down meta-modelling approach, where meta-models reside in a certain meta-level, and models are created one meta-level below by using types from the metamodel (cite MOF and EMF).
+Model-driven engineering (MDE) has traditionally adopted a strict hierarchical two-level approach, where metamodels reside in a certain meta-level, and models are created one meta-level below by using types from the metamodel.
+Notable examples of this approach are the widespread Eclipse Modeling Framework (EMF)~\citep{emf} and the Object Management Group Meta-Object Facility(MOF)~\citep{omg2013mof}.
 
-This strict approach shows its limitations when modeling complex domains which require more than one level of specialization, e.g., to adapt the model to application sub domains.
-Indeed, the two-level approach fails to acknowledge the existence of two different forms of classification (ontological vs linguistic) and the duality class-object, leading to the introduction of accidental complexity when dealing with the aforementioned complex domains.
+This strict approach shows its limitations when modeling complex domains requiring more than one level of specialization, e.g., to adapt the model to application sub domains.
+Indeed, the two-level approach fails to acknowledge and support: 1) the existence of different forms of classification and/or instantiation (e.g., ontological vs linguistic); and 2) the duality type-object for model elements.
+Therefore, while it may still be used for the aforementioned complex modeling tasks, the strict approach entails the introduction of accidental complexity in both the modeling process and the resulting modeling artefacts.
 
-In order to tackle this problem multi-level modeling have been introduced.
-Contrary to the strict approach, multi-level modeling advocates the use of a flexible number of levels as well as a flexiblzation of the instantiation concept.
+In order to tackle this problem, the \emph{multi-level} modeling paradigm have been introduced.
+Contrary to the strict approach, multi-level modeling advocates the use of a flexible number of levels as well as a flexibilization of the relations between levels.
 
-This approach is gaining interest lately within the modeling community as shown by the MULTI level-modeling workshop and the daghtul seminar and number of tools.
+This approach is gaining interest lately within the modeling community as shown by the MULTI level-modeling workshop and the daghtul seminar~\citep{almeida2018multi} and number of tools.
 Taking advantage of this vibrant state of affairs, and in order to foster discussion and enable comparison between competing approaches, a multi-level modeling challenge has been created.
@@
 
-Model federation is a modeling approach in which...
+Model federation is a modeling approach in which loosely couple reified models.
+Links among models can be within a same level of abstraction or across levels of abstractions which gives, as our answer to the challenge shows, a great flexibility.
 

### 2021-05-25 6-ModelPresentation.tex
Updates from Overleaf

@@
 Likewise, concept \textit{Actor} is defined as a specialization of \textit{MultiInstance} concept and references a set of \textit{ActorType} through the specialized \textit{/types} relation (\textbf{P15}).
+
+Semantics is unclear relatively to instantiation policy for artifacts.
+We assume that task execution implies that for each used and produced \textit{artifact type} defined in related \textit{task type}, it exists at least one \textit{artifact} declaring required ontologic instantiation (through \textit{type} relation).
+Following excerpt of FML code shows partial implementation of this.
+Same patterns applied for produced artifacts.
+
+\begin{lstlisting}[breaklines=true, language=java, basicstyle=\ttfamily\scriptsize, mathescape=true] 
+concept Task extends Instance 
+{
+  ...
+  boolean declaresRequiredUsedArtifacts() {
+    for (artifactType : type.usedArtifactType) {
+      boolean found = false;
+      for (artifact : usedArtifacts) {
+        if (artifact.isOfType(artifactType))
+          found = true;
+      }
+      if (!found) return false;
+    }
+    return true;
+  }
+  ...
+}    
+\end{lstlisting}
 
@@
   ...
-  public Process newTask(String name) {    
+  public Coding newTask(String name) {    
     return new Coding(name,this); 
@@
 \textbf{S5} is more ambigous as task type \textit{Coding} defines one or more programming languages but produces code which is expressed in one language only.
-Semantics is unclear relatively to instantiation policy for artifacts.
-We assume that task execution implies that for each used artifact type defined in related task type, at least one artifact
+This requirement is captured through the redefinition of \textit{declaresRequiredProducedArtifacts()} where programming language should also match \noteSylvain{Pas la place pour mettre le code, non nécessaire si ?}.
+\textbf{S6} is guaranteed though the \textit{language} relation defined in \textit{Code} and following invariant declared in \textit{Code}:
+
+\begin{lstlisting}[breaklines=true, language=java, basicstyle=\ttfamily\scriptsize, mathescape=true]
+forEach (artifactType : types) {
+    assert !(artifactType instanceof CodeArtifactType) | artifactType.isSenior
+}
+\end{lstlisting}
+
+%
 
@@
  \centering
-     \includegraphics[width=1.0 \textwidth]{Figures/ToolingArchitecture.pdf}
+     \includegraphics[width=0.9 \textwidth]{Figures/ToolingArchitecture.pdf}
      \caption{Tooling architecture}

### 2021-05-25 8-Discussion.tex
Updates from Overleaf
comments notes, remove crap, typos

@@
-\noteJC{notes en vrac
-  \begin{itemize}
-    \item qui ? jc, fabien
-    \item l'appel contraint plus ou monis la structure et le contenu de cette
-      section
-  \begin{itemize}
-    \item discussing choices  made (ex : héritage vs attr), pointing out potential compromises/deficiencies ;
-    \item \emph{Challenge respondents must discuss their multilevel
-model solution with regard to the following aspects, each
-of which should be treated in a specific sub-section of the
-“Assessment” section of the article}
-  \end{itemize}
-    \item pas d'explicitation de niveaux de « concrétude » : conséquences ? facile de tracer des liens entre les niveaux ; difficile d'avoir le niveau ;
-    \item outil adaptable
-    \item niveau d'abstraction
-    \item réutilisabilité des modèles ; 
-    \item ajouter quelques métriques "gros grains"
-    \item difficultés/facilités : contraintes faites à la main par rapport aux
-      langages où on peut spécifier le niveau d'extension lorsque l'on trace un
-      lien multi niveaux -> outil plus souple, mais moins de comportements
-      automatisés
-    \item + ouvert
-    \item moins de support (faudrait implémenter ce qui manque)
-    \item requiert de la méthode, ajouter des comportements, etc.
-    \item point saillant :  outillage "gratuit" = on fournit une solution + l'outil pour exécuter le processus
-\end{itemize}}
+%\noteJC{notes en vrac
+%  \begin{itemize}
+%    \item qui ? jc, fabien
+%    \item l'appel contraint plus ou moins la structure et le contenu de cette
+%      section
+%  \begin{itemize}
+%    \item discussing choices  made (ex : héritage vs attr), pointing out potential compromises/deficiencies ;
+%    \item \emph{Challenge respondents must discuss their multilevel
+%model solution with regard to the following aspects, each
+%of which should be treated in a specific sub-section of the
+%“Assessment” section of the article}
+%  \end{itemize}
+%    \item[\checkmark] pas d'explicitation de niveaux de « concrétude » : conséquences ? facile de tracer des liens entre les niveaux ; difficile d'avoir le niveau ;
+%    \item outil adaptable
+%    \item niveau d'abstraction
+%    \item[\checkmark] réutilisabilité des modèles ; 
+%    \item ajouter quelques métriques "gros grains"
+%    \item[\checkmark] difficultés/facilités : contraintes faites à la main par rapport aux
+%      langages où on peut spécifier le niveau d'extension lorsque l'on trace un
+%      lien multi niveaux -> outil plus souple, mais moins de comportements
+%      automatisés
+%    \item + ouvert
+%    \item[\checkmark] moins de support (faudrait implémenter ce qui manque)
+%    \item requiert de la méthode, ajouter des comportements, etc.
+%    \item[\checkmark] point saillant :  outillage "gratuit" = on fournit une solution + l'outil pour exécuter le processus
+%\end{itemize}}
 
-\todo[inline]{on pourrait aussi répondre ici aux 3 questions posées
-explicitement (copiées dans l'intro pour le moment), a priori non, le chef
-« trouve que c'est chiant »}
+%\todo[inline]{on pourrait aussi répondre ici aux 3 questions posées
+%explicitement (copiées dans l'intro pour le moment), a priori non, le chef
+%« trouve que c'est chiant »}
 
-%
-%%niveaux de ``concrétude'' -> level-blind solution
-%
-%Although our tool was not designed with a focus on multi-level modelling, we were able to propose a fully working solution. 
-%
-%Our approach is level-blind, therefore it is not easy to quickly retrieve the abstraction level of a concept.
-%
-%%métriques gros grains ?
-%
-%
-%%choices, compromises/deficiences
-%During the modeling phase, we had to choose 
-%(for instance requirement xx 
-%
-%%+/-
-%
-%%points saillants
+\noteJC{pour le moment, seulement réponse aux différents points, on ne voit pas
+de discussion sur les choix, sur les potentiels compromis/déficiencies ;
+sous-sections : bof, p-e plutôt des paragraphes ; très court/concis, faudrait
+étoffer}
 
   %plan plus ou moins imposé par l'appel
-\todo[inline]{mandatory discussion aspects (on verra plus tard si on fait des
-sous-sections ou pas) ↓}
+\todo[inline]{mandatory discussion aspects ↓}
 
@@
   %Explain the basic modeling constructs used in the solution.
+
+  \noteJC{pas clair pour moi ce qui est entendu par « basic modeling constructs »}
 
@@
   metamodels are built in a ad hoc way together with the models
-  (co-construction), threfore the user validates on a ad hoc basis.
+  (co-construction), therefore the user validates on a ad hoc basis.
 
@@
 
-  A strength of the approach of model federation we have adopted resides in its
+  A strength of the  model federation approach we have adopted resides in its
   flexibility. As we build metamodels and models together instead of fitting
@@
 \todo[inline]{recommended discussion aspects ↓}
-\subsection{formalisms and tools ?}
+\subsection{Formalisms and tools ?}
 

### 2021-05-25 6-ModelPresentation.tex

@@
 forEach (artifactType : types) {
-    assert !(artifactType instanceof CodeArtifactType) | artifactType.isSenior
+    assert !(artifactType instanceof CodeArtifactType) | artifactType.implements()
 }

### 2021-05-25 6-ModelPresentation.tex
Updates from Overleaf

--- 2021-05-25 6-ModelPresentation.tex
@@
 forEach (artifactType : types) {
-    assert !(artifactType instanceof CodeArtifactType) | artifactType.isSenior
+    assert !(artifactType instanceof CodeArtifactType) | artifactType.implements(programmingLanguages)
 }

### 2021-05-25 Figures/federation.tex
model federation figure

\documentclass{standalone}

\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{lmodern}
\usepackage{graphicx}

\usepackage{tikz}
\usetikzlibrary{calc,shapes.geometric,arrows.meta,backgrounds}

\begin{document}
  \begin{tikzpicture}[ >=Classical TikZ Rightarrow]
    \draw[use as bounding box] (-4.065,-4.065) rectangle (4.065,4.065);
    \foreach \n/\h in {0/1.2cm,1/1.3cm,2/1.2cm,3/1.1cm,4/1cm}{
      \node[inner sep=0pt] (m\n) at (90+\n*360/5: 3cm) {\includegraphics[height=\h]{house\n}};
      \node[minimum width=3mm,minimum height=2mm,draw=orange,fill=orange!30,inner sep=0pt,rotate=\n*360/5] (ms\n) at (90+\n*360/5:1.18cm) {};
      \draw[<->] (m\n) -- (ms\n);
    }
    \begin{pgfonlayer}{background}
      \foreach \n/\c in {0/black!15,1/black!20,2/black!25,3/black!20,4/black!25}{
        \node[inner sep=1pt,outer sep=0pt,circular sector,circular sector angle=360/5,fill=\c,minimum width=6.8cm,rotate=-90+\n*360/5] (ts\n) at (90+\n*360/5: 2.73cm) {};
      }
      \begin{scope}[blue]
        \node[circle,draw=orange,fill=orange!30,text width=2cm,align=center,inner sep=1mm] {conceptual model};
        \node[anchor=center,text width=2cm,align=center] (ts) at (m0-|m4)  {technological spaces};
        \node[align=center,text width=0.92cm] (ms) at ($(m3)!0.5!(m4)$)  {model slots};
        \node[anchor=south east,text width=1.4cm,align=center] (fm) at ($(m0)!0.5!(m1)$)  {federated models};
        \draw[->,thin] (ts.west) -- +(-0.4,0);
        \draw[->,thin] (ts.south) -- +(0,-0.4);
        \draw[->,thin] (ms) -- (ms3);
        \draw[->,thin] (ms) -- (ms4);
        \draw[->,thin] (fm) -- (m0);
        \draw[->,thin] (fm) -- (m1);
      \end{scope}
    \end{pgfonlayer}
  \end{tikzpicture}
\end{document}

### 2021-05-25 4-Technology.tex
Updates from Overleaf

@@
 % Inspiré de la section 10 du papier formose
+
+\noteFabien{metamodel ou meta-model ? perso je penche plus pour le premier}
+\noteJC{pareil, j'utilise plutôt metamodel, en général}
 
@@
 approach of model federation~\cite{Golra2016-federation}. \emph{Model
-  federation} is a way models are assembled together with some kind of
+  federation} is a way to assemble models using some kind of
 low-coupling links. It has been studied to answer the OMG' RFP
@@
 contrast to approaches that compose meta-models into a single large
-meta-model welcoming all needed entities, model federation build links
+meta-model grouping all needed entities, model federation build links
 among models and meta-models (even through levels) to make ``things''
@@
 serve as control components.
-% Figure~\ref{fig:space-ex} presents a
-% simplified version of the \textsf{Goal-xls} federation example that
-% groups these different models. The upper half of the figure
-% illustrates the design of a federation with a model coming from the
-% \smk technological space (the goal diagram), one from the Excel
-% technological space (the file) and one virtual model, named
-% \emph{ReqXls}, reifying the dependencies between the goal model and
-% the reporting file.
-
-A \emph{technology adapter} (TA) is a reusable library that defines
-connections between the FML execution engine and a particular
-technological space. The model federation framework provides ways to
-define TA.
-%The \emph{ReqXls} virtual model relies on two technology
-%adapters (Excel TA and \smk TA) for manipulating the models in their
-%respective technological spaces.
 
@@
 attributes of the class and its behaviors to the methods of the class.
-% In our example, a concept \texttt{Requirement} is presented with
-% three roles \texttt{row}, \texttt{goal} and \texttt{refines}. The
-% first two are represented as attributes whereas the third as an
-% association.
 These roles have types defining the kind of value the role will point
 at runtime.
-% Our example illustrates three forms of such types \ie another
-% concept for \textsf{refines}, a type defined in a TA for
-% \textsf{Row} or a type defined in an external model for
-% \textsf{Goal}.
 Whenever a type external to the federation space (from a TA or an
@@
 access to external elements of the corresponding technological space.
+
+A \emph{technology adapter} (TA) is a reusable library that defines
+connections between the FML execution engine and a particular
+technological space. The model federation framework provides ways to
+define TA.
 
@@
 for documents, EMF and OWL for modeling languages, JDBC for databases,
-REST and XMLRPC for external services and one for diagraming tools)
+REST and XMLRPC for external services and one for diagramming tools)
 and other more rudimentary (pdf, HTTP, XML, etc.).

### 2021-05-25 6-ModelPresentation.tex
Updates from Overleaf

--- 2021-05-25 6-ModelPresentation.tex
@@
 \textbf{S5} is more ambigous as task type \textit{Coding} defines one or more programming languages but produces code which is expressed in one language only.
-This requirement is captured through the redefinition of \textit{declaresRequiredProducedArtifacts()} where programming language should also match \noteSylvain{Pas la place pour mettre le code, non nécessaire si ?}.
-\textbf{S6} is guaranteed though the \textit{language} relation defined in \textit{Code} and following invariant declared in \textit{Code}:
+This requirement is captured through the redefinition of \textit{declaresRequiredProducedArtifacts()} where programming language should also match.
+\noteSylvain{Pas la place pour mettre le code, non nécessaire si ?} \textbf{S6} is guaranteed though the \textit{language} relation defined in \textit{Code} and following invariant declared in \textit{Code}:
 
@@
 forEach (artifactType : types) {
-    assert !(artifactType instanceof CodeArtifactType) | artifactType.isSenior
+    assert !(artifactType instanceof CodeArtifactType) | artifactType.implements(languages)
 }
@@
 
-%
-
-Code artifact must be produced for each programming language defined in a \textit{CodingTaskType}.
-If \textit{COBOL} is referenced in languages defined for a given Coding instance, then we must instantiate a \textit{Code} instance with \textit{COBOL} language (\textbf{S5} and \textbf{S6}).
+\textit{Ann Smith} is the only one allowed to perform coding in COBOL (\textbf{S7}). This is implemented through the redefinition of \textit{newTask(String)} in \textit{CodingTaskType}:
 
@@
   ...
-  public Artifact[] instantiateArtifacts() {
-    Artifact[] returned = {}; 
-    for (langage : languages) {
-      returned.add(new Code(langage); 
+  public Coding newTask(String name) {
+    Coding returned = new Coding(name,this); 
+    if (languages.contains(ProgrammingLanguage.COBOL)) {
+        returned.addToPerformingActors(getActor("Ann Smith"));
     }
@@
   }
-  ...
 }    
 \end{lstlisting}
+
+\textbf{S8} requirement is simply captured with the definition of \textit{Testing} instance of \textit{SETaskType}, \textit{Tester} instance of \textit{ActorType}, and \textit{Test report} instance of \textit{ArtifactType}.
+% TODO S9
+
+All software engineering artifacts defined in the context of Acme Software Engineering Process (figure \ref{fig:AcmeSoftwareDevelopmentProcess}) are all instances of \textit{SEArtifact} which defines both attributes \textit{responsible} (an \textit{Actor} instance), and \textit{versionNumber} (an \textit{integer} value), and thus fulfills \textbf{S10}.
+\textit{Bob Brown} is declared as an instance of \textit{Actor}, and references \textit{Analyst} and \textit{Tester} (\textit{ActorType} instances).
+He is also referenced by all instances of \textit{SETaskType} as the creator for related task types (\textbf{S11}).
+
+\textbf{S12} is somehow ambigous as it implies an implicit semantics regarding underlying business logic of process execution.
+We assume that all tasks may define an expected duration, which might be checked during process execution.
+This can be modelled through the addition of \textit{expectedDuration} attribute in \textit{TaskType}.
+Alternatively, this can be modelled through the definition of \textit{SETesting} concept, as a specialization of \textit{SETaskType}, and the instantiation of \textit{Testing} as an instance of \textit{SETesting}.
+The business logic expressed by \textbf{S12} requirement should be redefined in \textit{SETesting}.
+
+\textbf{S13} requirement has been previously partially fulfilled. This must be completed with...
+
+% TODO S13
 

### 2021-05-25 7-RequirementsSatisfaction.tex
Updates from Overleaf

@@
 Sylvain
+
+\noteJC{en fait, la section précédente fait déjà le travail, donc ça fait effectivement redandant.
+Dans ce cas, pourquoi le répéter ?
+On peut éventuellement fusionner les deux sections.
+L'appel impose plus ou moins le plan, mais tant que les différents ponits sont présents, ça ne doit pas trop déranger.
+Ton histoire de tableau pourrait être une sorte de résumé de la section précédente.}

### 2021-05-25 3-Introduction.tex

@@
 In order to tackle this problem, the \emph{multi-level} modeling paradigm have been introduced.
-Contrary to the strict approach, multi-level modeling advocates the use of a flexible number of levels as well as a flexibilization of the relations between levels.
+Contrary to the strict approach, multi-level modeling advocates the use of a flexible number of levels as well as a flexibilization of the relations between them.
+This paradigm is gaining traction lately within the modeling community as evidenced by the contribution of many new multi-level modeling approaches and tools and the celebration of topic specific workshops\footnote{https://homepages.ecs.vuw.ac.nz/Groups/MultiLevelModeling/WebHome} and seminars~\citep{almeida2018multi}.
+Taking advantage of this vibrant state of affairs, and in order to foster discussion and enable comparison between competing approaches, a multi-level modeling challenge has been created.
+This paper is a response to the latest multi-level modeling challenge
+which consists in providing a solution to the problem of specifying and enacting processes.
+Solutions to the \emph{Multi-level Process Challenge} must fulfill a number of requirements for a process representation defined at an abstract process-definition level and at various more concrete domain-specific levels, resulting in a multi-level hierarchy of related models.
+We provide a solution based on model federation~\citep{Golra2016-federation}.
 
-This approach is gaining interest lately within the modeling community as shown by the MULTI level-modeling workshop and the daghtul seminar~\citep{almeida2018multi} and number of tools.
-Taking advantage of this vibrant state of affairs, and in order to foster discussion and enable comparison between competing approaches, a multi-level modeling challenge has been created.
-This paper is a response to the latest multi-level modeling challenge which consist of modeling processes.
-We do so by using model federation and the OpenFlexo framework.
-
-Model federation is a modeling approach in which loosely couple reified models.
-Links among models can be within a same level of abstraction or across levels of abstractions which gives, as our answer to the challenge shows, a great flexibility.
-
+Model federation is a multi model management approach based on the use of virtual models and loosely coupled links.
+The models in a federation remain autonomous and represented in their original technological spaces whereas virtual models and links serve as control components used to present different views to the users and maintain synchronization.
+Our solution is based on this model federation infrastructure, namely, the use of virtual models and links.
+Concretely, we use OpenFlexo and its internal Flexo Modeling Language (FML), a language to create, link and manage virtual models, in order to achieve our goal.
 Our solution, which is fully implemented and executable, meets all the requirements.
 
+%Links among models can be within a same level of abstraction or across levels of abstractions which gives,
+
+%as our answer to the challenge shows, a great flexibility.
+
 The rest of the paper is organized as follows.
+Section~\ref{sec:technology} presents our approach and the technology it relies on.
+Section~\ref{sec:analysis} is the analysis of the problem.
+We detail our model in section~\ref{sec:model} and show it satisfies the challenge requirements in section~\ref{sec:requirements}.
+We assess the modeling solution in section~\ref{sec:discussion}.
+Section~\ref{sec:relatedwork} presents the related work before the conclusion~\ref{sec:conclusion}.
 

### 2021-05-25 5-Analysis.tex
Updates from Overleaf

@@
 But our goal during the challenge requirement analysis remains to lead to define a FML Virtual Models architecture as a federation of virtualmodels.
+As shown in figure~\ref{fig:MultilevelArchitecture}, we developped two metamodels (Base and ACME processes abstractions) instantiatied by two models (XSure and ACME processes).
+This follows the way the challenge is presented and this organisation allows for the possibility of easy extensions.
+
+\begin{figure}
+    \centering
+    \includegraphics[width=1.0 \columnwidth]{Figures/Instantiation.pdf}
+    \caption{Linguistic and ontologic instantiation}
+    \label{fig:LinguisticAndOntologicInstantiation}
+\end{figure}
 
@@
 \begin{itemize}
-    \item the horizontal axis is characterized as the an ontological instantiation axis, in the sense that the domain type definition (ProcessType as example) is referenced by an instance definition (Process, a.e).
-In our base  VirtualModel, each domain type definition is referenced by its instance definition, as developed in the section \ref{sec:ProcessEnactmen}
+    \item the horizontal axis is characterized as the ontological instantiation axis, in the sense that the domain type definition (ProcessType as example) is referenced by an instance definition (Process, a.e).
+In our base  VirtualModel, each domain type definition is referenced by its instance definition, as developed in the section \ref{sec:ProcessEnactment}
 
@@
 
-% il reste à décrire l'architecture générale des VirtualModels
-
-~\\
-~\\
-
-any disambiguations of the  case  description and assumptions made, any potentially added requirements
-
-quelles hypothèses supplémentaires ?
-
-+ architecture de la modélisation (ce qui se trouve dans le ppt), le détail se retrouvera dans la section suivante
-
-%Joel commence
-
-% Sylvain
-% Joel, je te propose les deux figures suivantes:
-
-\begin{figure}
-    \centering
-    \includegraphics[width=1.0 \columnwidth]{Figures/Instantiation.pdf}
-    \caption{Linguistic and ontologic instantiation}
-    \label{fig:LinguisticAndOntologicInstantiation}
-\end{figure}
-
 \begin{figure}
@@
 \end{figure}
+
+% il reste à décrire l'architecture générale des VirtualModels
+
+~\\
+~\\
+
+%\emph{any disambiguations of the  case  description and assumptions made, any potentially added requirements}
+
+\todo[inline]{quelles hypothèses supplémentaires ?}
+\todo[inline]{+ architecture de la modélisation (ce qui se trouve dans le ppt), le détail se retrouvera dans la section suivante}
+
+%Joel commence
+
+% Sylvain
+% Joel, je te propose les deux figures suivantes:

### 2021-05-26 5-Analysis.tex
Updates from Overleaf

@@
+%\emph{Analysis: any disambiguations of the  case  description and assumptions made, any potentially added requirements}
+
 During the requirement analysis of the challenge, the general model federation approach remains the foundation of our reflections and modeling intentions.
@@
     \centering
-    \includegraphics[width=1.0 \columnwidth]{Figures/Instantiation.pdf}
-    \caption{Linguistic and ontologic instantiation}
-    \label{fig:LinguisticAndOntologicInstantiation}
+    \includegraphics[width=0.7 \columnwidth]{Figures/MultilevelArchitecture.pdf}
+    \caption{Multilevel architecture for the Process Challenge}
+    \label{fig:MultilevelArchitecture}
 \end{figure}
@@
     \centering
-    \includegraphics[width=0.7 \columnwidth]{Figures/MultilevelArchitecture.pdf}
-    \caption{Multilevel architecture for the Process Challenge}
-    \label{fig:MultilevelArchitecture}
+    \includegraphics[width=1.0 \columnwidth]{Figures/Instantiation.pdf}
+    \caption{Linguistic and ontologic instantiation}
+    \label{fig:LinguisticAndOntologicInstantiation}
 \end{figure}
+
+Based on this approach, we organize our set of models as presented by the figure \ref{fig:MultilevelArchitecture}.
+The resulted multi-level architecture is organized as a set of VirtualModels.
+The first one is the core concepts model including the definition of Processes and Tasks.
+The second one, Acme VirtualModel, is extended from the previous model to integrate the Acme definitions.
+Our FML language defines the multiple inheritance concept between VirtualModels as illustrated in the section \ref{sec:AcmeSoftwareDevelopmentProcess}.
+
+Finally as explained previously, the defined level for the Acme model and the XSure is the result of the instantiation of VirtualModels of different levels.
+In our approach this level can't be specialized or extended but all the other VirtualModels could be extended  by any concepts, as a new federated model.
 
@@
 
-%\emph{any disambiguations of the  case  description and assumptions made, any potentially added requirements}
-
 \todo[inline]{quelles hypothèses supplémentaires ?}
-\todo[inline]{+ architecture de la modélisation (ce qui se trouve dans le ppt), le détail se retrouvera dans la section suivante}
-
-%Joel commence
-
-% Sylvain
-% Joel, je te propose les deux figures suivantes:
+%\todo[inline]{+ architecture de la modélisation (ce qui se trouve dans le ppt), le détail se retrouvera dans la section suivante}

### 2021-05-26 1-Main.tex
Updates from Overleaf

@@
     keepspaces=true,                 
-    numbers=left,                    
+    %numbers=left,                    
     numbersep=5pt,                  
@@
     \author{Jean-Christophe Bach}
-    \address{IMT Atlantique, Lab-STICC, UMR 6285, F-29238 Brest, France}
+    \address{IMT Atlantique, Lab-STICC, UMR 6285, 29238 Brest, France}
     % \ead{firstname.lastname@imt-atlantique.fr}
@@
     \abstract{\input{2-Abstract}}
-    \keywords{Multilevel Modeling Challenge, Federation Modeling Language, OpenFlexo}
+    \keywords{Multilevel Modeling Challenge, Federation Modeling Language, Openflexo}
     \bibliography{biblio.bib}
@@
 
-\noteSM{test}
-\noteJC{test}
-\noteSylvain{test}
-\noteAntoine{test}
-\noteFabien{test}
-\noteJoel{test}
+%\noteSM{test}
+%\noteJC{test}
+%\noteSylvain{test}
+%\noteAntoine{test}
+%\noteFabien{test}
+%\noteJoel{test}
 
@@
 \section{Conclusions}
+\label{sec:conclusions}
 \input{10-Conclusions.tex}

### 2021-05-26 6-ModelPresentation.tex

@@
-% detailed  presentation of a model, including justifications for design decisions
+%Model Presentation: detailed  presentation of a model, including justifications for design decisions
+
+\todo[inline]{phrase d'intro pour expliquer que les req étaient couverts au fur et à mesure (côté systématique)}
 
@@
  boolean isOrSpecializes(ActorType actorType) {    
-  if (actorType = this)   
+  if (actorType == this)   
    return true;      

### 2021-05-26 6-ModelPresentation.tex
Push for Fig 2

@@
 
-\begin{lstlisting}[breaklines=true, language=java, basicstyle=\ttfamily\scriptsize, mathescape=true]
+\begin{lstlisting}
 model MetaModel {
 
-  concept ModelingElement {
-    ...
-  }
+  concept ModelingElement { ... }
 
-  concept Type extends ModelingElement {
-    ...
-  }
+  concept Type extends ModelingElement { ... }
 
@@
     // Other core concepts
-    ...
   }
@@
 
-\begin{lstlisting}[breaklines=true, language=java, basicstyle=\ttfamily\scriptsize, mathescape=true]
+\begin{lstlisting}
 forEach (actor : allowedActors) {
@@
 
-\begin{lstlisting}[breaklines=true, language=java, basicstyle=\ttfamily\scriptsize, mathescape=true]
+\begin{lstlisting}
 public Process newProcess(String name) {    
@@
 
-\begin{lstlisting}[breaklines=true, language=java, basicstyle=\ttfamily\scriptsize, mathescape=true] 
-concept Task extends Instance 
-{
+\begin{lstlisting}
+concept Task extends Instance {
   ...
@@
 
-\begin{lstlisting}[breaklines=true, language=java, basicstyle=\ttfamily\scriptsize, mathescape=true]
+\begin{lstlisting}
 concept TaskType extends Type {
  ...
- // Check that an Actor is authorized to perform a task, according to allowed Actor and ActorTypes
+ // Check that an Actor is authorized to perform a task, using allowed Actor and ActorTypes
   boolean isAuthorizedActor(Actor actor) {      
@@
         return this.isValidActor(actor);
+    }
     return false;
@@
 
-\begin{lstlisting}[breaklines=true, language=java, basicstyle=\ttfamily\scriptsize, mathescape=true]
+\begin{lstlisting}
 concept Task extends Instance {

### 2021-05-26 5-Analysis.tex

@@
 First, the federation approach is mainly based on modeling relationships among several models, independently of abstraction levels and model architectures.
-The next step in our approach, we try to take into account the reusability of the relationships by identifying the meaning, or semantics, associated with each relationship, in a goal of concept reification with their behavior.
+During the next step of our approach, we try to take into account the reusability of the relationships by identifying the meaning, the semantics, associated with each relationship, in a goal of concept reification with their behavior.
 The last step is to organize or structure these concepts to improve reusability and extensibility, to create VirtualModels in our Openflexo framework.
 
-Like any modeling approaches, these different steps could also be achieved in any order.
-But our goal during the challenge requirement analysis remains to lead to define a FML Virtual Models architecture as a federation of virtualmodels.
-As shown in figure~\ref{fig:MultilevelArchitecture}, we developped two metamodels (Base and ACME processes abstractions) instantiatied by two models (XSure and ACME processes).
+Like any modeling approaches, these different steps could also be achieved in any order and iteratively.
+But our goal during the challenge's analysis remains to produce a FML VirtualModels architecture as a federation of VirtualModels.
+As shown in figure~\ref{fig:MultilevelArchitecture}, we defined two VirtualModels (Base and ACME processes abstractions) instantiatied by two VirtualModel instances (XSure and ACME processes).
 This follows the way the challenge is presented and this organisation allows for the possibility of easy extensions.

### 2021-05-26 8-Discussion.tex

@@
+% Assessment of the Modeling Solution: discussing choices made, pointing out potential compromises / deficiencies
+
 %\noteJC{notes en vrac
@@
 
-  \noteJC{pas trop clair pour moi pour le moment}
+  \noteJC{pas trop clair pour moi pour le moment -> N/A}
 
@@
 build the metamodels and the models.
+
+\todo[inline]{verification de cohérence syntaxique (cardinalités, type des instances) et des contraintes métier (comportements écrits par les humains) à la volée + validation manuelle ; typage ?
+; pas d'analyseur statique ; vérif de la structure des contraintes/bonne forme des contraintes par rapport à la structure des modèles (exemple binding pas valide)}

### 2021-05-26 1-Main.tex

@@
 
+\lstdefinelanguage{fml}{
+  keywords={
+    abstract,boolean,break,byte,case,catch,char,class,%
+    const,continue,default,do,double,else,extends,false,final,%
+    finally,float,for,goto,if,implements,import,instanceof,int,%
+    interface,label,long,native,new,null,package,private,protected,%
+    public,return,short,static,super,switch,this,throw,%
+    throws,transient,true,try,void,volatile,while,String},%
+  morekeywords={
+    [3]concept,model,forEach,assert,%
+  },%
+  sensitive,%
+%  alsoletter={+-!*=<>/\%|\&\{\}?},%
+%  ndkeywords={/,+,-,!,--,++,*,=,==,!=,<,>,<=,>=,\%,|,||,\&,\&\&,\{,\}},%
+  morecomment=[l]//,%
+  morecomment=[s]{/*}{*/},%
+  morestring=[b]",%
+  morestring=[b]',%
+}%[keywords,comments,strings]%
+
 \lstdefinestyle{mystyle}{
@@
     stringstyle=\color{codepurple},
-    basicstyle=\ttfamily\footnotesize,
+    basicstyle=\ttfamily\scriptsize,
     breakatwhitespace=false,         
     breaklines=true,
+    flexiblecolumns=true,
     captionpos=b,
-    keepspaces=true,                 
-    %numbers=left,                    
-    numbersep=5pt,                  
-    showspaces=false,                
-    showstringspaces=false,
-    showtabs=false,                  
+    emptylines=0,
+    %keepspaces=true,                 
+    %numbers=left,    numbersep=5pt,                  
+    %showspaces=false,                
+    %showstringspaces=false,
+    %showtabs=false,                  
     tabsize=2
@@
 
-\lstset{style=mystyle}
+\lstset{style=mystyle,language=fml}
 

### 2021-05-26 4-Technology.tex
Updates from Overleaf

@@
+%Technology: precise description of the technology / approach that is used
+
 % Inspiré de la section 10 du papier formose
-
-\noteFabien{metamodel ou meta-model ? perso je penche plus pour le premier}
-\noteJC{pareil, j'utilise plutôt metamodel, en général}
 
@@
 Semantic Modeling for Information Federation (SIMF)~\cite{simf}. In
-contrast to approaches that compose meta-models into a single large
-meta-model grouping all needed entities, model federation build links
-among models and meta-models (even through levels) to make ``things''
+contrast to approaches that compose metamodels into a single large
+metamodel grouping all needed entities, model federation build links
+among models and metamodels (even through levels) to make ``things''
 work together. As an illustration of this feature, we developed a
 free-modeling editor -- freeing oneself from the bonds of
-model/meta-model conformity -- that is presented
+model/metamodel conformity -- that is presented
 in~\cite{models2016-freemodel}. Another very interesting feature of
@@
 
-This framework relies on the following architecture. A federation
+\begin{figure}[t]
+    \centering
+    \includegraphics[width=\columnwidth]{Figures/federation.pdf}
+    \caption{The model federation approach}
+    \label{fig:mf}
+\end{figure}
+
+This framework relies on the architecture of Figure~\ref{fig:mf}. A federation
 gathers a set of conceptual models, named \emph{virtual models} and a
@@
 serve as control components.
+In this paper, we mainly use the conceptual models and more precisely FML its domain specific modeling language. The only use of the federation aspect is on the tooling made for the process language.
 
@@
 \emph{behaviors}. A parallel to object-oriented approach can be useful
-to understand FML\footnote{Beware, even though useful for
-  comprehension, this correspondence is not
-  reliable, as some aspects of FML do not map to object oriented
-  concepts}. A concept corresponds to a class, its roles to the
+to understand FML\footnote{Some aspects of FML do not exist in object oriented approach.}. A concept corresponds to a class, its roles to the
 attributes of the class and its behaviors to the methods of the class.

### 2021-05-26 9-Relatedworks.tex

@@
-%Positioning  and  contrasting  the  presented solution with related work
+%Related Work: Positioning  and  contrasting  the  presented solution with related work
 
@@
 
-More similar to us, \cite{deeptelos2019} used DeepTelos\citep{deeptelos2016}, an extension of the Telos language \citep{telos1990} in order to solve the process challenge.
+More similar to us, \cite{deeptelos2019} used DeepTelos \citep{deeptelos2016}, an extension of the Telos language \citep{telos1990} in order to solve the process challenge.
 DeepTelos is based on the use of the PowerType \citep{atkinson2001essence} pattern.
@@
 
-\cite{dmla2019} used the Dynamic Multi-Layer Algebra (DMLA)\citep{dmla2017} in order to solve the process challenge.
+\cite{dmla2019} used the Dynamic Multi-Layer Algebra (DMLA) \citep{dmla2017} in order to solve the process challenge.
 DMLA is a level-blind modeling framework which is fully customizable (e.g., different types of instantiaton may be implemented).

### 2021-05-26 6-ModelPresentation.tex
Updates from Overleaf

@@
     abstract concept Gateway extends Type {
-      abstract void executeGateway(Process process);
+      abstract void execute(Process process);
       ...
@@
 
-\begin{lstlisting}[breaklines=true, language=java, basicstyle=\ttfamily\scriptsize, mathescape=true]
+\begin{lstlisting}
 forEach (actor : performingActors) {
@@
 
-\begin{lstlisting}[breaklines=true, language=java, basicstyle=\ttfamily\scriptsize, mathescape=true]
+\begin{lstlisting}
 concept ActorType extends Type {

### 2021-05-26 5-Analysis.tex

@@
 As shown in figure~\ref{fig:MultilevelArchitecture}, we defined two VirtualModels (Base and ACME processes abstractions) instantiatied by two VirtualModel instances (XSure and ACME processes).
+As the figure shows the VirtualModels play the role of metamodels, in the sens of concept definition with a level blind approach.
+The VirtualModel instance plays the role of instance of metamodels with no
 This follows the way the challenge is presented and this organisation allows for the possibility of easy extensions.

### 2021-05-26 6-ModelPresentation.tex
Updates from Overleaf

@@
 
-\todo[inline]{phrase d'intro pour expliquer que les req étaient couverts au fur et à mesure (côté systématique)}
+The presentation of our solution to the \mpc follows a systematic methodology where modeling choices are introduced step by step, as the requirements are stated.
+Satisfaction of requirements are explained throughout this presentation.
+
+%\todo[inline]{phrase d'intro pour expliquer que les req étaient couverts au fur et à mesure (côté systématique)}
 
@@
         return true;
+    }
     return false;
@@
     for (type : types) {
-   if (type.isOrSpecializes(actType)) {
+      if (type.isOrSpecializes(actType))
         return true;
     }
-  }     
+    return false;
   }     
@@
 
-\begin{lstlisting}[breaklines=true, language=java, basicstyle=\ttfamily\scriptsize, mathescape=true]
-concept CodingTaskType extends SETaskType 
-{
+\begin{lstlisting}
+concept CodingTaskType extends SETaskType {
   ProgrammingLanguage[1..*] languages;
@@
 
-\begin{lstlisting}[breaklines=true, language=java, basicstyle=\ttfamily\scriptsize, mathescape=true]
+\begin{lstlisting}
 forEach (artifactType : types) {
-    assert !(artifactType instanceof CodeArtifactType) | artifactType.implements(languages)
+  assert !(artifactType instanceof CodeArtifactType) | artifactType.doesImplement(languages)
 }
@@
 
-\begin{lstlisting}[breaklines=true, language=java, basicstyle=\ttfamily\scriptsize, mathescape=true]
-concept CodingTaskType extends SETaskType 
-{
+\begin{lstlisting}
+concept CodingTaskType extends SETaskType {
   ...
@@
     Coding returned = new Coding(name,this); 
-    if (languages.contains(ProgrammingLanguage.COBOL)) {
+    if (languages.contains(ProgrammingLanguage.COBOL))
       returned.addToPerformingActors(getActor("Ann Smith"));
-    }
     return returned;
@@
 \textbf{S8} requirement is simply captured with the definition of \textit{Testing} instance of \textit{SETaskType}, \textit{Tester} instance of \textit{ActorType}, and \textit{Test report} instance of \textit{ArtifactType}.
-% TODO S9
+
+\todo[inline]{requirement S9 non couvert pour le moment: compléter}
 
@@
 
-% TODO S13
+\todo[inline]{requirement S13 pas complètement couvert : compléter}
 
 \subsection{Openflexo tooling}
+
+Our solution is fully implemented within Openflexo tool. Both usecases have been modelled in the interactive design environment and are executable by the FML execution engine.
+
+\todo[inline]{mettre ici le lien vers la page web qui montre la démo}
+
+Ici: un paragraphe qui décrit l'exécution, l'activation et la finalisation des taches, la business logic des gateways, etc...
+
+We took advantage of model federation and the availability of diagramming features though the \textit{Diagramming TA} to implement two interactive graphical tools built on top of conceptual levels detailed in previous section.
+A first tool offers a graphical edition of a Process type, while the second tool offers enactment feature (the instantiation of a process from its process type), the ability within an enacted process to assign tasks to some actors, and the execution of this process with a graphical visualization.
+
+Figure \ref{fig:ToolingArchitecture} shows the architecture of these two tools.
+Left of figure presents process type graphical editor.
+ProcessTypeEditor is implemented as a VirtualModel declaring two model slots (represented with bold circles).
+The first model slot references an instance of Acme metamodel, while the second model slot references an instance of diagram, conforms to the ProcessType diagram specification (a diagram metamodel
 

### 2021-05-26 5-Analysis.tex
Updates from Overleaf

--- 2021-05-26 5-Analysis.tex
@@
 As shown in figure~\ref{fig:MultilevelArchitecture}, we defined two VirtualModels (Base and ACME processes abstractions) instantiatied by two VirtualModel instances (XSure and ACME processes).
-This follows the way the challenge is presented and this organisation allows for the possibility of easy extensions.
+As the figure shows the VirtualModels play the role of metamodels, in the sens of concept definition with a level-blind approach.
+The VirtualModel instances play the role of models conforms to metamodel definitions.
+In the rest of the paper, we keep the the vocabulary metamodels and models as to simplify the presentation.
+The resulted architecture follows the way the challenge is presented and this organisation allows for the possibility of flexible extensions.
 
@@
     \item the horizontal axis is characterized as the ontological instantiation axis, in the sense that the domain type definition (ProcessType as example) is referenced by an instance definition (Process, a.e).
-In our base  VirtualModel, each domain type definition is referenced by its instance definition, as developed in the section \ref{sec:ProcessEnactment}
+In our base metamodel, each domain type definition is referenced by its instance definition, as developed in the section \ref{sec:ProcessEnactment}
 
 \item the vertical axis is viewed as the linguistic instantiation, relative to the use of the FML language.
-The VirtualModel, defined as the core definition model, generates a model founded on the instantiation mechanism of the FML Language.
-This mechanism is the classical Object/Instance mechanism of the object languages but without any constraint on the referenced VirtualModel.
+The VirtualModel, defined as the core definition metamodel, generates a model founded on the instantiation mechanism of the FML Language.
+This mechanism is the classical Object/Instance mechanism of the object languages but without any constraint on the referenced VirtualModel, i.e metamodel.
 The set of resulted instances comes from several VirtualModels
@@
 The resulted multi-level architecture is organized as a set of VirtualModels.
-The first one is the core concepts model including the definition of Processes and Tasks.
-The second one, Acme VirtualModel, is extended from the previous model to integrate the Acme definitions.
-Our FML language defines the multiple inheritance concept between VirtualModels as illustrated in the section \ref{sec:AcmeSoftwareDevelopmentProcess}.
+The first one is the core concepts metamodel including the definition of Processes and Tasks.
+The second one, Acme metamodel, is extended from the previous model to integrate the Acme definitions.
+Our FML language defines multiple inheritance concept between VirtualModels as illustrated in the section \ref{sec:AcmeSoftwareDevelopmentProcess}.
 
-Finally as explained previously, the defined level for the Acme model and the XSure is the result of the instantiation of VirtualModels of different levels.
+Finally as explained previously, the defined level for the Acme metamodel and the XSure is the result of the instantiation of VirtualModels.
 In our approach this level can't be specialized or extended but all the other VirtualModels could be extended  by any concepts, as a new federated model.
+Also, the instance level can be instantiated from any VirtualModel, which represent any metamodel level.
 
-% il reste à décrire l'architecture générale des VirtualModels
+% cross level constraints or added models as federated metamodels
 

### 2021-05-27 2-Abstract.tex
sync

@@
 We present a solution based on model federation and discuss the advantages and limitations of using this approach for multi-level modeling.
-Our solution is fully implemented with the Openflexo framework and its internal Flexo Modeling Language (FML).
+Concretely, we use virtual models and more precisely the Flexo Modeling Language (FML) that serves to build them as the main building block for the process modelling challenge whereas the federation aspect is used in order to provide editing tooling for the process language.
+Our solution is fully implemented with the Openflexo.
 It fulfills all the challenge requirements.
@@
   vendeur et les exigences sont bien remplies. La fin est très positive, mais
-afactuelle selon moi.}
+factuelle selon moi.}
 \noteSylvain{Ce qui m'embête un peu avec cet abstract c'est qu'on met en avant la fédération de modèles alors que sur ce cas précis on en fait pas vraiment (sauf pour la partie outillage avec les éditeurs diagrammatiques).
@@
 \noteSylvain{En revanche, on peut arguer que pouvoir au même niveau manipuler modèle et métamodèle nous simplifie grandement les choses pour adresser des problématiques multi-niveaux > un bout de la conclusion}.
-\noteSM{En fait l'idee de l'abstract c'est juste de nomer les approche utilisé. Les détails sont decrites dans la section 2}
+\noteSM{En fait l'idee de l'abstract c'est juste de nomer les approche utilisé.
+Les détails sont decrites dans la section 2-Tecnology.
+On peut nommer FML avant mais je crois que ça serait moins claire.
+Cela dit, n'hesitez pas a modifier.}
+
+%The models in a federation remain autonomous and represented in their original technological spaces %whereas virtual models and links serve as control components used to present different views to the %users and  maintain synchronization.

### 2021-05-27 6-ModelPresentation.tex
sync

@@
 
-Figure \ref{fig:ToolingArchitecture} shows the architecture of these two tools.
-Left of figure presents process type graphical editor.
-ProcessTypeEditor is implemented as a VirtualModel declaring two model slots (represented with bold circles).
-The first model slot references an instance of Acme metamodel, while the second model slot references an instance of diagram, conforms to the ProcessType diagram specification (a diagram metamodel
-
 \begin{figure*}
@@
 
+Figure \ref{fig:ToolingArchitecture} shows the architecture of these two tools.
+Left of figure presents \textit{process type graphical editor}.
+\textit{ProcessTypeEditor} is implemented as a \textit{VirtualModel} declaring two model slots (represented with bold circles).
+The first model slot references an instance of Acme metamodel, while the second model slot references an instance of diagram, conforms to the \textit{ProcessType diagram specification} (a diagram "metamodel" which defines and specify structure and graphical representations for edited items).
+While executed, this tool manages a graphical view for an instance of Acme metamodel (the \textit{Acme model}) and a specific \textit{ProcessType} instance.
+This tools allows to represent and edit a \textit{ProcessType}.
+One particular point should be noted regarding the highly reflective nature of FML and associated tooling : when drag and drop interactor is applied for a new item, the tool allows to choose the concept type to be instantiated (when a \textit{TaskType} is created for example, the user must choose the sub-concept of \textit{TaskType} to be selected - in Acme use case it can be \textit{SETaskTask} or \textit{CodingTaskType} or default value \textit{TaskType}).
+
+The other tool, an \textit{enacted process editor}, provides process edition and tasks assignations through a graphical visualization displaying process being executed.
+This tool is represented one the right side of the figure \ref{fig:ToolingArchitecture}.
+It follows the same architecture pattern as explicited for the \textit{process type graphical editor}, with two model slots referencing both the model and a diagram.
+Process enactment is operated from a \textit{ProcessType}, and must be defined using a name identifying the newly instantiated process.
+All tasks are created from their \textit{TaskType} definition, and required assignments applies (if for example "Coding" \textit{TaskType} defines COBOL as output programming language, related task is automatically assigned to "Ann Smith").
+All tasks get a status (\textit{Not startable} - when not assigned or when required input artifacts are not available - , \textit{Startable}, \textit{Started}, \textit{Completable}, \textit{Completed}), as well as a set of performing actors, a begin and end date, and used and produced artifacts.
+This tools also manages creation of artifacts, whose semantics follows rules defined in section \ref{sec:ProcessEnactment}}
+
 % Expliquer l'outillage construit et comment il marche

### 2021-05-27 6-ModelPresentation.tex
sync d'une version compilable
Updates from Overleaf

@@
 
-Ici: un paragraphe qui décrit l'exécution, l'activation et la finalisation des taches, la business logic des gateways, etc...
+\todo[inline]{Ici: un paragraphe qui décrit l'exécution, l'activation et la finalisation des taches, la business logic des gateways, etc...}
 
@@
 All tasks get a status (\textit{Not startable} - when not assigned or when required input artifacts are not available - , \textit{Startable}, \textit{Started}, \textit{Completable}, \textit{Completed}), as well as a set of performing actors, a begin and end date, and used and produced artifacts.
-This tools also manages creation of artifacts, whose semantics follows rules defined in section \ref{sec:ProcessEnactment}}
+This tool also manages creation of artifacts, whose semantics follows rules defined in section \ref{sec:ProcessEnactment} (\textbf{P13}).
 

### 2021-05-27 2-Abstract.tex

--- 2021-05-25 2-Abstract.tex
@@
 Model federation is a multi model \emph{management} approach based on the use of virtual models and loosely coupled links.
-The models in a federation remain autonomous and represented in their original technological spaces whereas virtual models and links serve as control components used to present different views to the users and  maintain synchronization.
+The models in a federation remain autonomous and represented in their original technological spaces whereas virtual models and links (which are not level bounded) serve as control components used to present different views to the users and  maintain synchronization.
 In this paper we tackle the \emph{process modeling challenge}, which consists in providing a solution to the problem of specifying and enacting processes.
@@
 We present a solution based on model federation and discuss the advantages and limitations of using this approach for multi-level modeling.
-Our solution is fully implemented with the Openflexo framework and its internal Flexo Modeling Language (FML).
+Concretely, we use virtual models and more precisely the Flexo Modeling Language (FML) that serves to describe them as the main building block in order to solve the process modeling challenge whereas the federation aspect is used in order to provide editing tooling for the resulting process language.
+Our solution is fully implemented with the Openflexo framework.
 It fulfills all the challenge requirements.
 
-\noteSM{J'ai fait un premier essai d'abstract... un peu trop vendeur peut-etre?}
-\noteJC{bof, pas de manière éhontée : il faut qu'un abstract soit un peu
-  vendeur et les exigences sont bien remplies. La fin est très positive, mais
-afactuelle selon moi.}
 \noteSylvain{Ce qui m'embête un peu avec cet abstract c'est qu'on met en avant la fédération de modèles alors que sur ce cas précis on en fait pas vraiment (sauf pour la partie outillage avec les éditeurs diagrammatiques).
@@
 \noteSylvain{En revanche, on peut arguer que pouvoir au même niveau manipuler modèle et métamodèle nous simplifie grandement les choses pour adresser des problématiques multi-niveaux > un bout de la conclusion}.
-\noteSM{En fait l'idee de l'abstract c'est juste de nomer les approche utilisé. Les détails sont decrites dans la section 2}
+%The models in a federation remain autonomous and represented in their original technological spaces %whereas virtual models and links serve as control components used to present different views to the %users and  maintain synchronization.

### 2021-05-27 3-Introduction.tex

@@
 Contrary to the strict approach, multi-level modeling advocates the use of a flexible number of levels as well as a flexibilization of the relations between them.
-This paradigm is gaining traction lately within the modeling community as evidenced by the contribution of many new multi-level modeling approaches and tools and the celebration of topic specific workshops\footnote{https://homepages.ecs.vuw.ac.nz/Groups/MultiLevelModeling/WebHome} and seminars~\citep{almeida2018multi}.
+This paradigm is gaining traction lately within the modeling community as evidenced by the contribution of many new multi-level modeling approaches and tools such as Melanee~\citep{melanee}, MetaDepth~\citep{metadepth}, MultEcore~\citep{multecore2016}, DeepTelos\citep{deeptelos2016} or DMLA~\citep{dmla2017}.
 Taking advantage of this vibrant state of affairs, and in order to foster discussion and enable comparison between competing approaches, a multi-level modeling challenge has been created.
@@
 We assess the modeling solution in section~\ref{sec:discussion}.
-Section~\ref{sec:relatedwork} presents the related work before the conclusion~\ref{sec:conclusion}.
+Section~\ref{sec:relatedwork} presents the related work before the conclusion.%~\ref{sec:conclusions}.
 

### 2021-05-27 4-Technology.tex

@@
 
-FML is a language designed to define virtual models. A virtual model
+\begin{figure}[t]
+    \centering
+    \includegraphics[width=\columnwidth]{Figures/FMLCoreModel.pdf}
+    \caption{The FML core metamodel}
+    \label{fig:mm}
+\end{figure}
+
+FML simplified metamodel is provided in Figure~\ref{fig:mm}, it is designed to define virtual models. A virtual model
 is composed of a set of \emph{concepts}, while itself being a concept.
@@
 at runtime.
-Whenever a type external to the federation space (from a TA or an
-external model) is used, one needs to use a \emph{model slot}. A model
-slot is a mediation entity, associated with a TA, in charge of giving
-access to external elements of the corresponding technological space.
-
-A \emph{technology adapter} (TA) is a reusable library that defines
+Whenever a type external to the federation space is used, one needs to use a \emph{model slot}. A model
+slot is a mediation entity in charge of giving
+access to external elements using
+a \emph{technology adapter}\footnote{It is a reusable library that defines
 connections between the FML execution engine and a particular
-technological space. The model federation framework provides ways to
-define TA.
-
-We should stress that building a virtual model is purely conceptual,
-model slots / TA are in charge of the technicalities. Additionally,
-when a virtual model or a concept point to an external model, it comes
-with an explicit interpretation of the external data. The virtual
-model can hide information, transform elements and does not need to
-cover the whole external metamodel.
-% In our example, it ignores all other elements of the \smk metamodel.
-We did not required TA in our proposal. The whole solution uses the
-internal (FML) language.
+technological space.}.
 
@@
 also to define the collection of actions an engineer can perform on
-them. These actions are called \emph{business behaviors} and may be
-compared to methods of object-oriented approach.
-% The rename operation already cited for the \textsf{Goal-xls}
-% federation is an example of such a behavior. One could also define
-% operations to add or remove requirements, to refine relationships,
-% \emph{etc.} The renaming operation is defined by a \texttt{rename}
-% behavior in the concept \texttt{Requirement}. It uses the
-% \texttt{setCell} action of the Excel TA and the \texttt{rnGoal} of
-% the \smk TA.
+them.
+These actions are called \emph{behaviors} and can either be called (like usual methods) or triggered by events.
+The reactive behaviors are mainly useful when models have to be synchronised.
+We did not
+exploit this possibility in the challenge.
+
 When the FML execution engine runs a federation, it creates virtual
@@
 are connected to external elements through model slot instances.
-% The lower part of the figure~\ref{fig:space-ex} illustrates this
-% runtime phase.
-
-FML also supports another kind of behavior which can be described as
-interactors. Such behavior is reactive and triggers when any of the
-federated models evolves. Two scenarios are then possible: (1) an
-action in the federation is propagated to the federated models or (2)
-a modification is detected in a federated model and triggers a
-behavioral interactor of the federation.
-% Hence the double arrows in the links of Figure~\ref{fig:space-ex}.
-This feature is useful when models have to be synchronised. We did not
-exploit this possibility in the challenge.
 

### 2021-05-27 8-Discussion.tex

@@
   Due to the fact our approach is level-blind, our solution could have had more
-  or fewer levels depending on the variations of the use case. %As levels are
-  %not specifically encoded,
+  or fewer levels depending on the variations of the use case.
+However, the number of levels is very related to the problem, making the approach fixed.
+For example, given a requirement, an instance cannot become a concept easily.
 
@@
 \noteJC{on a du tooling, un peu moins de formalisme}
-
+-
 Currently, there are no formalism to establish the semantics of the MLM
@@
 
-By default, our tooling does not include a lot of automated verifications.
+Our tooling includes mechanisms to verify the models.
+First, syntactical consistency is realised by cardinality checks and by typing.
+Second, the constraints the users have written are continuously checked in order to
+
 However, a user can add verification mechanisms by writing new behaviors

### 2021-05-27 6-ModelPresentation.tex

--- 2021-05-27 6-ModelPresentation.tex
@@
 
-Figure \ref{fig:MultilevelArchitecture} shows the overview of the multivel architecture designed to capture the two use cases described in the Process Challenge.
+The designed multilevel architecture is shown in Figure \ref{fig:MultilevelArchitecture} overview and captures the two use cases described in the Process Challenge.
 
@@
 Base conceptual element is called a \textit{FlexoConcept} in FML language, and is similar to the concept of class in object-oriented modeling.
-We use the UML representation of class to represent a \textit{FlexoConcept}, while cardinalities are shown using the UML-like classical rules.
+The UML representation of class is used to represent a \textit{FlexoConcept}, while cardinalities are shown using the UML-like classical rules.
+In the diagram, we choose to represent the role associated with the \textit{FlexoConcept} in the class and the role type is referenced by the arrow in front of the role name.
 
@@
 
-\todo{Augmenter la taille des fontes de la figure \ref{fig:AcmeFullArchitecture}}
+\todo[inline]{Augmenter la taille des fontes de la figure \ref{fig:AcmeFullArchitecture}}
+\todo[inline]{Ajouter les programming languages}
 
@@
 
-\todo[inline]{mettre ici le lien vers la page web qui montre la démo}
-
-Ici: un paragraphe qui décrit l'exécution, l'activation et la finalisation des taches, la business logic des gateways, etc...
+Base metamodel has been completed with some behaviours implementing execution semantic for executed processes.
+All tasks - instantiated from \textit{TaskType} for a given enacted process - manage a status which can be \textit{Not startable} (when not assigned to a performing actor or when required input artifacts are not available) , \textit{Startable}, \textit{Started}, \textit{Completable} (when all output artifacts are ready), \textit{Completed}).
+A \textit{Task} also manages a set of performing actors, a begin and end date, some used and produced artifacts.
+\textit{Gateways} business logic has also been implemented through the implementation of abstract \texttt{execute(Process)} behaviour.
+Management of artifacts whose semantics follows rules defined in section \ref{sec:ProcessEnactment} (\textbf{P13}) has also been implemented.
 
@@
 Figure \ref{fig:ToolingArchitecture} shows the architecture of these two tools.
-Left of figure presents \textit{process type graphical editor}.
-\textit{ProcessTypeEditor} is implemented as a \textit{VirtualModel} declaring two model slots (represented with bold circles).
+Left of figure presents \textit{ProcessType graphical editor}.
+\textit{ProcessTypeEditor} is modeled as a \textit{VirtualModel} declaring two model slots (represented with bold circles).
 The first model slot references an instance of Acme metamodel, while the second model slot references an instance of diagram, conforms to the \textit{ProcessType diagram specification} (a diagram "metamodel" which defines and specify structure and graphical representations for edited items).
@@
 
-The other tool, an \textit{enacted process editor}, provides process edition and tasks assignations through a graphical visualization displaying process being executed.
-This tool is represented one the right side of the figure \ref{fig:ToolingArchitecture}.
-It follows the same architecture pattern as explicited for the \textit{process type graphical editor}, with two model slots referencing both the model and a diagram.
+The other tool, called \textit{Enacted process graphical editor}, provides process edition and tasks assignations through a graphical visualization displaying process being executed.
+This tool is represented on the right side of the figure \ref{fig:ToolingArchitecture}.
+It follows the same architecture pattern as explicited for the \textit{ProcessType graphical editor}, with two model slots referencing both the model and a diagram.
 Process enactment is operated from a \textit{ProcessType}, and must be defined using a name identifying the newly instantiated process.
 All tasks are created from their \textit{TaskType} definition, and required assignments applies (if for example "Coding" \textit{TaskType} defines COBOL as output programming language, related task is automatically assigned to "Ann Smith").
-All tasks get a status (\textit{Not startable} - when not assigned or when required input artifacts are not available - , \textit{Startable}, \textit{Started}, \textit{Completable}, \textit{Completed}), as well as a set of performing actors, a begin and end date, and used and produced artifacts.
-This tools also manages creation of artifacts, whose semantics follows rules defined in section \ref{sec:ProcessEnactment}}
+All tasks get a status as well as a set of performing actors, a begin and end date, and used and produced artifacts.
 
-% Expliquer l'outillage construit et comment il marche
+\todo[inline]{mettre ici le lien vers la page web qui montre la démo}

### 2021-05-27 8-Discussion.tex

@@
   As for cross-level relationships, cross-level constraints do not have a
-  specific nature. They are like any constraint a user can define on a model
-  element. Therefore, if a constraint is mandatory when establishing a
+  specific nature. They are like any constraint a user can define by adding a behavior to a model element. Therefore, if a constraint is mandatory when establishing a
   relationship between elements from different levels, the user has to create
-  it manually. This can be a limit of our approach: the flexibility of our
-  tooling at the cost of a limited number of automated behaviors.
+  it manually. This can be a limit of our approach: the cost of the flexibility of our tooling is a limited number of automated behaviors.
+
+  %the flexibility of our tooling at the cost of a limited number of automated behaviors.
 
@@
 
-  In order to preserve the integrity of level contents when changes to level
-  contents occur, a validation has to be triggered. The tool reports all
-  inconsistencies that the user is free to fix. Note that in our approach, the
-  metamodels are built in a ad hoc way together with the models
-  (co-construction), therefore the user validates on a ad hoc basis.
+  Behaviors are continuously checked, ensuring the integrity of the models when changes occur. 
+  However, most of those behaviors have to be written by the users. 
+  Thus, the integrity of the contents relies on the users when they build the metamodels and the models.
+  Note that in our approach, the metamodels are built in a ad hoc way together with the models
+  (co-construction), therefore the user also validates on a ad hoc basis.
 
@@
 
-  \noteJC{pas trop clair pour moi pour le moment -> N/A}
+  \noteJC{pas trop clair pour moi pour le moment -> N/A ? }
+  Due to the nature of our approach and its level-blindness, \emph{deep characterization} does not apply to our solution.
+\todo[inline]{je suppose qu'en fait, on pourrait encoder dans un comportement quelque chose se rapprochant}
 
@@
 First, syntactical consistency is realised by cardinality checks and by typing.
-Second, the constraints the users have written are continuously checked in order to
-
-However, a user can add verification mechanisms by writing new behaviors
-attached to concepts. Presently, the verifications rely on the users when they
-build the metamodels and the models.
-
-\todo[inline]{verification de cohérence syntaxique (cardinalités, type des instances) et des contraintes métier (comportements écrits par les humains) à la volée + validation manuelle ; typage ?
-; pas d'analyseur statique ; vérif de la structure des contraintes/bonne forme des contraintes par rapport à la structure des modèles (exemple binding pas valide)}
+Second, the constraints the users have written are continuously checked.
+Thus, a part of the verification relies on the users when they build the metamodels and the models.

### 2021-05-27 3-Introduction.tex
sync overleaf->github

@@
 Model federation is a multi model management approach based on the use of virtual models and loosely coupled links.
-The models in a federation remain autonomous and represented in their original technological spaces whereas virtual models and links serve as control components used to present different views to the users and maintain synchronization.
-Our solution is based on this model federation infrastructure, namely, the use of virtual models and links.
-Concretely, we use OpenFlexo and its internal Flexo Modeling Language (FML), a language to create, link and manage virtual models, in order to achieve our goal.
+The models in a federation remain autonomous and represented in their original technological spaces whereas virtual models (also called conceptual models) and links serve as control components used to present different views to the users and maintain synchronization.
+Our solution is based on the model federation infrastructure, namely, the use of virtual models and links.
+Concretely, we use OpenFlexo and its internal Flexo Modeling Language (FML), a language to create, link and manage virtual models, in order to solve the MULTI process challenge while the federation aspect is used solely as a means to provide tooling for the resulting process language.
 Our solution, which is fully implemented and executable, meets all the requirements.
-
 %Links among models can be within a same level of abstraction or across levels of abstractions which gives,

### 2021-05-27 8-Discussion.tex
sync overleaf->github

@@
 
-\noteJC{pour le moment, seulement réponse aux différents points, on ne voit pas
-de discussion sur les choix, sur les potentiels compromis/déficiencies ;
-sous-sections : bof, p-e plutôt des paragraphes ; très court/concis, faudrait
-étoffer}
+%\noteJC{pour le moment, seulement réponse aux différents points, on ne voit pas
+%de discussion sur les choix, sur les potentiels compromis/déficiencies}
+
+In this section, we discuss our multilevel model solution with regard to the required aspects 
+mentioned in the challenge.
 
@@
 
-  Due to the fact our approach is level-blind, our solution could have had more
+  Due to the fact that our approach is level-blind, our solution could have had more
   or fewer levels depending on the variations of the use case.
 However, the number of levels is very related to the problem, making the approach fixed.
-For example, given a requirement, an instance cannot become a concept easily.
+For example, given a requirement, an instance cannot be changed into a concept easily.
 
@@
 
-  \noteJC{pas trop clair pour moi pour le moment -> N/A ? }
-  Due to the nature of our approach and its level-blindness, \emph{deep characterization} does not apply to our solution.
-\todo[inline]{je suppose qu'en fait, on pourrait encoder dans un comportement quelque chose se rapprochant}
+  %\noteJC{pas trop clair pour moi pour le moment -> N/A ? }
+  Due to the nature of our approach and its level-blindness, \emph{deep characterization} does not 
+  apply to our solution. Such a mechanism could probably be encoded by specific behaviors, however 
+  it would not be a generic mechanism, making it difficult to reuse for another problem.
 
@@
   Although we build ad hoc metamodels in our approach, our solution separates
-  the domain-specific elements from general purpose ones, making easily
-  possible to apply it to other domains.
+  the domain-specific elements from general purpose ones, making possible to 
+  apply it to other domains.
 
@@
   extensibility ability of our solution: it was decomposed into two steps
-  (generic process Xsure, then ACME process) which is kind of simulation of the
+  (Xsure process, then ACME process) which can be seen as a simulation of the
   evolution of the  specification. We observed that our approach has been
@@
   Due to the level-blind aspect of our approach, we could insert a new level,
-  for instance. It would consist in creating new concepts we would link to
+  for example. It would consist in creating new concepts we would link to
   other ones. Then, we would add any necessary constraints on those concepts
@@
 
-  \noteJC{comment étendre : prendre l'exemple donné ?; level insertion possible
-  : oui, vu que pas de notion de level}
+  %\noteJC{comment étendre : prendre l'exemple donné ?; level insertion possible
+  %: oui, vu que pas de notion de level}
 
 \todo[inline]{recommended discussion aspects ↓}
-\subsection{Formalisms and tools ?}
+\subsection{Formalism and tools}
 
@@
 
-\noteJC{on a du tooling, un peu moins de formalisme}
--
+\noteJC{on a du tooling, un peu moins de formalisme ; je me demande s'il ne vaut pas mieux ne pas traiter l'aspect formalisme}
+
 Currently, there are no formalism to establish the semantics of the MLM
@@
 
-Our solution is fully tooled by a ``free'' ad hoc tool. We use the Openflexo
-infrastructure to build models and metamodels together. In addition to the
-solution, it also provides a dedicated tool which allows processes execution. 
-\noteJC{dire qqch sur le fait que produire une solution avec son outil associé
-  est une force, genre ``Providing a solution and its associated tool can be
-…'' mais pour le moment je ne vois pas comment l'enchainer}
+As described in section~\ref{subsec:tooling}, our solution is fully tooled with Openflexo.
+We use the Openflexo infrastructure to build models and metamodels together, and to provide 
+dedicated tooling to edit a Process type, to enact a process and to execute it.
+Being able to provide a solution and its associated tools quickly and easily is an \textcolor{red}{interesting} feature.
+We could also have taken advantage of the model federation to connect our solution to other tools 
+dedicated to process edition and to process execution (\emph{e.g.} a BPMN engine). However we 
+already had all the necessary components to provide the aforementioned graphical tools.
 
@@
 Second, the constraints the users have written are continuously checked.
-Thus, a part of the verification relies on the users when they build the metamodels and the models.
+Thus, a part of the verification relies on the fact that users add behaviors when they build the metamodels and the models.

### 2021-05-27 9-Relatedworks.tex
Updates from Overleaf
sync overleaf->github

@@
 %Salvador commence
+A plethora of multi-level modeling approaches and tools with different foundations have appeared in the recent years (see \cite{multiTools})\todo{not sure how to cite a wiki page}.
+Comparing them lies out of the scope of this paper.
+In this sense, we limit the following discussion to the presentation and comparison of previous solutions to the MULTI Process Challenge.
 
-A first description of the process modeling challenge as a multi-level modeling problem was proposed in \cite{lara2018refactoring} in the context of a catalogue of refactoring for multi-level models.
-A solution is provided at a conceptual level using deep instantiation and linguistic extensions.
+A first description of process modeling as a multi-level modeling problem was proposed by \cite{lara2018refactoring} in the context of a catalogue of refactoring for multi-level models (in a simpler form with less constraints and requirements w.r.t.
+the challenge version).
+A solution is provided with MetaDepth~\citep{metadepth}, which supports modelling with any number of levels, dual ontological/linguistic typing and deep characterization through potency.
 
-\cite{multiecore2019} used MultiEcore \citep{multecore2016} to provide a solution to the challenge.
-MultEcore uses cascading and potency.
+\cite{multiecore2019} use MultiEcore \citep{multecore2016} to provide a solution to the challenge.
+MultEcore uses an extesion of the two-level cascading technique and potency.
 They do not need the synthetic typing relations we introduce.
@@
 
-More similar to us, \cite{deeptelos2019} used DeepTelos \citep{deeptelos2016}, an extension of the Telos language \citep{telos1990} in order to solve the process challenge.
+More similar to us, \cite{deeptelos2019} use DeepTelos \citep{deeptelos2016}, an extension of the Telos language \citep{telos1990} in order to solve the process challenge.
 DeepTelos is based on the use of the PowerType \citep{atkinson2001essence} pattern.
@@
 
-\cite{dmla2019} used the Dynamic Multi-Layer Algebra (DMLA) \citep{dmla2017} in order to solve the process challenge.
+Similarly, \cite{dmla2019} use the Dynamic Multi-Layer Algebra (DMLA) \citep{dmla2017} in order to solve the process challenge.
 DMLA is a level-blind modeling framework which is fully customizable (e.g., different types of instantiaton may be implemented).

### 2021-05-27 5-Analysis.tex

@@
 As the figure shows the VirtualModels play the role of metamodels, in the sens of concept definition with a level-blind approach.
-The VirtualModel instances play the role of models conforms to metamodel definitions.
+The VirtualModel instances play the role of models conform to metamodel definitions.
 In the rest of the paper, we keep the the vocabulary metamodels and models as to simplify the presentation.
@@
 
-Finally as explained previously, the defined level for the Acme metamodel and the XSure is the result of the instantiation of VirtualModels.
+Finally as explained previously, the defined level for the Acme  and the XSure models is the result of the instantiation of VirtualModels.
 In our approach this level can't be specialized or extended but all the other VirtualModels could be extended  by any concepts, as a new federated model.

### 2021-05-27 7-RequirementsSatisfaction.tex
sync overleaf->github

--- 2021-05-25 7-RequirementsSatisfaction.tex
@@
-Now we discuss how the solution satisfies the challenge requirements.
+%Satisfaction of Requirements: demonstration of how the solution satisfies the challenge requirements
+
+In the previous section~\ref{sec:model}, we demonstrated that our solution satisfies all the challenge requirements.
+In this section, we summarize the different techniques used to satisfy them, without repeating all the justifications.
+
+%Now we discuss how the solution satisfies the challenge requirements.
 
@@
 
-Sylvain
+\todo[inline]{OK pour le tableau synthétique +un petit peu de texte + une mini intro qui référence la section précédente (vu que les req ont été couverts dans la section précédente ; pas de fusion des deux sections}
+
+\begin{tabular}{|l|c|r|}
+   \hline
+   colonne 1 & \multicolumn{2}{c|}{colonnes 2 et 3} \\
+   \hline
+   1.1 & 1.2 & 1.3 \\
+   \hline
+   2.1 & 2.2 & \checkmark \\
+   \hline
+\end{tabular}

### 2021-05-28 6-ModelPresentation.tex
sync overleaf->github
synchro from overleaf

@@
 The UML representation of class is used to represent a \textit{FlexoConcept}, while cardinalities are shown using the UML-like classical rules.
-In the diagram, we choose to represent the role associated with the \textit{FlexoConcept} in the class and the role type is referenced by the arrow in front of the role name.
+In the diagram, we choose to represent the role associated with the \textit{FlexoConcept} in the class square and the role type is referenced by the arrow in front of the role name.
 
@@
 
-\todo{Supprimer ombres et dégradés de la figure \ref{fig:AcmeSoftwareDevelopmentProcess}, et faire en sorte qu'elle soit plus haute et moins large}
-
 \textit{"Requirements analysis"} is defined as an instance of \textit{SETaskType}, \textit{Analyst} as an instance of \textit{ActorType} and \textit{Requirement specifications} as an instance of \textit{SEArtifactType}.
@@
 \subsection{Openflexo tooling}
+\label{subsec:tooling}
 

### 2021-05-28 1-Main.tex
sync overleaf->github
cleaning comments + remove trailing spaces

--- 2021-05-26 1-Main.tex
@@
 
+\lstdefinelanguage{fml}{
+  keywords={
+    abstract,boolean,break,byte,case,catch,char,class,%
+    const,continue,default,do,double,else,extends,false,final,%
+    finally,float,for,goto,if,implements,import,instanceof,int,%
+    interface,label,long,native,new,null,package,private,protected,%
+    public,return,short,static,super,switch,this,throw,%
+    throws,transient,true,try,void,volatile,while,String},%
+  morekeywords={
+    [3]concept,model,forEach,assert,%
+  },%
+  sensitive,%
+  morecomment=[l]//,%
+  morecomment=[s]{/*}{*/},%
+  morestring=[b]",%
+  morestring=[b]',%
+}
+
 \lstdefinestyle{mystyle}{
@@
     stringstyle=\color{codepurple},
-    basicstyle=\ttfamily\footnotesize,
+    basicstyle=\ttfamily\scriptsize,
     breakatwhitespace=false,
     breaklines=true,
+    flexiblecolumns=true,
     captionpos=b,
-    keepspaces=true,                 
-    %numbers=left,                    
-    numbersep=5pt,                  
-    showspaces=false,                
-    showstringspaces=false,
-    showtabs=false,                  
+    emptylines=0,
     tabsize=2
@@
 
-\section{Satisfaction of Requirements}
+\section{Satisfaction of requirements}
 \label{sec:requirements}
@@
 
-\section{Assessment of the Modeling Solution}
+\section{Assessment of the modeling solution}
 \label{sec:discussion}

### 2021-05-28 3-Introduction.tex
replace eg by macro

@@
 
-This strict approach shows its limitations when modeling complex domains requiring more than one level of specialization, e.g., to adapt the model to application sub domains.
-Indeed, the two-level approach fails to acknowledge and support: 1) the existence of different forms of classification and/or instantiation (e.g., ontological vs linguistic); and 2) the duality type-object for model elements.
+This strict approach shows its limitations when modeling complex domains requiring more than one level of specialization, \eg to adapt the model to application sub domains.
+Indeed, the two-level approach fails to acknowledge and support: 1) the existence of different forms of classification and/or instantiation (\eg ontological vs linguistic); and 2) the duality type-object for model elements.
 Therefore, while it may still be used for the aforementioned complex modeling tasks, the strict approach entails the introduction of accidental complexity in both the modeling process and the resulting modeling artefacts.

### 2021-05-28 8-Discussion.tex
replace eg by macro

@@
 We could also have taken advantage of the model federation to connect our solution to other tools 
-dedicated to process edition and to process execution (\emph{e.g.} a BPMN engine). However we 
+dedicated to process edition and to process execution (\eg a BPMN engine). However we 
 already had all the necessary components to provide the aforementioned graphical tools.

### 2021-05-28 5-Analysis.tex
synchro from overleaf
replace ie by macro

@@
 As shown in figure~\ref{fig:MultilevelArchitecture}, we defined two VirtualModels (Base and ACME processes abstractions) instantiatied by two VirtualModel instances (XSure and ACME processes).
-As the figure shows the VirtualModels play the role of metamodels, in the sens of concept definition with a level-blind approach.
+As the figure shows the VirtualModels play the role of metamodels, in the sense of concept definition with a level-blind approach.
 The VirtualModel instances play the role of models conform to metamodel definitions.
@@
 The VirtualModel, defined as the core definition metamodel, generates a model founded on the instantiation mechanism of the FML Language.
-This mechanism is the classical Object/Instance mechanism of the object languages but without any constraint on the referenced VirtualModel, i.e metamodel.
+This mechanism is the classical Object/Instance mechanism of the object languages but without any constraint on the referenced VirtualModel, \ie metamodel.
 The set of resulted instances comes from several VirtualModels

### 2021-05-28 4-Technology.tex
Updates from Overleaf
synchro from overleaf
replace OMG by macro

@@
   federation} is a way to assemble models using some kind of
-low-coupling links. It has been studied to answer the OMG' RFP
+low-coupling links. It has been studied to answer the \OMG' RFP
 Semantic Modeling for Information Federation (SIMF)~\cite{simf}. In
@@
 \end{figure}
+%\noteAntoine{Dans la figure ne manque-t-il pas les containements ? (VM vers Concepts)}
 
-FML simplified metamodel is provided in Figure~\ref{fig:mm}, it is designed to define virtual models. A virtual model
+FML simplified metamodel is provided in Figure~\ref{fig:mm}. It is designed to reify models as virtual models. A virtual model
 is composed of a set of \emph{concepts}, while itself being a concept.

### 2021-05-28 6-ModelPresentation.tex
replace UML by macro

@@
 
-Figure \ref{fig:BaseMetamodel} represents this base metamodel with a UML-like formalism, well adapted to represent object-oriented FML concepts and their instances.
+Figure \ref{fig:BaseMetamodel} represents this base metamodel with a \UML-like formalism, well adapted to represent object-oriented FML concepts and their instances.
 Base conceptual element is called a \textit{FlexoConcept} in FML language, and is similar to the concept of class in object-oriented modeling.
-The UML representation of class is used to represent a \textit{FlexoConcept}, while cardinalities are shown using the UML-like classical rules.
+The \UML representation of class is used to represent a \textit{FlexoConcept}, while cardinalities are shown using the \UML-like classical rules.
 In the diagram, we choose to represent the role associated with the \textit{FlexoConcept} in the class square and the role type is referenced by the arrow in front of the role name.

### 2021-05-28 3-Introduction.tex
new initialisms + replacement

@@
-Model-driven engineering (MDE) has traditionally adopted a strict hierarchical two-level approach, where metamodels reside in a certain meta-level, and models are created one meta-level below by using types from the metamodel.
-Notable examples of this approach are the widespread Eclipse Modeling Framework (EMF)~\citep{emf} and the Object Management Group Meta-Object Facility(MOF)~\citep{omg2013mof}.
+Model-driven engineering (\MDE) has traditionally adopted a strict hierarchical two-level approach, where metamodels reside in a certain meta-level, and models are created one meta-level below by using types from the metamodel.
+Notable examples of this approach are the widespread Eclipse Modeling Framework (\EMF)~\citep{emf} and the Object Management Group Meta-Object Facility(\MOF)~\citep{omg2013mof}.
 
@@
 Our solution is based on the model federation infrastructure, namely, the use of virtual models and links.
-Concretely, we use OpenFlexo and its internal Flexo Modeling Language (FML), a language to create, link and manage virtual models, in order to solve the MULTI process challenge while the federation aspect is used solely as a means to provide tooling for the resulting process language.
+Concretely, we use OpenFlexo and its internal Flexo Modeling Language (\FML), a language to create, link and manage virtual models, in order to solve the MULTI process challenge while the federation aspect is used solely as a means to provide tooling for the resulting process language.
 Our solution, which is fully implemented and executable, meets all the requirements.

### 2021-05-28 4-Technology.tex
new initialisms + replacement

@@
 paradigm while a virtual model is built using the Flexo Modeling
-Language (FML). Each federated model can be viewed as an autonomous
+Language (\FML). Each federated model can be viewed as an autonomous
 component that may evolve with its own tooling. The virtual models
 serve as control components.
-In this paper, we mainly use the conceptual models and more precisely FML its domain specific modeling language. The only use of the federation aspect is on the tooling made for the process language.
+In this paper, we mainly use the conceptual models and more precisely \FML its domain specific modeling language. The only use of the federation aspect is on the tooling made for the process language.
 
@@
     \includegraphics[width=\columnwidth]{Figures/FMLCoreModel.pdf}
-    \caption{The FML core metamodel}
+    \caption{The \FML core metamodel}
     \label{fig:mm}
@@
 
-FML simplified metamodel is provided in Figure~\ref{fig:mm}. It is designed to reify models as virtual models. A virtual model
+\FML simplified metamodel is provided in Figure~\ref{fig:mm}. It is designed to reify models as virtual models. A virtual model
 is composed of a set of \emph{concepts}, while itself being a concept.
@@
 \emph{behaviors}. A parallel to object-oriented approach can be useful
-to understand FML\footnote{Some aspects of FML do not exist in object oriented approach.}. A concept corresponds to a class, its roles to the
+to understand \FML\footnote{Some aspects of \FML do not exist in object oriented approach.}. A concept corresponds to a class, its roles to the
 attributes of the class and its behaviors to the methods of the class.
@@
 a \emph{technology adapter}\footnote{It is a reusable library that defines
-connections between the FML execution engine and a particular
+connections between the \FML execution engine and a particular
 technological space.}.
 
-FML is designed to define not only the structure of virtual models but
+\FML is designed to define not only the structure of virtual models but
 also to define the collection of actions an engineer can perform on
@@
 
-When the FML execution engine runs a federation, it creates virtual
+When the \FML execution engine runs a federation, it creates virtual
 model instances containing concept instances. Some concept instances
@@
 Openflexo\footnote{\url{https://github.com/openflexo-team}}, is
-developed as an open source initiative. This tool offers a FML
+developed as an open source initiative. This tool offers a \FML
 execution engine with an interactive virtual model design environment.
@@
 offers some mature technology adapters (\emph{docx} and \emph{excel}
-for documents, EMF and OWL for modeling languages, JDBC for databases,
+for documents, \EMF and OWL for modeling languages, JDBC for databases,
 REST and XMLRPC for external services and one for diagramming tools)

### 2021-05-28 5-Analysis.tex
new initialisms + replacement

@@
 Like any modeling approaches, these different steps could also be achieved in any order and iteratively.
-But our goal during the challenge's analysis remains to produce a FML VirtualModels architecture as a federation of VirtualModels.
+But our goal during the challenge's analysis remains to produce a \FML VirtualModels architecture as a federation of VirtualModels.
 As shown in figure~\ref{fig:MultilevelArchitecture}, we defined two VirtualModels (Base and ACME processes abstractions) instantiatied by two VirtualModel instances (XSure and ACME processes).
@@
 
-\item the vertical axis is viewed as the linguistic instantiation, relative to the use of the FML language.
-The VirtualModel, defined as the core definition metamodel, generates a model founded on the instantiation mechanism of the FML Language.
+\item the vertical axis is viewed as the linguistic instantiation, relative to the use of the \FML language.
+The VirtualModel, defined as the core definition metamodel, generates a model founded on the instantiation mechanism of the \FML Language.
 This mechanism is the classical Object/Instance mechanism of the object languages but without any constraint on the referenced VirtualModel, \ie metamodel.
@@
 The second one, Acme metamodel, is extended from the previous model to integrate the Acme definitions.
-Our FML language defines multiple inheritance concept between VirtualModels as illustrated in the section \ref{sec:AcmeSoftwareDevelopmentProcess}.
+Our \FML language defines multiple inheritance concept between VirtualModels as illustrated in the section \ref{sec:AcmeSoftwareDevelopmentProcess}.
 

### 2021-05-28 6-ModelPresentation.tex
new initialisms + replacement

@@
 
-Figure \ref{fig:BaseMetamodel} represents this base metamodel with a \UML-like formalism, well adapted to represent object-oriented FML concepts and their instances.
-Base conceptual element is called a \textit{FlexoConcept} in FML language, and is similar to the concept of class in object-oriented modeling.
+Figure \ref{fig:BaseMetamodel} represents this base metamodel with a \UML-like formalism, well adapted to represent object-oriented \FML concepts and their instances.
+Base conceptual element is called a \textit{FlexoConcept} in \FML language, and is similar to the concept of class in object-oriented modeling.
 The \UML representation of class is used to represent a \textit{FlexoConcept}, while cardinalities are shown using the \UML-like classical rules.
@@
 
-Following listing shows an excerpt of FML code modeling some core concepts of process modeling base metamodel.
+Following listing shows an excerpt of \FML code modeling some core concepts of process modeling base metamodel.
 
@@
 Actor concept defines a boolean attribute called \textit{isSenior}, while \textit{TaskType} defines an additional \textit{isCritical} boolean attribute, indicating that some instance of \textit{TaskType} are flagged as critical and must be performed by senior actors.
-To fullfill \textbf{P9} requirement, a supplementary constraint is required for \textit{TaskType} and is captured through this invariant expressed in FML language:
+To fullfill \textbf{P9} requirement, a supplementary constraint is required for \textit{TaskType} and is captured through this invariant expressed in \FML language:
 

### 2021-05-28 8-Discussion.tex
new initialisms + replacement

@@
 We could also have taken advantage of the model federation to connect our solution to other tools 
-dedicated to process edition and to process execution (\eg a BPMN engine). However we 
+dedicated to process edition and to process execution (\eg a \BPMN engine). However we 
 already had all the necessary components to provide the aforementioned graphical tools.

### 2021-05-28 9-Relatedworks.tex
replace eg by macro
new initialisms + replacement

@@
 DeepTelos is based on the use of the PowerType \citep{atkinson2001essence} pattern.
-As with Openflexo and FML, they do not use explicit (numbered) levels nor potency.
+As with Openflexo and \FML, they do not use explicit (numbered) levels nor potency.
 
@@
 Similarly, \cite{dmla2019} use the Dynamic Multi-Layer Algebra (DMLA) \citep{dmla2017} in order to solve the process challenge.
-DMLA is a level-blind modeling framework which is fully customizable (e.g., different types of instantiaton may be implemented).
+DMLA is a level-blind modeling framework which is fully customizable (\eg different types of instantiaton may be implemented).
 Their proposed solution separates the process challenge in two separate domains, namely, the tasks definition domain and the process definition domain.
@@
 
-\noteSM{Say something about constraints in FML}
+\noteSM{Say something about constraints in \FML}

### 2021-05-28 3-Introduction.tex
replace citep by parencite

@@
 Model-driven engineering (\MDE) has traditionally adopted a strict hierarchical two-level approach, where metamodels reside in a certain meta-level, and models are created one meta-level below by using types from the metamodel.
-Notable examples of this approach are the widespread Eclipse Modeling Framework (\EMF)~\citep{emf} and the Object Management Group Meta-Object Facility(\MOF)~\citep{omg2013mof}.
+Notable examples of this approach are the widespread Eclipse Modeling Framework (\EMF)~\parencite{emf} and the Object Management Group Meta-Object Facility(\MOF)~\parencite{omg2013mof}.
 
@@
 Contrary to the strict approach, multi-level modeling advocates the use of a flexible number of levels as well as a flexibilization of the relations between them.
-This paradigm is gaining traction lately within the modeling community as evidenced by the contribution of many new multi-level modeling approaches and tools such as Melanee~\citep{melanee}, MetaDepth~\citep{metadepth}, MultEcore~\citep{multecore2016}, DeepTelos\citep{deeptelos2016} or DMLA~\citep{dmla2017}.
+This paradigm is gaining traction lately within the modeling community as evidenced by the contribution of many new multi-level modeling approaches and tools such as Melanee~\parencite{melanee}, MetaDepth~\parencite{metadepth}, MultEcore~\parencite{multecore2016}, DeepTelos\parencite{deeptelos2016} or DMLA~\parencite{dmla2017}.
 Taking advantage of this vibrant state of affairs, and in order to foster discussion and enable comparison between competing approaches, a multi-level modeling challenge has been created.
@@
 Solutions to the \emph{Multi-level Process Challenge} must fulfill a number of requirements for a process representation defined at an abstract process-definition level and at various more concrete domain-specific levels, resulting in a multi-level hierarchy of related models.
-We provide a solution based on model federation~\citep{Golra2016-federation}.
+We provide a solution based on model federation~\parencite{Golra2016-federation}.
 

### 2021-05-28 3-Introduction.tex
spacing

--- 2021-05-28 3-Introduction.tex
@@
 Model-driven engineering (\MDE) has traditionally adopted a strict hierarchical two-level approach, where metamodels reside in a certain meta-level, and models are created one meta-level below by using types from the metamodel.
-Notable examples of this approach are the widespread Eclipse Modeling Framework (\EMF)~\citep{emf} and the Object Management Group Meta-Object Facility(\MOF)~\citep{omg2013mof}.
+Notable examples of this approach are the widespread Eclipse Modeling Framework (\EMF)~\parencite{emf} and the Object Management Group Meta-Object Facility (\MOF)~\parencite{omg2013mof}.
 
@@
 In order to tackle this problem, the \emph{multi-level} modeling paradigm have been introduced.
-Contrary to the strict approach, multi-level modeling advocates the use of a flexible number of levels as well as a flexibilization of the relations between them.
-This paradigm is gaining traction lately within the modeling community as evidenced by the contribution of many new multi-level modeling approaches and tools such as Melanee~\citep{melanee}, MetaDepth~\citep{metadepth}, MultEcore~\citep{multecore2016}, DeepTelos\citep{deeptelos2016} or DMLA~\citep{dmla2017}.
+Contrary to the strict approach, multi-level modeling advocates the use of a flexible number of levels as well as more flexible relations between them.
+This paradigm is gaining traction lately within the modeling community as evidenced by the contribution of many new multi-level modeling approaches and tools such as Melanee~\parencite{melanee}, MetaDepth~\parencite{metadepth}, MultEcore~\parencite{multecore2016}, DeepTelos~\parencite{deeptelos2016} or DMLA~\parencite{dmla2017}.
 Taking advantage of this vibrant state of affairs, and in order to foster discussion and enable comparison between competing approaches, a multi-level modeling challenge has been created.
@@
 Solutions to the \emph{Multi-level Process Challenge} must fulfill a number of requirements for a process representation defined at an abstract process-definition level and at various more concrete domain-specific levels, resulting in a multi-level hierarchy of related models.
-We provide a solution based on model federation~\citep{Golra2016-federation}.
+We provide here a solution based on model federation~\parencite{Golra2016-federation}.
 

### 2021-05-28 2-Abstract.tex
synchro

@@
 We present a solution based on model federation and discuss the advantages and limitations of using this approach for multi-level modeling.
-Concretely, we use virtual models and more precisely the Flexo Modeling Language (FML) that serves to describe them as the main building block in order to solve the process modeling challenge whereas the federation aspect is used in order to provide editing tooling for the resulting process language.
+Concretely, we use virtual models and more precisely the Flexo Modeling Language (FML) that serves to describe them as the main building block in order to solve the process modeling challenge whereas the federation aspect is used as a means to provide editing tooling for the resulting process language.
 Our solution is fully implemented with the Openflexo framework.

### 2021-05-28 3-Introduction.tex
synchro

--- 2021-05-28 3-Introduction.tex
@@
-Model-driven engineering (\MDE) has traditionally adopted a strict hierarchical two-level approach, where metamodels reside in a certain meta-level, and models are created one meta-level below by using types from the metamodel.
-Notable examples of this approach are the widespread Eclipse Modeling Framework (\EMF)~\citep{emf} and the Object Management Group Meta-Object Facility(\MOF)~\citep{omg2013mof}.
+Model-driven engineering (MDE) has traditionally adopted a strict hierarchical two-level approach, where metamodels reside in a certain meta-level, and models are created one meta-level below by using types from the metamodel.
+Notable examples of this approach are the widespread Eclipse Modeling Framework (EMF)~\citep{emf} and the Object Management Group Meta-Object Facility(MOF)~\citep{omg2013mof}.
 
-This strict approach shows its limitations when modeling complex domains requiring more than one level of specialization, \eg to adapt the model to application sub domains.
-Indeed, the two-level approach fails to acknowledge and support: 1) the existence of different forms of classification and/or instantiation (\eg ontological vs linguistic); and 2) the duality type-object for model elements.
+This strict approach shows its limitations when modeling complex domains requiring more than one level of specialization, e.g., to adapt the model to application sub domains.
+Indeed, the two-level approach fails to acknowledge and support: 1) the existence of different forms of classification and/or instantiation (e.g., ontological vs linguistic); and 2) the duality type-object for model elements.
 Therefore, while it may still be used for the aforementioned complex modeling tasks, the strict approach entails the introduction of accidental complexity in both the modeling process and the resulting modeling artefacts.
@@
 Our solution is based on the model federation infrastructure, namely, the use of virtual models and links.
-Concretely, we use OpenFlexo and its internal Flexo Modeling Language (\FML), a language to create, link and manage virtual models, in order to solve the MULTI process challenge while the federation aspect is used solely as a means to provide tooling for the resulting process language.
+Concretely, we use OpenFlexo and its internal Flexo Modeling Language (FML), a language to create, link and manage virtual models, in order to solve the MULTI process challenge while the federation aspect is used solely so as to provide tooling for the resulting process language.
 Our solution, which is fully implemented and executable, meets all the requirements.

### 2021-05-28 6-ModelPresentation.tex
synchro

--- 2021-05-28 6-ModelPresentation.tex
@@
 Depending of its type, a \textit{Gateway} defines one or more inputs and one or more outputs, depending of its underlying execution semantics.
-A \textit{ProcessType} additionally exposes unique initial \textit{TaskType} and a collection of final \textit{TaskType} with both attributes \textit{initialTaskType} (single cardinality) and \textit{finalTaskType} (cardinality 0..*) (\textbf{P3}).
-TaskType exposes a creator attribute as an association relation to \textit{Actor} concept (\textbf{P4}), which is in the same conceptual level.
+A \textit{ProcessType} additionally exposes unique initial \textit{TaskType} and a collection of final \textit{TaskType} with both roles \textit{initialTaskType} (single cardinality) and \textit{finalTaskType} (cardinality 0..*) (\textbf{P3}).
+TaskType exposes a creator role as a reference to \textit{Actor} concept (\textbf{P4}), which is at the same conceptual level.
 
@@
 This scheme allows to rely on FML dynamic binding mechanism to delegate to types the responsability for instances creation.
-An instance of \textit{Process} references a unique \textit{ProcessType} trough the specialized \textit{/type} relation.
+An instance of \textit{Process} references a unique \textit{ProcessType} trough the specialized \textit{/type} role.
 Each instance of \textit{TaskType} is ontologically instantiated with a \textit{Task} (\textbf{P11}), using the same pattern where \textit{TaskType} has the responsability to manage this ontological instantiation.
@@
 
-Same pattern applies for artifacts used and produced with relations \textit{usedArtifacts}, \textit{producedArtifacts} and \textit{performingActors} defined in \textit{Task} concept (\textbf{P13}).
-An instance of \textit{Artifact}, defined as a specialization of \textit{MultiInstance} concept, references a set of \textit{ArtifactType} through the specialized \textit{/types} relation (\textbf{P14} and \textbf{P16}).
+Same pattern applies for artifacts used and produced with roles \textit{usedArtifacts}, \textit{producedArtifacts} and \textit{performingActors} defined in \textit{Task} concept (\textbf{P13}).
+An instance of \textit{Artifact}, defined as a specialization of \textit{MultiInstance} concept, references a set of \textit{ArtifactType} through the specialized \textit{/types} role (\textbf{P14} and \textbf{P16}).
 Likewise, concept \textit{Actor} is defined as a specialization of \textit{MultiInstance} concept and references a set of \textit{ActorType} through the specialized \textit{/types} relation (\textbf{P15}).
@@
 Semantics is unclear relatively to instantiation policy for artifacts.
-We assume that task execution implies that for each used and produced \textit{artifact type} defined in related \textit{task type}, it exists at least one \textit{artifact} declaring required ontologic instantiation (through \textit{type} relation).
+We assume that task execution implies that for each used and produced \textit{artifact type} defined in related \textit{task type}, it exists at least one \textit{artifact} declaring required ontologic instantiation (through \textit{type} role).
 Following excerpt of FML code shows partial implementation of this.
@@
 
-Authorization for an actor to perform a task (\textbf{P17}) is captured either through the relations \textit{allowedActors} and \textit{allowedActorTypes} defined in \textit{TaskType}.
+Authorization for an actor to perform a task (\textbf{P17}) is captured either through the roles \textit{allowedActors} and \textit{allowedActorTypes} defined in \textit{TaskType}.
 This mechanism is completed by following behaviours \textit{isAuthorizedActor(Actor)}, \textit{isValidActor(Actor)} and \textit{isValidActorType(ActorType)} defined in \textit{TaskType} :
@@
 Base metamodel as described in previous section is too generic to capture all domain-specific aspects of this use case.
-We choose to complete the architectual hierarchy with a specific metamodel, specific to Acme software development process, as shown in figure \ref{fig:AcmeArchitecture}.
-\textit{Acme metamodel} inherits and specializes base metamodel, and \textit{Acme model} is defined as an instance of \textit{Acme metamodel}.
+We choose to complete the architectual hierarchy with a specific VirtualModel, specific to Acme software development process metamodel, as shown in figure \ref{fig:AcmeArchitecture}.
+\textit{Acme} metamodel inherits and specializes base metamodel, and \textit{Acme model} is defined as an instance of \textit{Acme} metamodel.
+The metamodel inheritance is implemented as VirtualModel's inheritance.
 Any instance of Acme model is either an instance of a concept defined in base metamodel, or a concept defined in specialized Acme metamodel, which may or not inherit from a concept defined in base metamodel.

### 2021-05-28 7-RequirementsSatisfaction.tex
synchro from overleaf
synchro

@@
 
-\noteSylvain{
-Je voulais dire ici qu'on avait construit la solution pas à pas en suivant les requirements les uns après les autres, et qu'on avait à chaque fois veillé à traduire et respecter les requirements.
+%\noteSylvain{
+%Je voulais dire ici qu'on avait construit la solution pas à pas en suivant les requirements les uns après les autres, et qu'on avait à chaque fois veillé à traduire et respecter les requirements.
 Ca me semble très lourd de tout rejustifier ici.
 
-On peut cependant expliciter les différentes techniques employées:
-\begin{itemize}
-    \item Requirements satisfaits par conformance syntaxique: obligation de respecter le modèle structurel
-    \item Requirements satisfaits par ajout de contraintes: on peut instantier quelque chose de syntaxiquement correct, mais qui viole certaines contraintes (sur l'outillage les croix rouges)
-    \item Requirements satisfaits par l'outillage: ajout de comportements de manipulation du modèle qui "forcent" à respecter les contraintes d'intégrité resultant des requirements.
-\end{itemize}
+%On peut cependant expliciter les différentes techniques employées:
+%\begin{itemize}
+%    \item Requirements satisfaits par conformance syntaxique: obligation de respecter le modèle structurel
+%    \item Requirements satisfaits par ajout de contraintes: on peut instantier quelque chose de syntaxiquement correct, mais qui viole certaines contraintes (sur l'outillage les croix rouges)
+%    \item Requirements satisfaits par l'outillage: ajout de comportements de manipulation du modèle qui "forcent" à respecter les contraintes d'intégrité resultant des requirements.
+%\end{itemize}
 
-Peut-être peut-on faire un tableau qui rescence comment on a satisfait chaque contrainte ? Est ce que c'est intéressant ?
-}
+%Peut-être peut-on faire un tableau qui rescence comment on a satisfait chaque contrainte ? Est ce que c'est %intéressant ?
+%}
+%\todo[inline]{OK pour le tableau synthétique +un petit peu de texte + une mini intro qui référence la section précédente (vu que les req ont été couverts dans la section précédente ; pas de fusion des deux sections}
 
-\todo[inline]{OK pour le tableau synthétique +un petit peu de texte + une mini intro qui référence la section précédente (vu que les req ont été couverts dans la section précédente ; pas de fusion des deux sections}
+CT : Conceptualization
+SP : Specialization
+CP : Composition
+LI : Linguistic instantiation
+OI : Ontologic instantiation
+BM : Behavioural modeling
 
-\begin{tabular}{|l|c|r|}
+\begin{figure*}
+ \centering
+\begin{tabular}{|c|c|c|c|c|c|c|c|c|}
    \hline
-   colonne 1 & \multicolumn{2}{c|}{colonnes 2 et 3} \\
+    & \multicolumn{6}{c|}{Syntactic conformance} & Constraints checking & Tooling\\
    \hline
-   1.1 & 1.2 & 1.3 \\
+                 & Conc.      & Spec.      & Comp.      & L.Inst     & O.Inst     & B.Mod      &            & \\
   \hline
-   2.1 & 2.2 & \checkmark \\
+    \textbf{P1}  & \checkmark &            & \checkmark &            &            &            &            & \\
+    \textbf{P2}  & \checkmark & \checkmark & \checkmark &            &            &            &            & \\
+    \textbf{P3}  &            &            & \checkmark &            &            &            &            & \\
+    \textbf{P4}  & \checkmark &            &            &            & \checkmark &            &            & \\
+    \textbf{P5}  & \checkmark &            & \checkmark &            &            &            &            & \\
+    \textbf{P6}  & \checkmark &            & \checkmark &            & \checkmark &            &            & \\
+    \textbf{P7}  & \checkmark &            & \checkmark &            &            &            &            & \\
+    \textbf{P8}  &            &            & \checkmark &            &            &            &            & \\
+    \textbf{P9}  &            &            & \checkmark &            &            &            & \checkmark & \\
+    \textbf{P10} & \checkmark &            &            &            & \checkmark &            &            & \\
+    \textbf{P11} & \checkmark &            & \checkmark &            & \checkmark &            &            & \\
+    \textbf{P12} &            &            & \checkmark &            &            &            &            & \\
+    \textbf{P13} & \checkmark &            & \checkmark &            & \checkmark &            &            & \\
+    \textbf{P14} & \checkmark &            &            &            & \checkmark &            &            & \\
+    \textbf{P15} & \checkmark &            &            &            & \checkmark &            &            & \\
+    \textbf{P16} & \checkmark &            &            &            & \checkmark &            &            & \\
+    \textbf{P17} &            &            &            &            &            & \checkmark & \checkmark & \checkmark \\
+    \textbf{P18} &            &            &            &            &            & \checkmark &            & \\
+    \textbf{P19} &            & \checkmark & \checkmark &            &            & \checkmark &            & \\
   \hline
 \end{tabular}
+     \caption{Requirements satisfaction for base metamodel}
+    \label{tab:RequirementsSatisfactionBaseMetamodel}
+\end{figure*}
+
+\begin{figure*}
+ \centering
+\begin{tabular}{|c|c|c|c|c|c|c|c|c|}
+   \hline
+    & \multicolumn{6}{c|}{Syntactic conformance} & Constraints checking & Tooling\\
+   \hline
+                 & Conc.      & Spec.      & Comp.      & L.Inst     & O.Inst     & B.Mod      &            & \\
+  \hline
+    \textbf{S1}  &            &            &            & \checkmark &            &            &            & \\
+    \textbf{S2}  &            &            &            & \checkmark &            &            &            & \\
+    \textbf{S3}  & \checkmark & \checkmark & \checkmark & \checkmark &            &            &            & \checkmark \\
+    \textbf{S4}  & \checkmark & \checkmark & \checkmark & \checkmark &            &            &            & \\
+    \textbf{S5}  &            & \checkmark &            & \checkmark &            & \checkmark &            & \\
+    \textbf{S6}  &            & \checkmark &            & \checkmark &            & \checkmark & \checkmark & \\
+    \textbf{S7}  &            &            &            & \checkmark &            & \checkmark &            & \\
+    \textbf{S8}  &            &            &            & \checkmark &            &            &            & \\
+    \textbf{S9}  &            &            &            &            &            & \checkmark & \checkmark & \\
+    \textbf{S10} & \checkmark & \checkmark & \checkmark &            &            &            &            & \\
+    \textbf{S11} &            &            &            & \checkmark &            &            &            & \checkmark \\
+    \textbf{S12} &            &            &            & \checkmark &            &            &            & \checkmark \\
+    \textbf{S13} &            &            &            & \checkmark &            &            &            & \\
+  \hline
+\end{tabular}
+     \caption{Requirements satisfaction for Acme software engineering process}
+    \label{tab:RequirementsSatisfactionAcme}
+\end{figure*}

### 2021-05-28 3-Introduction.tex

--- 2021-05-28 3-Introduction.tex
@@
 Our solution is based on the model federation infrastructure, namely, the use of virtual models and links.
-Concretely, we use OpenFlexo and its internal Flexo Modeling Language (\FML), a language to create, link and manage virtual models, in order to solve the MULTI process challenge while the federation aspect is used solely as a means to provide tooling for the resulting process language.
+Concretely, we use OpenFlexo and its internal Flexo Modeling Language (\FML), a language to create, link and manage virtual models, in order to solve the MULTI process challenge while the federation aspect is used solely so as to provide tooling for the resulting process language.
 Our solution, which is fully implemented and executable, meets all the requirements.
+
 %Links among models can be within a same level of abstraction or across levels of abstractions which gives,

### 2021-05-28 6-ModelPresentation.tex

@@
 
-Figure \ref{fig:BaseMetamodel} represents this base metamodel with a UML-like formalism, well adapted to represent object-oriented FML concepts and their instances.
-Base conceptual element is called a \textit{FlexoConcept} in FML language, and is similar to the concept of class in object-oriented modeling.
-The UML representation of class is used to represent a \textit{FlexoConcept}, while cardinalities are shown using the UML-like classical rules.
+Figure \ref{fig:BaseMetamodel} represents this base metamodel with a \UML-like formalism, well adapted to represent object-oriented \FML concepts and their instances.
+Base conceptual element is called a \textit{FlexoConcept} in \FML language, and is similar to the concept of class in object-oriented modeling.
+The \UML representation of class is used to represent a \textit{FlexoConcept}, while cardinalities are shown using the \UML-like classical rules.
 In the diagram, we choose to represent the role associated with the \textit{FlexoConcept} in the class square and the role type is referenced by the arrow in front of the role name.

### 2021-05-28 5-Analysis.tex
sync vers GitHub

--- 2021-05-27 5-Analysis.tex
@@
 
-During the requirement analysis of the challenge, the general model federation approach remains the foundation of our reflections and modeling intentions.
+During the requirement analysis of the challenge, the foundation of our reflections and modeling intentions is guided by the general model federation approach.
 
-First, the federation approach is mainly based on modeling relationships among several models, independently of abstraction levels and model architectures.
-During the next step of our approach, we try to take into account the reusability of the relationships by identifying the meaning, the semantics, associated with each relationship, in a goal of concept reification with their behavior.
-The last step is to organize or structure these concepts to improve reusability and extensibility, to create VirtualModels in our Openflexo framework.
+In a first step, the federation approach is mainly based on modeling relationships between several models, independently of their abstraction level and the model architecture.
 
-Like any modeling approaches, these different steps could also be achieved in any order and iteratively.
-But our goal during the challenge's analysis remains to produce a FML VirtualModels architecture as a federation of VirtualModels.
-As shown in figure~\ref{fig:MultilevelArchitecture}, we defined two VirtualModels (Base and ACME processes abstractions) instantiatied by two VirtualModel instances (XSure and ACME processes).
-As the figure shows the VirtualModels play the role of metamodels, in the sens of concept definition with a level-blind approach.
-The VirtualModel instances play the role of models conform to metamodel definitions.
-In the rest of the paper, we keep the the vocabulary metamodels and models as to simplify the presentation.
-The resulted architecture follows the way the challenge is presented and this organisation allows for the possibility of flexible extensions.
+During the next step of our approach, we try to take into account the reusability of the relationships by identifying the semantics of each relationship.
+The goal is the identification of concepts with their behavior.
+The last step is to organize or structure these concepts to improve reusability and extensibility, to create virtual models in \FML terminology.
 
-\begin{figure}
+Like in lot of modeling approaches, these  steps could also be achieved in any order and iteratively.
+But our goal during the challenge's analysis remains to produce a \FML virtual model architecture for our federation.
+As shown in figure~\ref{fig:MultilevelArchitecture}, we defined two virtual models (Base and Acme processes abstractions) instantiated by two virtual model instances (XSure and Acme processes).
+As the figure shows the virtual models play the role of metamodels, in the sense of concept definition with a level-blind approach.
+The virtual model instances play the role of models conforming to the metamodel definitions.
+In the rest of the paper, we use the term metamodel and model to simplify the presentation.
+The resulting architecture follows the way the challenge is presented and this organisation allows for the possibility of flexible extensions.
+
+\begin{figure}[t]
     \centering
     \includegraphics[width=0.7 \columnwidth]{Figures/MultilevelArchitecture.pdf}
-    \caption{Multilevel architecture for the Process Challenge}
+    \caption{Multilevel architecture of our solution}
     \label{fig:MultilevelArchitecture}
@@
 
-Also as our understanding, the use case analysis leads to identify two main modeling axis (as presented in figure \ref{fig:LinguisticAndOntologicInstantiation}):
+Our analysis of the use case leads to identify two main modeling axis (as presented in figure \ref{fig:LinguisticAndOntologicInstantiation}):
 \begin{itemize}
-    \item the horizontal axis is characterized as the ontological instantiation axis, in the sense that the domain type definition (ProcessType as example) is referenced by an instance definition (Process, a.e).
+    \item The horizontal axis is characterized as the ontological instantiation axis, in the sense that the domain type definition (\ie \texttt{ProcessType}) is referenced by an instance definition (\ie \texttt{Process}).
 In our base metamodel, each domain type definition is referenced by its instance definition, as developed in the section \ref{sec:ProcessEnactment}
 
-\item the vertical axis is viewed as the linguistic instantiation, relative to the use of the FML language.
-The VirtualModel, defined as the core definition metamodel, generates a model founded on the instantiation mechanism of the FML Language.
-This mechanism is the classical Object/Instance mechanism of the object languages but without any constraint on the referenced VirtualModel, i.e metamodel.
-The set of resulted instances comes from several VirtualModels
-from any abstraction level as detailed in section \ref{sec:AcmeSoftwareDevelopmentProcess}
+\item The vertical axis is viewed as the linguistic instantiation, relative to the use of the \FML language.
+The virtual model, defined as the core definition metamodel, generates a model founded on the instantiation mechanism of the \FML Language.
+This mechanism is similar to the classical object/instance mechanism of the object languages but without any constraint on the referenced virtual model, \ie metamodel.
+The set of resulting instances comes from several virtual models
+of any abstraction level as detailed in section \ref{sec:AcmeSoftwareDevelopmentProcess}
 
@@
 
-Based on this approach, we organize our set of models as presented by the figure \ref{fig:MultilevelArchitecture}.
-The resulted multi-level architecture is organized as a set of VirtualModels.
-The first one is the core concepts metamodel including the definition of Processes and Tasks.
-The second one, Acme metamodel, is extended from the previous model to integrate the Acme definitions.
-Our FML language defines multiple inheritance concept between VirtualModels as illustrated in the section \ref{sec:AcmeSoftwareDevelopmentProcess}.
+Based on this approach, we organize our set of models following the architecture of the figure \ref{fig:MultilevelArchitecture}.
+The resulting multi-level architecture is organized in two virtual models, one for the core concepts metamodel containing the definition of Processes and Tasks, and one, the Acme metamodel, extending the previous model to integrate the Acme definitions.
+\FML language defines multiple inheritance concept between virtual models as illustrated in the section~\ref{sec:AcmeSoftwareDevelopmentProcess}.
 
-Finally as explained previously, the defined level for the Acme  and the XSure models is the result of the instantiation of VirtualModels.
-In our approach this level can't be specialized or extended but all the other VirtualModels could be extended  by any concepts, as a new federated model.
-Also, the instance level can be instantiated from any VirtualModel, which represent any metamodel level.
+Finally as explained previously, the defined level for the Acme  and the XSure models is the result of the instantiation of virtual models.
+In our approach this virtual model instance level can't be specialized or extended but all the other virtual models could be extended by any concepts, as a new federated model.
+Also, the instance level can be instantiated from any virtual model, which represent any metamodel level.
+
+\noteAntoine{Pourquoi S9 et S13 sont ambigus, quels choix on a fait...}
+\todo[inline]{noms pas dans la spec, complètement implicite ; }
+
+We now analysis a few specifications that deserve a comment for the choices we made.
+
+\begin{itemize}
+    \item We added an implicit specification (S0) saying that all things have names.
+    \item (S9) It is unclear when a tested artefact must be associated to its test report. At the creation of the artefact or of its report? \todo[inline]{We choose\dots}
+    \item (S13) [linked to P9] \todo[inline]{We choose ti interpret \dots}
+\end{itemize}
 
@@
 
-~\\
-~\\
-
-\todo[inline]{quelles hypothèses supplémentaires ?}
+%\todo[inline]{quelles hypothèses supplémentaires ?}
 %\todo[inline]{+ architecture de la modélisation (ce qui se trouve dans le ppt), le détail se retrouvera dans la section suivante}

### 2021-05-28 6-ModelPresentation.tex
sync vers GitHub

@@
 The presentation of our solution to the \mpc follows a systematic methodology where modeling choices are introduced step by step, as the requirements are stated.
-Satisfaction of requirements are explained throughout this presentation.
+Satisfaction of requirements is explained throughout this presentation.
 
@@
 
-The designed multilevel architecture is shown in Figure \ref{fig:MultilevelArchitecture} overview and captures the two use cases described in the Process Challenge.
+The designed multilevel architecture shown in Figure~\ref{fig:MultilevelArchitecture} captures the two use cases described in the Process Challenge.
 
@@
 
-In this section, we present and illustrate the base metamodel presented in figure \ref{fig:BaseMetamodel} with \textit{XSure} insurance domain use case, whose a partial description was provided in the challenge description.
-Note that the capture of the requirements \textbf{P1} to \textbf{P19} in the context of XSure insurance domain are straightforward implementable while instantiating \textit{XSure model}, as an instance of base metamodel (the left side of figure \ref{fig:MultilevelArchitecture}).
+In this section, we present the base metamodel presented in figure~\ref{fig:BaseMetamodel} with \textit{XSure} insurance domain use case, whose partial description was provided in the challenge description.
+The requirements \textbf{P1} to \textbf{P19} of XSure insurance domain are straightforwardly implemented by instantiating the \emph{XSure model}, as an instance of base metamodel (the left side of figure~\ref{fig:MultilevelArchitecture}).
 
-Figure \ref{fig:BaseMetamodel} represents this base metamodel with a \UML-like formalism, well adapted to represent object-oriented \FML concepts and their instances.
-Base conceptual element is called a \textit{FlexoConcept} in \FML language, and is similar to the concept of class in object-oriented modeling.
-The \UML representation of class is used to represent a \textit{FlexoConcept}, while cardinalities are shown using the \UML-like classical rules.
-In the diagram, we choose to represent the role associated with the \textit{FlexoConcept} in the class square and the role type is referenced by the arrow in front of the role name.
+Figure~\ref{fig:BaseMetamodel} represents this base metamodel with a \UML-like formalism, well adapted to represent \FML concepts and their instances.
+A \textit{Concept} of \FML is represented by a \UML class.
+The roles are represented by a combination of a name in their concept container and an arrow from their name to their type.
+The cardinality follow the \UML practice.
+For example, the role \texttt{parentActorTypes} of the concept \texttt{ActorType} has the type \texttt{ActorType} and the * cardinality.
 
-Our proposition relies on ontologic instantiation as presented in figure \ref{fig:LinguisticAndOntologicInstantiation}, with a common root concept \textit{ModelingElement}.
+Our proposition relies on ontologic instantiation as presented in figure \ref{fig:LinguisticAndOntologicInstantiation}, with a common root concept \texttt{ModelingElement}.
 Two types of ontological instantiation are needed to meet the requirements of the challenge.
-One provides that some instances conform to only one type (\textit{Process} and \textit{Task} do respectively conform to one \textit{ProcessType} and \textit{TaskType}) while some entities define their conformity to several types (\textit{Actor} and \textit{Artifact} do respectively conform to several \textit{ActorType} and \textit{ArtifactType}).
+One provides that some instances conform to only one type (\texttt{Process} and \texttt{Task} do respectively conform to one \texttt{ProcessType} and \texttt{TaskType}).
+While some entities define their conformity to several types (\textit{Actor} and \textit{Artifact} do respectively conform to several \textit{ActorType} and \textit{ArtifactType}).
 This instantiation is reified with the relations \textit{type} and \textit{types} between \textit{Type}, \textit{Instance} and \textit{MultiInstance} concepts.
@@
 \textit{ProcessType} is defined as a specialization of \textit{Type} \texttt{FlexoConcept}, and references a collection of \textit{TaskType} through the composition relation \textit{taskTypes} with (0..*) cardinality (\textbf{P1}).
-A \textit{TaskType} is embedded in a \textit{ProcessType} and inherits from it context.
-To illustrate this in \textit{XSure} insurance domain use case, \textit{XSure model} defines \textit{Claim Handling}, instance of \textit{ProcessType}, and \textit{Receive Claim}, \textit{Asses Claim} and \textit{Pay premium}, instances of \textit{TaskType}.
+A \textit{TaskType} is embedded in a \textit{ProcessType} and inherits from its context.
+To illustrate this in \textit{XSure} insurance domain use case, \textit{XSure model} defines \textit{Claim Handling}, instance of \textit{ProcessType}, and \textit{Receive Claim}, \textit{Assess Claim} and \textit{Pay premium}, instances of \textit{TaskType}.
 
@@
 
-Actor concept defines a boolean attribute called \textit{isSenior}, while \textit{TaskType} defines an additional \textit{isCritical} boolean attribute, indicating that some instance of \textit{TaskType} are flagged as critical and must be performed by senior actors.
+\textit{Actor} concept defines a boolean attribute called \textit{isSenior}, while \textit{TaskType} defines an additional \textit{isCritical} boolean attribute, indicating that some instance of \textit{TaskType} are flagged as critical and must be performed by senior actors.
 To fullfill \textbf{P9} requirement, a supplementary constraint is required for \textit{TaskType} and is captured through this invariant expressed in \FML language:
@@
 
-Figure \ref{fig:AcmeSoftwareDevelopmentProcess} shows a screen-capture from the tooling developed in the context of the challenge, and presents the \textit{Software Development Process} for Acme company.
+The \textit{Software Development Process} for Acme company is presented in the figure \ref{fig:AcmeSoftwareDevelopmentProcess} which is a screen-capture from the tooling developed in the context of the challenge and described in the next section.
 
@@
 This requirement is captured through the redefinition of \textit{declaresRequiredProducedArtifacts()} where programming language should also match.
-\noteSylvain{Pas la place pour mettre le code, non nécessaire si ?} \textbf{S6} is guaranteed though the \textit{language} relation defined in \textit{Code} and following invariant declared in \textit{Code}:
+\textbf{S6} is guaranteed though the \textit{language} relation defined in \textit{Code} and following invariant declared in \textit{Code}:
 
@@
 
-\textbf{S12} is somehow ambigous as it implies an implicit semantics regarding underlying business logic of process execution.
+\textbf{S12} is somehow ambiguous as it implies an implicit semantics regarding underlying business logic of process execution.
 We assume that all tasks may define an expected duration, which might be checked during process execution.
@@
 
-Base metamodel has been completed with some behaviours implementing execution semantic for executed processes.
+Base metamodel has been completed with some behaviors implementing execution semantic for executed processes.
 All tasks - instantiated from \textit{TaskType} for a given enacted process - manage a status which can be \textit{Not startable} (when not assigned to a performing actor or when required input artifacts are not available) , \textit{Startable}, \textit{Started}, \textit{Completable} (when all output artifacts are ready), \textit{Completed}).

### 2021-05-28 7-RequirementsSatisfaction.tex
synchro
sync vers GitHub

@@
 In this section, we summarize the different techniques used to satisfy them, without repeating all the justifications.
+We classify those techniques into three categories:
+\begin{itemize}
+    \item \textbf{syntactic conformance}: the requirements are structurally satisfied when the model conforms to its metamodel,
+    \item\textbf{constraints checking}: the requirements are fulfilled when all instances satisfy both syntactic conformance and the evaluation of invariants expressed for related concepts,
+    \item \textbf{tooling}: the requirements are enforced by the implementation of the associated tooling.
+\end{itemize}
+
+Table \ref{tab:RequirementsSatisfactionBaseMetamodel} summarizes requirement covering for the base metamodel illustrated by XSure insurance use case while table \ref{tab:RequirementsSatisfactionAcme} shows requirements satisfaction for Acme software engineering process.
+Syntactic conformance is more precisely classified into six subcategories:
+
+\begin{itemize}
+    \item \textbf{Conceptualization} (\textit{Conc.}): a concept carries the semantics exposed by the requirement (for example for \textbf{P1}: \textit{process type} is conceptualized in a \textit{FlexoConcept} \textit{ProcessType}).
+    \item \textbf{Specialization} (\textit{Spec.}): a concept is specialized in the sense of object-oriented modeling (for example for \textbf{P2}: \textit{Gateway} defines an abstract behaviour \texttt{execute()} and \textit{Sequencing}, \textit{AndSplit}, \textit{AndJoin}, \textit{OrSplit} and \textit{OrJoin} are defined as inherited concepts implementing specific business logic).
+    \item \textbf{Composition} (\textit{Comp.}): some concepts are in relation, or specific attributes were added to some concepts (for example for \textbf{P3}: a \textit{process type} has one \textit{initial task type}).
+    \item \textbf{Linguistic instanciation} (\textit{L.Inst.}): requirement is satisfied by the instantiation of one or more concepts (for example for \textbf{S1}: \textit{requirement analysis} is defined as an instance of \textit{SETasktype}).
+    \item \textbf{Ontologic instanciation} (\textit{O.Inst.}): fulfillment of requirement is obtained by the relation between an instance and its type definition instance (for example for \textbf{S3} : \textit{developer} has the dual concept nature in \textit{Acme metamodel} and instance nature in \textit{Acme model}).
+    \item \textbf{Behavioral modeling} (\textit{B.Mod.}): requirement is satisfied by the definition of one more object-oriented behavioural features, which may combined with constraints checking and associated tooling (for example \textbf{P17}).
+\end{itemize}
+
+%Table~\ref{tab:RequirementsSatisfactionBaseMetamodel} 
+%Table~\ref{tab:RequirementsSatisfactionAcme}
 
@@
 
-CT : Conceptualization
-SP : Specialization
-CP : Composition
-LI : Linguistic instantiation
-OI : Ontologic instantiation
-BM : Behavioural modeling
+%CT : Conceptualization
+%SP : Specialization
+%CP : Composition
+%LI : Linguistic instantiation
+%OI : Ontologic instantiation
+%BM : Behavioural modeling
 
@@
    \hline
-                 & Conc.      & Spec.      & Comp.      & L.Inst     & O.Inst     & B.Mod      &            & \\
+                 & Conc.      & Spec.      & Comp.      & L.Inst.     & O.Inst.     & B.Mod.      &            & \\
   \hline

### 2021-05-28 8-Discussion.tex
sync vers GitHub

@@
   %plan plus ou moins imposé par l'appel
-\todo[inline]{mandatory discussion aspects ↓}
+%\todo[inline]{mandatory discussion aspects ↓}
 
@@
 
-  \noteJC{pas clair pour moi ce qui est entendu par « basic modeling constructs »}
+  \noteJC{pas clair pour moi ce qui est entendu par « basic modeling constructs », ce qui est dans la partie 2 ?}
 
@@
 
-  The Openflexo approach is level-blind: ``levels'' have no specific
+  The Openflexo approach is level-agnostic: ``levels'' have no specific
   nature and there are no numbered levels.
+
+\todo[inline]{mais on peut dire que les niveaux sont empaquetés dans des virtual models ; pour passer d'un niveau à l'autre, on a des relations : héritage et instantiation}
 
@@
 
-  Due to the fact that our approach is level-blind, our solution could have had more
-  or fewer levels depending on the variations of the use case.
-However, the number of levels is very related to the problem, making the approach fixed.
-For example, given a requirement, an instance cannot be changed into a concept easily.
+  Due to the fact that our approach is level-agnostic, our solution could have had more
+  or fewer levels depending on the variations of the use case. However, the number of levels is related
+  to the problem, making the approach fixed. For example, given a requirement, an instance cannot be 
+  changed into a concept easily.
 
@@
 
-  Thanks to the level-blindness nature of the Openflexo approach, cross-level
+  Thanks to the level-agnosticity nature of the Openflexo approach, cross-level
   relashionships are not an issue. Model elements of different levels can be
@@
   %\noteJC{pas trop clair pour moi pour le moment -> N/A ? }
-  Due to the nature of our approach and its level-blindness, \emph{deep characterization} does not 
+  Due to the nature of our approach and its level-agnosticity, \emph{deep characterization} does not 
   apply to our solution. Such a mechanism could probably be encoded by specific behaviors, however 
@@
 
-  Due to the level-blind aspect of our approach, we could insert a new level,
+  Due to the level-agnostic aspect of our approach, we could insert a new level,
   for example. It would consist in creating new concepts we would link to
@@
 
-\todo[inline]{recommended discussion aspects ↓}
-\subsection{Formalism and tools}
+%\todo[inline]{recommended discussion aspects ↓}
+\subsection{Tools}
 
@@
 
-\noteJC{on a du tooling, un peu moins de formalisme ; je me demande s'il ne vaut pas mieux ne pas traiter l'aspect formalisme}
-
-Currently, there are no formalism to establish the semantics of the MLM
-technique in our solution. \textcolor{red}{To develop…}
-
 As described in section~\ref{subsec:tooling}, our solution is fully tooled with Openflexo.
 We use the Openflexo infrastructure to build models and metamodels together, and to provide 
-dedicated tooling to edit a Process type, to enact a process and to execute it.
-Being able to provide a solution and its associated tools quickly and easily is an \textcolor{red}{interesting} feature.
+dedicated tooling to edit a Process type, to enact a process and to execute it. Being able to provide a
+solution and its associated tools quickly and easily is a noticeable feature. 
 We could also have taken advantage of the model federation to connect our solution to other tools 

### 2021-05-28 8-Discussion.tex
few todos (assessment part)

@@
 
-  The Openflexo approach is level-agnostic: ``levels'' have no specific
-  nature and there are no numbered levels.
+  The Openflexo approach is level-agnostic: ``levels'' have no specific nature
+  and there are no numbered levels. In our solution, concepts of a given level
+  are grouped into a virtual model. Inheritance and instantiation allow the
+  establishment of relationships between concepts from different levels.
 
-\todo[inline]{mais on peut dire que les niveaux sont empaquetés dans des virtual models ; pour passer d'un niveau à l'autre, on a des relations : héritage et instantiation}
+%\todo[inline]{mais on peut dire que les niveaux sont empaquetés dans des virtual models ; pour passer d'un niveau à l'autre, on a des relations : héritage et instantiation}
 
@@
 
-  \subsection{Cross-level relashionships}
+  \subsection{Cross-level relationships}
 
@@
   Thanks to the level-agnosticity nature of the Openflexo approach, cross-level
-  relashionships are not an issue. Model elements of different levels can be
-  linked each other in a transparent way. %If the user has defined
-  %well-formedness constraints, then they are
+  relationships are not an issue. Model elements of different levels can be
+  linked each other in a transparent way, using inheritance or instantiation.
 
@@
   As for cross-level relationships, cross-level constraints do not have a
-  specific nature. They are like any constraint a user can define by adding a behavior to a model element. Therefore, if a constraint is mandatory when establishing a
-  relationship between elements from different levels, the user has to create
-  it manually. This can be a limit of our approach: the cost of the flexibility of our tooling is a limited number of automated behaviors.
+  specific nature. They are like any constraint a user can define by adding a
+  behavior to a model element. Therefore, if a constraint is mandatory when
+  establishing a relationship between elements from different levels, the user
+  has to create it manually. This can be a limit of our approach: the cost of
+  the flexibility of our tooling is a limited number of automated behaviors.
 
@@
 
-  Behaviors are continuously checked, ensuring the integrity of the models when changes occur. 
-  However, most of those behaviors have to be written by the users. 
-  Thus, the integrity of the contents relies on the users when they build the metamodels and the models.
-  Note that in our approach, the metamodels are built in a ad hoc way together with the models
+  Behaviors are continuously checked, ensuring the integrity of the models when
+  changes occur. However, most of these behaviors have to be written by the
+  users. Thus, the integrity of the contents relies on the users when they
+  build the metamodels and the models.  Note that in our approach, the
+  metamodels are built in a ad hoc way together with the models
   (co-construction), therefore the user also validates on a ad hoc basis.
@@
 
-  Due to the level-agnostic aspect of our approach, we could insert a new level,
-  for example. It would consist in creating new concepts we would link to
-  other ones. Then, we would add any necessary constraints on those concepts
-  and on the relashiontips linking them to the other existing concepts.
+  Due to the level-agnostic aspect of our approach, we could insert a new
+  level, for example. It would consist in creating new concepts we would link
+  to other ones. Then, we would add any necessary constraints on those concepts
+  and on the relationships linking them to the other existing concepts.
 
@@
 
-As described in section~\ref{subsec:tooling}, our solution is fully tooled with Openflexo.
-We use the Openflexo infrastructure to build models and metamodels together, and to provide 
-dedicated tooling to edit a Process type, to enact a process and to execute it. Being able to provide a
-solution and its associated tools quickly and easily is a noticeable feature. 
-We could also have taken advantage of the model federation to connect our solution to other tools 
-dedicated to process edition and to process execution (\eg a \BPMN engine). However we 
-already had all the necessary components to provide the aforementioned graphical tools.
+As described in section~\ref{subsec:tooling}, our solution is fully tooled with
+Openflexo.  We use the Openflexo infrastructure to build models and metamodels
+together, and to provide dedicated tooling to edit a Process type, to enact a
+process and to execute it. Being able to provide a solution and its associated
+tools quickly and easily is a noticeable feature.  We could also have taken
+advantage of the model federation to connect our solution to other tools
+dedicated to process edition and to process execution (\eg a \BPMN engine).
+However we already had all the necessary components to provide the
+aforementioned graphical tools.
 
@@
 
-Our tooling includes mechanisms to verify the models.
-First, syntactical consistency is realised by cardinality checks and by typing.
-Second, the constraints the users have written are continuously checked.
-Thus, a part of the verification relies on the fact that users add behaviors when they build the metamodels and the models.
+Our tooling includes mechanisms to verify the models. First, syntactical
+consistency is realised by cardinality checks and by typing. Second, the
+constraints the users have written are continuously checked. Thus, a part of
+the verification relies on the fact that users add behaviors when they build
+the metamodels and the models.

### 2021-05-28 9-Relatedworks.tex
sync vers GitHub
add section 6.1 (basic modeling constructs)

@@
 the challenge version).
-A solution is provided with MetaDepth~\citep{metadepth}, which supports modelling with any number of levels, dual ontological/linguistic typing and deep characterization through potency.
+A solution is provided with MetaDepth~\citep{metadepth}, which supports modeling with any number of levels, dual ontological/linguistic typing and deep characterization through potency.
 
@@
 More similar to us, \cite{deeptelos2019} use DeepTelos \citep{deeptelos2016}, an extension of the Telos language \citep{telos1990} in order to solve the process challenge.
-DeepTelos is based on the use of the PowerType \citep{atkinson2001essence} pattern.
 As with Openflexo and \FML, they do not use explicit (numbered) levels nor potency.
+However, DeepTelos is based on the use of the PowerType \citep{atkinson2001essence} pattern.
 
@@
 
-\noteSM{Say something about constraints in \FML}
+%\noteSM{Say something about constraints in \FML} Bah... everybody does multilevel constraints.

### 2021-05-28 1-Main.tex
new initialisms + replacement

@@
 \newcommand{\noteJoel}[1]{\todo[color=blue!20, author=\textbf{Joel}, inline, caption={}]{#1}}
+
+\emisainitialism{\EMF}{EMF}
+\emisainitialism{\MOF}{MOF}
+\emisainitialism{\MDE}{MDE}
+\emisainitialism{\FML}{FML}
 
@@
     % \title{\mpc: the FML solution}
-    \title{Multi-level modelling with Openflexo/FML - A contribution to the \mpc}
+    \title{Multi-level modeling with Openflexo/\FML - A contribution to the \mpc}
 
     %% or include affiliations in footnotes:
-    \author*{Sylvain Guérin}{firstname.lastname@ensta-bretagne.fr}
+    \author{Sylvain Guérin}%{firstname.lastname@ensta-bretagne.fr}
     \address{ENSTA Bretagne, Lab-STICC, UMR 6285, Brest, France}
@@
 
-    \author{Jean-Christophe Bach}
+    \author*{Jean-Christophe Bach}{jc.bach@imt-atlantique.fr}
     \address{IMT Atlantique, Lab-STICC, UMR 6285, 29238 Brest, France}

### 2021-05-28 10-Conclusions.tex
Updates from Overleaf

@@
-including  lessons  learned,  impulses  for future work, etc...
+%Conclusions: including  lessons  learned,  impulses  for future work, etc.
+
+%\todo[inline]{lessons learned, impulse for future work -> fait émerger la question d'ajouter des bibliothèques dédiées au multi-niveaux (avoir des behaviors)}
+
+We fulfill all the challenge requirements and we propose a tooling that demonstrates the usability of our solution.
+We developed an ad-hoc solution, including models and metamodels, thanks to the flexibility provided by the metametamodeling infrastructure offered by Openflexo.
+
+As a future work we envision to explore a number of alternative solutions: first, we intend to propose another approach based on a more agnostic level solution such as Clabjects.
+Instead of two levels defined by Type and Instance concepts, we could have a single concept mixing a type part and an instance part.
+This approach would probably be easier if specifications evolve.
+Secondly, another possible approach is the use of the free modeling tool proposed by Openflexo.
+The solution would be developed from examples (Acme and XSure processes for instance) from which models would be identified.
+Finally, the realisation of this challenge highlighted the interest of integrating specific behaviors for multi-level concepts management in Openflexo infrastructure, to avoid tailoring them, as in this exercise.
+We intend to do so.

### 2021-05-28 2-Abstract.tex

--- 2021-05-27 2-Abstract.tex
@@
 We present a solution based on model federation and discuss the advantages and limitations of using this approach for multi-level modeling.
-Concretely, we use virtual models and more precisely the Flexo Modeling Language (FML) that serves to describe them as the main building block in order to solve the process modeling challenge whereas the federation aspect is used in order to provide editing tooling for the resulting process language.
+Concretely, we use virtual models and more precisely the Flexo Modeling Language (\FML) that serves to describe them as the main building block in order to solve the process modeling challenge whereas the federation aspect is used in order to provide editing tooling for the resulting process language.
 Our solution is fully implemented with the Openflexo framework.
@@
 Concernant la fédération de modèle proprement dite, je crois qu'il faut qu'on montre qu'on a besoin d'une couche conceptuelle (qu'on relie à l'espace technique via les connecteurs techniques, les model slots et les rôles).
-C'est ce besoin qu'adresse le language FML.
-On montre ici plutôt que le language FML - utilisé comme language conceptuel - est assez intéressant, assez expressif pour capturer les exigences du challenge d'une part, et va jusqu'à permettre l'exécution des processus.
+C'est ce besoin qu'adresse le language \FML.
+On montre ici plutôt que le language \FML - utilisé comme language conceptuel - est assez intéressant, assez expressif pour capturer les exigences du challenge d'une part, et va jusqu'à permettre l'exécution des processus.
 Et cerise sur le gâteau: on peut construire un outillage graphique pour editer et exécuter ce use case.}

### 2021-05-28 3-Introduction.tex

@@
 The models in a federation remain autonomous and represented in their original technological spaces whereas virtual models (also called conceptual models) and links serve as control components used to present different views to the users and maintain synchronization.
-Our solution is based on the model federation infrastructure, namely, the use of virtual models and links.
-Concretely, we use OpenFlexo and its internal Flexo Modeling Language (\FML), a language to create, link and manage virtual models, in order to solve the MULTI process challenge while the federation aspect is used solely so as to provide tooling for the resulting process language.
+Our solution is based on the model federation infrastructure.
+Concretely, we use Openflexo and its internal Flexo Modeling Language (\FML), a language to create, link and manage virtual models, in order to solve the MULTI process challenge while the federation aspect is used solely so as to provide tooling for the resulting process language.
 Our solution, which is fully implemented and executable, meets all the requirements.
+%\noteFabien{Le paragraphe qui précède me semble redire des choses de la section 2, comment fait-on ?}
 
@@
 Section~\ref{sec:analysis} is the analysis of the problem.
-We detail our model in section~\ref{sec:model} and show it satisfies the challenge requirements in section~\ref{sec:requirements}.
+We detail our model in section~\ref{sec:model} and show how it satisfies the challenge requirements in section~\ref{sec:requirements}.
 We assess the modeling solution in section~\ref{sec:discussion}.
-Section~\ref{sec:relatedwork} presents the related work before the conclusion.%~\ref{sec:conclusions}.
+Section~\ref{sec:relatedwork} presents the related work before the conclusions.%~\ref{sec:conclusions}.
 

### 2021-05-28 6-ModelPresentation.tex

@@
 \textit{Process} represents an enacted \textit{ProcessType}, as defined in previous subsection (\textbf{P10}).
-FML defines behavioural features (operations in object-oriented modelling), called \texttt{FlexoBehaviour}.
-\textit{ProcessType} defines following behaviour \texttt{newProcess(String)}, taking a \texttt{String} argument (the name of the process to enact):
+FML defines behavioral features (operations in object-oriented modeling), called \texttt{FlexoBehavior}.
+\textit{ProcessType} defines following behavior \texttt{newProcess(String)}, taking a \texttt{String} argument (the name of the process to enact):
 
@@
 Authorization for an actor to perform a task (\textbf{P17}) is captured either through the roles \textit{allowedActors} and \textit{allowedActorTypes} defined in \textit{TaskType}.
-This mechanism is completed by following behaviours \textit{isAuthorizedActor(Actor)}, \textit{isValidActor(Actor)} and \textit{isValidActorType(ActorType)} defined in \textit{TaskType} :
+This mechanism is completed by following behaviors \textit{isAuthorizedActor(Actor)}, \textit{isValidActor(Actor)} and \textit{isValidActorType(ActorType)} defined in \textit{TaskType} :
 
@@
 
-Default behaviour states that all actors and actor types are valid for all task types.
-This modelling scheme offers many extension points, by the redefinition of one or some behaviours in inherited concepts (although none were required in the context of XSure insurance use case).
+Default behavior states that all actors and actor types are valid for all task types.
+This modeling scheme offers many extension points, by the redefinition of one or some behaviors in inherited concepts (although none were required in the context of XSure insurance use case).
 
 Actor types specialization is captured by the \textit{parentActorTypes} relation defined in \textit{ActorType} (\textbf{P18}).
-This is completed by both the definition of \textit{hasActorType(ActorType)} behaviour in \textit{Actor} concept and recursive behaviour \textit{isOrSpecializes(ActorType)} in \textit{ActorType}:
+This is completed by both the definition of \textit{hasActorType(ActorType)} behavior in \textit{Actor} concept and recursive behavior \textit{isOrSpecializes(ActorType)} in \textit{ActorType}:
 
@@
 The capture of \textbf{S3} requirement is performed while defining \textit{Coding} as an instance of concept \textit{CodingTaskType} extending \textit{SETaskType} and defining a relation \textit{languages} to \textit{ProgrammingLanguage} with 1..* cardinality.
-\textit{newTask(String)} behaviour overrides generic implementation while instantiating a \textit{Coding} concept, specializing \textit{SEArtifact}, as shown in following excerpt of FML code:
+\textit{newTask(String)} behavior overrides generic implementation while instantiating a \textit{Coding} concept, specializing \textit{SEArtifact}, as shown in following excerpt of FML code:
 
@@
 We assume that all tasks may define an expected duration, which might be checked during process execution.
-This can be modelled through the addition of \textit{expectedDuration} attribute in \textit{TaskType}.
-Alternatively, this can be modelled through the definition of \textit{SETesting} concept, as a specialization of \textit{SETaskType}, and the instantiation of \textit{Testing} as an instance of \textit{SETesting}.
+This can be modeled through the addition of \textit{expectedDuration} attribute in \textit{TaskType}.
+Alternatively, this can be modeled through the definition of \textit{SETesting} concept, as a specialization of \textit{SETaskType}, and the instantiation of \textit{Testing} as an instance of \textit{SETesting}.
 The business logic expressed by \textbf{S12} requirement should be redefined in \textit{SETesting}.
@@
 
-Our solution is fully implemented within Openflexo tool. Both usecases have been modelled in the interactive design environment and are executable by the FML execution engine.
+Our solution is fully implemented within Openflexo tool. Both usecases have been modeled in the interactive design environment and are executable by the FML execution engine.
 
@@
 A \textit{Task} also manages a set of performing actors, a begin and end date, some used and produced artifacts.
-\textit{Gateways} business logic has also been implemented through the implementation of abstract \texttt{execute(Process)} behaviour.
+\textit{Gateways} business logic has also been implemented through the implementation of abstract \texttt{execute(Process)} behavior.
 Management of artifacts whose semantics follows rules defined in section \ref{sec:ProcessEnactment} (\textbf{P13}) has also been implemented.

### 2021-05-28 6-ModelPresentation.tex
Updates from Overleaf

@@
 Figure~\ref{fig:BaseMetamodel} represents this base metamodel with a \UML-like formalism, well adapted to represent \FML concepts and their instances.
-A \textit{Concept} of \FML is represented by a \UML class.
-The roles are represented by a combination of a name in their concept container and an arrow from their name to their type.
+A \textit{Concept} of \FML is represented by a \UML class where roles of basic types are attributes.
+The roles whose types are concepts are represented by a combination of a name in their containing concept and an arrow from their name to their type.
 The cardinality follow the \UML practice.
@@
 Two types of ontological instantiation are needed to meet the requirements of the challenge.
-One provides that some instances conform to only one type (\texttt{Process} and \texttt{Task} do respectively conform to one \texttt{ProcessType} and \texttt{TaskType}).
-While some entities define their conformity to several types (\textit{Actor} and \textit{Artifact} do respectively conform to several \textit{ActorType} and \textit{ArtifactType}).
-This instantiation is reified with the relations \textit{type} and \textit{types} between \textit{Type}, \textit{Instance} and \textit{MultiInstance} concepts.
+One provides that some instances conform to only one type (\texttt{Process} and \texttt{Task} do respectively conform to \texttt{ProcessType} and \texttt{TaskType}).
+While some entities define their conformity to several types (\texttt{Actor} and \texttt{Artifact} do respectively conform to several \texttt{ActorType} and \texttt{ArtifactType}).
+These instantiations are respectively expressed using the relations \texttt{type} and \texttt{types} between \texttt{Type}, \texttt{Instance} and \texttt{MultiInstance} concepts.
+In the diagram, the realization of these relations are indicated by a derived relation, \texttt{/type} or \texttt{/types}, as for example between \texttt{Process} and \texttt{ProcessType} or between \texttt{Actor} and \texttt{ActorType}.
 
@@
 
-\textit{ProcessType} is defined as a specialization of \textit{Type} \texttt{FlexoConcept}, and references a collection of \textit{TaskType} through the composition relation \textit{taskTypes} with (0..*) cardinality (\textbf{P1}).
-A \textit{TaskType} is embedded in a \textit{ProcessType} and inherits from its context.
-To illustrate this in \textit{XSure} insurance domain use case, \textit{XSure model} defines \textit{Claim Handling}, instance of \textit{ProcessType}, and \textit{Receive Claim}, \textit{Assess Claim} and \textit{Pay premium}, instances of \textit{TaskType}.
+\texttt{ProcessType} is a specialization of the \texttt{Type} concept, and references a collection of \texttt{TaskType} through the composition relation \texttt{taskTypes} with (0..*) cardinality (\textbf{P1}).
+A \texttt{TaskType} is embedded in a \texttt{ProcessType} and inherits from its context.
+To illustrate this in \textit{XSure} insurance domain use case, \textit{XSure model} defines \textsf{Claim Handling}, instance of \texttt{ProcessType}, and \textsf{Receive Claim}, \textsf{Assess Claim} and \textsf{Pay premium}, instances of \texttt{TaskType}.
 
-\textit{ProcessType} also references a collection of gateways, reified with the \textit{Gateway} concept hierarchy.
-\textit{Gateway} is defined as a specialization of \textit{Type} and is specialized with \textit{Sequencing}, \textit{AndSplit}, \textit{AndJoin}, \textit{OrSplit} and \textit{OrJoin} concepts (\textbf{P2}).
-Depending of its type, a \textit{Gateway} defines one or more inputs and one or more outputs, depending of its underlying execution semantics.
-A \textit{ProcessType} additionally exposes unique initial \textit{TaskType} and a collection of final \textit{TaskType} with both roles \textit{initialTaskType} (single cardinality) and \textit{finalTaskType} (cardinality 0..*) (\textbf{P3}).
-TaskType exposes a creator role as a reference to \textit{Actor} concept (\textbf{P4}), which is at the same conceptual level.
+\texttt{ProcessType} also references a collection of gateways, reified with the \texttt{Gateway} concept hierarchy.
+\texttt{Gateway} is a specialization of \texttt{Type} and is specialized by \texttt{Sequencing}, \texttt{AndSplit}, \texttt{AndJoin}, \texttt{OrSplit} and \texttt{OrJoin} concepts (\textbf{P2}).
+Depending on its type and following its underlying operational semantics, a gateway defines one or more inputs and one or more outputs.
+A \texttt{ProcessType} additionally exposes a unique initial \texttt{TaskType} and a collection of final \texttt{TaskType} with both roles \texttt{initialTaskType} (single cardinality) and \texttt{finalTaskType} (cardinality 0..*) (\textbf{P3}).
+\textt{TaskType} exposes a creator role as a reference to an \texttt{Actor} concept (\textbf{P4}), which is at the same conceptual level.
 
-Following listing shows an excerpt of \FML code modeling some core concepts of process modeling base metamodel.
+The following listing shows an excerpt of the \FML code modeling some core concepts of process modeling base metamodel.
 
@@
 
-\textit{ActorType} is defined as a sub-concept of \textit{Type} and the \textit{allowedActorTypes} relation to \textit{ActorType} which is defined in \textit{TaskType} (with 0..* cardinality) captures \textbf{P5} requirement.
-Requirement \textbf{P6} is symmetrically satisfied with \textit{allowedActors} relation to \textit{Actor} also defined in \textit{TaskType} (with same 0..* cardinality).
-Same modeling pattern applies for \textit{ArtefactType} defined as a sub-concept of \textit{Type}, and both relations \textit{usedArtifactTypes} and \textit{producedArtifactTypes} defined in \textit{TaskType} (\textbf{P7}).
-\textit{TaskType} additionnaly exposes an \textit{expectedDuration} attribute (expressed in number of days), satisfying \textbf{P8}.
+The \texttt{ActorType} concept is a sub-concept of \texttt{Type} and the \texttt{allowedActorTypes} relation to \texttt{ActorType} defined in \texttt{TaskType} (with 0..* cardinality) captures \textbf{P5} requirement.
+Requirement \textbf{P6} is symmetrically satisfied with \texttt{allowedActors} relation to \texttt{Actor} also defined in \texttt{TaskType} (with 0..* cardinality).
+The same modeling pattern applies to \texttt{ArtefactType} defined as a sub-concept of \texttt{Type}, and both relations \texttt{usedArtifactTypes} and \texttt{producedArtifactTypes} defined in \texttt{TaskType} (\textbf{P7}).
+\texttt{TaskType} additionally exposes an \texttt{expectedDuration} role (expressed in number of days), satisfying \textbf{P8}.
+\noteFabien{expectedDuration pas dans la figure}
 
-\textit{Actor} concept defines a boolean attribute called \textit{isSenior}, while \textit{TaskType} defines an additional \textit{isCritical} boolean attribute, indicating that some instance of \textit{TaskType} are flagged as critical and must be performed by senior actors.
-To fullfill \textbf{P9} requirement, a supplementary constraint is required for \textit{TaskType} and is captured through this invariant expressed in \FML language:
+The \texttt{Actor} concept defines a boolean attribute called \texttt{isSenior}, while \texttt{TaskType} defines an additional \texttt{isCritical} boolean attribute, indicating that some instance are flagged as critical and must be performed by senior actors.
+To fulfil \textbf{P9} requirement, a supplementary constraint is required for \texttt{TaskType} and is captured through the following invariant expressed in the \FML language:
 
@@
 
-This invariant should be completed with additional constraints defined in \textit{Task} \textit{FlexoConcept}, which apply to performing actors actually assigned to enacted tasks.
+This invariant should be completed with additional constraints defined in the \texttt{Task} concept, which apply to performing actors assigned to enact tasks.
 
@@
 \label{sec:ProcessEnactment}
-We now present the process enactment part of base metamodel, located right of figure \ref{fig:BaseMetamodel}.
-All concepts defined in this subsection are either specialization of \textit{Instance} concept (if they are associated with exactly one \textit{Type}) or \textit{MultiInstance} concept (for those which have several types).
+We now present the process enactment part of base metamodel, located right of figure~\ref{fig:BaseMetamodel}.
+All concepts defined in this subsection are either specialization of the \texttt{Instance} concept (if they have exactly one type) or the \texttt{MultiInstance} concept (when they have several types).
 
-\textit{Process} represents an enacted \textit{ProcessType}, as defined in previous subsection (\textbf{P10}).
-FML defines behavioral features (operations in object-oriented modeling), called \texttt{FlexoBehavior}.
-\textit{ProcessType} defines following behavior \texttt{newProcess(String)}, taking a \texttt{String} argument (the name of the process to enact):
+\texttt{Process} represents an enacted \texttt{ProcessType}, as defined in previous subsection (\textbf{P10}).
+FML defines behavioral features called behavior.
+\texttt{ProcessType} defines the behavior \texttt{newProcess(String)}, taking the name of the process to enact as argument:
 
@@
 
-This scheme allows to rely on FML dynamic binding mechanism to delegate to types the responsability for instances creation.
-An instance of \textit{Process} references a unique \textit{ProcessType} trough the specialized \textit{/type} role.
-Each instance of \textit{TaskType} is ontologically instantiated with a \textit{Task} (\textbf{P11}), using the same pattern where \textit{TaskType} has the responsability to manage this ontological instantiation.
+This scheme rely on \FML dynamic binding mechanism to delegate to task types the responsibility of instances creation.
+An instance of \texttt{Process} references its unique type \texttt{ProcessType} by the specialized \texttt{/type} role.
+Each instance of \texttt{TaskType} is ontologically instantiated with a \textit{Task} (\textbf{P11}), using the same pattern where \textit{TaskType} has the responsability to manage this ontological instantiation.
 A \textit{Task} references its unique \textit{TaskType}, and defines a \textit{begin date} and an \textit{end date} as basic attributes of the concept (\textbf{P12}).

### 2021-05-28 6-ModelPresentation.tex
synchro vendredi soir

@@
 A \texttt{ProcessType} additionally exposes a unique initial \texttt{TaskType} and a collection of final \texttt{TaskType} with both roles \texttt{initialTaskType} (single cardinality) and \texttt{finalTaskType} (cardinality 0..*) (\textbf{P3}).
-\textt{TaskType} exposes a creator role as a reference to an \texttt{Actor} concept (\textbf{P4}), which is at the same conceptual level.
+\texttt{TaskType} exposes a creator role as a reference to an \texttt{Actor} concept (\textbf{P4}), which is at the same conceptual level.
 
@@
 An instance of \texttt{Process} references its unique type \texttt{ProcessType} by the specialized \texttt{/type} role.
-Each instance of \texttt{TaskType} is ontologically instantiated with a \textit{Task} (\textbf{P11}), using the same pattern where \textit{TaskType} has the responsability to manage this ontological instantiation.
-A \textit{Task} references its unique \textit{TaskType}, and defines a \textit{begin date} and an \textit{end date} as basic attributes of the concept (\textbf{P12}).
+Each instance of \texttt{TaskType} is ontologically instantiated with a \texttt{Task} (\textbf{P11}), using the same pattern where \texttt{TaskType} has the responsibility to manage the ontological instantiation.
+A \texttt{Task} references its unique \texttt{TaskType}, and defines a \texttt{begin date} and an \texttt{end date} basic role (\textbf{P12}).
+\noteFabien{endDate en double}
 
-Same pattern applies for artifacts used and produced with roles \textit{usedArtifacts}, \textit{producedArtifacts} and \textit{performingActors} defined in \textit{Task} concept (\textbf{P13}).
-An instance of \textit{Artifact}, defined as a specialization of \textit{MultiInstance} concept, references a set of \textit{ArtifactType} through the specialized \textit{/types} role (\textbf{P14} and \textbf{P16}).
-Likewise, concept \textit{Actor} is defined as a specialization of \textit{MultiInstance} concept and references a set of \textit{ActorType} through the specialized \textit{/types} relation (\textbf{P15}).
+The same pattern applies for the used and produced artifacts by roles \texttt{usedArtifacts}, \texttt{producedArtifacts} and \texttt{performingActors} defined in \texttt{Task} (\textbf{P13}).
+An instance of \texttt{Artifact} specializes \textit{MultiInstance} and references a set of \texttt{ArtifactType} through the specialized \texttt{/types} role (\textbf{P14} and \textbf{P16}).
+Likewise, \texttt{Actor} specializes \texttt{MultiInstance} and references a set of \texttt{ActorType} through the specialized \texttt{/types} relation (\textbf{P15}).
 
-Semantics is unclear relatively to instantiation policy for artifacts.
-We assume that task execution implies that for each used and produced \textit{artifact type} defined in related \textit{task type}, it exists at least one \textit{artifact} declaring required ontologic instantiation (through \textit{type} role).
-Following excerpt of FML code shows partial implementation of this.
-Same patterns applied for produced artifacts.
+The semantics is unclear relatively to the instantiation policy for artifacts.
+We assume that task execution implies that for each used and produced \emph{artifact type} defined in a related \emph{task type}, it exists at least one \emph{artifact} declaring required ontologic instantiation (through \textit{type} role).
+The following excerpt of \FML code shows a partial implementation of this.
+The same pattern applies to produced artifacts.
+\noteFabien{Je ne suis pas bien sûr de comprendre le paragraphe}
 
@@
 
-Authorization for an actor to perform a task (\textbf{P17}) is captured either through the roles \textit{allowedActors} and \textit{allowedActorTypes} defined in \textit{TaskType}.
-This mechanism is completed by following behaviors \textit{isAuthorizedActor(Actor)}, \textit{isValidActor(Actor)} and \textit{isValidActorType(ActorType)} defined in \textit{TaskType} :
+Authorization for an actor to perform a task (\textbf{P17}) is captured either by the role \texttt{allowedActors} or the role \texttt{allowedActorTypes} defined in \texttt{TaskType}.
+This mechanism is completed by the behaviors \texttt{isAuthorizedActor}, \texttt{isValidActor} and \texttt{isValidActorType} defined in \texttt{TaskType}:
 
@@
 
-\textit{Task} concept basically delegates this authorization to its related \textit{TaskType}, as shown in following FML excerpt:
+The \texttt{Task} concept delegates this authorization to its task type, as shown in the following \FML code:
 
@@
 
-Enforcing those constraints is finally performed by the definition of this invariant in \textit{Task} concept:
+Enforcing those constraints is finally performed by the definition of this invariant in the \texttt{Task} concept:
 
@@
 
-Default behavior states that all actors and actor types are valid for all task types.
-This modeling scheme offers many extension points, by the redefinition of one or some behaviors in inherited concepts (although none were required in the context of XSure insurance use case).
+The default behavior states that all actors and actor types are valid for all task types.
+This modeling scheme offers many extension points, by the redefinition of some behaviors in the inherited concepts (although none were required in the context of \emph{XSure} use case).
 
-Actor types specialization is captured by the \textit{parentActorTypes} relation defined in \textit{ActorType} (\textbf{P18}).
-This is completed by both the definition of \textit{hasActorType(ActorType)} behavior in \textit{Actor} concept and recursive behavior \textit{isOrSpecializes(ActorType)} in \textit{ActorType}:
+Actor types specialization is captured by the \texttt{parentActorTypes} relation defined in \texttt{ActorType} (\textbf{P18}).
+This is completed by both the definition of the \texttt{hasActorType(ActorType)} behavior in \texttt{Actor} and the recursive behavior \texttt{isOrSpecializes(ActorType)} in \texttt{ActorType}:
 
@@
 
-All concepts inherits from \textit{ModelingElement}, which defines a \textit{lastUpdated} attribute with \textit{Date} type, and thus satisfies \textbf{P19} requirement.
+All concepts inherits from \texttt{ModelingElement}, which defines a \texttt{lastUpdated} attribute with \texttt{Date} type, and thus satisfies \textbf{P19} requirement.
 
@@
 
-The other tool, called \textit{Enacted process graphical editor}, provides process edition and tasks assignations through a graphical visualization displaying process being executed.
+The other tool, called \textit{Enacted process graphical editor} and shown on figure \ref{fig:ScreenshotEnactedProcessEditor}, provides process edition and tasks assignations through a graphical visualization displaying process being executed.
 This tool is represented on the right side of the figure \ref{fig:ToolingArchitecture}.
@@
 
+\begin{figure*}
+ \centering
+     \includegraphics[width=1.0 \textwidth]{Figures/ScreenshotEnactedProcessEditor.png}
+     \caption{Screenshot of the enacted process graphical editor}
+    \label{fig:ScreenshotEnactedProcessEditor}
+\end{figure*}
+
 \todo[inline]{mettre ici le lien vers la page web qui montre la démo}

### 2021-05-31 2-Abstract.tex
Updates from Overleaf

@@
 The models in a federation remain autonomous and represented in their original technological spaces whereas virtual models and links (which are not level bounded) serve as control components used to present different views to the users and  maintain synchronization.
-In this paper we tackle the \emph{process modeling challenge}, which consists in providing a solution to the problem of specifying and enacting processes.
+In this paper we tackle the \emph{MULTI process modeling challenge}, which consists in providing a solution to the problem of specifying and enacting processes.
 Solutions must fulfill a number of requirements for a process representation defined at an abstract process-definition level and at various more concrete domain-specific levels, resulting in a multi-level hierarchy of related models.
 We present a solution based on model federation and discuss the advantages and limitations of using this approach for multi-level modeling.
-Concretely, we use virtual models and more precisely the Flexo Modeling Language (\FML) that serves to describe them as the main building block in order to solve the process modeling challenge whereas the federation aspect is used in order to provide editing tooling for the resulting process language.
+Concretely, we use virtual models and more precisely the Flexo Modeling Language (\FML) that serves to describe them as the main building block in order to solve the process modeling challenge whereas the federation aspect is used as a means to provide editing tooling for the resulting process language.
 Our solution is fully implemented with the Openflexo framework.

### 2021-05-31 10-conclusions.tex
remove all capital letters from files, as specified by the style

--- 2021-05-28 10-Conclusions.tex
@@
 
-As a future work we envision to explore a number of alternative solutions: first, we intend to propose another approach based on a more agnostic level solution such as Clabjects.
+As a future work we envision to explore a number of alternative solutions: first, we intend to propose another solution based on a more level-agnostic approach such as Clabjects.
 Instead of two levels defined by Type and Instance concepts, we could have a single concept mixing a type part and an instance part.
-This approach would probably be easier if specifications evolve.
+This approach would probably ease adaptation if the problem specifications evolve.
 Secondly, another possible approach is the use of the free modeling tool proposed by Openflexo.
 The solution would be developed from examples (Acme and XSure processes for instance) from which models would be identified.
-Finally, the realisation of this challenge highlighted the interest of integrating specific behaviors for multi-level concepts management in Openflexo infrastructure, to avoid tailoring them, as in this exercise.
+Finally, the realisation of this challenge highlighted the interest of integrating specific behaviors for the management of multi-level concepts in the Openflexo infrastructure (as opposed to tailoring them as we did in this exercise).
 We intend to do so.

### 2021-05-31 4-technology.tex
remove all capital letters from files, as specified by the style

--- 2021-05-28 4-Technology.tex
@@
 To meet the \mpc, we have decided to use the general
-approach of model federation~\cite{Golra2016-federation}. \emph{Model
+approach of model federation~\parencite{Golra2016-federation}. \emph{Model
   federation} is a way to assemble models using some kind of
 low-coupling links. It has been studied to answer the \OMG' RFP
-Semantic Modeling for Information Federation (SIMF)~\cite{simf}. In
+Semantic Modeling for Information Federation~\parencite{simf}. In
 contrast to approaches that compose metamodels into a single large
@@
 model/metamodel conformity -- that is presented
-in~\cite{models2016-freemodel}. Another very interesting feature of
+in~\parencite{models2016-freemodel}. Another notable feature of
 this approach is the strong decoupling among tools that remain usable
@@
 architecture in next section, here are some key concepts implemented
-in the Openflexo~\cite{openflexo_link} framework.
+in the Openflexo~\parencite{openflexo_link} framework.
 
@@
 paradigm while a virtual model is built using the Flexo Modeling
-Language (\FML). Each federated model can be viewed as an autonomous
-component that may evolve with its own tooling. The virtual models
-serve as control components.
-In this paper, we mainly use the conceptual models and more precisely \FML its domain specific modeling language. The only use of the federation aspect is on the tooling made for the process language.
+Language (\FML). Each federated model is an autonomous
+element that may evolve with its own tooling. The virtual models
+serve as control elements binding the federated models together.
+In this paper, we mainly use conceptual models and more precisely \FML its domain specific modeling language. The only use of the federation aspect is on the tooling made for the process language.
 
@@
 
-\FML simplified metamodel is provided in Figure~\ref{fig:mm}. It is designed to reify models as virtual models. A virtual model
+\FML simplified metamodel is provided in Figure~\ref{fig:mm}. It is designed to define models as virtual models. A virtual model
 is composed of a set of \emph{concepts}, while itself being a concept.
-Hence, virtual models are structuring units while concepts are the
+Hence, virtual models are structuring units forming architectures while concepts are the
 core entities. A concept has a set of \emph{roles} and
@@
 These actions are called \emph{behaviors} and can either be called (like usual methods) or triggered by events.
-The reactive behaviors are mainly useful when models have to be synchronised.
-We did not
+The reactive behaviors are mainly useful when a federated model evolution needs to trigger a computation.
+We do not
 exploit this possibility in the challenge.
@@
 execution engine with an interactive virtual model design environment.
-It has been used in several case studies including model mapping,
-multi-paradigm process modeling and enterprise architecting. It has
-also been used to build a tool, the freemodeling editor that has been
-used in industrial projects~\cite{models2016-freemodel}. As of today, this tool
-offers some mature technology adapters (\emph{docx} and \emph{excel}
-for documents, \EMF and OWL for modeling languages, JDBC for databases,
-REST and XMLRPC for external services and one for diagramming tools)
-and other more rudimentary (pdf, HTTP, XML, etc.).
 
 Finally, tools have their own model. We have taken advantage of
-Openflexo features to build, in parallel with the models and virtual
-models required by the challenge, a drawing tool that makes our
+Openflexo features to build, in parallel with the models  required by the challenge, a drawing tool that makes our
 solution (partially) executable.

### 2021-05-31 7-requirementssatisfaction.tex
remove all capital letters from files, as specified by the style

--- 2021-05-28 7-RequirementsSatisfaction.tex
@@
     \item \textbf{Conceptualization} (\textit{Conc.}): a concept carries the semantics exposed by the requirement (for example for \textbf{P1}: \textit{process type} is conceptualized in a \textit{FlexoConcept} \textit{ProcessType}).
-    \item \textbf{Specialization} (\textit{Spec.}): a concept is specialized in the sense of object-oriented modeling (for example for \textbf{P2}: \textit{Gateway} defines an abstract behaviour \texttt{execute()} and \textit{Sequencing}, \textit{AndSplit}, \textit{AndJoin}, \textit{OrSplit} and \textit{OrJoin} are defined as inherited concepts implementing specific business logic).
+    \item \textbf{Specialization} (\textit{Spec.}): a concept is specialized in the sense of object-oriented modeling (for example for \textbf{P2}: \textit{Gateway} defines an abstract behavior \texttt{execute()} and \textit{Sequencing}, \textit{AndSplit}, \textit{AndJoin}, \textit{OrSplit} and \textit{OrJoin} are defined as inherited concepts implementing specific business logic).
     \item \textbf{Composition} (\textit{Comp.}): some concepts are in relation, or specific attributes were added to some concepts (for example for \textbf{P3}: a \textit{process type} has one \textit{initial task type}).
@@
     \item \textbf{Ontologic instanciation} (\textit{O.Inst.}): fulfillment of requirement is obtained by the relation between an instance and its type definition instance (for example for \textbf{S3} : \textit{developer} has the dual concept nature in \textit{Acme metamodel} and instance nature in \textit{Acme model}).
-    \item \textbf{Behavioral modeling} (\textit{B.Mod.}): requirement is satisfied by the definition of one more object-oriented behavioural features, which may combined with constraints checking and associated tooling (for example \textbf{P17}).
+    \item \textbf{Behavioral modeling} (\textit{B.Mod.}): requirement is satisfied by the definition of one more object-oriented behavioral features, which may combined with constraints checking and associated tooling (for example \textbf{P17}).
 \end{itemize}
@@
 %OI : Ontologic instantiation
-%BM : Behavioural modeling
+%BM : behavioral modeling
 

### 2021-05-31 8-discussion.tex
remove all capital letters from files, as specified by the style

--- 2021-05-28 8-Discussion.tex
@@
 
-  \noteJC{pas clair pour moi ce qui est entendu par « basic modeling constructs », ce qui est dans la partie 2 ?}
+%  \noteJC{pas clair pour moi ce qui est entendu par « basic modeling constructs », ce qui est dans la partie 2 ?}
+
+  Our solution uses the basic modeling constructs depicted by the \FML core
+  metamodel (Figure~\ref{fig:mm}) and described in
+  section~\ref{sec:technology}. Models are \emph{virtual models}. They are
+  composed of \emph{concepts}, being themselves concepts. Concepts may have
+  \emph{roles} and \emph{behaviors} (actions one can perform).
+
+  In order to provide the graphical tools, our solution also uses \emph{model
+  slots} to connect some concept instances to external elements (in our case:
+  instance of diagram from our diagramming TA). The use of slots has been
+  described in section~\ref{sec:model} and is illustrated by bold circles in
+  the Figure~\ref{fig:ToolingArchitecture}).
 

### 2021-05-31 9-relatedworks.tex
remove all capital letters from files, as specified by the style

--- 2021-05-28 9-Relatedworks.tex
@@
 %Salvador commence
-A plethora of multi-level modeling approaches and tools with different foundations have appeared in the recent years (see \cite{multiTools})\todo{not sure how to cite a wiki page}.
+A plethora of multi-level modeling approaches and tools with different foundations have appeared in the recent years~\footnote{\url{https://homepages.ecs.vuw.ac.nz/Groups/MultiLevelModeling/MultiTools}}.
 Comparing them lies out of the scope of this paper.
@@
 \cite{multiecore2019} use MultiEcore \citep{multecore2016} to provide a solution to the challenge.
-MultEcore uses an extesion of the two-level cascading technique and potency.
+MultEcore uses an extension of the two-level cascading technique and potency.
 They do not need the synthetic typing relations we introduce.
 However they need to do frequent model transformations to transform instance models into instantiable models.
-With respect to them, we use less models as we do not include a generic software engineering process model nor a generic assurance process model.
-Finally, they use multi-level couple model transformations for the definition and enforcement of constraints.
+With respect to the proposed solution, we use less models as we do not include a generic software engineering process model nor a generic assurance process model.
+%Finally, they use multi-level couple model transformations for the definition and enforcement of constraints.
 
@@
 As with Openflexo and \FML, they do not use explicit (numbered) levels nor potency.
-However, DeepTelos is based on the use of the PowerType \citep{atkinson2001essence} pattern.
+However, unlike Openflexo, DeepTelos integrates a multi-level modeling specific construct similar to the PowerType \citep{atkinson2001essence} pattern they call \emph{most general instances}.
 

### 2021-05-31 5-analysis.tex
remove all capital letters from files, as specified by the style
Typos

--- 2021-05-28 5-Analysis.tex
@@
 
-Like in lot of modeling approaches, these  steps could also be achieved in any order and iteratively.
+Like in a lot of modeling approaches, these  steps could also be achieved in any order and iteratively.
 But our goal during the challenge's analysis remains to produce a \FML virtual model architecture for our federation.
 As shown in figure~\ref{fig:MultilevelArchitecture}, we defined two virtual models (Base and Acme processes abstractions) instantiated by two virtual model instances (XSure and Acme processes).
-As the figure shows the virtual models play the role of metamodels, in the sense of concept definition with a level-blind approach.
+As the figure shows the virtual models play the role of metamodels, in the sense of concept definition with a level-agnostic approach.
 The virtual model instances play the role of models conforming to the metamodel definitions.
 In the rest of the paper, we use the term metamodel and model to simplify the presentation.
-The resulting architecture follows the way the challenge is presented and this organisation allows for the possibility of flexible extensions.
+The resulting architecture follows the way the challenge is presented and this
+organisation allows for the possibility of flexible extensions.
 
@@
 
-Based on this approach, we organize our set of models following the architecture of the figure \ref{fig:MultilevelArchitecture}.
+Based on this approach, we organize our set of models following the architecture of the aforementioned Figure \ref{fig:MultilevelArchitecture}.
 The resulting multi-level architecture is organized in two virtual models, one for the core concepts metamodel containing the definition of Processes and Tasks, and one, the Acme metamodel, extending the previous model to integrate the Acme definitions.
-\FML language defines multiple inheritance concept between virtual models as illustrated in the section~\ref{sec:AcmeSoftwareDevelopmentProcess}.
+The \FML language defines multiple inheritance concept between virtual models as illustrated in the section~\ref{sec:AcmeSoftwareDevelopmentProcess}.
 
@@
 
-We now analysis a few specifications that deserve a comment for the choices we made.
+We now analyse a few specifications that deserve a comment for the choices we made.
 

### 2021-05-31 6-modelpresentation.tex
Typos

--- 2021-05-28 6-ModelPresentation.tex
@@
 The presentation of our solution to the \mpc follows a systematic methodology where modeling choices are introduced step by step, as the requirements are stated.
-Satisfaction of requirements is explained throughout this presentation.
+The satisfaction of requirements is explained throughout this presentation.
 
@@
 
-We first present the process definition part of base metamodel, located left of figure \ref{fig:BaseMetamodel}.
+We first present the process definition part of the base metamodel, located on the left of Figure \ref{fig:BaseMetamodel}.
 
@@
 
-The following listing shows an excerpt of the \FML code modeling some core concepts of process modeling base metamodel.
+The following listing shows an excerpt of the \FML code modeling some core concepts of the process modeling base metamodel.
 
@@
 \texttt{TaskType} additionally exposes an \texttt{expectedDuration} role (expressed in number of days), satisfying \textbf{P8}.
-\noteFabien{expectedDuration pas dans la figure}
 
@@
 A \texttt{Task} references its unique \texttt{TaskType}, and defines a \texttt{begin date} and an \texttt{end date} basic role (\textbf{P12}).
-\noteFabien{endDate en double}
 
@@
 \noteFabien{Je ne suis pas bien sûr de comprendre le paragraphe}
+\noteJC{autre remarque qui n'a rien à voir: si on suit les recommandations de l'appel, ce paragraphe devrait se retrouver dans la partie analyse étant donné qu'il établit un choix/une interprétation.
+Mais c'est délicat à intégrer sans le contexte du modèle de cette section.
+Ou alors il faudrait une liste, avec des références en avant}
 
@@
 The challenge describes in a second part a Software engineering process for a fictional Acme company.
-Base metamodel as described in previous section is too generic to capture all domain-specific aspects of this use case.
-We choose to complete the architectual hierarchy with a specific VirtualModel, specific to Acme software development process metamodel, as shown in figure \ref{fig:AcmeArchitecture}.
-\textit{Acme} metamodel inherits and specializes base metamodel, and \textit{Acme model} is defined as an instance of \textit{Acme} metamodel.
-The metamodel inheritance is implemented as VirtualModel's inheritance.
-Any instance of Acme model is either an instance of a concept defined in base metamodel, or a concept defined in specialized Acme metamodel, which may or not inherit from a concept defined in base metamodel.
+The Base metamodel described previously is too generic to capture all domain-specific aspects of this use case.
+We choose to complete the architectural hierarchy with a specific virtual model, specific to the Acme software development process metamodel, shown in figure~\ref{fig:AcmeFullArchitecture}.
+The \textit{Acme} metamodel specializes the Base metamodel, and \textit{Acme model} is an instance of the \textit{Acme} metamodel.
+The metamodel inheritance is implemented by virtual model's inheritance.
+Any instance of the Acme model is either an instance of a concept defined in the Base metamodel, or a concept defined in the specialized \textit{Acme} metamodel.
 % FML follows a classical object-oriented semantics. ???
 
-\begin{figure}
- \centering
-    \includegraphics[width=1.0 \columnwidth]{Figures/AcmeArchitecture.pdf}
-     \caption{Acme software development process architecture}
-    \label{fig:AcmeArchitecture}
-\end{figure}
+%\begin{figure}[t]
+% \centering
+%    \includegraphics[width=1.0 \columnwidth]{Figures/AcmeArchitecture.pdf}
+%     \caption{Acme software development process architecture}
+%    \label{fig:AcmeArchitecture}
+%\end{figure}
 
@@
 
-\todo[inline]{Augmenter la taille des fontes de la figure \ref{fig:AcmeFullArchitecture}}
-\todo[inline]{Ajouter les programming languages}
+\todo[inline]{Augmenter la taille des fontes de la figure \ref{fig:AcmeFullArchitecture} <- a priori non}
+\todo[inline]{Ajouter les programming languages <- je les vois apparaitre dans la figure, ou alors c'était un todo pour autre chose ?}
 
-Figure \ref{fig:AcmeFullArchitecture} shows the full architecture capturing Acme software development process, with an instance of enacted \textit{Software Engineering Process}, and highlights all conceptual levels.
-Acme metamodel enriches base metamodel by offering specializing concepts: \textit{SETaskType} extending \textit{TaskType}, a more specialized concept \textit{CodingTaskType} extending \textit{SETaskType}, \textit{SEArtifactType} extending \textit{ArtifactType}, a more specialized concept \textit{CodeArtifactType} extending \textit{SEArtifactType}.
-Acme metamodel also defines task type \textit{Coding} as a specialized concept of \textit{Task}, and provides concept \textit{Developer} extending \textit{ActorType}.
-This metamodel is completed with \textit{ProgrammingLanguage} concept, defined as an enumeration (\textit{Java}, \textit{C}, \textit{COBOL}).
-All instances required to capture challenge use case are still defined in final Acme model itself instance of Acme metamodel.
-Instances are represented with rounded boxes and linguistic instantiation are represented with dashed connectors.
+The Figure~\ref{fig:AcmeFullArchitecture} highlights all conceptual levels of the full architecture.
+It also contains an instance of enacted a \textit{Software Engineering Process} at the bottom.
+The Acme metamodel specializes some concepts of the Base metamodel, for example, \texttt{SETaskType} extends \texttt{TaskType} and \texttt{SEArtifactType} extends \texttt{ArtifactType}.
+These concepts are further specialized  \texttt{CodingTaskType} extends \texttt{SETaskType} and \texttt{CodeArtifactType} extends \texttt{SEArtifactType}.
+The Acme metamodel also defines a specialized task type \texttt{Coding} extending \texttt{Task}, and provides the \texttt{Developer} concept extending \texttt{ActorType}.
+This metamodel is completed with the \texttt{ProgrammingLanguage} concept, defined as an enumeration (\texttt{Java}, \texttt{C}, \texttt{COBOL}).
+All the instances required to capture the challenge use case are defined in the final Acme model (lower part of the figure).
+Instances are represented with rounded boxes and linguistic instantiation is represented using dashed connectors.
 
-The \textit{Software Development Process} for Acme company is presented in the figure \ref{fig:AcmeSoftwareDevelopmentProcess} which is a screen-capture from the tooling developed in the context of the challenge and described in the next section.
+The \textit{Software Development Process} for Acme company is presented in the figure~\ref{fig:AcmeSoftwareDevelopmentProcess}.
+It is a screen-capture of the tool developed for the challenge and described in the next section.
 
-\begin{figure}
+\begin{figure}[t]
  \centering
@@
 
-\textit{"Requirements analysis"} is defined as an instance of \textit{SETaskType}, \textit{Analyst} as an instance of \textit{ActorType} and \textit{Requirement specifications} as an instance of \textit{SEArtifactType}.
-\textit{"Requirements analysis"} instance defines property values \textit{producedArtifactTypes=\{Requirement specifications\}} and \textit{allowedActorTypes=\{Analyst\}} (\textbf{S1}).
-In the same way, \textit{"Test case design"} is defined as an instance of \textit{SETaskType}, marked as critical with property value \textit{isCritical=\{true\}}, and bound to \textit{Analyst} with \textit{allowedActorTypes} relation.
-\textit{"Test case design"} produces \textit{"Test cases"} defined as an instance of \textit{SETaskType}.
+% J'utilise sf pour les instances et '
+
+\textsf{'Requirements analysis'} is defined as an instance of \texttt{SETaskType}, \textsf{'Analyst'} is an instance of \texttt{ActorType} and \textsf{'Requirement specifications'} is an instance of \texttt{SEArtifactType}.
+The value of the property \texttt{producedArtifactTypes} of \textsf{'Requirements analysis'} is \texttt{\{}\textsf{Requirement specifications}\texttt{\}} and is \texttt{\{}\textsf{Analyst}\texttt{\}} for \texttt{allowedActorTypes} (\textbf{S1}).
+Similarly, \textsf{Test case design} is an instance of \texttt{SETaskType}, with a value of \texttt{true} for its property \texttt{isCritical}, and bound to \texttt{Analyst} by the \texttt{allowedActorTypes} relation.
+\textit{"Test case design"} produces \textsf{"Test cases"} defined as an instance of \textit{SETaskType}.
 The latter is used in "Test design review" (\textit{producedArtifactTypes} relation) (\textbf{S1} and \textbf{S13}, satisfied with \textbf{P9}).
+\noteFabien{pas de Test case design dans la figure}
 

### 2021-05-31 5-analysis.tex

@@
 
-\noteAntoine{Pourquoi S9 et S13 sont ambigus, quels choix on a fait...}
-\todo[inline]{noms pas dans la spec, complètement implicite ; }
+%\noteAntoine{Pourquoi S9 et S13 sont ambigus, quels choix on a fait...}
+%\todo[inline]{noms pas dans la spec, complètement implicite ; }
 
-We now analyse a few specifications that deserve a comment for the choices we made.
+%We now analyse a few specifications that deserve a comment for the choices we made.
 
-\begin{itemize}
-    \item We added an implicit specification (S0) saying that all things have names.
-    \item (S9) It is unclear when a tested artefact must be associated to its test report. At the creation of the artefact or of its report? \todo[inline]{We choose\dots}
-    \item (S13) [linked to P9] \todo[inline]{We choose ti interpret \dots}
-\end{itemize}
+%\begin{itemize}
+%    \item We added an implicit specification (S0) saying that all things have names.
+%    \item (S9) It is unclear when a tested artefact must be associated to its test report. At the creation of the artefact or of its report? \todo[inline]{We choose\dots}
+%    \item (S13) [linked to P9] \todo[inline]{We choose to interpret \dots}
+%\end{itemize}
+
+%The analysis of the specification should be presented in this section. However, t
+The choices we made are complicated to justify out of context. That is why our choices are presented within the next section model presentation.
 

### 2021-05-31 2-abstract.tex
Updates from Overleaf

--- 2021-05-31 2-Abstract.tex
@@
 
-\noteSylvain{Ce qui m'embête un peu avec cet abstract c'est qu'on met en avant la fédération de modèles alors que sur ce cas précis on en fait pas vraiment (sauf pour la partie outillage avec les éditeurs diagrammatiques).
+%\noteSylvain{Ce qui m'embête un peu avec cet abstract c'est qu'on met en avant la fédération de modèles alors que sur ce cas précis on en fait pas vraiment (sauf pour la partie outillage avec les éditeurs diagrammatiques).
 Concernant la fédération de modèle proprement dite, je crois qu'il faut qu'on montre qu'on a besoin d'une couche conceptuelle (qu'on relie à l'espace technique via les connecteurs techniques, les model slots et les rôles).
@@
 Et cerise sur le gâteau: on peut construire un outillage graphique pour editer et exécuter ce use case.}
-\noteSylvain{En revanche, on peut arguer que pouvoir au même niveau manipuler modèle et métamodèle nous simplifie grandement les choses pour adresser des problématiques multi-niveaux > un bout de la conclusion}.
+%\noteSylvain{En revanche, on peut arguer que pouvoir au même niveau manipuler modèle et métamodèle nous simplifie grandement les choses pour adresser des problématiques multi-niveaux > un bout de la conclusion}.
 %The models in a federation remain autonomous and represented in their original technological spaces %whereas virtual models and links serve as control components used to present different views to the %users and  maintain synchronization.

### 2021-05-31 6-modelpresentation.tex
Updates from Overleaf

@@
 
-In this section, we present the base metamodel presented in figure~\ref{fig:BaseMetamodel} with \textit{XSure} insurance domain use case, whose partial description was provided in the challenge description.
+In this section, we present the base metamodel presented in Figure~\ref{fig:BaseMetamodel} with \textit{XSure} insurance domain use case, whose partial description was provided in the challenge description.
 The requirements \textbf{P1} to \textbf{P19} of XSure insurance domain are straightforwardly implemented by instantiating the \emph{XSure model}, as an instance of base metamodel (the left side of figure~\ref{fig:MultilevelArchitecture}).
@@
 
-\todo{Gérer également la fin de P9, non traité pour le moment: artifact they produced must be associated with a validation task}
-
 \subsubsection{Process enactment}
@@
 The semantics is unclear relatively to the instantiation policy for artifacts.
+In the requirements, nothing is explicitly stated about the link between the \emph{artifact type} of a \emph{task type} and the type of the corresponding artifact
+
 We assume that task execution implies that for each used and produced \emph{artifact type} defined in a related \emph{task type}, it exists at least one \emph{artifact} declaring required ontologic instantiation (through \textit{type} role).
@@
 \noteFabien{Je ne suis pas bien sûr de comprendre le paragraphe}
-\noteJC{autre remarque qui n'a rien à voir: si on suit les recommandations de l'appel, ce paragraphe devrait se retrouver dans la partie analyse étant donné qu'il établit un choix/une interprétation.
-Mais c'est délicat à intégrer sans le contexte du modèle de cette section.
-Ou alors il faudrait une liste, avec des références en avant}
 
@@
      \includegraphics[width=1.0 \textwidth]{Figures/AcmeFullArchitecture.pdf}
-     \caption{Acme software development process architecture}
+     \caption{Acme software development process architecture (not all instances are shown for readability reasons)}
     \label{fig:AcmeFullArchitecture}
@@
 
-\todo[inline]{Augmenter la taille des fontes de la figure \ref{fig:AcmeFullArchitecture} <- a priori non}
 \todo[inline]{Ajouter les programming languages <- je les vois apparaitre dans la figure, ou alors c'était un todo pour autre chose ?}
@@
 The Figure~\ref{fig:AcmeFullArchitecture} highlights all conceptual levels of the full architecture.
+This figure is only partial at instance level and not all instances are represented.
 It also contains an instance of enacted a \textit{Software Engineering Process} at the bottom.
@@
 
-\textsf{'Requirements analysis'} is defined as an instance of \texttt{SETaskType}, \textsf{'Analyst'} is an instance of \texttt{ActorType} and \textsf{'Requirement specifications'} is an instance of \texttt{SEArtifactType}.
+\textsf{'Requirements analysis'} is an instance of \texttt{SETaskType}, \textsf{'Analyst'} is an instance of \texttt{ActorType} and \textsf{'Requirement specifications'} is an instance of \texttt{SEArtifactType}.
 The value of the property \texttt{producedArtifactTypes} of \textsf{'Requirements analysis'} is \texttt{\{}\textsf{Requirement specifications}\texttt{\}} and is \texttt{\{}\textsf{Analyst}\texttt{\}} for \texttt{allowedActorTypes} (\textbf{S1}).
-Similarly, \textsf{Test case design} is an instance of \texttt{SETaskType}, with a value of \texttt{true} for its property \texttt{isCritical}, and bound to \texttt{Analyst} by the \texttt{allowedActorTypes} relation.
-\textit{"Test case design"} produces \textsf{"Test cases"} defined as an instance of \textit{SETaskType}.
-The latter is used in "Test design review" (\textit{producedArtifactTypes} relation) (\textbf{S1} and \textbf{S13}, satisfied with \textbf{P9}).
-\noteFabien{pas de Test case design dans la figure}
+Similarly, \textsf{'Test case design'} is an instance of \texttt{SETaskType}, with a value of \texttt{true} for its property \texttt{isCritical}, and bound to \textsf{'Analyst'} by the \texttt{allowedActorTypes} relation.
+\textsf{'Test case design'} produces \textsf{'Test cases'} instance of \texttt{SETaskType}.
+The latter is used in \textsf{'Test design review'} (\texttt{producedArtifactTypes} relation) (\textbf{S1} and \textbf{S13}, satisfied with \textbf{P9}).
 
-The capture of \textbf{S3} requirement is performed while defining \textit{Coding} as an instance of concept \textit{CodingTaskType} extending \textit{SETaskType} and defining a relation \textit{languages} to \textit{ProgrammingLanguage} with 1..* cardinality.
-\textit{newTask(String)} behavior overrides generic implementation while instantiating a \textit{Coding} concept, specializing \textit{SEArtifact}, as shown in following excerpt of FML code:
+The capture of \textbf{S3} requirement is performed while defining \texttt{Coding} as an instance of concept \texttt{CodingTaskType} and defining a relation \texttt{languages} to \texttt{ProgrammingLanguage} with 1..* cardinality.
+The \texttt{newTask(String)} behavior overrides the generic implementation by instantiating a \texttt{Coding} concept, specializing \texttt{Task}, as shown in following excerpt of FML code:
 
@@
 
-\textit{Code} concept extends \textit{SEArtifact}, and defines a relation \textit{language} to \textit{ProgrammingLanguage} with single cardinality (\textbf{S4}).
-\textbf{S5} is more ambigous as task type \textit{Coding} defines one or more programming languages but produces code which is expressed in one language only.
-This requirement is captured through the redefinition of \textit{declaresRequiredProducedArtifacts()} where programming language should also match.
-\textbf{S6} is guaranteed though the \textit{language} relation defined in \textit{Code} and following invariant declared in \textit{Code}:
-
+The \texttt{Code} concept extends \texttt{SEArtifact}, and defines a relation \texttt{language} to \textit{ProgrammingLanguage} with single cardinality (\textbf{S4}). 
+\noteFabien{Code devient CodeArtifact}
+\textbf{S5} is more ambiguous as a task type \texttt{Coding} defines one or more programming languages but produces code which is expressed in one language only.
+This requirement is captured through the redefinition of the behavior \texttt{declaresRequiredProducedArtifacts} where programming language should also match.
+\textbf{S6} is guaranteed though the \texttt{language} relation defined in \texttt{Code} and the following invariant declared in \texttt{Code}:  
+\noteFabien{Coding devient CodingTask}
 \begin{lstlisting}
@@
 
-\textit{Ann Smith} is the only one allowed to perform coding in COBOL (\textbf{S7}). This is implemented through the redefinition of \textit{newTask(String)} in \textit{CodingTaskType}:
+\textsf{Ann Smith} is the only one allowed to perform coding in COBOL (\textbf{S7}). This is implemented through the redefinition of \texttt{newTask(String)} in \texttt{CodingTaskType}:
 
@@
 
-\textbf{S8} requirement is simply captured with the definition of \textit{Testing} instance of \textit{SETaskType}, \textit{Tester} instance of \textit{ActorType}, and \textit{Test report} instance of \textit{ArtifactType}.
+The requirement \textbf{S8} is simply expressed by the definition of the \textsf{'Testing'} instance of \texttt{SETaskType}, \textsf{'Tester'} instance of \texttt{ActorType}, and \textsf{'Test report'} instance of \texttt{ArtifactType}.
 
-\todo[inline]{requirement S9 non couvert pour le moment: compléter}
+Critical task must additionnaly produce artifacts that must be associated with a validation task.
+This is modelled with a supplementary relation \texttt{validatedWith} defined in \texttt{Artifact} and referencing \texttt{Artifact} validating it (\textbf{S9}).
 
@@
 
-\textbf{S13} requirement has been previously partially fulfilled. This must be completed with...
-
-\todo[inline]{requirement S13 pas complètement couvert : compléter}
+\textbf{S13} requirement has been previously partially fulfilled.
+This must be completed with an association between artifact produced and the task that validates it.
+This is modelled by the reference to 'Test Cases' as a produced artifact type in 'Test Case design' and the reference to 'Test Cases' as a used artifact in 'Test Design review', as shown in figure \ref{fig:AcmeSoftwareDevelopmentProcess}.
 

### 2021-05-31 8-discussion.tex
Updates from Overleaf

@@
   or fewer levels depending on the variations of the use case. However, the number of levels is related
-  to the problem, making the approach fixed. For example, given a requirement, an instance cannot be 
-  changed into a concept easily.
+  to the problem, making the approach fixed.
 
@@
   build the metamodels and the models.  Note that in our approach, the
-  metamodels are built in a ad hoc way together with the models
+  metamodels are built in an ad hoc way together with the models
   (co-construction), therefore the user also validates on a ad hoc basis.
@@
   to other ones. Then, we would add any necessary constraints on those concepts
-  and on the relationships linking them to the other existing concepts.
+  and on the relationships linking them to the other existing concepts. This type of change can be easily done.
+
+  On the other hand, making changes to the existing models could be difficult.
+For example, given a requirement, an instance cannot be changed into a concept easily.
+In the context of the process described in the challenge, this could
+
+  Design Task ->
 

### 2021-05-31 9-relatedworks.tex
Updates from Overleaf

@@
 %Salvador commence
-A plethora of multi-level modeling approaches and tools with different foundations have appeared in the recent years~\footnote{\url{https://homepages.ecs.vuw.ac.nz/Groups/MultiLevelModeling/MultiTools}}.
+A plethora of multi-level modeling approaches and tools with different foundations have appeared in the recent years\footnote{\url{https://homepages.ecs.vuw.ac.nz/Groups/MultiLevelModeling/MultiTools}}.
 Comparing them lies out of the scope of this paper.

### 2021-05-31 8-discussion.tex

@@
   to other ones. Then, we would add any necessary constraints on those concepts
-  and on the relationships linking them to the other existing concepts. This type of change can be easily done.
-
-  On the other hand, making changes to the existing models could be difficult.
-For example, given a requirement, an instance cannot be changed into a concept easily.
-In the context of the process described in the challenge, this could
-
-  Design Task ->
+  and on the relationships linking them to the other existing concepts. This type 
+  of change can be easily done. On the other hand, making changes to the existing 
+  models could be difficult. For example, given a requirement, an instance cannot 
+  be changed into a concept easily. This case could occur if the challenge had 
+  refined the specification by requiring specific design tasks. In this context, 
+  we would have to transform \emph{Design Task} into a concept whose instances 
+  could have been "using agile methodology".
 

### 2021-05-31 6-modelpresentation.tex
Updates from Overleaf

@@
 A \texttt{TaskType} is embedded in a \texttt{ProcessType} and inherits from its context.
-To illustrate this in \textit{XSure} insurance domain use case, \textit{XSure model} defines \textsf{Claim Handling}, instance of \texttt{ProcessType}, and \textsf{Receive Claim}, \textsf{Assess Claim} and \textsf{Pay premium}, instances of \texttt{TaskType}.
+To illustrate this in \textit{XSure} insurance domain use case, \textit{XSure model} defines \inst{Claim Handling}, instance of \texttt{ProcessType}, and \inst{Receive Claim}, \inst{Assess Claim} and \inst{Pay premium}, instances of \texttt{TaskType}.
 
@@
 The same pattern applies for the used and produced artifacts by roles \texttt{usedArtifacts}, \texttt{producedArtifacts} and \texttt{performingActors} defined in \texttt{Task} (\textbf{P13}).
-An instance of \texttt{Artifact} specializes \textit{MultiInstance} and references a set of \texttt{ArtifactType} through the specialized \texttt{/types} role (\textbf{P14} and \textbf{P16}).
+An instance of \texttt{Artifact} specializes \texttt{MultiInstance} and references a set of \texttt{ArtifactType} through the specialized \texttt{/types} role (\textbf{P14} and \textbf{P16}).
 Likewise, \texttt{Actor} specializes \texttt{MultiInstance} and references a set of \texttt{ActorType} through the specialized \texttt{/types} relation (\textbf{P15}).
@@
 The semantics is unclear relatively to the instantiation policy for artifacts.
-In the requirements, nothing is explicitly stated about the link between the \emph{artifact type} of a \emph{task type} and the type of the corresponding artifact
-
-We assume that task execution implies that for each used and produced \emph{artifact type} defined in a related \emph{task type}, it exists at least one \emph{artifact} declaring required ontologic instantiation (through \textit{type} role).
+In the requirements, nothing is explicitly stated about the link between the \emph{artifact type} of a \emph{task type} and the types of the artifacts of a corresponding task.
+Here, we assume that task execution implies that for each used and produced \emph{artifact type} defined in a related \emph{task type}, it exists at least one artifact of the right type.
 The following excerpt of \FML code shows a partial implementation of this.
 The same pattern applies to produced artifacts.
-\noteFabien{Je ne suis pas bien sûr de comprendre le paragraphe}
 
@@
 
-\textsf{'Requirements analysis'} is an instance of \texttt{SETaskType}, \textsf{'Analyst'} is an instance of \texttt{ActorType} and \textsf{'Requirement specifications'} is an instance of \texttt{SEArtifactType}.
-The value of the property \texttt{producedArtifactTypes} of \textsf{'Requirements analysis'} is \texttt{\{}\textsf{Requirement specifications}\texttt{\}} and is \texttt{\{}\textsf{Analyst}\texttt{\}} for \texttt{allowedActorTypes} (\textbf{S1}).
-Similarly, \textsf{'Test case design'} is an instance of \texttt{SETaskType}, with a value of \texttt{true} for its property \texttt{isCritical}, and bound to \textsf{'Analyst'} by the \texttt{allowedActorTypes} relation.
-\textsf{'Test case design'} produces \textsf{'Test cases'} instance of \texttt{SETaskType}.
-The latter is used in \textsf{'Test design review'} (\texttt{producedArtifactTypes} relation) (\textbf{S1} and \textbf{S13}, satisfied with \textbf{P9}).
+\inst{Requirements analysis} is an instance of \texttt{SETaskType}, \inst{Analyst} is an instance of \texttt{ActorType} and \inst{Requirement specifications} is an instance of \texttt{SEArtifactType}.
+The value of the property \texttt{producedArtifactTypes} of \inst{Requirements analysis} is \texttt{\{}\inst{Requirement specifications}\texttt{\}} and is \texttt{\{}\inst{Analyst}\texttt{\}} for \texttt{allowedActorTypes} (\textbf{S1}).
+Similarly, \inst{Test case design} is an instance of \texttt{SETaskType}, with a value of \texttt{true} for its property \texttt{isCritical}, and bound to \inst{Analyst} by the \texttt{allowedActorTypes} relation.
+\inst{Test case design} produces \inst{Test cases} instance of \texttt{SETaskType}.
+The latter is used in \inst{Test design review} (\texttt{producedArtifactTypes} relation) (\textbf{S1} and \textbf{S13}, satisfied with \textbf{P9}).
 
@@
 
-\textsf{Ann Smith} is the only one allowed to perform coding in COBOL (\textbf{S7}). This is implemented through the redefinition of \texttt{newTask(String)} in \texttt{CodingTaskType}:
+\inst{Ann Smith} is the only one allowed to perform coding in COBOL (\textbf{S7}). This is implemented through the redefinition of \texttt{newTask(String)} in \texttt{CodingTaskType}:
 
@@
 
-The requirement \textbf{S8} is simply expressed by the definition of the \textsf{'Testing'} instance of \texttt{SETaskType}, \textsf{'Tester'} instance of \texttt{ActorType}, and \textsf{'Test report'} instance of \texttt{ArtifactType}.
+The requirement \textbf{S8} is simply expressed by the definition of the \inst{Testing} instance of \texttt{SETaskType}, \inst{Tester} instance of \texttt{ActorType}, and \inst{Test report} instance of \texttt{ArtifactType}.
 

### 2021-05-31 6-modelpresentation.tex
sync Overleaf -> Github

@@
 
-\todo[inline]{Ajouter les programming languages <- je les vois apparaitre dans la figure, ou alors c'était un todo pour autre chose ?}
-
-The Figure~\ref{fig:AcmeFullArchitecture} highlights all conceptual levels of the full architecture.
+The Figure~\ref{fig:AcmeFullArchitecture} highlights all conceptual levels of the architecture.
 This figure is only partial at instance level and not all instances are represented.
@@
 Similarly, \inst{Test case design} is an instance of \texttt{SETaskType}, with a value of \texttt{true} for its property \texttt{isCritical}, and bound to \inst{Analyst} by the \texttt{allowedActorTypes} relation.
-\inst{Test case design} produces \inst{Test cases} instance of \texttt{SETaskType}.
+\inst{Test case design} produces \inst{Test cases} instance of \texttt{SETaskType} (\textbf{S2}).
 The latter is used in \inst{Test design review} (\texttt{producedArtifactTypes} relation) (\textbf{S1} and \textbf{S13}, satisfied with \textbf{P9}).
 
-The capture of \textbf{S3} requirement is performed while defining \texttt{Coding} as an instance of concept \texttt{CodingTaskType} and defining a relation \texttt{languages} to \texttt{ProgrammingLanguage} with 1..* cardinality.
-The \texttt{newTask(String)} behavior overrides the generic implementation by instantiating a \texttt{Coding} concept, specializing \texttt{Task}, as shown in following excerpt of FML code:
+The capture of \textbf{S3} requirement is performed while defining \inst{Coding} as an instance of concept \texttt{CodingTaskType} and defining a relation \texttt{languages} to \texttt{ProgrammingLanguage} with 1..* cardinality.
+\noteFabien{* dans la figure} The \texttt{newTask(String)} behavior overrides the generic implementation by instantiating a \texttt{CodingTask} concept, specializing \texttt{Task}, as shown in following excerpt of FML code:
 
@@
   ...
-  public Coding newTask(String name) { 
-    return new Coding(name,this); 
+  public CodingTask newTask(String name) { 
+    return new CodingTask(name,this); 
   }
@@
 
-The \texttt{Code} concept extends \texttt{SEArtifact}, and defines a relation \texttt{language} to \textit{ProgrammingLanguage} with single cardinality (\textbf{S4}). 
-\noteFabien{Code devient CodeArtifact}
+The \texttt{CodeArtifact} concept extends \texttt{SEArtifact}, and defines a relation \texttt{language} to \textit{ProgrammingLanguage} with single cardinality (\textbf{S4}). 
+\noteFabien{pas la cardinalité de la figure}
 \textbf{S5} is more ambiguous as a task type \texttt{Coding} defines one or more programming languages but produces code which is expressed in one language only.
 This requirement is captured through the redefinition of the behavior \texttt{declaresRequiredProducedArtifacts} where programming language should also match.
-\textbf{S6} is guaranteed though the \texttt{language} relation defined in \texttt{Code} and the following invariant declared in \texttt{Code}:  
-\noteFabien{Coding devient CodingTask}
+\textbf{S6} is guaranteed though the \texttt{language} relation defined in \texttt{CodeArtifact} and the following invariant declared in \texttt{CodeArtifact}: 
 \begin{lstlisting}

### 2021-05-31 7-requirementssatisfaction.tex
Updates from Overleaf
sync Overleaf -> Github

@@
 \begin{itemize}
-    \item \textbf{Conceptualization} (\textit{Conc.}): a concept carries the semantics exposed by the requirement (for example for \textbf{P1}: \textit{process type} is conceptualized in a \textit{FlexoConcept} \textit{ProcessType}).
-    \item \textbf{Specialization} (\textit{Spec.}): a concept is specialized in the sense of object-oriented modeling (for example for \textbf{P2}: \textit{Gateway} defines an abstract behavior \texttt{execute()} and \textit{Sequencing}, \textit{AndSplit}, \textit{AndJoin}, \textit{OrSplit} and \textit{OrJoin} are defined as inherited concepts implementing specific business logic).
+    \item \textbf{Conceptualization} (\textit{Conc.}): a concept carries the semantics exposed by the requirement (for example for \textbf{P1}: \textit{process type} is conceptualized in a \textit{FlexoConcept} \texttt{ProcessType}).
+    \item \textbf{Specialization} (\textit{Spec.}): a concept is specialized in the sense of object-oriented modeling (for example for \textbf{P2}: \texttt{Gateway} defines an abstract behavior \texttt{execute()} and \texttt{Sequencing}, \texttt{AndSplit}, \texttt{AndJoin}, \texttt{OrSplit} and \texttt{OrJoin} are defined as inherited concepts implementing specific business logic).
     \item \textbf{Composition} (\textit{Comp.}): some concepts are in relation, or specific attributes were added to some concepts (for example for \textbf{P3}: a \textit{process type} has one \textit{initial task type}).
-    \item \textbf{Linguistic instanciation} (\textit{L.Inst.}): requirement is satisfied by the instantiation of one or more concepts (for example for \textbf{S1}: \textit{requirement analysis} is defined as an instance of \textit{SETasktype}).
+    \item \textbf{Linguistic instanciation} (\textit{L.Inst.}): requirement is satisfied by the instantiation of one or more concepts (for example for \textbf{S1}: \textit{requirement analysis} is defined as an instance of \texttt{SETasktype}).
     \item \textbf{Ontologic instanciation} (\textit{O.Inst.}): fulfillment of requirement is obtained by the relation between an instance and its type definition instance (for example for \textbf{S3} : \textit{developer} has the dual concept nature in \textit{Acme metamodel} and instance nature in \textit{Acme model}).
@@
 
-\begin{figure*}
+\begin{table*}
  \centering
+ \rowcolors{4}{tableShade}{white}  %% start alternating shades from 3rd row
 \begin{tabular}{|c|c|c|c|c|c|c|c|c|}
    \hline
-    & \multicolumn{6}{c|}{Syntactic conformance} & Constraints checking & Tooling\\
-   \hline
+    & \multicolumn{6}{c|}{Syntactic conformance} & \multirow{2}{*}{Constraints checking} & \multirow{2}{*}{Tooling}\\
+   \cline{1-7}
                  & Conc.      & Spec.      & Comp.      & L.Inst.     & O.Inst.     & B.Mod.      &            & \\
@@
     \textbf{P5}  & \checkmark &            & \checkmark &            &            &            &            & \\
-    \textbf{P6}  & \checkmark &            & \checkmark &            & \checkmark &            &            & \\
+    \textbf{P6}  & \checkmark &            & \checkmark &            &  &            &            & \\
     \textbf{P7}  & \checkmark &            & \checkmark &            &            &            &            & \\
@@
     \textbf{P16} & \checkmark &            &            &            & \checkmark &            &            & \\
-    \textbf{P17} &            &            &            &            &            & \checkmark & \checkmark & \checkmark \\
-    \textbf{P18} &            &            &            &            &            & \checkmark &            & \\
+    \textbf{P17} & \checkmark           &            &            &            &            & \checkmark & \checkmark & \checkmark \\
+    \textbf{P18} & \checkmark           &            &            &            &            & \checkmark &            & \\
     \textbf{P19} &            & \checkmark & \checkmark &            &            & \checkmark &            & \\
@@
     \label{tab:RequirementsSatisfactionBaseMetamodel}
-\end{figure*}
+\end{table*}
 
-\begin{figure*}
+\begin{table*}
  \centering
+  \rowcolors{4}{tableShade}{white}  %% start alternating shades from 3rd row
 \begin{tabular}{|c|c|c|c|c|c|c|c|c|}
    \hline
-    & \multicolumn{6}{c|}{Syntactic conformance} & Constraints checking & Tooling\\
-   \hline
+    & \multicolumn{6}{c|}{Syntactic conformance} & \multirow{2}{*}{Constraints checking} & \multirow{2}{*}{Tooling}\\
+   \cline{1-7}
                  & Conc.      & Spec.      & Comp.      & L.Inst     & O.Inst     & B.Mod      &            & \\
@@
     \label{tab:RequirementsSatisfactionAcme}
-\end{figure*}
+\end{table*}

### 2021-05-31 6-modelpresentation.tex
updating table 2

@@
 \noteFabien{pas la cardinalité de la figure}
-\textbf{S5} is more ambiguous as a task type \texttt{Coding} defines one or more programming languages but produces code which is expressed in one language only.
+\textbf{S5} is more ambiguous as a task type \texttt{CodingTask} defines one or more programming languages but produces code which is expressed in one language only.
 This requirement is captured through the redefinition of the behavior \texttt{declaresRequiredProducedArtifacts} where programming language should also match.
@@
   ...
-  public Coding newTask(String name) {
-    Coding returned = new Coding(name,this); 
+  public CodingTask newTask(String name) {
+    CodingTask result = new CodingTask(name,this); 
     if (languages.contains(ProgrammingLanguage.COBOL))
-      returned.addToPerformingActors(getActor("Ann Smith"));
-    return returned;
+      result.addToPerformingActors(getActor("Ann Smith"));
+    return result;
   }
@@
 
-Critical task must additionnaly produce artifacts that must be associated with a validation task.
+A critical task must additionally produce artifacts that must be associated with a validation task.
 This is modelled with a supplementary relation \texttt{validatedWith} defined in \texttt{Artifact} and referencing \texttt{Artifact} validating it (\textbf{S9}).
 
-All software engineering artifacts defined in the context of Acme Software Engineering Process (figure \ref{fig:AcmeSoftwareDevelopmentProcess}) are all instances of \textit{SEArtifact} which defines both attributes \textit{responsible} (an \textit{Actor} instance), and \textit{versionNumber} (an \textit{integer} value), and thus fulfills \textbf{S10}.
+All software engineering artifacts defined in the context of Acme Software Engineering Process are  instances of \textit{SEArtifact} which defines both attributes \textit{responsible} (an \textit{Actor} instance), and \textit{versionNumber} (an \textit{integer} value), and thus fulfills \textbf{S10}.
 \textit{Bob Brown} is declared as an instance of \textit{Actor}, and references \textit{Analyst} and \textit{Tester} (\textit{ActorType} instances).

### 2021-05-31 3-introduction.tex
antidote introduction

--- 2021-05-28 3-Introduction.tex
@@
 
-In order to tackle this problem, the \emph{multi-level} modeling paradigm have been introduced.
+In order to tackle this problem, the \emph{multi-level} modeling paradigm has been introduced.
 Contrary to the strict approach, multi-level modeling advocates the use of a flexible number of levels as well as more flexible relations between them.
@@
 
-Model federation is a multi model management approach based on the use of virtual models and loosely coupled links.
+Model federation is a multi-model management approach based on the use of virtual models and loosely coupled links.
 The models in a federation remain autonomous and represented in their original technological spaces whereas virtual models (also called conceptual models) and links serve as control components used to present different views to the users and maintain synchronization.

### 2021-05-31 4-technology.tex
antidote technology

@@
 metamodel grouping all needed entities, model federation build links
-among models and metamodels (even through levels) to make ``things''
+between models and metamodels (even through levels) to make ``things''
 work together. As an illustration of this feature, we developed a

### 2021-05-31 8-discussion.tex
sync Overleaf -> Github
antidote discussion

@@
   %plan plus ou moins imposé par l'appel
-%\todo[inline]{mandatory discussion aspects ↓}
+%\todo[inline]{mandatory discussion aspects}
 
@@
   metamodel (Figure~\ref{fig:mm}) and described in
-  section~\ref{sec:technology}. Models are \emph{virtual models}. They are
+  the section~\ref{sec:technology}. Models are \emph{virtual models}. They are
   composed of \emph{concepts}, being themselves concepts. Concepts may have
@@
   slots} to connect some concept instances to external elements (in our case:
-  instance of diagram from our diagramming TA). The use of slots has been
-  described in section~\ref{sec:model} and is illustrated by bold circles in
+  instances of diagram from our diagraming TA). The use of slots has been
+  described in the section~\ref{sec:model} and is illustrated by bold circles in
   the Figure~\ref{fig:ToolingArchitecture}).
@@
 
-  Thanks to the level-agnosticity nature of the Openflexo approach, cross-level
+  Thanks to the level-agnosticism nature of the Openflexo approach, cross-level
   relationships are not an issue. Model elements of different levels can be
@@
   metamodels are built in an ad hoc way together with the models
-  (co-construction), therefore the user also validates on a ad hoc basis.
+  (co-construction), therefore the user also validates on an ad hoc basis.
 
@@
   %\noteJC{pas trop clair pour moi pour le moment -> N/A ? }
-  Due to the nature of our approach and its level-agnosticity, \emph{deep characterization} does not 
-  apply to our solution. Such a mechanism could probably be encoded by specific behaviors, however 
+  Due to the nature of our approach and its level-agnosticism, \emph{deep characterization} does not 
+  apply to our solution. Such a mechanism could probably be encoded by specific behaviors, however, 
   it would not be a generic mechanism, making it difficult to reuse for another problem.
@@
   and on the relationships linking them to the other existing concepts. This type 
-  of change can be easily done. On the other hand, making changes to the existing 
-  models could be difficult. For example, given a requirement, an instance cannot 
-  be changed into a concept easily. This case could occur if the challenge had 
-  refined the specification by requiring specific design tasks. In this context, 
-  we would have to transform \emph{Design Task} into a concept whose instances 
-  could have been "using agile methodology".
+  of change can be done without much effort. On the other hand, making changes 
+  to the existing models could be difficult. For example, given a requirement, 
+  an instance cannot be changed into a concept easily. This case could occur if 
+  the challenge had refined the specification by requiring specific design tasks. 
+  In this context, we would have to transform \emph{Design Task} into a concept 
+  whose instances could have been "using agile methodology".
 
@@
 
-%\todo[inline]{recommended discussion aspects ↓}
+%\todo[inline]{recommended discussion aspects}
 \subsection{Tools}
@@
 
-As described in section~\ref{subsec:tooling}, our solution is fully tooled with
+As described in the section~\ref{subsec:tooling}, our solution is fully tooled with
 Openflexo.  We use the Openflexo infrastructure to build models and metamodels
@@
 Our tooling includes mechanisms to verify the models. First, syntactical
-consistency is realised by cardinality checks and by typing. Second, the
+consistency is realized by cardinality checks and by typing. Second, the
 constraints the users have written are continuously checked. Thus, a part of

### 2021-05-31 5-analysis.tex
antidote analysis

@@
 
-During the next step of our approach, we try to take into account the reusability of the relationships by identifying the semantics of each relationship.
+During the next step of our approach, we try to take into account the reusability of the relationships by identifying the semantics of them.
 The goal is the identification of concepts with their behavior.
@@
 But our goal during the challenge's analysis remains to produce a \FML virtual model architecture for our federation.
-As shown in figure~\ref{fig:MultilevelArchitecture}, we defined two virtual models (Base and Acme processes abstractions) instantiated by two virtual model instances (XSure and Acme processes).
+As shown in the figure~\ref{fig:MultilevelArchitecture}, we defined two virtual models (Base and Acme processes abstractions) instantiated by two virtual model instances (XSure and Acme processes).
 As the figure shows the virtual models play the role of metamodels, in the sense of concept definition with a level-agnostic approach.
@@
 The resulting architecture follows the way the challenge is presented and this
-organisation allows for the possibility of flexible extensions.
+organization allows for the possibility of flexible extensions.
 
@@
 
-Our analysis of the use case leads to identify two main modeling axis (as presented in figure \ref{fig:LinguisticAndOntologicInstantiation}):
+Our analysis of the use case leads to identify two main modeling axis (as presented in the figure \ref{fig:LinguisticAndOntologicInstantiation}):
 \begin{itemize}
@@
 The set of resulting instances comes from several virtual models
-of any abstraction level as detailed in section \ref{sec:AcmeSoftwareDevelopmentProcess}
+of any abstraction level as detailed in the section. \ref{sec:AcmeSoftwareDevelopmentProcess}
 
@@
 
-Finally as explained previously, the defined level for the Acme  and the XSure models is the result of the instantiation of virtual models.
+Finally, as explained previously, the defined level for the Acme  and the XSure models is the result of the instantiation of virtual models.
 In our approach this virtual model instance level can't be specialized or extended but all the other virtual models could be extended by any concepts, as a new federated model.
-Also, the instance level can be instantiated from any virtual model, which represent any metamodel level.
+Also, the instance level can be instantiated from any virtual model, which represents any metamodel level.
 

### 2021-05-31 9-relatedworks.tex
antidote related work

@@
 %Salvador commence
-A plethora of multi-level modeling approaches and tools with different foundations have appeared in the recent years\footnote{\url{https://homepages.ecs.vuw.ac.nz/Groups/MultiLevelModeling/MultiTools}}.
+A plethora of multi-level modeling approaches and tools with different foundations have appeared in recent years\footnote{\url{https://homepages.ecs.vuw.ac.nz/Groups/MultiLevelModeling/MultiTools}}.
 Comparing them lies out of the scope of this paper.
@@
 
-A first description of process modeling as a multi-level modeling problem was proposed by \cite{lara2018refactoring} in the context of a catalogue of refactoring for multi-level models (in a simpler form with less constraints and requirements w.r.t.
+A first description of process modeling as a multi-level modeling problem was proposed by \cite{lara2018refactoring} in the context of a catalogue of refactoring for multi-level models (in a simpler form with fewer constraints and requirements w.r.t.
 the challenge version).
@@
 However they need to do frequent model transformations to transform instance models into instantiable models.
-With respect to the proposed solution, we use less models as we do not include a generic software engineering process model nor a generic assurance process model.
+With respect to the proposed solution, we use fewer models as we do not include a generic software engineering process model nor a generic assurance process model.
 %Finally, they use multi-level couple model transformations for the definition and enforcement of constraints.
@@
 Similarly, \cite{dmla2019} use the Dynamic Multi-Layer Algebra (DMLA) \citep{dmla2017} in order to solve the process challenge.
-DMLA is a level-blind modeling framework which is fully customizable (\eg different types of instantiaton may be implemented).
-Their proposed solution separates the process challenge in two separate domains, namely, the tasks definition domain and the process definition domain.
+DMLA is a level-blind modeling framework which is fully customizable (\eg different types of instantiation may be implemented).
+Their proposed solution separates the process challenge in two separate domains, namely, the task definition domain and the process definition domain.
 Wrappers are used in order to reuse tasks in processes.

### 2021-05-31 10-conclusions.tex
sync Overleaf -> Github
antidote conclusions

@@
 We fulfill all the challenge requirements and we propose a tooling that demonstrates the usability of our solution.
-We developed an ad-hoc solution, including models and metamodels, thanks to the flexibility provided by the metametamodeling infrastructure offered by Openflexo.
+We developed an ad hoc solution, including models and metamodels, thanks to the flexibility provided by the metametamodeling infrastructure offered by Openflexo.
 
@@
 The solution would be developed from examples (Acme and XSure processes for instance) from which models would be identified.
-Finally, the realisation of this challenge highlighted the interest of integrating specific behaviors for the management of multi-level concepts in the Openflexo infrastructure (as opposed to tailoring them as we did in this exercise).
-We intend to do so.
+Finally, the realization of this challenge highlighted the interest of integrating specific behaviors for the management of multi-level concepts in the Openflexo infrastructure.
+%We intend to do so.

### 2021-05-31 6-modelpresentation.tex

@@
 
-In this section, we present the base metamodel presented in Figure~\ref{fig:BaseMetamodel} with \textit{XSure} insurance domain use case, whose partial description was provided in the challenge description.
+In this section, we present the base metamodel presented in Figure~\ref{fig:BaseMetamodel} with the \textit{XSure} insurance domain use case, whose partial description was provided in the challenge description.
 The requirements \textbf{P1} to \textbf{P19} of XSure insurance domain are straightforwardly implemented by instantiating the \emph{XSure model}, as an instance of base metamodel (the left side of figure~\ref{fig:MultilevelArchitecture}).
@@
 
-All software engineering artifacts defined in the context of Acme Software Engineering Process are  instances of \textit{SEArtifact} which defines both attributes \textit{responsible} (an \textit{Actor} instance), and \textit{versionNumber} (an \textit{integer} value), and thus fulfills \textbf{S10}.
-\textit{Bob Brown} is declared as an instance of \textit{Actor}, and references \textit{Analyst} and \textit{Tester} (\textit{ActorType} instances).
-He is also referenced by all instances of \textit{SETaskType} as the creator for related task types (\textbf{S11}).
+All software engineering artifacts defined in the context of Acme Software Engineering Process are  instances of \texttt{SEArtifact}.
+This concept defines two attributes: \texttt{responsible} (an \texttt{Actor} instance), and \texttt{versionNumber} (an integer).
+It thus fulfills \textbf{S10}.
+\inst{Bob Brown} is declared as an instance of \texttt{Actor}, and references \inst{Analyst} and \inst{Tester} (\texttt{ActorType} instances).
+He is also referenced by all instances of \texttt{SETaskType} as the creator for related task types (\textbf{S11}).
 
-\textbf{S12} is somehow ambiguous as it implies an implicit semantics regarding underlying business logic of process execution.
-We assume that all tasks may define an expected duration, which might be checked during process execution.
-This can be modeled through the addition of \textit{expectedDuration} attribute in \textit{TaskType}.
-Alternatively, this can be modeled through the definition of \textit{SETesting} concept, as a specialization of \textit{SETaskType}, and the instantiation of \textit{Testing} as an instance of \textit{SETesting}.
-The business logic expressed by \textbf{S12} requirement should be redefined in \textit{SETesting}.
+For \textbf{S12}, we assume that all tasks may define an expected duration, which might be checked during process execution.
+This is modeled by the \texttt{expectedDuration} attribute in \texttt{TaskType}.
+Alternatively, this could be modeled through the definition of a \texttt{SETesting} concept, as a specialization of \texttt{SETaskType}, and the instantiation of \inst{Testing} as an instance of \texttt{SETesting}.
+The business logic expressed by \textbf{S12} requirement should then be redefined in \texttt{SETesting}.
 
 \textbf{S13} requirement has been previously partially fulfilled.
-This must be completed with an association between artifact produced and the task that validates it.
-This is modelled by the reference to 'Test Cases' as a produced artifact type in 'Test Case design' and the reference to 'Test Cases' as a used artifact in 'Test Design review', as shown in figure \ref{fig:AcmeSoftwareDevelopmentProcess}.
+This must be completed with an association between a artifact produced and the task that validates it.
+This is modelled by the reference to \inst{Test Cases} as a produced artifact type in \inst{Test Case design} and the reference to \inst{Test Cases} as a used artifact in \inst{Test Design review}, as shown in figure~\ref{fig:AcmeSoftwareDevelopmentProcess}.
 
@@
 
-Our solution is fully implemented within Openflexo tool. Both usecases have been modeled in the interactive design environment and are executable by the FML execution engine.
+Our solution is fully implemented within the Openflexo tool. Both use cases have been modeled in the interactive design environment and are executable by the \FML execution engine.
 
-Base metamodel has been completed with some behaviors implementing execution semantic for executed processes.
-All tasks - instantiated from \textit{TaskType} for a given enacted process - manage a status which can be \textit{Not startable} (when not assigned to a performing actor or when required input artifacts are not available) , \textit{Startable}, \textit{Started}, \textit{Completable} (when all output artifacts are ready), \textit{Completed}).
-A \textit{Task} also manages a set of performing actors, a begin and end date, some used and produced artifacts.
-\textit{Gateways} business logic has also been implemented through the implementation of abstract \texttt{execute(Process)} behavior.
-Management of artifacts whose semantics follows rules defined in section \ref{sec:ProcessEnactment} (\textbf{P13}) has also been implemented.
+The Base metamodel is completed with some behaviors implementing an execution semantics for the executed processes.
+All tasks -- instantiated from \texttt{TaskType} for a given enacted process -- manage a status.
+This status is either \texttt{Not startable} (when not assigned to a performing actor or when required input artifacts are not available) , \texttt{Startable}, \texttt{Started}, \texttt{Completable} (when all output artifacts are ready) or \texttt{Completed}.
+A \texttt{Task} also manages a set of performing actors, a begin and end date, some used and produced artifacts.
+The \texttt{Gateway} business logic has also been implemented through the implementation of an abstract \texttt{execute(Process)} behavior.
+The management of artifacts whose semantics follows rules defined in section~\ref{sec:ProcessEnactment} (\textbf{P13}) is also implemented.
 
-We took advantage of model federation and the availability of diagramming features though the \textit{Diagramming TA} to implement two interactive graphical tools built on top of conceptual levels detailed in previous section.
-A first tool offers a graphical edition of a Process type, while the second tool offers enactment feature (the instantiation of a process from its process type), the ability within an enacted process to assign tasks to some actors, and the execution of this process with a graphical visualization.
+We took advantage of model federation and the availability of diagramming features though the \textit{Diagramming TA} to implement two interactive graphical tools built on top of the conceptual levels detailed in previous sections.
+A first tool offers a graphical edition of a Process type.
+The second one offers enactment feature (the instantiation of a process from its process type), the ability within an enacted process to assign tasks to actors, and a graphical visualization of this process execution.
 
@@
 Figure \ref{fig:ToolingArchitecture} shows the architecture of these two tools.
-Left of figure presents \textit{ProcessType graphical editor}.
-\textit{ProcessTypeEditor} is modeled as a \textit{VirtualModel} declaring two model slots (represented with bold circles).
+The left of the figure presents the \textit{ProcessType graphical editor}.
+\texttt{ProcessTypeEditor} is modeled as a virtual model declaring two model slots (represented with bold circles).
 The first model slot references an instance of Acme metamodel, while the second model slot references an instance of diagram, conforms to the \textit{ProcessType diagram specification} (a diagram "metamodel" which defines and specify structure and graphical representations for edited items).

### 2021-05-31 6-modelpresentation.tex
Updates from Overleaf

@@
 \texttt{ProcessTypeEditor} is modeled as a virtual model declaring two model slots (represented with bold circles).
-The first model slot references an instance of Acme metamodel, while the second model slot references an instance of diagram, conforms to the \textit{ProcessType diagram specification} (a diagram "metamodel" which defines and specify structure and graphical representations for edited items).
-While executed, this tool manages a graphical view for an instance of Acme metamodel (the \textit{Acme model}) and a specific \textit{ProcessType} instance.
-This tools allows to represent and edit a \textit{ProcessType}.
-One particular point should be noted regarding the highly reflective nature of FML and associated tooling : when drag and drop interactor is applied for a new item, the tool allows to choose the concept type to be instantiated (when a \textit{TaskType} is created for example, the user must choose the sub-concept of \textit{TaskType} to be selected - in Acme use case it can be \textit{SETaskTask} or \textit{CodingTaskType} or default value \textit{TaskType}).
+The first model slot references an instance of Acme metamodel, while the second model slot references an instance of diagram, conforms to the \textit{ProcessType diagram specification}\footnote{a diagram "metamodel" which defines and specify structure and graphical representations for edited items}.
+When executed, this tool manages a graphical view for an instance of Acme metamodel (the \textit{Acme model}) and a specific \texttt{ProcessType} instance.
+This tools allows to represent and edit a \texttt{ProcessType}.
+The highly reflective nature of \FML and its tooling should be noted.
+When the drag and drop interactor is applied for a new item, the tool allows to choose the concept type to be instantiated.
+For example, when a \texttt{TaskType} is created, the user must choose a sub-concept of \texttt{TaskType} -- in Acme use case it can be \texttt{SETaskTask} or \texttt{CodingTaskType} or default value \texttt{TaskType}).
 
-The other tool, called \textit{Enacted process graphical editor} and shown on figure \ref{fig:ScreenshotEnactedProcessEditor}, provides process edition and tasks assignations through a graphical visualization displaying process being executed.
-This tool is represented on the right side of the figure \ref{fig:ToolingArchitecture}.
-It follows the same architecture pattern as explicited for the \textit{ProcessType graphical editor}, with two model slots referencing both the model and a diagram.
-Process enactment is operated from a \textit{ProcessType}, and must be defined using a name identifying the newly instantiated process.
-All tasks are created from their \textit{TaskType} definition, and required assignments applies (if for example "Coding" \textit{TaskType} defines COBOL as output programming language, related task is automatically assigned to "Ann Smith").
+The other tool, called \textit{Enacted process graphical editor} and shown on figure~\ref{fig:ScreenshotEnactedProcessEditor}, provides process edition and tasks assignations through a graphical visualization displaying the process being executed.
+This tool is represented on the right side of the figure~\ref{fig:ToolingArchitecture}.
+It follows the same architecture pattern of the \textit{ProcessType graphical editor}, with two model slots referencing both the model and a diagram.
+Process enactment is operated from a \texttt{ProcessType}, and must be defined using a name identifying the newly instantiated process.
+All tasks are created from their \texttt{TaskType} definition, and required assignments applies (if for example "Coding" \textit{TaskType} defines COBOL as output programming language, related task is automatically assigned to "Ann Smith").
 All tasks get a status as well as a set of performing actors, a begin and end date, and used and produced artifacts.

### 2021-05-31 6-modelpresentation.tex
sync

@@
 
-Our solution is fully implemented within the Openflexo tool. Both use cases have been modeled in the interactive design environment and are executable by the \FML execution engine.
+\begin{figure*}
+ \centering
+     \includegraphics[width=0.9 \textwidth]{Figures/ToolingArchitecture.pdf}
+     \caption{Tooling architecture}
+    \label{fig:ToolingArchitecture}
+\end{figure*}
+
+Our solution is fully implemented within the Openflexo tool. Both use cases have been modeled in the interactive design environment and can be executed by the \FML execution engine.
 
@@
 
-We took advantage of model federation and the availability of diagramming features though the \textit{Diagramming TA} to implement two interactive graphical tools built on top of the conceptual levels detailed in previous sections.
+We took advantage of model federation and the availability of diagramming features though the \textit{Diagramming Technology Adapter} to implement two interactive graphical tools built on top of the conceptual levels detailed in previous sections.
 A first tool offers a graphical edition of a Process type.
 The second one offers enactment feature (the instantiation of a process from its process type), the ability within an enacted process to assign tasks to actors, and a graphical visualization of this process execution.
-
-\begin{figure*}
- \centering
-     \includegraphics[width=0.9 \textwidth]{Figures/ToolingArchitecture.pdf}
-     \caption{Tooling architecture}
-    \label{fig:ToolingArchitecture}
-\end{figure*}
 
@@
 Process enactment is operated from a \texttt{ProcessType}, and must be defined using a name identifying the newly instantiated process.
-All tasks are created from their \texttt{TaskType} definition, and required assignments applies (if for example "Coding" \textit{TaskType} defines COBOL as output programming language, related task is automatically assigned to "Ann Smith").
-All tasks get a status as well as a set of performing actors, a begin and end date, and used and produced artifacts.
+All tasks are created from their \texttt{TaskType} definition, and required assignments applies (if for example \inst{Coding} \texttt{TaskType} defines COBOL as output programming language, related task is automatically assigned to \inst{Ann Smith}).
+Each task has a status as well as a set of performing actors, a begin and end date, and used and produced artifacts.
 
@@
  \centering
-     \includegraphics[width=1.0 \textwidth]{Figures/ScreenshotEnactedProcessEditor.png}
+     \includegraphics[width=\textwidth]{Figures/ScreenshotEnactedProcessEditor.png}
      \caption{Screenshot of the enacted process graphical editor}

### 2021-05-31 7-requirementssatisfaction.tex
sync

@@
 In this section, we summarize the different techniques used to satisfy them, without repeating all the justifications.
-We classify those techniques into three categories:
+We classify these techniques into three categories:
 \begin{itemize}
-    \item \textbf{syntactic conformance}: the requirements are structurally satisfied when the model conforms to its metamodel,
-    \item\textbf{constraints checking}: the requirements are fulfilled when all instances satisfy both syntactic conformance and the evaluation of invariants expressed for related concepts,
-    \item \textbf{tooling}: the requirements are enforced by the implementation of the associated tooling.
+    \item \emph{syntactic conformance}: the requirements are structurally satisfied when the model conforms to its metamodel,
+    \item \emph{constraints checking}: the requirements are fulfilled when all instances satisfy both syntactic conformance and the evaluation of invariants expressed for related concepts,
+    \item \emph{tooling}: the requirements are enforced by the implementation of the associated tooling.
 \end{itemize}
@@
 \begin{itemize}
-    \item \textbf{Conceptualization} (\textit{Conc.}): a concept carries the semantics exposed by the requirement (for example for \textbf{P1}: \textit{process type} is conceptualized in a \textit{FlexoConcept} \texttt{ProcessType}).
-    \item \textbf{Specialization} (\textit{Spec.}): a concept is specialized in the sense of object-oriented modeling (for example for \textbf{P2}: \texttt{Gateway} defines an abstract behavior \texttt{execute()} and \texttt{Sequencing}, \texttt{AndSplit}, \texttt{AndJoin}, \texttt{OrSplit} and \texttt{OrJoin} are defined as inherited concepts implementing specific business logic).
-    \item \textbf{Composition} (\textit{Comp.}): some concepts are in relation, or specific attributes were added to some concepts (for example for \textbf{P3}: a \textit{process type} has one \textit{initial task type}).
-    \item \textbf{Linguistic instanciation} (\textit{L.Inst.}): requirement is satisfied by the instantiation of one or more concepts (for example for \textbf{S1}: \textit{requirement analysis} is defined as an instance of \texttt{SETasktype}).
-    \item \textbf{Ontologic instanciation} (\textit{O.Inst.}): fulfillment of requirement is obtained by the relation between an instance and its type definition instance (for example for \textbf{S3} : \textit{developer} has the dual concept nature in \textit{Acme metamodel} and instance nature in \textit{Acme model}).
-    \item \textbf{Behavioral modeling} (\textit{B.Mod.}): requirement is satisfied by the definition of one more object-oriented behavioral features, which may combined with constraints checking and associated tooling (for example \textbf{P17}).
+    \item \emph{Conceptualization} (\textit{Conc.}): a concept carries the semantics exposed by the requirement (for example for \textbf{P1}: \textit{process type} is conceptualized in a \textit{FlexoConcept} \texttt{ProcessType}).
+    \item \emph{Specialization} (\textit{Spec.}): a concept is specialized in the sense of object-oriented modeling (for example for \textbf{P2}: \texttt{Gateway} defines an abstract behavior \texttt{execute()} and \texttt{Sequencing}, \texttt{AndSplit}, \texttt{AndJoin}, \texttt{OrSplit} and \texttt{OrJoin} are defined as inherited concepts implementing specific business logic).
+    \item \emph{Composition} (\textit{Comp.}): some concepts are in relation, or specific attributes were added to some concepts (for example for \textbf{P3}: a \textit{process type} has one \textit{initial task type}).
+    \item \emph{Linguistic instanciation} (\textit{L.Inst.}): requirement is satisfied by the instantiation of one or more concepts (for example for \textbf{S1}: \textit{requirement analysis} is defined as an instance of \texttt{SETasktype}).
+    \item \emph{Ontologic instanciation} (\textit{O.Inst.}): fulfillment of requirement is obtained by the relation between an instance and its type definition instance (for example for \textbf{S3} : \textit{developer} has the dual concept nature in \textit{Acme metamodel} and instance nature in \textit{Acme model}).
+    \item \emph{Behavioral modeling} (\textit{B.Mod.}): requirement is satisfied by the definition of one more object-oriented behavioral features, which may combined with constraints checking and associated tooling (for example \textbf{P17}).
 \end{itemize}

### 2021-05-31 7-requirementssatisfaction.tex
synchro

@@
 
-Table \ref{tab:RequirementsSatisfactionBaseMetamodel} summarizes requirement covering for the base metamodel illustrated by XSure insurance use case while table \ref{tab:RequirementsSatisfactionAcme} shows requirements satisfaction for Acme software engineering process.
+\begin{table*}
+ \centering
+ \rowcolors{4}{tableShade}{white}  %% start alternating shades from 3rd row
+\begin{tabular}{|c|c|c|c|c|c|c|c|c|}
+   \hline
+    & \multicolumn{6}{c|}{Syntactic conformance} & Constraints & \multirow{2}{*}{Tooling}\\
+   \cline{1-7}
+                 & Conc.      & Spec.      & Comp.      & L.Inst.     & O.Inst.     & B.Mod.   &  checking & \\
+  \hline
+    \textbf{P1}  & \checkmark &            & \checkmark &            &            &            &            & \\
+    \textbf{P2}  & \checkmark & \checkmark & \checkmark &            &            &            &            & \\
+    \textbf{P3}  &            &            & \checkmark &            &            &            &            & \\
+    \textbf{P4}  & \checkmark &            &            &            & \checkmark &            &            & \\
+    \textbf{P5}  & \checkmark &            & \checkmark &            &            &            &            & \\
+    \textbf{P6}  & \checkmark &            & \checkmark &            &  &            &            & \\
+    \textbf{P7}  & \checkmark &            & \checkmark &            &            &            &            & \\
+    \textbf{P8}  &            &            & \checkmark &            &            &            &            & \\
+    \textbf{P9}  &            &            & \checkmark &            &            &            & \checkmark & \\
+    \textbf{P10} & \checkmark &            &            &            & \checkmark &            &            & \\
+    \textbf{P11} & \checkmark &            & \checkmark &            & \checkmark &            &            & \\
+    \textbf{P12} &            &            & \checkmark &            &            &            &            & \\
+    \textbf{P13} & \checkmark &            & \checkmark &            & \checkmark &            &            & \\
+    \textbf{P14} & \checkmark &            &            &            & \checkmark &            &            & \\
+    \textbf{P15} & \checkmark &            &            &            & \checkmark &            &            & \\
+    \textbf{P16} & \checkmark &            &            &            & \checkmark &            &            & \\
+    \textbf{P17} & \checkmark           &            &            &            &            & \checkmark & \checkmark & \checkmark \\
+    \textbf{P18} & \checkmark           &            &            &            &            & \checkmark &            & \\
+    \textbf{P19} &            & \checkmark & \checkmark &            &            & \checkmark &            & \\
+  \hline
+\end{tabular}
+     \caption{Requirements satisfaction for base metamodel}
+    \label{tab:RequirementsSatisfactionBaseMetamodel}
+\end{table*}
+
+Table~\ref{tab:RequirementsSatisfactionBaseMetamodel} summarizes the requirement coverage for the base metamodel illustrated by XSure insurance use case while table~\ref{tab:RequirementsSatisfactionAcme} shows requirements satisfaction for Acme software engineering process.
 Syntactic conformance is more precisely classified into six subcategories:
@@
 \begin{itemize}
-    \item \emph{Conceptualization} (\textit{Conc.}): a concept carries the semantics exposed by the requirement (for example for \textbf{P1}: \textit{process type} is conceptualized in a \textit{FlexoConcept} \texttt{ProcessType}).
+    \item \emph{Conceptualization} (\textit{Conc.}): a concept carries the semantics exposed by the requirement (for example for \textbf{P1}: \textit{process type} is conceptualized in a concept \texttt{ProcessType}).
+
     \item \emph{Specialization} (\textit{Spec.}): a concept is specialized in the sense of object-oriented modeling (for example for \textbf{P2}: \texttt{Gateway} defines an abstract behavior \texttt{execute()} and \texttt{Sequencing}, \texttt{AndSplit}, \texttt{AndJoin}, \texttt{OrSplit} and \texttt{OrJoin} are defined as inherited concepts implementing specific business logic).
+
     \item \emph{Composition} (\textit{Comp.}): some concepts are in relation, or specific attributes were added to some concepts (for example for \textbf{P3}: a \textit{process type} has one \textit{initial task type}).
-    \item \emph{Linguistic instanciation} (\textit{L.Inst.}): requirement is satisfied by the instantiation of one or more concepts (for example for \textbf{S1}: \textit{requirement analysis} is defined as an instance of \texttt{SETasktype}).
-    \item \emph{Ontologic instanciation} (\textit{O.Inst.}): fulfillment of requirement is obtained by the relation between an instance and its type definition instance (for example for \textbf{S3} : \textit{developer} has the dual concept nature in \textit{Acme metamodel} and instance nature in \textit{Acme model}).
-    \item \emph{Behavioral modeling} (\textit{B.Mod.}): requirement is satisfied by the definition of one more object-oriented behavioral features, which may combined with constraints checking and associated tooling (for example \textbf{P17}).
+
+    \item \emph{Linguistic instantiation} (\textit{L.Inst.}): requirement is satisfied by the instantiation of one or more concepts (for example for \textbf{S1}: \inst{Requirement analysis} is defined as an instance of \texttt{SETaskType}).
+
+    \item \emph{Ontologic instantiation} (\textit{O.Inst.}): fulfillment of requirement is obtained by the relation between an instance and its type definition instance (for example for \textbf{S3} : \textit{developer} has the dual concept nature in \textit{Acme metamodel} and instance nature in \textit{Acme model}).
+
+    \item \emph{Behavioral modeling} (\textit{B.Mod.}): requirement is satisfied by the definition of one or more behaviors, which may be combined with constraints checking and associated tooling (for example \textbf{P17}).
 \end{itemize}
@@
    \hline
-    & \multicolumn{6}{c|}{Syntactic conformance} & \multirow{2}{*}{Constraints checking} & \multirow{2}{*}{Tooling}\\
+    & \multicolumn{6}{c|}{Syntactic conformance} & Constraints & \multirow{2}{*}{Tooling}\\
    \cline{1-7}
-                 & Conc.      & Spec.      & Comp.      & L.Inst.     & O.Inst.     & B.Mod.      &            & \\
-  \hline
-    \textbf{P1}  & \checkmark &            & \checkmark &            &            &            &            & \\
-    \textbf{P2}  & \checkmark & \checkmark & \checkmark &            &            &            &            & \\
-    \textbf{P3}  &            &            & \checkmark &            &            &            &            & \\
-    \textbf{P4}  & \checkmark &            &            &            & \checkmark &            &            & \\
-    \textbf{P5}  & \checkmark &            & \checkmark &            &            &            &            & \\
-    \textbf{P6}  & \checkmark &            & \checkmark &            &  &            &            & \\
-    \textbf{P7}  & \checkmark &            & \checkmark &            &            &            &            & \\
-    \textbf{P8}  &            &            & \checkmark &            &            &            &            & \\
-    \textbf{P9}  &            &            & \checkmark &            &            &            & \checkmark & \\
-    \textbf{P10} & \checkmark &            &            &            & \checkmark &            &            & \\
-    \textbf{P11} & \checkmark &            & \checkmark &            & \checkmark &            &            & \\
-    \textbf{P12} &            &            & \checkmark &            &            &            &            & \\
-    \textbf{P13} & \checkmark &            & \checkmark &            & \checkmark &            &            & \\
-    \textbf{P14} & \checkmark &            &            &            & \checkmark &            &            & \\
-    \textbf{P15} & \checkmark &            &            &            & \checkmark &            &            & \\
-    \textbf{P16} & \checkmark &            &            &            & \checkmark &            &            & \\
-    \textbf{P17} & \checkmark           &            &            &            &            & \checkmark & \checkmark & \checkmark \\
-    \textbf{P18} & \checkmark           &            &            &            &            & \checkmark &            & \\
-    \textbf{P19} &            & \checkmark & \checkmark &            &            & \checkmark &            & \\
-  \hline
-\end{tabular}
-     \caption{Requirements satisfaction for base metamodel}
-    \label{tab:RequirementsSatisfactionBaseMetamodel}
-\end{table*}
-
-\begin{table*}
- \centering
-  \rowcolors{4}{tableShade}{white}  %% start alternating shades from 3rd row
-\begin{tabular}{|c|c|c|c|c|c|c|c|c|}
-   \hline
-    & \multicolumn{6}{c|}{Syntactic conformance} & \multirow{2}{*}{Constraints checking} & \multirow{2}{*}{Tooling}\\
-   \cline{1-7}
-                 & Conc.      & Spec.      & Comp.      & L.Inst     & O.Inst     & B.Mod      &            & \\
+                 & Conc.      & Spec.      & Comp.      & L.Inst     & O.Inst     & B.Mod      & checking   & \\
   \hline

### 2021-05-31 8-discussion.tex
sync
synchro

@@
   described in the section~\ref{sec:model} and is illustrated by bold circles in
-  the Figure~\ref{fig:ToolingArchitecture}).
+  the Figure~\ref{fig:ToolingArchitecture}.
 
@@
   the challenge had refined the specification by requiring specific design tasks. 
-  In this context, we would have to transform \emph{Design Task} into a concept 
+  In this context, we would have to transform the \inst{Design Task} instance into a concept 
   whose instances could have been "using agile methodology".
@@
 
-As described in the section~\ref{subsec:tooling}, our solution is fully tooled with
-Openflexo.  We use the Openflexo infrastructure to build models and metamodels
+We use the Openflexo infrastructure to build models and metamodels
 together, and to provide dedicated tooling to edit a Process type, to enact a

### 2021-05-31 9-relatedworks.tex
synchro

@@
 
-A first description of process modeling as a multi-level modeling problem was proposed by \cite{lara2018refactoring} in the context of a catalogue of refactoring for multi-level models (in a simpler form with fewer constraints and requirements w.r.t.
+A first description of process modeling as a multi-level modeling problem was proposed by \parencite{lara2018refactoring} in the context of a catalogue of refactoring for multi-level models (in a simpler form with fewer constraints and requirements w.r.t.
 the challenge version).
-A solution is provided with MetaDepth~\citep{metadepth}, which supports modeling with any number of levels, dual ontological/linguistic typing and deep characterization through potency.
+A solution is provided with MetaDepth~\parencite{metadepth}, which supports modeling with any number of levels, dual ontological/linguistic typing and deep characterization through potency.
 
@@
 
-More similar to us, \cite{deeptelos2019} use DeepTelos \citep{deeptelos2016}, an extension of the Telos language \citep{telos1990} in order to solve the process challenge.
+More similar to us, \cite{deeptelos2019} use DeepTelos \parencite{deeptelos2016}, an extension of the Telos language \parencite{telos1990} in order to solve the process challenge.
 As with Openflexo and \FML, they do not use explicit (numbered) levels nor potency.
-However, unlike Openflexo, DeepTelos integrates a multi-level modeling specific construct similar to the PowerType \citep{atkinson2001essence} pattern they call \emph{most general instances}.
+However, unlike Openflexo, DeepTelos integrates a multi-level modeling specific construct similar to the PowerType \parencite{atkinson2001essence} pattern they call \emph{most general instances}.
 
@@
 
-Similarly, \cite{dmla2019} use the Dynamic Multi-Layer Algebra (DMLA) \citep{dmla2017} in order to solve the process challenge.
+Similarly, \cite{dmla2019} use the Dynamic Multi-Layer Algebra (DMLA) \parencite{dmla2017} in order to solve the process challenge.
 DMLA is a level-blind modeling framework which is fully customizable (\eg different types of instantiation may be implemented).

### 2021-05-31 6-modelpresentation.tex
synchro
replace FML by \FML

@@
 \texttt{Process} represents an enacted \texttt{ProcessType}, as defined in previous subsection (\textbf{P10}).
-FML defines behavioral features called behavior.
+\FML defines behavioral features called behavior.
 \texttt{ProcessType} defines the behavior \texttt{newProcess(String)}, taking the name of the process to enact as argument:
@@
 The capture of \textbf{S3} requirement is performed while defining \inst{Coding} as an instance of concept \texttt{CodingTaskType} and defining a relation \texttt{languages} to \texttt{ProgrammingLanguage} with 1..* cardinality.
-\noteFabien{* dans la figure} The \texttt{newTask(String)} behavior overrides the generic implementation by instantiating a \texttt{CodingTask} concept, specializing \texttt{Task}, as shown in following excerpt of FML code:
+\noteFabien{* dans la figure} The \texttt{newTask(String)} behavior overrides the generic implementation by instantiating a \texttt{CodingTask} concept, specializing \texttt{Task}, as shown in following excerpt of \FML code:
 
@@
 
+\begin{figure*}
+ \centering
+     \includegraphics[width=\textwidth]{Figures/ScreenshotEnactedProcessEditor.png}
+     \caption{Screenshot of the enacted process graphical editor}
+    \label{fig:ScreenshotEnactedProcessEditor}
+\end{figure*}
+
 Our solution is fully implemented within the Openflexo tool. Both use cases have been modeled in the interactive design environment and can be executed by the \FML execution engine.
+
+We took advantage of model federation and the availability of diagramming features though the \textit{Diagramming Technology Adapter} to implement two interactive graphical tools built on top of the conceptual levels detailed in previous sections.
+A first tool offers a graphical edition of a Process type.
+The second one offers enactment feature (the instantiation of a process from its process type), the ability within an enacted process to assign tasks to actors, and a graphical visualization of this process execution.
 
@@
 The management of artifacts whose semantics follows rules defined in section~\ref{sec:ProcessEnactment} (\textbf{P13}) is also implemented.
-
-We took advantage of model federation and the availability of diagramming features though the \textit{Diagramming Technology Adapter} to implement two interactive graphical tools built on top of the conceptual levels detailed in previous sections.
-A first tool offers a graphical edition of a Process type.
-The second one offers enactment feature (the instantiation of a process from its process type), the ability within an enacted process to assign tasks to actors, and a graphical visualization of this process execution.
 
@@
 
-\begin{figure*}
- \centering
-     \includegraphics[width=\textwidth]{Figures/ScreenshotEnactedProcessEditor.png}
-     \caption{Screenshot of the enacted process graphical editor}
-    \label{fig:ScreenshotEnactedProcessEditor}
-\end{figure*}
-
 \todo[inline]{mettre ici le lien vers la page web qui montre la démo}

### 2021-05-31 6-modelpresentation.tex
removing trailing whitespace
antidote

@@
 The roles whose types are concepts are represented by a combination of a name in their containing concept and an arrow from their name to their type.
-The cardinality follow the \UML practice.
+The cardinality follows the \UML practice.
 For example, the role \texttt{parentActorTypes} of the concept \texttt{ActorType} has the type \texttt{ActorType} and the * cardinality.
 
-Our proposition relies on ontologic instantiation as presented in figure \ref{fig:LinguisticAndOntologicInstantiation}, with a common root concept \texttt{ModelingElement}.
+Our proposition relies on ontological instantiation as presented in figure \ref{fig:LinguisticAndOntologicInstantiation}, with a common root concept \texttt{ModelingElement}.
 Two types of ontological instantiation are needed to meet the requirements of the challenge.
@@
 
-The \texttt{Actor} concept defines a boolean attribute called \texttt{isSenior}, while \texttt{TaskType} defines an additional \texttt{isCritical} boolean attribute, indicating that some instance are flagged as critical and must be performed by senior actors.
-To fulfil \textbf{P9} requirement, a supplementary constraint is required for \texttt{TaskType} and is captured through the following invariant expressed in the \FML language:
+The \texttt{Actor} concept defines a boolean attribute called \texttt{isSenior}, while \texttt{TaskType} defines an additional \texttt{isCritical} boolean attribute, indicating that some instances are flagged as critical and must be performed by senior actors.
+To fulfill \textbf{P9} requirement, a supplementary constraint is required for \texttt{TaskType} and is captured through the following invariant expressed in the \FML language:
 
@@
 
-This scheme rely on \FML dynamic binding mechanism to delegate to task types the responsibility of instances creation.
+This scheme relies on \FML dynamic binding mechanism to delegate to task types the responsibility of instances creation.
 An instance of \texttt{Process} references its unique type \texttt{ProcessType} by the specialized \texttt{/type} role.
@@
 
-All concepts inherits from \texttt{ModelingElement}, which defines a \texttt{lastUpdated} attribute with \texttt{Date} type, and thus satisfies \textbf{P19} requirement.
+Each concept inherits from \texttt{ModelingElement}, which defines a \texttt{lastUpdated} attribute with \texttt{Date} type, and thus satisfies \textbf{P19} requirement.
 
@@
 The \textit{Software Development Process} for Acme company is presented in the figure~\ref{fig:AcmeSoftwareDevelopmentProcess}.
-It is a screen-capture of the tool developed for the challenge and described in the next section.
+It is a screen capture of the tool developed for the challenge and described in the next section.
 
@@
 The capture of \textbf{S3} requirement is performed while defining \inst{Coding} as an instance of concept \texttt{CodingTaskType} and defining a relation \texttt{languages} to \texttt{ProgrammingLanguage} with 1..* cardinality.
-\noteFabien{* dans la figure} The \texttt{newTask(String)} behavior overrides the generic implementation by instantiating a \texttt{CodingTask} concept, specializing \texttt{Task}, as shown in following excerpt of \FML code:
+\noteFabien{* dans la figure} The \texttt{newTask(String)} behavior overrides the generic implementation by instantiating a \texttt{CodingTask} concept, specializing \texttt{Task}, as shown in the following excerpt of \FML code:
 
@@
 A critical task must additionally produce artifacts that must be associated with a validation task.
-This is modelled with a supplementary relation \texttt{validatedWith} defined in \texttt{Artifact} and referencing \texttt{Artifact} validating it (\textbf{S9}).
+This is modeled with a supplementary relation \texttt{validatedWith} defined in \texttt{Artifact} and referencing \texttt{Artifact} validating it (\textbf{S9}).
 
@@
 \textbf{S13} requirement has been previously partially fulfilled.
-This must be completed with an association between a artifact produced and the task that validates it.
-This is modelled by the reference to \inst{Test Cases} as a produced artifact type in \inst{Test Case design} and the reference to \inst{Test Cases} as a used artifact in \inst{Test Design review}, as shown in figure~\ref{fig:AcmeSoftwareDevelopmentProcess}.
+This must be completed with an association between an artifact produced and the task that validates it.
+This is modeled by the reference to \inst{Test Cases} as a produced artifact type in \inst{Test Case design} and the reference to \inst{Test Cases} as a used artifact in \inst{Test Design review}, as shown in figure~\ref{fig:AcmeSoftwareDevelopmentProcess}.
 
@@
 
-We took advantage of model federation and the availability of diagramming features though the \textit{Diagramming Technology Adapter} to implement two interactive graphical tools built on top of the conceptual levels detailed in previous sections.
+We took advantage of model federation and the availability of diagraming features though the \textit{Diagramming Technology Adapter} to implement two interactive graphical tools built on top of the conceptual levels detailed in previous sections.
 A first tool offers a graphical edition of a Process type.
-The second one offers enactment feature (the instantiation of a process from its process type), the ability within an enacted process to assign tasks to actors, and a graphical visualization of this process execution.
+The second one offers an enactment feature (the instantiation of a process from its process type), the ability within an enacted process to assign tasks to actors, and a graphical visualization of this process execution.
 
@@
 When executed, this tool manages a graphical view for an instance of Acme metamodel (the \textit{Acme model}) and a specific \texttt{ProcessType} instance.
-This tools allows to represent and edit a \texttt{ProcessType}.
+This tool allows representing and editing a \texttt{ProcessType}.
 The highly reflective nature of \FML and its tooling should be noted.
-When the drag and drop interactor is applied for a new item, the tool allows to choose the concept type to be instantiated.
+When the drag and drop interactor is applied for a new item, the tool allows choosing the concept type to be instantiated.
 For example, when a \texttt{TaskType} is created, the user must choose a sub-concept of \texttt{TaskType} -- in Acme use case it can be \texttt{SETaskTask} or \texttt{CodingTaskType} or default value \texttt{TaskType}).
@@
 Process enactment is operated from a \texttt{ProcessType}, and must be defined using a name identifying the newly instantiated process.
-All tasks are created from their \texttt{TaskType} definition, and required assignments applies (if for example \inst{Coding} \texttt{TaskType} defines COBOL as output programming language, related task is automatically assigned to \inst{Ann Smith}).
+All tasks are created from their \texttt{TaskType} definition, and required assignments apply (if for example \inst{Coding} \texttt{TaskType} defines COBOL as output programming language, the related task is automatically assigned to \inst{Ann Smith}).
 Each task has a status as well as a set of performing actors, a begin and end date, and used and produced artifacts.

### 2021-05-31 6-modelpresentation.tex

@@
 The capture of \textbf{S3} requirement is performed while defining \inst{Coding} as an instance of concept \texttt{CodingTaskType} and defining a relation \texttt{languages} to \texttt{ProgrammingLanguage} with 1..* cardinality.
-\noteFabien{* dans la figure} The \texttt{newTask(String)} behavior overrides the generic implementation by instantiating a \texttt{CodingTask} concept, specializing \texttt{Task}, as shown in the following excerpt of \FML code:
+The \texttt{newTask(String)} behavior overrides the generic implementation by instantiating a \texttt{CodingTask} concept, specializing \texttt{Task}, as shown in the following excerpt of \FML code:
 
@@
 
-The \texttt{CodeArtifact} concept extends \texttt{SEArtifact}, and defines a relation \texttt{language} to \textit{ProgrammingLanguage} with single cardinality (\textbf{S4}).
-\noteFabien{pas la cardinalité de la figure}
+The \texttt{CodeArtifact} concept extends \texttt{SEArtifact}, and defines a relation \texttt{language} to \textit{ProgrammingLanguage} with 1..* cardinality (\textbf{S4}).
+
 \textbf{S5} is more ambiguous as a task type \texttt{CodingTask} defines one or more programming languages but produces code which is expressed in one language only.

### 2021-05-31 7-requirementssatisfaction.tex
removing trailing whitespace
synchro
antidote requirements satisfaction

@@
 
+Table~\ref{tab:RequirementsSatisfactionBaseMetamodel} summarizes the requirement coverage for the base metamodel illustrated by XSure insurance use case while table~\ref{tab:RequirementsSatisfactionAcme} shows requirements satisfaction for Acme software engineering process.
+Syntactic conformance is more precisely classified into six subcategories:
+
+\begin{itemize}
+    \item \emph{Conceptualization} (\textit{Conc.}): a concept carries the semantics exposed by the requirement (for example for \textbf{P1}: \textit{process type} is conceptualized in a concept \texttt{ProcessType}).
+
+    \item \emph{Specialization} (\textit{Spec.}): a concept is specialized in the sense of object-oriented modeling (for example for \textbf{P2}: \texttt{Gateway} defines an abstract behavior \texttt{execute()} and \texttt{Sequencing}, \texttt{AndSplit}, \texttt{AndJoin}, \texttt{OrSplit} and \texttt{OrJoin} are defined as inherited concepts implementing specific business logic).
+
+    \item \emph{Composition} (\textit{Comp.}): some concepts are in relation, or specific attributes were added to some concepts (for example for \textbf{P3}: a \textit{process type} has one \textit{initial task type}).
+
+    \item \emph{Linguistic instantiation} (\textit{L.Inst.}): a requirement is satisfied by the instantiation of one or more concepts (for example for \textbf{S1}: \inst{Requirement analysis} is defined as an instance of \texttt{SETaskType}).
+
+    \item \emph{Ontological instantiation} (\textit{O.Inst.}): fulfillment of a requirement is obtained by the relation between an instance and its type definition instance (for example for \textbf{S3} : \textit{developer} has the dual concept nature in \textit{Acme metamodel} and instance nature in \textit{Acme model}).
+
+    \item \emph{Behavioral modeling} (\textit{B.Mod.}): a requirement is satisfied by the definition of one or more behaviors, which may be combined with constraints checking and associated tooling (for example \textbf{P17}).
+\end{itemize}
+
+%Table~\ref{tab:RequirementsSatisfactionBaseMetamodel}
+%Table~\ref{tab:RequirementsSatisfactionAcme}
+
+%Now we discuss how the solution satisfies the challenge requirements.
+
+%\noteSylvain{
+%Je voulais dire ici qu'on avait construit la solution pas à pas en suivant les requirements les uns après les autres, et qu'on avait à chaque fois veillé à traduire et respecter les requirements.
+Ca me semble très lourd de tout rejustifier ici.
+
+%On peut cependant expliciter les différentes techniques employées:
+%\begin{itemize}
+%    \item Requirements satisfaits par conformance syntaxique: obligation de respecter le modèle structurel
+%    \item Requirements satisfaits par ajout de contraintes: on peut instantier quelque chose de syntaxiquement correct, mais qui viole certaines contraintes (sur l'outillage les croix rouges)
+%    \item Requirements satisfaits par l'outillage: ajout de comportements de manipulation du modèle qui "forcent" à respecter les contraintes d'intégrité resultant des requirements.
+%\end{itemize}
+
+%Peut-être peut-on faire un tableau qui rescence comment on a satisfait chaque contrainte ? Est ce que c'est %intéressant ?
+%}
+%\todo[inline]{OK pour le tableau synthétique +un petit peu de texte + une mini intro qui référence la section précédente (vu que les req ont été couverts dans la section précédente ; pas de fusion des deux sections}
+
+%CT : Conceptualization
+%SP : Specialization
+%CP : Composition
+%LI : Linguistic instantiation
+%OI : Ontologic instantiation
+%BM : behavioral modeling
+
+\renewcommand{\arraystretch}{1.1}
 \begin{table*}
@@
 
-Table~\ref{tab:RequirementsSatisfactionBaseMetamodel} summarizes the requirement coverage for the base metamodel illustrated by XSure insurance use case while table~\ref{tab:RequirementsSatisfactionAcme} shows requirements satisfaction for Acme software engineering process.
-Syntactic conformance is more precisely classified into six subcategories:
-
-\begin{itemize}
-    \item \emph{Conceptualization} (\textit{Conc.}): a concept carries the semantics exposed by the requirement (for example for \textbf{P1}: \textit{process type} is conceptualized in a concept \texttt{ProcessType}).
-
-    \item \emph{Specialization} (\textit{Spec.}): a concept is specialized in the sense of object-oriented modeling (for example for \textbf{P2}: \texttt{Gateway} defines an abstract behavior \texttt{execute()} and \texttt{Sequencing}, \texttt{AndSplit}, \texttt{AndJoin}, \texttt{OrSplit} and \texttt{OrJoin} are defined as inherited concepts implementing specific business logic).
-
-    \item \emph{Composition} (\textit{Comp.}): some concepts are in relation, or specific attributes were added to some concepts (for example for \textbf{P3}: a \textit{process type} has one \textit{initial task type}).
-
-    \item \emph{Linguistic instantiation} (\textit{L.Inst.}): requirement is satisfied by the instantiation of one or more concepts (for example for \textbf{S1}: \inst{Requirement analysis} is defined as an instance of \texttt{SETaskType}).
-
-    \item \emph{Ontologic instantiation} (\textit{O.Inst.}): fulfillment of requirement is obtained by the relation between an instance and its type definition instance (for example for \textbf{S3} : \textit{developer} has the dual concept nature in \textit{Acme metamodel} and instance nature in \textit{Acme model}).
-
-    \item \emph{Behavioral modeling} (\textit{B.Mod.}): requirement is satisfied by the definition of one or more behaviors, which may be combined with constraints checking and associated tooling (for example \textbf{P17}).
-\end{itemize}
-
-%Table~\ref{tab:RequirementsSatisfactionBaseMetamodel} 
-%Table~\ref{tab:RequirementsSatisfactionAcme}
-
-%Now we discuss how the solution satisfies the challenge requirements.
-
-%\noteSylvain{
-%Je voulais dire ici qu'on avait construit la solution pas à pas en suivant les requirements les uns après les autres, et qu'on avait à chaque fois veillé à traduire et respecter les requirements.
-Ca me semble très lourd de tout rejustifier ici.
-
-%On peut cependant expliciter les différentes techniques employées:
-%\begin{itemize}
-%    \item Requirements satisfaits par conformance syntaxique: obligation de respecter le modèle structurel
-%    \item Requirements satisfaits par ajout de contraintes: on peut instantier quelque chose de syntaxiquement correct, mais qui viole certaines contraintes (sur l'outillage les croix rouges)
-%    \item Requirements satisfaits par l'outillage: ajout de comportements de manipulation du modèle qui "forcent" à respecter les contraintes d'intégrité resultant des requirements.
-%\end{itemize}
-
-%Peut-être peut-on faire un tableau qui rescence comment on a satisfait chaque contrainte ? Est ce que c'est %intéressant ?
-%}
-%\todo[inline]{OK pour le tableau synthétique +un petit peu de texte + une mini intro qui référence la section précédente (vu que les req ont été couverts dans la section précédente ; pas de fusion des deux sections}
-
-%CT : Conceptualization
-%SP : Specialization
-%CP : Composition
-%LI : Linguistic instantiation
-%OI : Ontologic instantiation
-%BM : behavioral modeling
-
 \begin{table*}

### 2021-08-18 responseLetter.tex
Adding a template for the response letter

% Title:    A LaTeX Template For Responses To a Referees' Reports
% Author:   Petr Zemek <s3rvac@gmail.com>
% Homepage: https://blog.petrzemek.net/2016/07/17/latex-template-for-responses-to-referees-reports/
% License:  CC BY 4.0 (https://creativecommons.org/licenses/by/4.0/)
\documentclass[10pt]{article}

% Allow Unicode input (alternatively, you can use XeLaTeX or LuaLaTeX)
\usepackage[utf8]{inputenc}
\usepackage{hyperref}
\usepackage{xcolor}
\usepackage{todonotes}

\usepackage{microtype,xparse,tcolorbox}
\newenvironment{reviewer-comment }{}{}
\tcbuselibrary{skins}
\tcolorboxenvironment{reviewer-comment }{empty,
  left = 1em, top = 1ex, bottom = 1ex,
  borderline west = {2pt} {0pt} {black!20},
}
\ExplSyntaxOn
\NewDocumentEnvironment {response} { +m O{black!20} } {
  \IfValueT {#1} {
    \begin{reviewer-comment~ }
      \setlength\parindent{2em}
      \noindent
      \ttfamily #1
    \end{reviewer-comment~ }
  }
  \par\noindent\ignorespaces\color{blue}
} { \bigskip\par }

\NewDocumentCommand \Reviewer { m } {
  \section*{Comments~by~Reviewer~#1}
}

\NewDocumentCommand \Editor { m } {
  \section*{Comments~by~Editor}
}

\NewDocumentCommand \ToolComments { m } {
  \section*{Tool~Comments~by~Reviewer~#1}
}

\ExplSyntaxOff
\AtBeginDocument{\maketitle\thispagestyle{empty}\noindent}

% You can get probably get rid of these definitions:
\newcommand\meta[1]{$\langle\hbox{#1}\rangle$}
\newcommand\PaperTitle[1]{``\textit{#1}''}

\title{Response letter to EMISAJ Manuscript entitled "Multi-level modeling with Openflexo/FML- A contribution to the MULTI process challenge"}
%\author{Author1 \and Author2 \and Author3}
\date{\today}

\begin{document}

Dear Editor,

\bigskip
Thanks for your message of August 16th, 2021, in which you suggested preparing a revision of our manuscript  \textbf{Multi-level modeling with Openflexo/FML- A contribution to the MULTI process challenge}.

\bigskip
We are grateful to the reviewers for their interest, effort, and suggestions. We detail in the following pages how we have addressed the reviewers' concerns; we are confident that the paper is now improved, and hope that you will agree. We have organized the responses to the comments on a per-reviewer basis, focusing, as suggested, on:

\begin{itemize}
\item Improving the presentation.
\item Provide a more explicit description of how the multi-level modelling emulation is achieved.
\item Discussing the general advantages and limitations of the multi-level emulation, including a clarification of what the implications would be of supporting more "levels" than the challenge required.
\end{itemize}

Moreover, the paper has been thoroughly revised to fix minor issues and to give the additional clarifications required by the reviewers. Please let us know if you need any further information or clarification.

\bigskip
Sincerely yours,

\bigskip
Sylvain, Guillaume, Caine, Joel, Jean-Christophe, Salvador, Fabien and Antoine

\pagebreak

\Editor{}

\begin{response}{the reviewers have made suggestions for improving the presentation but also raised a number of issues requiring clarification. An overarching theme is the request for a more explicit description of how the multi-level modelling emulation is achieved and what its general advantages and limitations are, including a clarification of what the implications would be of supporting more "levels" than the challenge required.} TODO.
\end{response}

\begin{response}{We very much welcome a "conventional" response to the challenge but would like to see the respective reviewer questions addressed in a revised version which is explicit as possible about the advantages and the disadvantages of the emulation. You may find the references listed below useful to keep explanations concise by referencing related work (note that reviewer 3’s “Type-Instance pattern” is also known as the “Type Object” pattern). You should also respond to other concerns reviewers raise, either by making changes to the submission or in the form of an authors' response.} TODO. We may have to create a new section.
\end{response}

\begin{response}{Finally, to achieve consistency with other submissions, could you please change the subtitle to "A Contribution to the Multi-Level Process Challenge"?} Of course we can.
\end{response}

\pagebreak

\Reviewer{\#1}
\begin{response}{Here goes the question} And here goes the answer.
\end{response}

\pagebreak

\Reviewer{\#2}

\begin{response}{Here goes the question} And here goes the answer.
\end{response}

\pagebreak

\Reviewer{\#3}

\begin{response}{Here goes the question} And here goes the answer.
\end{response}

\end{document}

### 2021-08-18 responseLetter.tex
Adding reviewer 1 comments

@@
 \Reviewer{\#1}
-\begin{response}{Here goes the question} And here goes the answer.
+\begin{response}{In section 2 Technology the authors describe the foundations of the Openflexo approach.
+Here it would greatly help the reader to see an example how the approach works for traditional modeling languages.
+For example, how would you realize a small subset of BPMN with the approach?
+This could briefly be illustrated so as to give the reader an immediate understanding.
+Especially, the 'metamodel'-agnostic approach seems very interesting but I would like to see how that is used for traditional modeling approaches.}
+TODO.
+\end{response}
+
+\begin{response}{Also, at the stage of section 2, the characterization of the FML language should be made clearer.
+The metamodel in Figure 2 seems mainly conceptual - but at the same time you refer to it as a basis for implementation.
+Could that be clarified?
+eg you have the class "Behavior" - which is a concept and I did not understand right away what it is composed of.
+On the other hand you have for example the class "ModelSlot" which seems like a very technical realization.
+Also, you note that roles have types but I could not find any types in the metamodel (?).
+Thus, I propose to include a more detailed version of your metamodel at this stage.
+Please note that EMISAJ does not have a page limitation, so no worries if that should require more space.}
+TODO.
+\end{response}
+
+\begin{response}{In Figure 5, the base metamodel for representing processes is shown - however it is not linked to the concepts you introduced before.
+Would it be possible to maybe graphically highlight how it relates to the metamodel in Figure 2?
+Besides, as multi-level modeling is much different to traditional modeling, you may want to consider using a different notation rather than UML class diagrams - as these are commonly known for traditional modeling approaches.
+When reading the paper, I actually marked the references to types in yellow in my copy - ie from ProcessType to TaskType and from ActorType to ActorType (on the attribute levels) - this helped me to see more clearly where you have references from attributes to types - but that's just my way of doing it.}
+TODO.
+\end{response}
+
+\begin{response}{Regarding Figure 6 and Figure 7: In Figure 7 you show the process incl.
+sequence connectors - however in Figure 6 these seem to be missing - maybe to reduce complexity?
+I would suggest to probably focus on a subset of the process from Figure 7 and also show how the connectors are realized.
+To me this would increase the comprehensibility / consistency.
+Ie I could then infer exactly, how elements from the concrete model have been realized in the background.}
+TODO.
+\end{response}
+
+\begin{response}{in Figure 9 you show the model editor. Could you please also add a screenshot of the metamodel editor?} 
+TODO.
+\end{response}
+
+\begin{response}{There are some minor typos throughout the manuscript (eg also regarding the reference Jeusfeld2019; choose -> chose; axis -> axes; aN FML virtual model...; etc.)} 
+TODO.
 \end{response}
@@
 
-\begin{response}{Here goes the question} And here goes the answer.
+\begin{response}{Here goes the question} 
+TODO.
 \end{response}

### 2021-08-18 responseLetter.tex
Reviewers comments added.

@@
 
-\begin{response}{Here goes the question} 
+\begin{response}{The basic idea of model federation is interesting, its primary focus is not on creating multi-level models.
+However, as the approach is quite flexible, multi-level behavior can be emulated by model federation.
+This fact is possibly the most serious advantage and disadvantage of the paper at the same time.
+Advantage, since it brings us a completely new, different way of thinking, and disadvantage since it does not really help in categorizing/unifying multi-level methods and – for me – it also feels a little bit off-topic in this special issue.
+At this point, it should be mentioned that the authors managed to cover/solve all requirements of the challenge, thus the paper, as a solution to the challenge fits well.}
+TODO.\todo{Answering this is going to be tricky!}
+\end{response}
+
+\begin{response}{The paper gives us a short introduction (somewhat overlapping with the abstract, even at the level of sentences) and then the technology is detailed.
+The section is a bit hard to understand at first, the difference and the relationship between federated and virtual models could be explained in more detail especially considered that the challenge solution uses virtual models only.
+It is not clear, how federation works in this case.}
 TODO.
+\end{response}
+
+\begin{response}{The third section contains an overview of the solution.
+Here, virtual models are instantiated to virtual model instances conforming to the virtual models acting as the metamodel for the virtual model instances.
+The question naturally arises: how does it work, if we have more abstraction levels?
+In this section, the authors also discuss linguistic and ontologic instantiation.
+The explanation refers to Section 4 resulting in a hardly understandable analysis of different instantiation methods.
+Figure 4.
+does not help in clarifying the concepts either (e.g.
+using the name InstanceInstance seems to be a very bad decision).
+From the point of the muli-level behavior of view, the interpretation of the instantiation relationship is the most crucial one that is why this section must be clarified.
+Where are the rules of ontological instantiation defined, how are they enforced?
+The text mentions the (linguistic) instantiation mechanism of the FML Language… how does it work exactly?}
+TODO.
+\end{response}
+
+\begin{response}{Section 4 elaborates the solution in detail.
+Figure 5 presents an overview of the model elements handling P1 to P19.
+The diagram uses a UML-like annotation, which helps in making it understandable, but it also has a strange decision: roles (attributes) having a modeled concept as their type are visualized by links, but instead of using rolenames on the links, the concepts (classes) contain them.
+I strongly suggest following the UML guidelines here and use rolenames (similarly to multiplicity).}
+TODO.
+\end{response}
+
+\begin{response}{Overall, the solutions for the requirements are correctly presented in this section, but there are some spots not clear enough.
+E.
+g.
+regarding ontological instantiation, the solution uses two kinds: single and multiple.
+More details would be welcome to see, what their exact semantics are and where do we enforce these semantic rules (is it enforced by the engine?
+Is the semantic hard-coded?).
+What are the contexts of the constraints (to which concept are they attached to)?}
+TODO.
+\end{response}
+
+\begin{response}{In Section 4.1.2 a behavior called newProcess is defined.
+Is it called automatically (by the framework) when creating a new process, or should the users call it?
+Is this a convention, or are such methods configured somehow to replace the usual constructors?
+According to the text, “same pattern applies” in several cases, but these patterns could be also included, because it is not completely clear to what part of the behavior “same” refers.
+In same cases, the diagram uses “/type”, while in the code it is “type” – why?
+The specialization of Actors is encoded by FML, wouldn’t be easier to use the built-in specialization instead?}
+TODO.
+\end{response}
+
+\begin{response}{In Section 4.2 it is mentioned, where the solution uses linguistic instantiation, but it is not clear by which component it is handled and what rules are followed when instantiating the concepts.
+In Section 4.3, tooling is elaborated.
+I suggest removing redundant information about the two editors and add more details on how the concrete syntax (e.g.
+the texts in the box of Fig.
+9.) is defined.}
+TODO.
+\end{response}
+
+\begin{response}{In Section 5, the meaning of the category Conceptualization is hard to understand, does this mean that a concept is created that covers the requirement completely?
+In the case of ontological instantiation, the term “dual concept nature” and “instance nature” is mentioned… what do these terms mean?}
+TODO.
+\end{response}
+
+\begin{response}{Section 7 contains a rather short elaboration on related approaches.
+Although I agree that it is enough to focus on previous solutions here, much more detail and a comparison would be required.}
+TODO.
+\end{response}
+
+\begin{response}{As Section 6 shows, the authors know exactly, why their solution may have difficulties as a multi-level modeling approach despite solving the given challenge.
+The solution is basically a smart, two-level modeling solution of the MULTI challenge.
+Multi-level behavior is emulated by constraints and additional mechanisms particularly written to this challenge.
+The result is a working workbench fulfilling the requirements but having limited reusability and extensibility.
+For example, it would be hard to increase the number of abstraction levels as class-like and object-like concepts are handled differently.
+Another problem may arise if we would use the framework as a prototyping tool and change the metamodel and the models at the same time.}
+TODO.
+\end{response}
+
+\begin{response}{Minor issues/questions:} 
+\begin{itemize}
+    \item Are there any tools supporting model federation besides Openflexo?
+    \item Section 2. “In contrast to approaches that compose metamodels into a single large metamodel \item grouping all needed entities,…” - I do not see, why this is relevant here
+    \item Figure 5 – concept Sequencing – the roles “in” and “out” are not represented by lines… why?
+    \item What does the term “sub-concept” mean exactly? Please explain it before the first use
+    \item ProgrammingLanguage concept is an enumeration… how is it modelled?
+    \item “The Acme metamodel also defines a specialized task type Coding” – isn’t it CodingTask?
+    \item “The value of the property producedArtifactTypes” – is it property, or role?
+
+\end{itemize}
+TODO. \todo{The idea is to group all the answers to the minor questions here.}
 \end{response}
@@
 
-\begin{response}{Here goes the question} And here goes the answer.
+\begin{response}{(1) The Type-Instance pattern is key to the authors approach to the challenge.
+However, the benefits of this pattern are not explored in the context of MLM.
+For example, the authors do not really explain approaches such as Clabjects or potency levels as they compare to the Type-Instance pattern (or the challenge).
+Why it the pattern better?
+Is it just a different approach?
+What can it do that other approaches cannot?
+Perhaps the benefit is that the Type-Instance pattern can be implemented in most modelling and implementation technologies whereas other approaches need special facilities?
+Finally, in terms of the pattern, why should anyone care - why would should they use the pattern?}
+TODO.
+\end{response}
+
+\begin{response}{(2) In section 1 the authors list 2 key reasons why a strict approach to modelling is sub-optimal.
+However, they do not establish the case that their approach uses these 2 features to any concrete benefit.
+It would be appropriate (perhaps at the end of the article) to revisit these features and to clearly demonstrate that their approach supports them and that they have concrete utility in some way in the context of the challenge.}
+TODO.
+\end{response}
+
+\begin{response}{(3) One of the key benefits of the Type-Instance pattern is that it provides a 'hook' on which semantics can be defined using constraints.
+There is a reified relationships between a type and its instance and therefore conditions can be defined that establish both static and dynamic semantics.
+I am confused regaring why this is not a key feature of the semantics for the challenge as it is implemented by the authors.
+Constraints are provided very sparingly compared to implementation of behaviour in FML.
+Unfortunately, since the dynamic semantics of FML is not provided (and is bespoke) this is less clear than it might otherwise be for the reader.
+Perhaps it would be better to use constraints in the context of the Type-Instance pattern to specify semantics?
+It would be more abstract than the current 'method' style semantics.}
+TODO.
+\end{response}
+
+\begin{response}{(4) Section 6 lists a number of benefits of the approach.
+Unfortunately, the presentation in the section is very general and none of the claims are justified in terms of concrete descriptions and/or examples.
+This section seems to be very important and could be of great interest to the reader.
+Perhaps it would be appropriate to reduce some of the rather implementation heavy descriptions earlier in the paper and increase the analysis in this section?
+In addition section 4.3 notes that MLM is key to the ability of Openflexo to create tooling.
+Again, this is very interesting, but no concrete details are provided as to how Openflexo uses the level agnotic features of the models to create tools.
+Can the essence of this be captured for Openflexo and then described in terms of how it facilitiates the tooling arising from the challenge?}
+TODO.
+\end{response}
+
+\begin{response}{(5) The paper makes an issue of the model federation approach supported by Openflexo, including the use of model slots.
+However, after introducing these concepts they do not seem to be important to the rest of the article.
+Perhaps this is an implementation concern that could be elided as its inclusion does not seem key to the contributions of the article.
+}
+TODO.
+\end{response}
+
+\begin{response}{Some minor issues:
+
+(a) In fig 5: why are the in/out gateway features shown as properties and not as associations?
+
+(b) On p. 7 'specialization' -> 'specializations'.} 
+TODO.
 \end{response}

### 2021-08-26 1-main.tex
remove all capital letters from files, as specified by the style
fix keywords command
sync Overleaf -> Github
synchro
make subtitle compliant, replace \mpc by \mlpc, update response

--- 2021-05-28 1-Main.tex
@@
 
-\newcommand{\mpc}{MULTI process challenge\xspace}
+\newcommand{\mpc}{MULTI process challenge\xspace}%keep it for the moment
+\newcommand{\mlpc}{Multi-Level Process Challenge\xspace}
 
+\usepackage{multirow}
 \usepackage{listings}
@@
 \usepackage{tikz-uml}
+%\usepackage{float}
+
+\newcommand{\inst}[1]{\textsf{'#1'}}
 
@@
 \definecolor{backcolour}{rgb}{0.95,0.95,0.92}
+\definecolor{tableShade}{gray}{0.9}
 
@@
     % \title{\mpc: the FML solution}
-    \title{Multi-level modeling with Openflexo/\FML - A contribution to the \mpc}
+    \title{Multi-level modeling with Openflexo/\FML}
+    \subtitle{A Contribution to the \mlpc}
 
@@
     \author*{Jean-Christophe Bach}{jc.bach@imt-atlantique.fr}
-    \address{IMT Atlantique, Lab-STICC, UMR 6285, 29238 Brest, France}
+    \address{IMT Atlantique, Lab-STICC, UMR 6285, Brest, France}
     % \ead{firstname.lastname@imt-atlantique.fr}
-
-    \author{Salvador Mart\'inez}
-    \address[b]{}
 
@@
 
-    \abstract{\input{2-Abstract}}
-    \keywords{Multilevel Modeling Challenge, Federation Modeling Language, Openflexo}
+    \author{Salvador Mart\'inez}
+    \address[b]{}
+
+    \abstract{\input{2-abstract}}
+    \keywords{Multilevel Modeling Challenge \and Federation Modeling Language
+    \and Openflexo}
     \bibliography{biblio.bib}
@@
 \label{sec:introduction}
-\input{3-Introduction.tex}
+\input{3-introduction.tex}
 
@@
 \label{sec:technology}
-\input{4-Technology.tex}
+\input{4-technology.tex}
 
@@
 \label{sec:analysis}
-\input{5-Analysis.tex}
+\input{5-analysis.tex}
 
@@
 \label{sec:model}
-\input{6-ModelPresentation.tex}
+\input{6-modelpresentation.tex}
 
@@
 \label{sec:requirements}
-\input{7-RequirementsSatisfaction.tex}
+\input{7-requirementssatisfaction.tex}
 
@@
 \label{sec:discussion}
-\input{8-Discussion.tex}
+\input{8-discussion.tex}
 
@@
 \label{sec:relatedwork}
-\input{9-Relatedworks.tex}
+\input{9-relatedworks.tex}
 
@@
 \label{sec:conclusions}
-\input{10-Conclusions.tex}
+\input{10-conclusions.tex}
 

### 2021-08-26 6-modelpresentation.tex
sync overleaf->github
make subtitle compliant, replace \mpc by \mlpc, update response

@@
 
-The presentation of our solution to the \mpc follows a systematic methodology where modeling choices are introduced step by step, as the requirements are stated.
+The presentation of our solution to the \mlpc follows a systematic methodology where modeling choices are introduced step by step, as the requirements are stated.
 The satisfaction of requirements is explained throughout this presentation.
@@
 
-\todo[inline]{mettre ici le lien vers la page web qui montre la démo}
+A demonstrative video for the proposed tooling is available on our website\footnote{\url{https://research.openflexo.org/MLMChallenge.html}}, as well as download and installation instructions.

### 2021-08-26 6-modelpresentation.tex
typos

@@
 We now present the process enactment part of base metamodel, located right of figure~\ref{fig:BaseMetamodel}.
-All concepts defined in this subsection are either specialization of the \texttt{Instance} concept (if they have exactly one type) or the \texttt{MultiInstance} concept (when they have several types).
+All concepts defined in this subsection are either specializations of the \texttt{Instance} concept (if they have exactly one type) or the \texttt{MultiInstance} concept (when they have several types).
 
@@
 The Base metamodel described previously is too generic to capture all domain-specific aspects of this use case.
-We choose to complete the architectural hierarchy with a specific virtual model, specific to the Acme software development process metamodel, shown in figure~\ref{fig:AcmeFullArchitecture}.
+We chose to complete the architectural hierarchy with a specific virtual model, specific to the Acme software development process metamodel, shown in figure~\ref{fig:AcmeFullArchitecture}.
 The \textit{Acme} metamodel specializes the Base metamodel, and \textit{Acme model} is an instance of the \textit{Acme} metamodel.

### 2021-09-20 7-requirementssatisfaction.tex

@@
 
-    \item \emph{Ontological instantiation} (\textit{O.Inst.}): fulfillment of a requirement is obtained by the relation between an instance and its type definition instance (for example for \textbf{S3} : \textit{developer} has the dual concept nature in \textit{Acme metamodel} and instance nature in \textit{Acme model}).
+    \item \emph{Ontological instantiation} (\textit{O.Inst.}): fulfillment of a requirement is obtained by the relation between an instance and its type definition instance (for example \todo[inline]{the notion of \textit{developer} of \textbf{S3} is implemented by both a concept in \textit{Acme metamodel} and an instance in \textit{Acme model}).}
 

### 2021-09-20 9-relatedworks.tex

@@
 
-More similar to us, \cite{deeptelos2019} use DeepTelos \parencite{deeptelos2016}, an extension of the Telos language \parencite{telos1990} in order to solve the process challenge.
+More similar to us, in the submission \parencite{deeptelos2019}, the author uses DeepTelos \parencite{deeptelos2016}, an extension of the Telos language \parencite{telos1990} in order to solve the process challenge.
 As with Openflexo and \FML, they do not use explicit (numbered) levels nor potency.

### 2021-09-20 1-main.tex
Updates from Overleaf

@@
     public,return,short,static,super,switch,this,throw,%
-    throws,transient,true,try,void,volatile,while,String},%
+    throws,transient,true,try,use,void,values,volatile,while,with,String},%
   morekeywords={

### 2021-09-20 4-technology.tex
make subtitle compliant, replace \mpc by \mlpc, update response
Updates from Overleaf

@@
 
-To meet the \mpc, we have decided to use the general
+\begin{figure}[t]
+    \centering
+    \includegraphics[width=\columnwidth]{Figures/federation.pdf}
+    \caption{The model federation approach}
+    \label{fig:mf}
+\end{figure}
+
+To meet the \mlpc, we have decided to use the general
 approach of model federation~\parencite{Golra2016-federation}. \emph{Model
@@
 
-\begin{figure}[t]
+\begin{figure*}
     \centering
-    \includegraphics[width=\columnwidth]{Figures/federation.pdf}
-    \caption{The model federation approach}
-    \label{fig:mf}
-\end{figure}
+    \includegraphics[width=1.0 \textwidth]{Figures/FMLMetaModel.pdf}
+    \caption{Internal representation of \FML language}
+    \label{fig:mm}
+\end{figure*}
+%\noteAntoine{Dans la figure ne manque-t-il pas les containements ? (VM vers Concepts)}
+%\noteSylvain{C'est le cas dans cette nouvelle figure}
 
@@
 
-\begin{figure}[t]
-    \centering
-    \includegraphics[width=\columnwidth]{Figures/FMLCoreModel.pdf}
-    \caption{The \FML core metamodel}
-    \label{fig:mm}
-\end{figure}
-%\noteAntoine{Dans la figure ne manque-t-il pas les containements ? (VM vers Concepts)}
-
+\noteSylvain{Parler plutot ici de représentation interne du language FML (plutot que metamodel qui embrouille le lecteur avec ce qui suit ?}
 \FML simplified metamodel is provided in Figure~\ref{fig:mm}. It is designed to define models as virtual models. A virtual model
@@
 
+\noteSylvain{Tentative de représenter ici un sous-ensemble de BPMN en montrant la capacité du language FML à faire de la modélisation conceptuelle d'une part, mais aussi à fédérer des nouveaux concepts avec un modèle et un métamodèle existant.}
+
+\noteSylvain{On explique dans la suite que dans le challenge, on ne va pas exploiter l'aspect fédération de modèle (sauf pour la partie diagramme - c'est expliqué plus haut également - ).}
+
+Figure \ref{fig:BBPMNSubsetExample} shows...
+
+\begin{figure}[t]
+    \centering
+    \includegraphics[width=\columnwidth]{Figures/BPMNSubsetExample.pdf}
+    \caption{An example of BPMN subset built on top of a BPMN/EMF model}
+    \label{fig:BPMNSubsetExample}
+\end{figure}
+
+\begin{lstlisting}
+use org.openflexo.technologyadapter.emf.EMFModelSlot as EMF;
+import ["http://www.omg.org/spec/BPMN/20100524/MODEL-XMI"] as BPMN_METAMODEL;
+
+// A simplified BPMN metamodel exposing
+// both concepts BPMNProcess and BPMNTask
+model SimplifiedBPMN {
+
+  // Connexion to a .bpmn file
+  EMFModel bpmnModel 
+    with EMF::EMFModelSlot(
+        metaModel = BPMN_METAMODEL);
+
+  // Abstract concept BPMNElement defining
+  // a name and a link to an eventual BPMNElement
+  // in .bpmn file
+  abstract concept BPMNElement {
+  	EMFObjectIndividual bpmnElement 
+  	    with EMF::EMFObjectIndividualRole (
+			source = bpmnModel
+		);
+    String name values bpmnElement.name;
+  }
+
+  concept BPMNProcess extends BPMNElement {
+
+    // Describe here properties and
+    // behaviours for BPMNProcess concept
+    ...
+
+    concept BPMNTask extends BPMNElement {
+
+        // Describe here properties and
+        // behaviours for BPMNTask concept
+        ...
+
+    }
+
+    // Other core concepts
+  }
+}
+\end{lstlisting}
+
+\noteSylvain{Faut-il montrer des exemples de comportement aussi ?}
+
 When the \FML execution engine runs a federation, it creates virtual

### 2021-09-22 responseLetter.tex
typos
Updates from Overleaf
small change to the conclusions

@@
 \usepackage{microtype,xparse,tcolorbox}
-\newenvironment{reviewer-comment }{}{}
+\newenvironment{reviewer-commentX}{}{}
 \tcbuselibrary{skins}
@@
   \IfValueT {#1} {
-    \begin{reviewer-comment~ }
+    \begin{reviewer-commentX}
       \setlength\parindent{2em}
@@
       \ttfamily #1
-    \end{reviewer-comment~ }
+    \end{reviewer-commentX}
   }
@@
 \begin{response}{Finally, to achieve consistency with other submissions, could you please change the subtitle to "A Contribution to the Multi-Level Process Challenge"?} Of course we can.
+  DONE.
 \end{response}
@@
 TODO.
+choose $\rightarrow$ chose OK ; axis $\rightarrow$  axes OK ; je ne comprends pas pour Jeusfeld2019
+(deeptelos2019) ni pour aN FML (je dirais ``a FML'', mais je puis me tromper)
 \end{response}
@@
 It would be appropriate (perhaps at the end of the article) to revisit these features and to clearly demonstrate that their approach supports them and that they have concrete utility in some way in the context of the challenge.}
-TODO.
+
+We have added a paragraph in the Conclusions section indicating how Openflexo/FML permitted us to overcome the mentioned limitations of the strict modeling paradigm and how this was used in order to build our solution to the Multi-Level Process challenge.
+Notably, we mention how the type/object pattern allowed us to use model element instances as types in order to achieve ontological instantiation.
 \end{response}

### 2021-09-22 10-conclusions.tex
small change to the conclusions

@@
 
-We fulfill all the challenge requirements and we propose a tooling that demonstrates the usability of our solution.
-We developed an ad hoc solution, including models and metamodels, thanks to the flexibility provided by the metametamodeling infrastructure offered by Openflexo.
+We fulfill all the challenge requirements and we propose a tooling that
+demonstrates the usability of our solution. We developed an ad hoc solution,
+including models and metamodels, thanks to the flexibility provided by the
+metametamodeling infrastructure offered by Openflexo.
+\noteSM{This infrastructure helped us to overcome the limitations of the strict modeling paradigm mentioned in Section \ref{sec:introduction}, this is, the lack of support for different forms of classifications and for the duality type/object.
+In that sense, our solution seamlessly integrates the type-object pattern which allows us to dynamically create different \emph{instances} of model elements (e.g., \emph{TaskType}, among others) and use them for typing other model elements.
+This, together with the FML core support for (level agnostic) classical linguistic instantiation permitted us to merge linguistic and ontological instantiation in order to provide a satisfactory solution to the Multi-Level Process challenge.}
 
-As a future work we envision to explore a number of alternative solutions: first, we intend to propose another solution based on a more level-agnostic approach such as Clabjects.
-Instead of two levels defined by Type and Instance concepts, we could have a single concept mixing a type part and an instance part.
-This approach would probably ease adaptation if the problem specifications evolve.
-Secondly, another possible approach is the use of the free modeling tool proposed by Openflexo.
-The solution would be developed from examples (Acme and XSure processes for instance) from which models would be identified.
-Finally, the realization of this challenge highlighted the interest of integrating specific behaviors for the management of multi-level concepts in the Openflexo infrastructure.
-%We intend to do so.
+As a future work we envision to explore a number of alternative solutions:
+first, we intend to propose another solution based on a more level-agnostic
+approach such as Clabjects. Instead of two levels defined by Type and Instance
+concepts, we could have a single concept mixing a type part and an instance
+part. This approach would probably ease adaptation if the problem
+specifications evolve. Secondly, another possible approach is the use of the
+free modeling tool proposed by Openflexo. The solution would be developed from
+examples (Acme and XSure processes for instance) from which models would be
+identified. Finally, the realization of this challenge highlighted the interest
+of integrating specific behaviors for the management of multi-level concepts in
+the Openflexo infrastructure. %We intend to do so.

### 2021-09-22 3-introduction.tex
removing trailing whitespace
small change to the conclusions

@@
 %\noteFabien{Le paragraphe qui précède me semble redire des choses de la section 2, comment fait-on ?}
+
+Our modeling approach is based on two levels.
+The meta level (\FML) provides the (linguistic) concepts that are used to define the model level (ontological).
+The ontological level is used to define the models and meta-models that are needed.
+\FML allows to define models as first class entities (which we call VirtualModel) and to link them together (by extension).
+In the \mpc case study, we exploit this possibility to:
+\begin{itemize}
+    \item Extend and adapt existing models, keeping the previous ones, to meet changing modeling needs,
+    \item Align this hierarchy of models with the process of their evolution.
+\end{itemize}
+This methodological choice allows us to adapt the models and keep track of their evolution each time the needs evolve.
+
+Contrary to a classical multi-level approach, in which the notion of multi-level is integrated in the meta-model and must therefore be respected, our approach allows to build multi-levels on demand, and with the appropriate form.
+As a limitation, without multilevel-specific operators/concepts, the "custom" construction must be done explicitly during the model development and evolution process.
+
+Our answer to the challenge is therefore not an answer using a classical multilevel approach that would allow an efficiency comparison with another classical multilevel answer, but an approach that proposes another organization of multilevels through the translation of needs into levels over time and the possibility to build an evolving hierarchy of adapted models.
+This is made possible by (1) the reification of the notion of model and the possibility to interconnect them (specialization) and (2) the freedom of meta-modeling (low model conformity, flexible meta-modeling).
 

### 2021-09-22 5-analysis.tex
removing trailing whitespace
typos
small change to the conclusions

@@
 
-Our analysis of the use case leads to identify two main modeling axis (as presented in the figure \ref{fig:LinguisticAndOntologicInstantiation}):
+Our analysis of the use case leads to identify two main modeling axes (as presented in the figure \ref{fig:LinguisticAndOntologicInstantiation}):
 \begin{itemize}
@@
     \includegraphics[width=1.0 \columnwidth]{Figures/Instantiation.pdf}
-    \caption{Linguistic and ontologic instantiation}
+    \caption{Problem  Domain Linguistic and ontologic instantiation}
     \label{fig:LinguisticAndOntologicInstantiation}

### 2021-09-22 9-relatedworks.tex
adding comments to the related work section...

@@
 In this sense, we limit the following discussion to the presentation and comparison of previous solutions to the MULTI Process Challenge.
+
+\noteSM{so... to please reviewer 2 we should improve this section, notably better compare approaches. I am thinking about a table...}
+
+\noteSM{Our solution relies on the use of the type/object pattern for achieving ontological instantiation and the core capacity of Openflexo/FML to reference models disregarding abstraction levels.
+Our solution uses a metamodel to group the required concepts for describing universal processes and an extending metamodel that adapts it to the domain of software engineering processes.
+Process instances reference elements from both metamodels.}
+
+\noteSM{Add some infor about how constraints are managed in each of the approaches.}
 

### 2021-09-24 responseLetter.tex
Extending related work.

@@
 Advantage, since it brings us a completely new, different way of thinking, and disadvantage since it does not really help in categorizing/unifying multi-level methods and – for me – it also feels a little bit off-topic in this special issue.
-At this point, it should be mentioned that the authors managed to cover/solve all requirements of the challenge, thus the paper, as a solution to the challenge fits well.}
-TODO.\todo{Answering this is going to be tricky!}
+At this point, it should be mentioned that the authors managed to cover/solve all requirements of the challenge, thus the paper, as a solution to the challenge fits well.} We introduced at the end of the introduction a description of our approach.
+In a word, we use FML (linguistic level) as a tool to build models and meta-models that allows multi-level modelling.
+The evolution of models (multi-levels) is encoded as a hierarchy of models and meta-models.
+We do multi-level modelling on demands, and propose an organization guided by models evolution.
+\todo{Answering this is going to be tricky!}
 \end{response}

### 2021-09-24 9-relatedworks.tex
Extending related work.

--- 2021-09-20 9-relatedworks.tex
@@
 %Salvador commence
+
+\textcolor{red}{Our solution to the Multi-Level process challenge relies on the use of the type/object pattern~\citep{typeObject} for achieving ontological instantiation and the core capacity of Openflexo/FML to reference models disregarding abstraction levels.
+Concretely, we use a metamodel to group the required concepts for describing universal processes and an extending metamodel that adapts it to the domain of software engineering processes.
+Process instances may reference elements from both metamodels.}
+
 A plethora of multi-level modeling approaches and tools with different foundations have appeared in recent years\footnote{\url{https://homepages.ecs.vuw.ac.nz/Groups/MultiLevelModeling/MultiTools}}.
@@
 the challenge version).
-A solution is provided with MetaDepth~\parencite{metadepth}, which supports modeling with any number of levels, dual ontological/linguistic typing and deep characterization through potency.
+A solution is provided with MetaDepth~\parencite{metadepth}, which supports modeling with any number of levels, dual ontological/linguistic typing and deep characterization through potency~\citep{potency}.
+\textcolor{red}{For their solution to the multi-level process problem they use three levels.
+The first level describes generic processes, the second level, software engineering processes and the third level, software engineering enacted processes.
+The use \emph{linguistic extensions}\citep{metadepth}, a mechanism to  linguistically extend ontological instance models, in order to introduce artefact types and tasks duration in the second level.}
 
-\cite{multiecore2019} use MultiEcore \citep{multecore2016} to provide a solution to the challenge.
-MultEcore uses an extension of the two-level cascading technique and potency.
-They do not need the synthetic typing relations we introduce.
-However they need to do frequent model transformations to transform instance models into instantiable models.
-With respect to the proposed solution, we use fewer models as we do not include a generic software engineering process model nor a generic assurance process model.
-%Finally, they use multi-level couple model transformations for the definition and enforcement of constraints.
+\cite{multiecore2019} use MultEcore \citep{multecore2016} to provide a solution to the challenge.
+\textcolor{red}{MultEcore uses, (un)pluggable linguistic levels (e.g, there is not a \emph{fixed} core metamodel defining the concept of clabjet), an extension of the two-level cascading technique \citep{atkinson2005concepts} (this is achieved with model transformations to transform instance models into instantiable models) and potency.
+With respect to the challenge, their solution uses 4 levels.
+The first level represents generic processes and constitutes the root level for two ontological hierarchies, one for the software engineering domain and the other for the insurance domain.
+Each hierarchy contains three other ontological levels.
+The second level refines generic process to software engineering and insurance processes.
+The third level refines the aforementioned models to adapt them to the ACME and Xsure processes.
+Finally, process instances lie in level four.
+Additionally, the authors use an orthogonal linguistic hierarchy to fulfill requirement \textbf{P19} about alternative names.
+With respect to our solution, the authors use more models, as we require less refinement steps.
+However accidental complexity is introduced by our approach as we need additional constructs and constraints for the typing relations which are otherwise built-in in MultEcore.} \noteSM{How do we deal with P19?}
 
@@
 %\noteSM{Say something about constraints in \FML} Bah... everybody does multilevel constraints.
+
+\noteSM{Add some infor about how constraints are managed in each of the approaches.}
+\noteSM{e.g., built-in constraints for typing relations}

### 2021-09-24 9-relatedworks.tex
Still extending related work.

@@
 More similar to us, in the submission \parencite{deeptelos2019}, the author uses DeepTelos \parencite{deeptelos2016}, an extension of the Telos language \parencite{telos1990} in order to solve the process challenge.
-As with Openflexo and \FML, they do not use explicit (numbered) levels nor potency.
-However, unlike Openflexo, DeepTelos integrates a multi-level modeling specific construct similar to the PowerType \parencite{atkinson2001essence} pattern they call \emph{most general instances}.
+As with Openflexo/\FML, they do not use explicit (numbered) levels nor potency.
+However, unlike Openflexo, DeepTelos integrates a multi-level modeling specific construct similar to the PowerType \parencite{atkinson2001essence} pattern they call \emph{most general instances} \textcolor{red}{which the authors use to simulate potency.
+Models in DeepTelos are organized in (tree-like) module hierarchies, where submodules can see all the concepts defined in parent modules.
+The authors use this module system in order to \emph{organize} their solution for the Multi-level process challenge.
+Concretely they created a hierarchy of modules such as the top module contains base concepts and formulas required for multi-level modeling (e.g., the support for \emph{most general instances}), and a sub-module contains process definitions fulfilling requirements P1 to P19.
+This sub-module contains in turn a sub-module representing the coding process type whereas concrete coding processes are represented in subsequent sub-modules}.
 

### 2021-09-24 10-conclusions.tex
Finishing related work. A word on constraint is missing.

@@
 metametamodeling infrastructure offered by Openflexo.
-\noteSM{This infrastructure helped us to overcome the limitations of the strict modeling paradigm mentioned in Section \ref{sec:introduction}, this is, the lack of support for different forms of classifications and for the duality type/object.
+\textcolor{red}{This infrastructure helped us to overcome the limitations of the strict modeling paradigm mentioned in Section \ref{sec:introduction}, this is, the lack of support for different forms of classifications and for the duality type/object.
 In that sense, our solution seamlessly integrates the type-object pattern which allows us to dynamically create different \emph{instances} of model elements (e.g., \emph{TaskType}, among others) and use them for typing other model elements.

### 2021-09-24 9-relatedworks.tex
Finishing related work. A word on constraint is missing.

@@
 This sub-module contains in turn a sub-module representing the coding process type whereas concrete coding processes are represented in subsequent sub-modules}.
+There are two main differences between the DeepTelos and Openflexo/\FML: 1) we use the type-object pattern instead of the powertype pattern; 2) DeepTelos reify their support for multi-level modeling while ours remains ad-hoc.
 
@@
 Similarly, \cite{dmla2019} use the Dynamic Multi-Layer Algebra (DMLA) \parencite{dmla2017} in order to solve the process challenge.
-DMLA is a level-blind modeling framework which is fully customizable (\eg different types of instantiation may be implemented).
+DMLA is a level-blind modeling framework which is fully customizable (\eg different types of instantiation may be implemented) \textcolor{red}{and includes support for deep characterization (instantiation may refer to any other element disregarding hierarchy levels and a sort of potency in the form of fluid metamodeling at the entity level exists).}
+
 Their proposed solution separates the process challenge in two separate domains, namely, the task definition domain and the process definition domain.

### 2021-09-27 9-relatedworks.tex
Updates from Overleaf

@@
 \textcolor{red}{For their solution to the multi-level process problem they use three levels.
-The first level describes generic processes, the second level, software engineering processes and the third level, software engineering enacted processes.
-The use \emph{linguistic extensions}\citep{metadepth}, a mechanism to  linguistically extend ontological instance models, in order to introduce artefact types and tasks duration in the second level.}
+The first level describes generic processes, the second level software engineering processes, and the third level, software engineering enacted processes.
+Additionally, the authors use~\emph{linguistic extensions}~\citep{metadepth}, a mechanism to  linguistically extend ontological instance models, in order to introduce artefact types and tasks duration in the second level.}
 

### 2021-09-27 4-technology.tex
Updates from Overleaf
small change to the conclusions

@@
     \includegraphics[width=1.0 \textwidth]{Figures/FMLMetaModel.pdf}
-    \caption{Internal representation of \FML language}
+    \caption{Representation of the main concepts of \FML language}
     \label{fig:mm}
@@
 
-\noteSylvain{Parler plutot ici de représentation interne du language FML (plutot que metamodel qui embrouille le lecteur avec ce qui suit ?}
 \FML simplified metamodel is provided in Figure~\ref{fig:mm}. It is designed to define models as virtual models. A virtual model
@@
 
+\noteSylvain{Retravailler le texte pour parler de tous les concepts qui sont sur la figure}
+
 \FML is designed to define not only the structure of virtual models but
@@
 
-Figure \ref{fig:BBPMNSubsetExample} shows...
+Figure \ref{fig:BPMNSubsetExample} shows...
+
+\noteSM{We use the current Figure 3 to answer reviewer 1 in the response letter just as to show how the model federation works, but not here.}
+
+\todo{Faire l'explication une fois que l'on aura décidé ce qu'on y montre...}
 

### 2021-09-27 responseLetter.tex
Updates from Overleaf

@@
 TODO.
+
+Voici la figure avec la connexion externe
+
+\begin{figure}[t]
+    \centering
+    \includegraphics[width=1.0 \textwidth]{Figures/BPMNSubsetExampleWithExternalConnexion.pdf}
+    \caption{An example of BPMN subset built on top of a BPMN/EMF model}
+    \label{fig:BPMNSubsetExample}
+\end{figure}
+
 \end{response}
@@
 Although I agree that it is enough to focus on previous solutions here, much more detail and a comparison would be required.}
-TODO.
+
+We have extended Section 7. It discusses now previously published solutions to the challenge in more detail. Comparisons with our approach have been also extended.
+
 \end{response}

### 2021-09-27 9-relatedworks.tex

@@
 The first level describes generic processes, the second level software engineering processes, and the third level, software engineering enacted processes.
-Additionally, the authors use~\emph{linguistic extensions}~\citep{metadepth}, a mechanism to  linguistically extend ontological instance models, in order to introduce artefact types and tasks duration in the second level.}
+Additionally, the authors use~\emph{linguistic extensions}~\citep{metadepth}, a mechanism to  linguistically extend ontological instance models at any level, in order to introduce artefact types and tasks duration in the second level.
+Our solution is similar to theirs w.r.t.
+the grouping and organization of concepts (e.g., a model for general process, concepts, an extension for software engineering process) but it replaces the use of potency with the use of the type-object pattern.
+The so called \emph{linguistic extensions} are natively supported in Openflexo/FML.}
 
@@
 Concretely they created a hierarchy of modules such as the top module contains base concepts and formulas required for multi-level modeling (e.g., the support for \emph{most general instances}), and a sub-module contains process definitions fulfilling requirements P1 to P19.
-This sub-module contains in turn a sub-module representing the coding process type whereas concrete coding processes are represented in subsequent sub-modules}.
-There are two main differences between the DeepTelos and Openflexo/\FML: 1) we use the type-object pattern instead of the powertype pattern; 2) DeepTelos reify their support for multi-level modeling while ours remains ad-hoc.
+This sub-module contains in turn a sub-module representing the coding process type whereas concrete coding processes are represented in subsequent sub-modules.
+There are two main differences between the DeepTelos and Openflexo/\FML: 1) we use the type-object pattern instead of the powertype pattern; 2) DeepTelos reify their support for multi-level modeling while ours remains ad-hoc.}
 

### 2021-09-27 responseLetter.tex
Somme more changes in the related work.

@@
 \end{figure}
+
+\begin{lstlisting}
+use org.openflexo.technologyadapter.emf.EMFModelSlot as EMF;
+import ["http://www.omg.org/spec/BPMN/20100524/MODEL-XMI"] as BPMN_METAMODEL;
+
+// A simplified BPMN metamodel exposing
+// both concepts BPMNProcess and BPMNTask
+model SimplifiedBPMN {
+
+  // Connexion to a .bpmn file
+  EMFModel bpmnModel 
+    with EMF::EMFModelSlot(
+        metaModel = BPMN_METAMODEL);
+
+  // Abstract concept BPMNElement defining
+  // a name and a link to an eventual BPMNElement
+  // in .bpmn file
+  abstract concept BPMNElement {
+  	EMFObjectIndividual bpmnElement 
+  	    with EMF::EMFObjectIndividualRole (
+			source = bpmnModel
+		);
+    String name values bpmnElement.name;
+  }
+
+  concept BPMNProcess extends BPMNElement {
+
+    // Describe here properties and
+    // behaviours for BPMNProcess concept
+    ...
+
+    concept BPMNTask extends BPMNElement {
+
+        // Describe here properties and
+        // behaviours for BPMNTask concept
+        ...
+
+    }
+
+    // Other core concepts
+  }
+}
+\end{lstlisting}
 

### 2021-09-27 4-technology.tex
Somme more changes in the related work.

--- 2021-05-31 4-technology.tex
@@
 
-\begin{figure}[t]
+\begin{figure*}
     \centering
-    \includegraphics[width=\columnwidth]{Figures/federation.pdf}
-    \caption{The model federation approach}
-    \label{fig:mf}
-\end{figure}
+    \includegraphics[width=1.0 \textwidth]{Figures/FMLMetaModel.pdf}
+    \caption{Representation of the main concepts of \FML language}
+    \label{fig:mm}
+\end{figure*}
+%\noteAntoine{Dans la figure ne manque-t-il pas les containements ? (VM vers Concepts)}
+%\noteSylvain{C'est le cas dans cette nouvelle figure}
 
@@
 In this paper, we mainly use conceptual models and more precisely \FML its domain specific modeling language. The only use of the federation aspect is on the tooling made for the process language.
-
-\begin{figure}[t]
-    \centering
-    \includegraphics[width=\columnwidth]{Figures/FMLCoreModel.pdf}
-    \caption{The \FML core metamodel}
-    \label{fig:mm}
-\end{figure}
-%\noteAntoine{Dans la figure ne manque-t-il pas les containements ? (VM vers Concepts)}
 
@@
 exploit this possibility in the challenge.
+
+%\noteSylvain{Tentative de représenter ici un sous-ensemble de BPMN en montrant la capacité du language FML à faire de la modélisation conceptuelle d'une part, mais aussi à fédérer des nouveaux concepts avec un modèle et un métamodèle existant.}
+
+%\noteSylvain{On explique dans la suite que dans le challenge, on ne va pas exploiter l'aspect fédération de modèle (sauf pour la partie diagramme - c'est expliqué plus haut également - ).}
+
+Figure \ref{fig:BPMNSubsetExample} shows...
+
+%\noteSM{We use the current Figure 3 to answer reviewer 1 in the response letter just as to show how the model federation works, but not here.}
+
+\noteSylvain{Rédiger l'explication...}
+
+\begin{figure}[t]
+    \centering
+    \includegraphics[width=\columnwidth]{Figures/BPMNSubsetExample.pdf}
+    \caption{An example of BPMN subset showing both the model and an instance of that model}
+    \label{fig:BPMNSubsetExample}
+\end{figure}
+
+\begin{lstlisting}
+// A simplified BPMN metamodel exposing
+// both concepts BPMNProcess and BPMNTask
+model SimplifiedBPMN {
+
+  // Abstract concept BPMNElement defining
+  // a name
+  abstract concept BPMNElement {
+    String name;
+  }
+
+  concept BPMNProcess extends BPMNElement {
+
+    // Describe here properties and
+    // behaviours for BPMNProcess concept
+    ...
+
+    concept BPMNTask extends BPMNElement {
+
+        // Describe here properties and
+        // behaviours for BPMNTask concept
+        ...
+
+    }
+
+    // Other core concepts
+  }
+}
+\end{lstlisting}
+
+\noteSylvain{Faut-il montrer des exemples de comportement aussi ?}
 

### 2021-09-27 9-relatedworks.tex
Updates from Overleaf

@@
 Our solution is similar to theirs w.r.t.
-the grouping and organization of concepts (e.g., a model for general process, concepts, an extension for software engineering process) but it replaces the use of potency with the use of the type-object pattern.
+the grouping and organization of concepts (e.g., a model for general process, concepts, an extension for software engineering process) but it replaces the use of clabjets and potency with the use of the type-object pattern.
 The so called \emph{linguistic extensions} are natively supported in Openflexo/FML.}

### 2021-09-27 4-technology.tex

@@
     String name;
+    // A basic constructor
+    create(String name) {
+      this.name = name;
+    }
   }
@@
 
-    // Describe here properties and
+    // A basic constructor
+    create(String name) {
+      super(name);
+    }
+    // Describe here other properties and
     // behaviours for BPMNProcess concept
@@
 
-        // Describe here properties and
+      // A basic constructor
+      create(String name) {
+        super(name);
+      }
+      // Encodes execution of task
+      execute() {
+        log("Execute task "+name);
+        ...
+      }
+      // Describe here other properties and
       // behaviours for BPMNTask concept
@@
 \end{lstlisting}
-
-\noteSylvain{Faut-il montrer des exemples de comportement aussi ?}
 

### 2021-09-27 9-relatedworks.tex

@@
 
-A plethora of multi-level modeling approaches and tools with different foundations have appeared in recent years\footnote{\url{https://homepages.ecs.vuw.ac.nz/Groups/MultiLevelModeling/MultiTools}}.
+A plethora of multi-level modeling approaches and tools\footnote{\url{https://homepages.ecs.vuw.ac.nz/Groups/MultiLevelModeling/MultiTools}} with different foundations have appeared in recent years~\citep{somogyi2021playground}.
 Comparing them lies out of the scope of this paper.

### 2021-09-27 3-introduction.tex

--- 2021-05-31 3-introduction.tex
@@
 %\noteFabien{Le paragraphe qui précède me semble redire des choses de la section 2, comment fait-on ?}
+
+\textcolor{red}{Our modeling approach is based on two levels.
+The meta level (\FML) provides the (linguistic) concepts that are used to define the model level (ontological).
+The ontological level is used to define the models and meta-models that are needed.
+\FML allows to define models as first class entities (which we call VirtualModel) and to link them together (by extension).
+In the \mpc case study, we exploit this possibility to:
+\begin{itemize}
+    \item Extend and adapt existing models, keeping the previous ones, to meet changing modeling needs,
+    \item Align this hierarchy of models with the process of their evolution.
+\end{itemize}
+This methodological choice allows us to adapt the models and keep track of their evolution each time the needs evolve.}
+
+\textcolor{red}{Contrary to a classical multi-level approach, in which the notion of multi-level is integrated in the meta-model and must therefore be respected, our approach allows to build multi-levels on demand, and with the appropriate form.
+As a limitation, without multilevel-specific operators/concepts, the "custom" construction must be done explicitly during the model development and evolution process.}
+
+\textcolor{red}{Our answer to the challenge is therefore not an answer using a classical multilevel approach that would allow an efficiency comparison with another classical multilevel answer, but an approach that proposes another organization of multilevels through the translation of needs into levels over time and the possibility to build an evolving hierarchy of adapted models.
+This is made possible by (1) the reification of the notion of model and the possibility to interconnect them (specialization) and (2) the freedom of meta-modeling (low model conformity, flexible meta-modeling).}
 

### 2021-09-27 8-discussion.tex
removing trailing whitespace
Updates from Overleaf

@@
   or fewer levels depending on the variations of the use case. However, the number of levels is related
-  to the problem, making the approach fixed.
+  to the problem, making the approach fixed.\textcolor{red}{AB: Fixed?}
 
@@
   In this context, we would have to transform the \inst{Design Task} instance into a concept
-  whose instances could have been "using agile methodology".
+  whose instances could have been "using agile methodology"
 

### 2021-09-27 3-introduction.tex
Updates from Overleaf

@@
 
-\textcolor{red}{Our modeling approach is based on two levels.
-The meta level (\FML) provides the (linguistic) concepts that are used to define the model level (ontological).
-The ontological level is used to define the models and meta-models that are needed.
+\textcolor{red}{Our modeling approach is based on a language (\FML) which provides the (linguistic) concepts that are used to define  the multi-levels models and meta-models that are needed.
 \FML allows to define models as first class entities (which we call VirtualModel) and to link them together (by extension).
@@
 \begin{itemize}
-    \item Extend and adapt existing models, keeping the previous ones, to meet changing modeling needs,
-    \item Align this hierarchy of models with the process of their evolution.
+    \item Define, extend and adapt models, keeping the history, to meet changing modeling requirements of the case study,
+    \item Align this hierarchy of models with their specification evolution.
 \end{itemize}
-This methodological choice allows us to adapt the models and keep track of their evolution each time the needs evolve.}
+This methodological choice allows us to adapt the models and keep track of their evolution each time the requirements evolve.}
 

### 2021-09-27 6-modelpresentation.tex
Updates from Overleaf

@@
     \includegraphics[width=1.0 \textwidth]{Figures/Metamodel.pdf}
-     \caption{Process management base metamodel}
+     \caption{Process management base metamodel (\texttt{Gateway}'s associations with \texttt{TaskType} are represented with attributes)}
     \label{fig:BaseMetamodel}
@@
      \includegraphics[width=1.0 \textwidth]{Figures/AcmeFullArchitecture.pdf}
-     \caption{Acme software development process architecture (not all instances are shown for readability reasons)}
+     \caption{Acme software development process architecture (not all instances are shown for the sake of readability)}
     \label{fig:AcmeFullArchitecture}

### 2021-09-27 9-relatedworks.tex

@@
 With respect to our solution, the authors use more models, as we require less refinement steps.
-However accidental complexity is introduced by our approach as we need additional constructs and constraints for the typing relations which are otherwise built-in in MultEcore.} \noteSM{How do we deal with P19?}
+However accidental complexity is introduced by our approach as we need additional constructs and constraints for the typing relations which are otherwise built-in in MultEcore.}
+
+\noteSM{P19 has changed in the new challenge... to remove references to it}.
+
+%\noteSM{How do we deal with P19}
 
@@
 %\noteSM{Say something about constraints in \FML} Bah... everybody does multilevel constraints.
-
-\noteSM{Add some infor about how constraints are managed in each of the approaches.}
-\noteSM{e.g., built-in constraints for typing relations}

### 2021-09-27 6-modelpresentation.tex
Updates from Overleaf

@@
 
-The \texttt{CodeArtifact} concept extends \texttt{SEArtifact}, and defines a relation \texttt{language} to \textit{ProgrammingLanguage} with 1..* cardinality (\textbf{S4}).
+The \texttt{CodeArtifact} concept extends \texttt{SEArtifact}, and defines a relation \texttt{language} to \texttt{ProgrammingLanguage} with 1..* cardinality (\textbf{S4}).
 
@@
 All tasks -- instantiated from \texttt{TaskType} for a given enacted process -- manage a status.
-This status is either \texttt{Not startable} (when not assigned to a performing actor or when required input artifacts are not available) , \texttt{Startable}, \texttt{Started}, \texttt{Completable} (when all output artifacts are ready) or \texttt{Completed}.
+This status is either \texttt{Not startable} (when not assigned to a performing actor or when required input artifacts are not available), \texttt{Startable}, \texttt{Started}, \texttt{Completable} (when all output artifacts are ready) or \texttt{Completed}.
 A \texttt{Task} also manages a set of performing actors, a begin and end date, some used and produced artifacts.

### 2021-09-27 2-abstract.tex
antidote abstract.tex
removing trailing whitespace
sync overleaf -> github

@@
-Model federation is a multi model \emph{management} approach based on the use of virtual models and loosely coupled links.
+Model federation is a multi-model \emph{management} approach based on the use of virtual models and loosely coupled links.
 The models in a federation remain autonomous and represented in their original technological spaces whereas virtual models and links (which are not level bounded) serve as control components used to present different views to the users and  maintain synchronization.
@@
 We present a solution based on model federation and discuss the advantages and limitations of using this approach for multi-level modeling.
-Concretely, we use virtual models and more precisely the Flexo Modeling Language (\FML) that serves to describe them as the main building block in order to solve the process modeling challenge whereas the federation aspect is used as a means to provide editing tooling for the resulting process language.
+Concretely, we use virtual models and more precisely the Flexo Modeling Language ({FML}) that serves to describe them as the main building block in order to solve the process modeling challenge whereas the federation aspect is used as a means to provide editing tooling for the resulting process language.
 Our solution is fully implemented with the Openflexo framework.

### 2021-09-27 4-technology.tex
sync overleaf -> github

@@
     \includegraphics[width=1.0 \textwidth]{Figures/FMLMetaModel.pdf}
-    \caption{Representation of the main concepts of \FML language}
+    \caption{Representation of the main concepts of {FML} language}
     \label{fig:mm}
@@
 
-\FML simplified metamodel is provided in Figure~\ref{fig:mm}. It is designed to define models as virtual models. A virtual model
+A UML like representation of a simplified metamodel of \FML is provided in Figure~\ref{fig:mm}. \FML is language designed to define models as virtual models. A virtual model
 is composed of a set of \emph{concepts}, while itself being a concept.
@@
 
-  // Abstract concept BPMNElement defining
-  // a name
+  // Abstract concept BPMNElement defining a name
   abstract concept BPMNElement {

### 2021-10-05 responseLetter.tex
Commented out listing
Polishing related work.

--- 2021-09-27 responseLetter.tex
@@
 \begin{response}{the reviewers have made suggestions for improving the presentation but also raised a number of issues requiring clarification.
-An overarching theme is the request for a more explicit description of how the multi-level modelling emulation is achieved and what its general advantages and limitations are, including a clarification of what the implications would be of supporting more "levels" than the challenge required.} TODO.
+An overarching theme is the request for a more explicit description of how the multi-level modelling emulation is achieved and what its general advantages and limitations are, including a clarification of what the implications would be of supporting more "levels" than the challenge required.} We have been more explicit to explain how we use the FML language at the end of the introduction, and in various figures.
+TODO.
 \end{response}
@@
 You may find the references listed below useful to keep explanations concise by referencing related work (note that reviewer 3’s “Type-Instance pattern” is also known as the “Type Object” pattern).
-You should also respond to other concerns reviewers raise, either by making changes to the submission or in the form of an authors' response.} TODO.
+You should also respond to other concerns reviewers raise, either by making changes to the submission or in the form of an authors' response.} The related work section has been modified to show that we are not the first conventional response with multi-level emulation, and to improve comparisons and limitations of different approaches.
+TODO.
 We may have to create a new section.
@@
 Especially, the 'metamodel'-agnostic approach seems very interesting but I would like to see how that is used for traditional modeling approaches.}
+The previous figure illustrates the connection with external sources.
+However, in the article, we choose not to give this level of details and to keep focus on the left part of the figure.
+It shows how a VirtualModel (SimplifiedBPMN) can filter a model (BPMN) and be instantiated.
 TODO.
-
-Voici la figure avec la connexion externe
 
@@
 \end{figure}
+
+%\begin{lstlisting}
+%use org.openflexo.technologyadapter.emf.EMFModelSlot as EMF;
+%import ["http://www.omg.org/spec/BPMN/20100524/MODEL-XMI"] as BPMN_METAMODEL;
+%
+%// A simplified BPMN metamodel exposing
+%// both concepts BPMNProcess and BPMNTask
+%model SimplifiedBPMN {
+%
+%  // Connexion to a .bpmn file
+%  EMFModel bpmnModel 
+%    with EMF::EMFModelSlot(
+%        metaModel = BPMN_METAMODEL);
+%
+%  // Abstract concept BPMNElement defining
+%  // a name and a link to an eventual BPMNElement
+%  // in .bpmn file
+%  abstract concept BPMNElement {
+%  	EMFObjectIndividual bpmnElement 
+%  	    with EMF::EMFObjectIndividualRole (
+%			source = bpmnModel
+%		);
+%    String name values bpmnElement.name;
+%  }
+%
+%  concept BPMNProcess extends BPMNElement {
+%
+%    // Describe here properties and
+%    // behaviours for BPMNProcess concept
+%    ...
+%
+%    concept BPMNTask extends BPMNElement {
+%
+%        // Describe here properties and
+%        // behaviours for BPMNTask concept
+%        ...
+%
+%    }
+%
+%    // Other core concepts
+%  }
+%}
+%\textbf{}\end{lstlisting}
 
@@
 Please note that EMISAJ does not have a page limitation, so no worries if that should require more space.}
-TODO.
+We've changed figure 2. It shows \texttt{Type} and the main concepts used by FML. TODO. Explanation of figure has been improved. 
 \end{response}
@@
 TODO.
+Faire le lien entre figure 2 et les représentations des figures 3, 4, 5, 6 et 7.
+Package = VirtualModel, Classes bleues = FlexoConcept, attribut =(~) Get/GetSetProperty, association = FlexoRole, no FlexoBehavior visible.
+(!!
+Dans la figure 2 il n'y a pas d'héritage possible visible entre FlexoConcept.) Le niveau orange est l'instantiation / interprétation du FML.
 \end{response}
@@
 It is not clear, how federation works in this case.}
-TODO.
+We have improved the description and we hope this is much clearer. TODO.
 \end{response}
@@
 The text mentions the (linguistic) instantiation mechanism of the FML Language… how does it work exactly?}
-TODO.
+We hope we have improved the explanation. \texttt{InstanceInstance} has been replace with \texttt{ObjectInstance}. TODO.
 \end{response}

### 2021-10-05 3-introduction.tex
sync overleaf -> github
Polishing related work.

@@
 \begin{itemize}
-    \item Define, extend and adapt models, keeping the history, to meet changing modeling requirements of the case study,
-    \item Align this hierarchy of models with their specification evolution.
+    \item Define, extend and adapt models, keeping the history, to meet {\color{blue}the} %changing
+    modeling requirements of the case study,
+    \item Align this hierarchy of models with requirements evolution ($P_1-P_{19}$ then $S_1-S_{13}$).
 \end{itemize}
@@
 
-\textcolor{red}{Contrary to a classical multi-level approach, in which the notion of multi-level is integrated in the meta-model and must therefore be respected, our approach allows to build multi-levels on demand, and with the appropriate form.
-As a limitation, without multilevel-specific operators/concepts, the "custom" construction must be done explicitly during the model development and evolution process.}
+\textcolor{red}{Contrary to a classical multi-level approach, in which the notion of multi-level is integrated in the language and must therefore be respected, our approach allows to build multi-levels on demand, and with the appropriate form.
+Without multilevel-specific operators/concepts, the "custom" construction must be done explicitly during the model development and evolution process.}
 

### 2021-10-05 4-technology.tex
Polishing related work.

@@
 
-To meet the \mlpc, we have decided to use the general
-approach of model federation~\parencite{Golra2016-federation}. \emph{Model
+To meet the \mlpc, we have decided to use the language infrastructure of
+model federation approach~\parencite{Golra2016-federation}. \emph{Model
   federation} is a way to assemble models using some kind of
@@
 after federations are made.
+So to provide an adaptable framework, we have developed a flexible modeling language to provide modeling features without strict modeling levels.
+In a federation, the targeted modelling approaches can be unknown at the beginning and also can integrate several modeling levels.
 
@@
 \FML is designed to define not only the structure of virtual models but
-also to define the collection of actions an engineer can perform on
-them.
+also also the collection of actions that can be performed on them.
 These actions are called \emph{behaviors} and can either be called (like usual methods) or triggered by events.
 The reactive behaviors are mainly useful when a federated model evolution needs to trigger a computation.
-We do not
-exploit this possibility in the challenge.
+%We do not exploit this possibility in the challenge.
 

### 2021-10-05 5-analysis.tex
Commented out listing
Polishing related work.

@@
 The goal is the identification of concepts with their behavior.
+
 The last step is to organize or structure these concepts to improve reusability and extensibility, to create virtual models in \FML terminology.
@@
 Like in a lot of modeling approaches, these  steps could also be achieved in any order and iteratively.
-But our goal during the challenge's analysis remains to produce a \FML virtual model architecture for our federation.
+But our goal during the challenge's analysis remains to produce a \FML virtual model architecture.
+%for our federation.
 As shown in the figure~\ref{fig:MultilevelArchitecture}, we defined two virtual models (Base and Acme processes abstractions) instantiated by two virtual model instances (XSure and Acme processes).
@@
     \centering
-    \includegraphics[width=0.7 \columnwidth]{Figures/MultilevelArchitecture.pdf}
+    \includegraphics[width=0.9 \columnwidth]{Figures/MultilevelArchitecture.pdf}
     \caption{Multilevel architecture of our solution}
@@
 
+%{Version Joel, à prendre ou à laisser} \item The vertical axis is viewed as the linguistic specialization, relative to the use of the \FML language.
+The virtual model, defined as the core definition metamodel, generates a model founded on the instantiation mechanism of the \FML Language or can be extended by another virtual model.
+The instantiation mechanism is similar to the classical object/instance mechanism of the object languages but without any constraint on the referenced virtual model, \ie metamodel.
+The set of resulting instances comes from several virtual models of any abstraction level as detailed in the section.
+\ref{sec:AcmeSoftwareDevelopmentProcess}
+
 \end{itemize}
@@
     \includegraphics[width=1.0 \columnwidth]{Figures/Instantiation.pdf}
-    \caption{Problem  Domain Linguistic and ontologic instantiation}
+    \caption{Problem  domain linguistic and ontologic instantiation}
     \label{fig:LinguisticAndOntologicInstantiation}
@@
 Based on this approach, we organize our set of models following the architecture of the aforementioned Figure \ref{fig:MultilevelArchitecture}.
-The resulting multi-level architecture is organized in two virtual models, one for the core concepts metamodel containing the definition of Processes and Tasks, and one, the Acme metamodel, extending the previous model to integrate the Acme definitions.
+The resulting multi-level architecture is organized in two virtual models, one for the core concepts metamodel containing the definition of Processes and Tasks, and one, the Acme metamodel, extending the previous model to specialize the Acme definitions.
 The \FML language defines multiple inheritance concept between virtual models as illustrated in the section~\ref{sec:AcmeSoftwareDevelopmentProcess}.

### 2021-10-05 6-modelpresentation.tex
Polishing related work.

@@
 A \textit{Concept} of \FML is represented by a \UML class where roles of basic types are attributes.
-The roles whose types are concepts are represented by a combination of a name in their containing concept and an arrow from their name to their type.
+The roles whose types are concepts are represented by an arrow from their concept to their type, with their name on the arrow.
 The cardinality follows the \UML practice.
@@
 
-The requirement \textbf{S8} is simply expressed by the definition of the \inst{Testing} instance of \texttt{SETaskType}, \inst{Tester} instance of \texttt{ActorType}, and \inst{Test report} instance of \texttt{ArtifactType}.
+The requirement \textbf{S8} is simply expressed by the definition of the \inst{Testing} instance of \texttt{SETaskType}, \inst{Tester} instance of \texttt{ActorType}, and \inst{Test report} instance of \texttt{ArtifactType}, as showed in the figure \ref{fig:AcmeSoftwareDevelopmentProcess}.
 
@@
 We took advantage of model federation and the availability of diagraming features though the \textit{Diagramming Technology Adapter} to implement two interactive graphical tools built on top of the conceptual levels detailed in previous sections.
+The figure \ref{fig:ToolingArchitecture} shows the overview of the tooling architecture and is detailed in the next section.
 A first tool offers a graphical edition of a Process type.
@@
 
-Figure \ref{fig:ToolingArchitecture} shows the architecture of these two tools.
+Figure \ref{fig:ToolingArchitecture} details the architecture of these two tools.
  The left of the figure presents the \textit{ProcessType graphical editor}.

### 2021-10-05 8-discussion.tex
Commented out listing
Polishing related work.

--- 2021-05-31 8-discussion.tex
@@
   are grouped into a virtual model. Inheritance and instantiation allow the
-  establishment of relationships between concepts from different levels.
+  establishment of relationships between concepts from different levels. See figure~\ref{fig:BPMNSubsetExample}
+  and \ref{fig:AcmeFullArchitecture}.
 
@@
 
-  Due to the fact that our approach is level-agnostic, our solution could have had more
+  Due to the fact that our approach is level-agnostic, our solution could have more
   or fewer levels depending on the variations of the use case. However, the number of levels is related
-  to the problem, making the approach fixed.
+  to the problem, making the approach fixed (to 3).
 
@@
   linked each other in a transparent way, using inheritance or instantiation.
+For instance the \texttt{Developer} concept, from \texttt{Acme} metamodel, specializes the \texttt{ActorType} of the base metamodel, and the \texttt{Developer} of the \texttt{Acme} model, is an instance of the metamodel concept.
+\textcolor{red}{Phrase faite a relire sur Developer\dots}
 
@@
   Although we build ad hoc metamodels in our approach, our solution separates
-  the domain-specific elements from general purpose ones, making possible to 
-  apply it to other domains.
+  the domain-specific elements from general purpose ones, making possible to reuse the general concepts to other domains.
 
@@
 consistency is realized by cardinality checks and by typing. Second, the
-constraints the users have written are continuously checked. Thus, a part of
-the verification relies on the fact that users add behaviors when they build
-the metamodels and the models.
+constraints the users have written are continuously checked during model edition and enactment. 
+Thus, a part of the verification relies on the fact that users add behaviors when they build
+the metamodels and the models. \textcolor{red}{An example of constraint\dots}

### 2021-10-05 9-relatedworks.tex
Polishing related work.

@@
 %Salvador commence
-
-\textcolor{red}{Our solution to the Multi-Level process challenge relies on the use of the type/object pattern~\citep{typeObject} for achieving ontological instantiation and the core capacity of Openflexo/FML to reference models disregarding abstraction levels.
-Concretely, we use a metamodel to group the required concepts for describing universal processes and an extending metamodel that adapts it to the domain of software engineering processes.
-Process instances may reference elements from both metamodels.}
 
@@
 
-A first description of process modeling as a multi-level modeling problem was proposed by \parencite{lara2018refactoring} in the context of a catalogue of refactoring for multi-level models (in a simpler form with fewer constraints and requirements w.r.t.
+A first description of process modeling as a multi-level modeling problem was proposed by~\citet{lara2018refactoring} in the context of a catalogue of refactoring for multi-level models (in a simpler form with fewer constraints and requirements w.r.t.
 the challenge version).
@@
 Our solution is similar to theirs w.r.t.
-the grouping and organization of concepts (e.g., a model for general process, concepts, an extension for software engineering process) but it replaces the use of clabjets and potency with the use of the type-object pattern.
+the grouping and organization of concepts (e.g., a model for general process, concepts, an extension for software engineering process) but it replaces the use of clabjets and potency with the use of the type-object pattern~\citep{typeObject}.
 The so called \emph{linguistic extensions} are natively supported in Openflexo/FML.}
+
+%\textcolor{red}{Our solution to the Multi-Level process challenge relies on the use of %the type/object pattern~\citep{typeObject} for achieving ontological instantiation and %the core capacity of Openflexo/FML to reference models disregarding abstraction levels.
+%Concretely, we use a metamodel to group the required concepts for describing universal %processes and an extending metamodel that adapts it to the domain of software %engineering processes.
+Process instances may reference elements from both metamodels.}
 
@@
 \textcolor{red}{MultEcore uses, (un)pluggable linguistic levels (e.g, there is not a \emph{fixed} core metamodel defining the concept of clabjet), an extension of the two-level cascading technique \citep{atkinson2005concepts} (this is achieved with model transformations to transform instance models into instantiable models) and potency.
-With respect to the challenge, their solution uses 4 levels.
+Regarding the challenge, their solution uses 4 levels.
 The first level represents generic processes and constitutes the root level for two ontological hierarchies, one for the software engineering domain and the other for the insurance domain.
@@
 Finally, process instances lie in level four.
-Additionally, the authors use an orthogonal linguistic hierarchy to fulfill requirement \textbf{P19} about alternative names.
+Additionally, the authors use an orthogonal linguistic hierarchy to support alternative names for every model element.
 With respect to our solution, the authors use more models, as we require less refinement steps.
@@
 
-\noteSM{P19 has changed in the new challenge... to remove references to it}.
+%\noteSM{P19 has changed in the new challenge... to remove references to it.
+%JOel : pour simplifier la description j'ai juste mis en commentaire le détail des 4 %niveaux et ca supprime P19. A valider}.
+%\noteSM{How do we deal with P19 - actually P19 does not exist anymore}
 
-%\noteSM{How do we deal with P19}
-
-More similar to us, in the submission \parencite{deeptelos2019}, the author uses DeepTelos \parencite{deeptelos2016}, an extension of the Telos language \parencite{telos1990} in order to solve the process challenge.
+More similar to us,~\citet{deeptelos2019} uses DeepTelos \parencite{deeptelos2016}, an extension of the Telos language \parencite{telos1990} in order to solve the process challenge.
 As with Openflexo/\FML, they do not use explicit (numbered) levels nor potency.
@@
 
-Similarly, \cite{dmla2019} use the Dynamic Multi-Layer Algebra (DMLA) \parencite{dmla2017} in order to solve the process challenge.
-DMLA is a level-blind modeling framework which is fully customizable (\eg different types of instantiation may be implemented) \textcolor{red}{and includes support for deep characterization (instantiation may refer to any other element disregarding hierarchy levels and a sort of potency in the form of fluid metamodeling at the entity level exists).}
-
-Their proposed solution separates the process challenge in two separate domains, namely, the task definition domain and the process definition domain.
+Finally, \cite{dmla2019} use the Dynamic Multi-Layer Algebra (DMLA) \parencite{dmla2017} in order to solve the process challenge.
+DMLA is a level-blind modeling framework which is fully customizable (\eg different types of instantiation may be implemented) \textcolor{red}{and includes support for deep characterization (instantiation may refer to any other element disregarding hierarchy levels and a sort of potency in the form of fluid metamodeling at the entity level exists).} Their proposed solution separates the process challenge in two separate domains, namely, the task definition domain and the process definition domain.
 Wrappers are used in order to reuse tasks in processes.

### 2021-10-05 responseLetter.tex
adding notes to the response letter

@@
 \begin{response}{the reviewers have made suggestions for improving the presentation but also raised a number of issues requiring clarification.
-An overarching theme is the request for a more explicit description of how the multi-level modelling emulation is achieved and what its general advantages and limitations are, including a clarification of what the implications would be of supporting more "levels" than the challenge required.} We have been more explicit to explain how we use the FML language at the end of the introduction, and in various figures.
-TODO.
+An overarching theme is the request for a more explicit description of how the multi-level modelling emulation is achieved and what its general advantages and limitations are, including a clarification of what the implications would be of supporting more "levels" than the challenge required.}
+
+We have been more explicit to explain how we use the FML language at the end of the introduction, and in various figures. TODO.
+
+\todo[inline]{We have enhanced the presentation of both, Openflexo/FML and our solution to the challenge so that: 1) the techniques we use to \emph{achieve} multi-level capabilities are more explicit; and, 2) the advantages and disadvantages of using them in order to fulfil the challenge requirements become clearer.
+Notably, we have added many details in the final paragraphs of the introduction section, improved the Section 2, devoted to the description of OpenFlexo, refactored figures X, Y, Z.}
 \end{response}
@@
 You may find the references listed below useful to keep explanations concise by referencing related work (note that reviewer 3’s “Type-Instance pattern” is also known as the “Type Object” pattern).
-You should also respond to other concerns reviewers raise, either by making changes to the submission or in the form of an authors' response.} The related work section has been modified to show that we are not the first conventional response with multi-level emulation, and to improve comparisons and limitations of different approaches.
-TODO.
-We may have to create a new section.
+You should also respond to other concerns reviewers raise, either by making changes to the submission or in the form of an authors' response.}
+The related work section has been modified to show that we are not the first conventional response with multi-level emulation, and to improve comparisons and limitations of different approaches.\todo[inline]{I do not understand this answer.
+We are saying that other approaches are also unconventional?}
 \end{response}
 
-\begin{response}{Finally, to achieve consistency with other submissions, could you please change the subtitle to "A Contribution to the Multi-Level Process Challenge"?} Of course we can.
-  DONE.
+\begin{response}{Finally, to achieve consistency with other submissions, could you please change the subtitle to "A Contribution to the Multi-Level Process Challenge"?}
+
+Done.
 \end{response}
@@
 Especially, the 'metamodel'-agnostic approach seems very interesting but I would like to see how that is used for traditional modeling approaches.}
-The previous figure illustrates the connection with external sources.
-However, in the article, we choose not to give this level of details and to keep focus on the left part of the figure.
-It shows how a VirtualModel (SimplifiedBPMN) can filter a model (BPMN) and be instantiated.
-TODO.
+
+The Figure \ref{fig:BPMNSubsetExample} (above) shows how to realize a small subset of BPMN with Openflexo/FML (assuming a metamodel of the full specification of BPMN exists).
+Concretely, we show  how a VirtualModel (SimplifiedBPMN) can filter an existing model (BPMN) and be instantiated.
+Note that connection with external sources is not required for our solution to the challenge and is only represented in order to illustrate the capabilities of Openflexo/FML.
+Therefore, in the article we choose not to give this level of details and focus on the left part of the figure.
+
+\todo[inline]{I've refactored the paragraph above. To be verified.}
 

### 2021-10-06 8-discussion.tex
add examples in section 6 (discussion)

@@
   are grouped into a virtual model. Inheritance and instantiation allow the
-  establishment of relationships between concepts from different levels. See figure~\ref{fig:BPMNSubsetExample}
-  and \ref{fig:AcmeFullArchitecture}.
+  establishment of relationships between concepts from different levels.
+%\ref{fig:BPMNSubsetExample}
+%  Figure~\ref{fig:AcmeFullArchitecture} shows three levels: the metamodel, the
+%  ACME metamodel and the ACME model.
+Figure~\ref{fig:MultilevelArchitecture} shows the multilevel architecture of
+our solution.
+\noteJC{updated}
 
@@
 
-  Due to the fact that our approach is level-agnostic, our solution could have more
-  or fewer levels depending on the variations of the use case. However, the number of levels is related
-  to the problem, making the approach fixed (to 3).
+  Due to the fact that our approach is level-agnostic, our solution could have
+  more or fewer levels depending on the variations of the use case. However,
+  the number of levels is related to the problem, making the approach fixed (to
+  3).
 
@@
   linked each other in a transparent way, using inheritance or instantiation.
-For instance the \texttt{Developer} concept, from \texttt{Acme} metamodel, specializes the \texttt{ActorType} of the base metamodel, and the \texttt{Developer} of the \texttt{Acme} model, is an instance of the metamodel concept.
-\textcolor{red}{Phrase faite a relire sur Developer\dots}
+  For instance in the figure~\ref{fig:AcmeFullArchitecture}, the
+  \texttt{Developer} concept from \texttt{Acme} metamodel specializes the
+  \texttt{ActorType} of the base metamodel, and the \texttt{Developer} of the
+  \texttt{Acme} model is an instance of the metamodel concept.
+%  \textcolor{red}{Phrase faite a relire sur Developer\dots}
+\noteJC{updated}
 
@@
   users. Thus, the integrity of the contents relies on the users when they
-  build the metamodels and the models.  Note that in our approach, the
-  metamodels are built in an ad hoc way together with the models
-  (co-construction), therefore the user also validates on an ad hoc basis.
+  build the metamodels and the models. The section~\ref{sec:model} describes
+  how the requirements are captured and how behaviors are implemented with
+  \FML: for instance, \texttt{isAuthorizedActor} behavior is needed to enforce
+  \textbf{P17} requirement, leading to writing \FML code in \texttt{TaskType}
+  and \texttt{Task}. Note that in our approach, the metamodels are built in an
+  ad hoc way together with the models (co-construction), therefore the user
+  also validates on an ad hoc basis.
+\noteJC{updated}
 
@@
   %\noteJC{pas trop clair pour moi pour le moment -> N/A ? }
-  Due to the nature of our approach and its level-agnosticism, \emph{deep characterization} does not
-  apply to our solution. Such a mechanism could probably be encoded by specific behaviors, however,
-  it would not be a generic mechanism, making it difficult to reuse for another problem.
+
+  Due to the nature of our approach and its level-agnosticism, \emph{deep
+  characterization} does not apply to our solution. Such a mechanism could
+  probably be encoded by specific concepts and constraints (behaviors).
+  However, it would not be a generic mechanism, making it difficult to reuse
+  for another problem.  Plus, programming the notion of \emph{depth} would
+  probably imply that a virtual model knows its sub-elements, which is not
+  desired and considered as bad design.
+\noteJC{updated}
 
@@
   Although we build ad hoc metamodels in our approach, our solution separates
-  the domain-specific elements from general purpose ones, making possible to reuse the general concepts to other domains.
+  the domain-specific elements from general purpose ones, making possible to
+  reuse the general concepts to other domains.
 
@@
   to other ones. Then, we would add any necessary constraints on those concepts
-  and on the relationships linking them to the other existing concepts. This type
-  of change can be done without much effort. On the other hand, making changes
-  to the existing models could be difficult. For example, given a requirement,
-  an instance cannot be changed into a concept easily. This case could occur if
-  the challenge had refined the specification by requiring specific design tasks.
-  In this context, we would have to transform the \inst{Design Task} instance into a concept
-  whose instances could have been "using agile methodology".
+  and on the relationships linking them to the other existing concepts. This
+  type of change can be done without much effort. On the other hand, making
+  changes to the existing models could be difficult. For example, given a
+  requirement, an instance cannot be changed into a concept easily. This case
+  could occur if the challenge had refined the specification by requiring
+  specific design tasks.  In this context, we would have to transform the
+  \inst{Design Task} instance into a concept whose instances could have been
+  "using agile methodology".
 
@@
 
-We use the Openflexo infrastructure to build models and metamodels
-together, and to provide dedicated tooling to edit a Process type, to enact a
-process and to execute it. Being able to provide a solution and its associated
-tools quickly and easily is a noticeable feature.  We could also have taken
-advantage of the model federation to connect our solution to other tools
-dedicated to process edition and to process execution (\eg a \BPMN engine).
-However we already had all the necessary components to provide the
-aforementioned graphical tools.
+We use the Openflexo infrastructure to build models and metamodels together,
+and to provide dedicated tooling to edit a Process type, to enact a process and
+to execute it. Being able to provide a solution and its associated tools
+quickly and easily is a noticeable feature.  We could also have taken advantage
+of the model federation to connect our solution to other tools dedicated to
+process edition and to process execution (\eg a \BPMN engine). However we
+already had all the necessary components to provide the aforementioned
+graphical tools.
 
@@
 consistency is realized by cardinality checks and by typing. Second, the
-constraints the users have written are continuously checked during model edition and enactment. 
-Thus, a part of the verification relies on the fact that users add behaviors when they build
-the metamodels and the models. \textcolor{red}{An example of constraint\dots}
+constraints the users have written are continuously checked during model
+edition and enactment. Thus, a part of the verification relies on the fact that
+users add behaviors when they build the metamodels and the models. For
+instance, \textbf{S6} requirement is constrained by an invariant implemented in
+\texttt{CodeArtifact}. It ensures that \texttt{artifactType} has the right type
+and that it implements the \texttt{languages} relation.
+\noteJC{updated}

### 2021-10-06 6-modelpresentation.tex
consistency between figure and text: the relation between CodeArtifact and ProgrammingLanguage is "languages", but text was saying "language". Updated the text.

@@
 
-The \texttt{CodeArtifact} concept extends \texttt{SEArtifact}, and defines a relation \texttt{language} to \texttt{ProgrammingLanguage} with 1..* cardinality (\textbf{S4}).
+The \texttt{CodeArtifact} concept extends \texttt{SEArtifact}, and defines a relation \texttt{languages} to \texttt{ProgrammingLanguage} with 1..* cardinality (\textbf{S4}).
 
@@
 This requirement is captured through the redefinition of the behavior \texttt{declaresRequiredProducedArtifacts} where programming language should also match.
-\textbf{S6} is guaranteed though the \texttt{language} relation defined in \texttt{CodeArtifact} and the following invariant declared in \texttt{CodeArtifact}:
+\textbf{S6} is guaranteed though the \texttt{languages} relation defined in \texttt{CodeArtifact} and the following invariant declared in \texttt{CodeArtifact}:
 \begin{lstlisting}

### 2021-10-06 responseLetter.tex
sync overleaf -> github

@@
 Can the essence of this be captured for Openflexo and then described in terms of how it facilitiates the tooling arising from the challenge?}
-TODO.
+
+We tried to illustrate subsections of section 6 by giving examples, by referencing figures and sections.
+
+Concerning the fact that "MLM is the key to the ability of Openflexo to create tooling", the reviewer probably read "MLM" instead of "FML" (Flexo Modeling Language).
 \end{response}

### 2021-10-07 4-technology.tex

@@
 
-A UML like representation of a simplified metamodel of \FML is provided in Figure~\ref{fig:mm}. \FML is language designed to define models as virtual models. A virtual model
-is composed of a set of \emph{concepts}, while itself being a concept.
-Hence, virtual models are structuring units forming architectures while concepts are the
-core entities. A concept has a set of \emph{roles} and
-\emph{behaviors}. A parallel to object-oriented approach can be useful
-to understand \FML\footnote{Some aspects of \FML do not exist in object oriented approach.}. A concept corresponds to a class, its roles to the
-attributes of the class and its behaviors to the methods of the class.
-These roles have types defining the kind of value the role will point
-at runtime.
-Whenever a type external to the federation space is used, one needs to use a \emph{model slot}. A model
-slot is a mediation entity in charge of giving
-access to external elements using
-a \emph{technology adapter}\footnote{It is a reusable library that defines
-connections between the \FML execution engine and a particular
-technological space.}.
+A UML like representation of a simplified metamodel of \FML is provided in Figure~\ref{fig:mm}.
+\FML is language designed to define models as virtual models.
+A \emph{virtual model} is composed of a set of \emph{concepts}, while itself being a \emph{concept}.
+Hence, \emph{virtual models} are structuring units forming architectures while \emph{concepts} are the core entities.
+A concept has a set of \emph{properties} and \emph{behaviors}.
+\emph{Properties} can be either simple variables, roles (pointers to a modeling element in a \emph{virtual model} or an external technology-specific model), or properties bound to complex control graphs.
+\FML is designed to define not only the structure of virtual models but also also the collection of actions that can be performed on them.
+These actions are called \emph{behaviors} and are composed with behavioural primitives called \emph{EditionAction} (core \FML primitives or behavioural primitives exposed by a specific technology).
+These \emph{behaviors} can either be called or triggered by events.
+The reactive behaviors are mainly useful when a federated model evolution needs to trigger a computation.
+Note that we do not exploit this possibility in the challenge.
 
-\noteSylvain{Retravailler le texte pour parler de tous les concepts qui sont sur la figure}
+A parallel to object-oriented approach can be useful to understand \FML\footnote{Some aspects of \FML do not exist in object oriented approach.}.
+A concept corresponds to a class, its properties to the attributes of the class and its behaviors to the methods of the class.
+These properties have types defining the kind of value the role will point at runtime.
+Whenever a type external to the federation space is used, one needs to use a \emph{model slot}.
+A model slot is a mediation entity in charge of giving access to external elements using a \emph{technology adapter}\footnote{It is a reusable library that defines connections between the \FML execution engine and a particular technological space.}.
 
-\FML is designed to define not only the structure of virtual models but
-also also the collection of actions that can be performed on them.
-These actions are called \emph{behaviors} and can either be called (like usual methods) or triggered by events.
-The reactive behaviors are mainly useful when a federated model evolution needs to trigger a computation.
-%We do not exploit this possibility in the challenge.
+% Sylvain : relire et vérifier
 
-%\noteSylvain{Tentative de représenter ici un sous-ensemble de BPMN en montrant la capacité du language FML à faire de la modélisation conceptuelle d'une part, mais aussi à fédérer des nouveaux concepts avec un modèle et un métamodèle existant.}
-
-%\noteSylvain{On explique dans la suite que dans le challenge, on ne va pas exploiter l'aspect fédération de modèle (sauf pour la partie diagramme - c'est expliqué plus haut également - ).}
-
-Figure \ref{fig:BPMNSubsetExample} shows...
-
-%\noteSM{We use the current Figure 3 to answer reviewer 1 in the response letter just as to show how the model federation works, but not here.}
-
-\noteSylvain{Rédiger l'explication...}
+\noteSylvain{A la fin s'arranger pour que la figure 3 (fig:BPMNSubsetExample) et le listing soient cote a cote}
 
@@
 \end{figure}
+
+Figure \ref{fig:BPMNSubsetExample} shows how to realize a small subset of BPMN with Openflexo/FML.
+Concretely, we have designed a \textit{virtual model} (SimplifiedBPMN) with 3 \textit{concepts} representing a process and its contained tasks.
+An instance of this \textit{virtual model} is presented in the bottom of the figure, showing instances with their instanceOf relationships with their model.
+
+Following listing shows the textual representation of the above described SimplifiedBPMN \textit{virtual model}.
+
+% Sylvain : relire et vérifier
 

### 2021-10-07 responseLetter.tex
Updates from Overleaf
synchro

@@
 \bigskip
-Sylvain, Guillaume, Caine, Joel, Jean-Christophe, Salvador, Fabien and Antoine
+Sylvain, Joel, Jean-Christophe, Antoine, Fabien and Salvador
 
@@
 \begin{response}{in Figure 9 you show the model editor. Could you please also add a screenshot of the metamodel editor?} 
-TODO.
+TODO. Figure 9 now shows a screen capture of the FML metamodel editor.
 \end{response}
@@
 The text mentions the (linguistic) instantiation mechanism of the FML Language… how does it work exactly?}
-We hope we have improved the explanation. \texttt{InstanceInstance} has been replace with \texttt{ObjectInstance}. TODO.
+We hope we have improved the explanation. \texttt{InstanceInstance} has been replaced with \texttt{ObjectInstance}. TODO.
 \end{response}
@@
 
-Concerning the fact that "MLM is the key to the ability of Openflexo to create tooling", the reviewer probably read "MLM" instead of "FML" (Flexo Modeling Language).
+Concerning the fact that "MLM is the key to the ability of Openflexo to create tooling", the reviewer probably read "MLM" instead of "FML" (Federation Modeling Language).
 \end{response}

### 2021-10-07 3-introduction.tex
synchro

@@
 Our solution is based on the model federation infrastructure.
-Concretely, we use Openflexo and its internal Flexo Modeling Language (\FML), a language to create, link and manage virtual models, in order to solve the MULTI process challenge while the federation aspect is used solely so as to provide tooling for the resulting process language.
+Concretely, we use Openflexo and its internal Federation Modeling Language (\FML), a language to create, link and manage virtual models, in order to solve the MULTI process challenge while the federation aspect is used solely so as to provide tooling for the resulting process language.
 Our solution, which is fully implemented and executable, meets all the requirements.
@@
 
-\textcolor{red}{Our modeling approach is based on a language (\FML) which provides the (linguistic) concepts that are used to define  the multi-levels models and meta-models that are needed.
+Our modeling approach is based on a language (\FML) which provides the (linguistic) concepts that are used to define  the multi-levels models and meta-models that are needed.
 \FML allows to define models as first class entities (which we call VirtualModel) and to link them together (by extension).
@@
 \begin{itemize}
-    \item Define, extend and adapt models, keeping the history, to meet {\color{blue}the} %changing
+    \item Define, extend and adapt models, keeping the history, to meet the %changing
     modeling requirements of the case study,
@@
 \end{itemize}
-This methodological choice allows us to adapt the models and keep track of their evolution each time the requirements evolve.}
+This methodological choice allows us to adapt the models and keep track of their evolution each time the requirements evolve.
 
-\textcolor{red}{Contrary to a classical multi-level approach, in which the notion of multi-level is integrated in the language and must therefore be respected, our approach allows to build multi-levels on demand, and with the appropriate form.
-Without multilevel-specific operators/concepts, the "custom" construction must be done explicitly during the model development and evolution process.}
+Contrary to a classical multi-level approach, in which the notion of multi-level is integrated in the language and must therefore be respected, our approach allows to build multi-levels on demand, and with the appropriate form.
+Without multilevel-specific operators/concepts, the "custom" construction must be done explicitly during the model development and evolution process.
 
-\textcolor{red}{Our answer to the challenge is therefore not an answer using a classical multilevel approach that would allow an efficiency comparison with another classical multilevel answer, but an approach that proposes another organization of multilevels through the translation of needs into levels over time and the possibility to build an evolving hierarchy of adapted models.
-This is made possible by (1) the reification of the notion of model and the possibility to interconnect them (specialization) and (2) the freedom of meta-modeling (low model conformity, flexible meta-modeling).}
+Our answer to the challenge is therefore not an answer using a classical multilevel approach that would allow an efficiency comparison with another classical multilevel answer, but an approach that proposes another organization of multilevels through the translation of needs into levels over time and the possibility to build an evolving hierarchy of adapted models.
+This is made possible by (1) the reification of the notion of model and the possibility to interconnect them (specialization) and (2) the freedom of meta-modeling (low model conformity, flexible meta-modeling).
 

### 2021-10-07 4-technology.tex
synchro

@@
 \emph{Properties} can be either simple variables, roles (pointers to a modeling element in a \emph{virtual model} or an external technology-specific model), or properties bound to complex control graphs.
-\FML is designed to define not only the structure of virtual models but also also the collection of actions that can be performed on them.
-These actions are called \emph{behaviors} and are composed with behavioural primitives called \emph{EditionAction} (core \FML primitives or behavioural primitives exposed by a specific technology).
+\FML is designed to define not only the structure of virtual models but also the collection of actions that can be performed on them.
+These actions are called \emph{behaviors} and rely on behavioural primitives called \emph{EditionAction}.
 These \emph{behaviors} can either be called or triggered by events.
-The reactive behaviors are mainly useful when a federated model evolution needs to trigger a computation.
+The reactive behaviors are useful when a federated model evolution needs to trigger a computation.
 Note that we do not exploit this possibility in the challenge.
@@
 These properties have types defining the kind of value the role will point at runtime.
-Whenever a type external to the federation space is used, one needs to use a \emph{model slot}.
+Whenever an element external to the federation space is used, one needs to use a \emph{model slot}.
 A model slot is a mediation entity in charge of giving access to external elements using a \emph{technology adapter}\footnote{It is a reusable library that defines connections between the \FML execution engine and a particular technological space.}.
 
+When the \FML execution engine runs a federation, it creates virtual
+model instances containing concept instances. Some concept instances
+are connected to external elements through model slot instances.
+
 % Sylvain : relire et vérifier
-
-\noteSylvain{A la fin s'arranger pour que la figure 3 (fig:BPMNSubsetExample) et le listing soient cote a cote}
 
@@
 
-Figure \ref{fig:BPMNSubsetExample} shows how to realize a small subset of BPMN with Openflexo/FML.
-Concretely, we have designed a \textit{virtual model} (SimplifiedBPMN) with 3 \textit{concepts} representing a process and its contained tasks.
-An instance of this \textit{virtual model} is presented in the bottom of the figure, showing instances with their instanceOf relationships with their model.
-
-Following listing shows the textual representation of the above described SimplifiedBPMN \textit{virtual model}.
-
-% Sylvain : relire et vérifier
+To illustrate this, figure \ref{fig:BPMNSubsetExample} shows how to realize a small subset of BPMN with Openflexo/FML.
+Concretely, we have designed a \textit{virtual model} (\texttt{SimplifiedBPMN}) with 3 \textit{concepts} representing a process and its contained tasks.
+An instance of this \textit{virtual model} is presented in the bottom of the figure, showing instances with their \textsf{instanceOf} relationships with their model.
+%
+The following listing shows the \FML code of this \textit{virtual model}.
 
 \begin{lstlisting}
-// A simplified BPMN metamodel exposing
-// both concepts BPMNProcess and BPMNTask
+// A simplified BPMN metamodel exposing the
+// concepts BPMNProcess and BPMNTask
 model SimplifiedBPMN {
@@
     // A basic constructor
-    create(String name) {
-      this.name = name;
-    }
+    create(String name) { this.name = name; }
   }
@@
     // A basic constructor
-    create(String name) {
-      super(name);
-    }
-    // Describe here other properties and
-    // behaviours for BPMNProcess concept
+    create(String name) { super(name); }
+    // Other properties and behaviours of the
+    // BPMNProcess concept
     ...
@@
       // A basic constructor
-      create(String name) {
-        super(name);
-      }
+      create(String name) { super(name); }
       // Encodes execution of task
@@
       }
-      // Describe here other properties and
-      // behaviours for BPMNTask concept
+      // Other properties and behaviours of the
+      // BPMNTask concept
       ...
@@
 
-    // Other core concepts
+    // Other concepts
   }
@@
 
-When the \FML execution engine runs a federation, it creates virtual
-model instances containing concept instances. Some concept instances
-are connected to external elements through model slot instances.
+\noteSylvain{A la fin de la rédaction, s'arranger pour que la figure 3 (fig:BPMNSubsetExample) et le listing soient cote a cote, et aussi que le listing tienne sur une seule page}
 
@@
 
-Finally, tools have their own model. We have taken advantage of
-Openflexo features to build, in parallel with the models  required by the challenge, a drawing tool that makes our
+Finally, tools have their own model. We have taken advantage of Openflexo features to build, in addition to the models  required by the challenge, a drawing tool that makes our
 solution (partially) executable.

### 2021-10-07 5-analysis.tex
synchro

@@
 
-\item The vertical axis is viewed as the linguistic instantiation, relative to the use of the \FML language.
-The virtual model, defined as the core definition metamodel, generates a model founded on the instantiation mechanism of the \FML Language.
+\item The vertical axis is viewed as the linguistic instantiation.
+The Metamodel (a virtual model) can be instantiated as a model thanks to the instantiation mechanism of \FML.
 This mechanism is similar to the classical object/instance mechanism of the object languages but without any constraint on the referenced virtual model, \ie metamodel.
-The set of resulting instances comes from several virtual models
-of any abstraction level as detailed in the section. \ref{sec:AcmeSoftwareDevelopmentProcess}
+The set of resulting instances comes from several virtual models of any abstraction level as detailed in the section~\ref{sec:AcmeSoftwareDevelopmentProcess}
 
-%{Version Joel, à prendre ou à laisser} \item The vertical axis is viewed as the linguistic specialization, relative to the use of the \FML language.
+%{Version Joel, à prendre ou à laisser} 
+%\item The vertical axis is viewed as the linguistic specialization, relative to the use of the \FML language.
 The virtual model, defined as the core definition metamodel, generates a model founded on the instantiation mechanism of the \FML Language or can be extended by another virtual model.
@@
 Finally, as explained previously, the defined level for the Acme  and the XSure models is the result of the instantiation of virtual models.
-In our approach this virtual model instance level can't be specialized or extended but all the other virtual models could be extended by any concepts, as a new federated model.
+In our approach this virtual model instance level can not be specialized or extended but all the other virtual models could be extended by any concepts, as a new federated model.
 Also, the instance level can be instantiated from any virtual model, which represents any metamodel level.

### 2021-10-07 7-requirementssatisfaction.tex
adding notes to the response letter
synchro

--- 2021-05-31 7-requirementssatisfaction.tex
@@
 \begin{itemize}
-    \item \emph{Conceptualization} (\textit{Conc.}): a concept carries the semantics exposed by the requirement (for example for \textbf{P1}: \textit{process type} is conceptualized in a concept \texttt{ProcessType}).
+    \item \emph{Conceptualization} (\textit{Conc.}): a concept is created to fulfill or meet the requirement (for example for \textbf{P1}: \textit{process type} is conceptualized in a concept \texttt{ProcessType}).
 
@@
 
-    \item \emph{Ontological instantiation} (\textit{O.Inst.}): fulfillment of a requirement is obtained by the relation between an instance and its type definition instance (for example for \textbf{S3} : \textit{developer} has the dual concept nature in \textit{Acme metamodel} and instance nature in \textit{Acme model}).
+    \item \emph{Ontological instantiation} (\textit{O.Inst.}): fulfillment of a requirement is obtained by the relation between an instance and its type definition instance (for example the notion of \textit{developer} of \textbf{S3} is implemented by both a concept in \textit{Acme metamodel} and an instance in \textit{Acme model}).
 
@@
 \end{itemize}
+
+\noteSylvain{updated}
 

### 2021-10-07 7-requirementssatisfaction.tex
synchro

@@
 
-    \item \emph{Ontological instantiation} (\textit{O.Inst.}): fulfillment of a requirement is obtained by the relation between an instance and its type definition instance (for example the notion of \textit{developer} of \textbf{S3} is implemented by both a concept in \textit{Acme metamodel} and an instance in \textit{Acme model}).
+    \item \emph{Ontological instantiation} (\textit{O.Inst.}): fulfillment of a requirement is obtained by the relation between an instance and its type definition instance (for example the notion of \textit{developer} of \textbf{S3} is implemented by both a concept in \textit{Acme metamodel} and an instance in \textit{Acme model}, see bottom left of Figure~\ref{fig:AcmeFullArchitecture}).
 
@@
 
-\noteSylvain{updated}
+%\noteSylvain{updated}
 

### 2021-10-07 8-discussion.tex
synchro
des questions + maj

@@
 our solution.
-\noteJC{updated}
 
@@
   more or fewer levels depending on the variations of the use case. However,
-  the number of levels is related to the problem, making the approach fixed (to
-  3).
+  the number of levels is related to the problem, our solution uses 3 levels.
 
@@
   \texttt{Developer} concept from \texttt{Acme} metamodel specializes the
-  \texttt{ActorType} of the base metamodel, and the \texttt{Developer} of the
+  \texttt{ActorType} of the base metamodel, and the \texttt{Developer} {\color{red}?} of the
   \texttt{Acme} model is an instance of the metamodel concept.
 %  \textcolor{red}{Phrase faite a relire sur Developer\dots}
-\noteJC{updated}
 
@@
   how the requirements are captured and how behaviors are implemented with
-  \FML: for instance, \texttt{isAuthorizedActor} behavior is needed to enforce
+  \FML. For instance, \texttt{isAuthorizedActor} behavior is needed to enforce
   \textbf{P17} requirement, leading to writing \FML code in \texttt{TaskType}
   and \texttt{Task}. Note that in our approach, the metamodels are built in an
-  ad hoc way together with the models (co-construction), therefore the user
-  also validates on an ad hoc basis.
-\noteJC{updated}
+  {\color{red}ad hoc} way together with the models (co-construction), therefore the user
+  also validates on an {\color{red}ad hoc} basis.
 
@@
   desired and considered as bad design.
-\noteJC{updated}
 

### 2021-10-07 10-conclusions.tex
Polishing related work.
Updates from Overleaf

--- 2021-05-31 10-conclusions.tex
@@
 
-We fulfill all the challenge requirements and we propose a tooling that demonstrates the usability of our solution.
-We developed an ad hoc solution, including models and metamodels, thanks to the flexibility provided by the metametamodeling infrastructure offered by Openflexo.
+We fulfill all the challenge requirements and we propose a tooling that
+demonstrates the usability of our solution. We developed an ad hoc solution,
+including models and metamodels, thanks to the flexibility provided by the
+metametamodeling infrastructure offered by Openflexo.
+This infrastructure helped us to overcome the limitations of the strict modeling paradigm mentioned in Section \ref{sec:introduction}, this is, the lack of support for different forms of classifications and for the duality type/object.
+In that sense, our solution seamlessly integrates the type-object pattern which allows us to dynamically create different \emph{instances} of model elements (e.g., \emph{TaskType}, among others) and use them for typing other model elements.
+This, together with the FML core support for (level agnostic) classical linguistic instantiation, permitted us to merge linguistic and ontological instantiation in order to provide a satisfactory solution to the Multi-Level Process challenge.
 
-As a future work we envision to explore a number of alternative solutions: first, we intend to propose another solution based on a more level-agnostic approach such as Clabjects.
-Instead of two levels defined by Type and Instance concepts, we could have a single concept mixing a type part and an instance part.
-This approach would probably ease adaptation if the problem specifications evolve.
-Secondly, another possible approach is the use of the free modeling tool proposed by Openflexo.
-The solution would be developed from examples (Acme and XSure processes for instance) from which models would be identified.
-Finally, the realization of this challenge highlighted the interest of integrating specific behaviors for the management of multi-level concepts in the Openflexo infrastructure.
-%We intend to do so.
+\noteSM{Say that the flexibility of our approach coupled with a methodology allows us to achieve multi/level capabilities without specific tooling in order to solve problems which require it.}
+
+As a future work we envision to explore a number of alternative solutions:
+first, we intend to propose another solution based on a more level-agnostic
+approach such as Clabjects. Instead of two levels defined by Type and Instance
+concepts, we could have a single concept mixing a type part and an instance
+part. This approach would probably ease adaptation if the problem
+specifications evolve. Secondly, another possible approach is the use of the
+free modeling tool proposed by Openflexo. The solution would be developed from
+examples (Acme and XSure processes for instance) from which models would be
+identified.
+
+Finally, the realization of this challenge highlighted the interest
+of integrating specific behaviors for the management of multi-level concepts in
+the Openflexo infrastructure. %We intend to do so.

### 2021-10-07 8-discussion.tex
Updates from Overleaf

--- 2021-10-06 8-discussion.tex
@@
   However, it would not be a generic mechanism, making it difficult to reuse
-  for another problem.  Plus, programming the notion of \emph{depth} would
-  probably imply that a virtual model knows its sub-elements, which is not
-  desired and considered as bad design.
-\noteJC{updated}
+  for another problem.
 

### 2021-10-07 9-relatedworks.tex
adding notes to the response letter
Updates from Overleaf

@@
 A solution is provided with MetaDepth~\parencite{metadepth}, which supports modeling with any number of levels, dual ontological/linguistic typing and deep characterization through potency~\citep{potency}.
-\textcolor{red}{For their solution to the multi-level process problem they use three levels.
+For their solution to the multi-level process problem they use three levels.
 The first level describes generic processes, the second level software engineering processes, and the third level, software engineering enacted processes.
@@
 the grouping and organization of concepts (e.g., a model for general process, concepts, an extension for software engineering process) but it replaces the use of clabjets and potency with the use of the type-object pattern~\citep{typeObject}.
-The so called \emph{linguistic extensions} are natively supported in Openflexo/FML.}
+The so called \emph{linguistic extensions} are natively supported in Openflexo/FML.
 
@@
 \cite{multiecore2019} use MultEcore \citep{multecore2016} to provide a solution to the challenge.
-\textcolor{red}{MultEcore uses, (un)pluggable linguistic levels (e.g, there is not a \emph{fixed} core metamodel defining the concept of clabjet), an extension of the two-level cascading technique \citep{atkinson2005concepts} (this is achieved with model transformations to transform instance models into instantiable models) and potency.
+MultEcore uses, (un)pluggable linguistic levels (e.g, there is not a \emph{fixed} core metamodel defining the concept of clabjet), an extension of the two-level cascading technique \citep{atkinson2005concepts} (this is achieved with model transformations to transform instance models into instantiable models) and potency.
 Regarding the challenge, their solution uses 4 levels.
@@
 With respect to our solution, the authors use more models, as we require less refinement steps.
-However accidental complexity is introduced by our approach as we need additional constructs and constraints for the typing relations which are otherwise built-in in MultEcore.}
+However accidental complexity is introduced by our approach as we need additional constructs and constraints for the typing relations which are otherwise built-in in MultEcore.
 
@@
 %JOel : pour simplifier la description j'ai juste mis en commentaire le détail des 4 %niveaux et ca supprime P19. A valider}.
-%\noteSM{How do we deal with P19 - actually P19 does not exist anymore}
+% Vu. J'ai effacé la mention du P19.
 
@@
 As with Openflexo/\FML, they do not use explicit (numbered) levels nor potency.
-However, unlike Openflexo, DeepTelos integrates a multi-level modeling specific construct similar to the PowerType \parencite{atkinson2001essence} pattern they call \emph{most general instances} \textcolor{red}{which the authors use to simulate potency.
+However, unlike Openflexo, DeepTelos integrates a multi-level modeling specific construct similar to the PowerType \parencite{atkinson2001essence} pattern they call \emph{most general instances} which the authors use to simulate potency.
 Models in DeepTelos are organized in (tree-like) module hierarchies, where submodules can see all the concepts defined in parent modules.
@@
 This sub-module contains in turn a sub-module representing the coding process type whereas concrete coding processes are represented in subsequent sub-modules.
-There are two main differences between the DeepTelos and Openflexo/\FML: 1) we use the type-object pattern instead of the powertype pattern; 2) DeepTelos reify their support for multi-level modeling while ours remains ad-hoc.}
+There are two main differences between the DeepTelos and Openflexo/\FML: 1) we use the type-object pattern instead of the powertype pattern; 2) DeepTelos reify their support for multi-level modeling while ours remains ad-hoc.
 
@@
 Finally, \cite{dmla2019} use the Dynamic Multi-Layer Algebra (DMLA) \parencite{dmla2017} in order to solve the process challenge.
-DMLA is a level-blind modeling framework which is fully customizable (\eg different types of instantiation may be implemented) \textcolor{red}{and includes support for deep characterization (instantiation may refer to any other element disregarding hierarchy levels and a sort of potency in the form of fluid metamodeling at the entity level exists).} Their proposed solution separates the process challenge in two separate domains, namely, the task definition domain and the process definition domain.
+DMLA is a level-blind modeling framework which is fully customizable (\eg different types of instantiation may be implemented) and includes support for deep characterization (instantiation may refer to any other element disregarding hierarchy levels and a sort of potency in the form of fluid metamodeling at the entity level exists).
+Their proposed solution separates the process challenge in two separate domains, namely, the task definition domain and the process definition domain.
 Wrappers are used in order to reuse tasks in processes.

### 2021-10-13 8-discussion.tex
Updates from Overleaf

@@
   \texttt{Developer} concept from \texttt{Acme} metamodel specializes the
-  \texttt{ActorType} of the base metamodel, and the \texttt{Developer} of the
-  \texttt{Acme} model is an instance of the metamodel concept.
+  \texttt{ActorType} of the base metamodel. The \texttt{Software Engineering Process} of the
+  \texttt{Acme} model is an instance of \texttt{ProcessType} in the metamodel.
 %  \textcolor{red}{Phrase faite a relire sur Developer\dots}
-\noteJC{updated}
 
@@
   how the requirements are captured and how behaviors are implemented with
-  \FML: for instance, \texttt{isAuthorizedActor} behavior is needed to enforce
+  \FML. For instance, \texttt{isAuthorizedActor} behavior is needed to enforce
   \textbf{P17} requirement, leading to writing \FML code in \texttt{TaskType}
   and \texttt{Task}. Note that in our approach, the metamodels are built in an
-  ad hoc way together with the models (co-construction), therefore the user
-  also validates on an ad hoc basis.
-\noteJC{updated}
+  {\color{red}ad hoc} way together with the models (co-construction), therefore the user
+  also validates on an on-demand basis.
 

### 2021-10-13 10-conclusions.tex

@@
 
-\noteSM{Say that the flexibility of our approach coupled with a methodology allows us to achieve multi/level capabilities without specific tooling in order to solve problems which require it.}
+\textcolor{red}{The features of \FML enable a great modelling flexibility which, coupled with an appropriate methodology, allow us to achieve
+the multi/level capabilities without dedicated tooling in order to solve problems which require it. We use virtual models to "implement" levels on demand.}
+%\noteSM{Say that the flexibility of our approach coupled with a methodology allows us to achieve multi/level capabilities without specific tooling in order to solve problems which require it.}
 

### 2021-10-13 2-abstract.tex
Updates from Overleaf

@@
 We present a solution based on model federation and discuss the advantages and limitations of using this approach for multi-level modeling.
-Concretely, we use virtual models and more precisely the Flexo Modeling Language ({FML}) that serves to describe them as the main building block in order to solve the process modeling challenge whereas the federation aspect is used as a means to provide editing tooling for the resulting process language.
+Concretely, we use virtual models and more precisely the Federation Modeling Language ({FML}) that serves to describe them as the main building block in order to solve the process modeling challenge whereas the federation aspect is used as a means to provide editing tooling for the resulting process language.
 Our solution is fully implemented with the Openflexo framework.

### 2021-10-13 responseLetter.tex

@@
 
-We have been more explicit to explain how we use the FML language at the end of the introduction, and in various figures. TODO.
+%We have been more explicit to explain how we use the FML language at the end of the introduction, and in various figures. TODO.
 
-\todo[inline]{We have enhanced the presentation of both, Openflexo/FML and our solution to the challenge so that: 1) the techniques we use to \emph{achieve} multi-level capabilities are more explicit; and, 2) the advantages and disadvantages of using them in order to fulfil the challenge requirements become clearer.
-Notably, we have added many details in the final paragraphs of the introduction section, improved the Section 2, devoted to the description of OpenFlexo, refactored figures X, Y, Z.}
+We have enhanced the presentation of both, Openflexo/FML and our solution to the challenge so that: 1) the techniques we use to \emph{achieve} multi-level capabilities are more explicit; and, 2) the advantages and disadvantages of using them in order to fulfil the challenge requirements become clearer.
+Notably, we have added many details in the final paragraphs of the introduction section, improved the Section 2, devoted to the description of OpenFlexo, refactored figures 2, 4, 5, 6, 7 and 9 and, added figures 3 and 10.
 \end{response}
@@
 You should also respond to other concerns reviewers raise, either by making changes to the submission or in the form of an authors' response.}
-The related work section has been modified to show that we are not the first conventional response with multi-level emulation, and to improve comparisons and limitations of different approaches.\todo[inline]{I do not understand this answer.
-We are saying that other approaches are also unconventional?}
+The previous answer describes the improvements of our revision.
+Additionally, the related work section has been modified to show that we are not the first "conventional" response with multi-level emulation, and to improve comparisons and limitations of different approaches ("conventional" or not).
+
+%\todo[inline]{I do not understand this answer. We are saying that other approaches are also unconventional?}
 \end{response}
@@
 
-\todo[inline]{I've refactored the paragraph above. To be verified.}
+%\todo[inline]{I've refactored the paragraph above. To be verified.}
 
@@
 Please note that EMISAJ does not have a page limitation, so no worries if that should require more space.}
-We've changed figure 2. It shows \texttt{Type} and the main concepts used by FML. TODO. Explanation of figure has been improved. 
+We've changed figure 2. It shows \texttt{Type} and the main concepts used by FML. Explanation of figure has been improved. 
 \end{response}
@@
 When reading the paper, I actually marked the references to types in yellow in my copy - ie from ProcessType to TaskType and from ActorType to ActorType (on the attribute levels) - this helped me to see more clearly where you have references from attributes to types - but that's just my way of doing it.}
+
 TODO.
@@
 
-We tried to illustrate subsections of section 6 by giving examples, by referencing figures and sections.
+We tried to illustrate subsections of section 6 by giving examples, by referencing figures and sections (subsections 6.2, 6.3, 6.4, 6.6, 6.7 and 6.11 have been updated).
 
@@
 }
-TODO.
+\todo[inline]{While it is true that model slots and other features of Openflexo/FML are not used in order to build our solution,  we believe including them in our description help readers understand the framework.
+We therefore choose to keep the description as is.}
 \end{response}

### 2021-10-13 10-conclusions.tex

@@
 
-\textcolor{red}{The features of \FML enable a great modelling flexibility which, coupled with an appropriate methodology, allow us to achieve
-the multi/level capabilities without dedicated tooling in order to solve problems which require it. We use virtual models to "implement" levels on demand.}
+The features of \FML enable a great modelling flexibility which, coupled with an appropriate methodology, allow us to achieve
+the multi/level capabilities without dedicated tooling in order to solve problems which require it. We use virtual models to "implement" levels on demand.
 %\noteSM{Say that the flexibility of our approach coupled with a methodology allows us to achieve multi/level capabilities without specific tooling in order to solve problems which require it.}

### 2021-10-13 8-discussion.tex

@@
   \textbf{P17} requirement, leading to writing \FML code in \texttt{TaskType}
-  and \texttt{Task}. Note that in our approach, the metamodels are built in an
-  {\color{red}ad hoc} way together with the models (co-construction), therefore the user
+  and \texttt{Task}. Note that in our approach, metamodels are built in an
+ ad hoc way together with the models (co-construction), therefore the user
   also validates on an on-demand basis.
@@
 and that it implements the \texttt{languages} relation.
-\noteJC{updated}
+%\noteJC{updated}

### 2021-10-14 responseLetter.tex
Updates from Overleaf

@@
 An overarching theme is the request for a more explicit description of how the multi-level modelling emulation is achieved and what its general advantages and limitations are, including a clarification of what the implications would be of supporting more "levels" than the challenge required.}
-
 %We have been more explicit to explain how we use the FML language at the end of the introduction, and in various figures. TODO.
 
-We have enhanced the presentation of both, Openflexo/FML and our solution to the challenge so that: 1) the techniques we use to \emph{achieve} multi-level capabilities are more explicit; and, 2) the advantages and disadvantages of using them in order to fulfil the challenge requirements become clearer.
+We have improved the presentation of both, Openflexo/FML and our solution to the challenge so that: 1) the techniques we use to \emph{achieve} multi-level capabilities are more explicit; and, 2) the advantages and disadvantages of using them in order to fulfil the challenge requirements become clearer.
 Notably, we have added many details in the final paragraphs of the introduction section, improved the Section 2, devoted to the description of OpenFlexo, refactored figures 2, 4, 5, 6, 7 and 9 and, added figures 3 and 10.
@@
 You should also respond to other concerns reviewers raise, either by making changes to the submission or in the form of an authors' response.}
-The previous answer describes the improvements of our revision.
-Additionally, the related work section has been modified to show that we are not the first "conventional" response with multi-level emulation, and to improve comparisons and limitations of different approaches ("conventional" or not).
+
+\textcolor{orange}{In the answer to the previous editor's comment we describe the improvements of our revision, which includes an emphasis in discussing trade-offs related to the emulation of multi-level modeling by using the Openflexo/FML and the type-object pattern.
+Additionally, the related work section have been extended in order to enhance the comparison between our approach and previous solutions to the challenge.}
 
@@
 Therefore, in the article we choose not to give this level of details and focus on the left part of the figure.
-
-%\todo[inline]{I've refactored the paragraph above. To be verified.}
+% Fabien OK pour moi
 
@@
 Please note that EMISAJ does not have a page limitation, so no worries if that should require more space.}
-We've changed figure 2. It shows \texttt{Type} and the main concepts used by FML. Explanation of figure has been improved. 
+
+\textcolor{orange}{We agree with the reviewer.
+The previous version of the figure depicting the main concepts of the FML language (Figure 2) was too succinct.
+We have fully refactored it so that now important concepts such as \emph{Type} are shown.
+We have correspondingly improved the text explaining the figure too.
+We hope this changes make understanding the core concepts of the FML language easier.}
+
+%We've changed figure 2. It shows \texttt{Type} and the main concepts used by FML. %Explanation of figure has been improved. 
 \end{response}
@@
 
-TODO.
+\textcolor{orange}{The only two concepts from Figured 2 represented elsewhere in the paper are the \emph{VirtualModel} and \emph{FlexoConcept} elements, which are represented in blue in figures 3, 5, 6, 7 and 9.
+Then, we use the brown color to represent the instantiation of these concepts.
+We hope this makes the connection between the Figure 2 and the other representations in the paper clearer.
+Regarding the notation, it is indeed similar to UML but it has different semantics as explained in the paper.}
+
+%In the paper, we only represent VirtualModel and FlexoConcept visible in figure 2. %Both are represented in blue in figures 3, 5, 6, 7 and 9.
+%We use brown to represent instantiations of concepts. Doing so, we hope the %connection between FML and representations are more understandable.
+
+%Note also that we use diagram notation close to UML to be easily understood, but that %it is not UML.
+
+%TODO.
 Faire le lien entre figure 2 et les représentations des figures 3, 4, 5, 6 et 7.
@@
 Ie I could then infer exactly, how elements from the concrete model have been realized in the background.}
-TODO. 
+
+\textcolor{orange}{Thanks for pointing out this inconsistency. We have completed Figure 6 (now Figure 7) so that it contains all the necessary elements for Figure 7 (now Figure 8).}
+
+%We have completed the figure to make it consistent with the example process.
+
 \end{response}
@@
 \begin{response}{in Figure 9 you show the model editor. Could you please also add a screenshot of the metamodel editor?}
-TODO. Figure 9 now shows a screen capture of the FML metamodel editor.
+
+\textcolor{orange}{Our manuscript includes now a new figure (Figure 10) that shos a screen capture of the FML metamodel editor.}
+%Figure 10 now shows a screen capture of the FML metamodel editor.
 \end{response}
@@
 \begin{response}{There are some minor typos throughout the manuscript (eg also regarding the reference Jeusfeld2019; choose -> chose; axis -> axes; aN FML virtual model...; etc.)} 
-TODO.
-choose $\rightarrow$ chose OK ; axis $\rightarrow$  axes OK ; je ne comprends pas pour Jeusfeld2019
-(deeptelos2019) ni pour aN FML (je dirais ``a FML'', mais je puis me tromper)
+Done.
 \end{response}
@@
 This fact is possibly the most serious advantage and disadvantage of the paper at the same time.
-Advantage, since it brings us a completely new, different way of thinking, and disadvantage since it does not really help in categorizing/unifying multi-level methods and – for me – it also feels a little bit off-topic in this special issue.
-At this point, it should be mentioned that the authors managed to cover/solve all requirements of the challenge, thus the paper, as a solution to the challenge fits well.} We introduced at the end of the introduction a description of our approach.
-In a word, we use FML (linguistic level) as a tool to build models and meta-models that allows multi-level modelling.
+Advantage, since it brings us a completely new, different way of thinking, and disadvantage since it does not really help in categorizing/unifying multi-level methods and -- for me -- it also feels a little bit off-topic in this special issue.
+At this point, it should be mentioned that the authors managed to cover/solve all requirements of the challenge, thus the paper, as a solution to the challenge fits well.}
+
+\textcolor{red}{We introduced at the end of the introduction a description of our approach.
+In a word, we use FML (linguistic level) as a tool to build models and meta-models that allows multi-level modelling emulation.
 The evolution of models (multi-levels) is encoded as a hierarchy of models and meta-models.
 We do multi-level modelling on demands, and propose an organization guided by models evolution.
-\todo{Answering this is going to be tricky!}
+}
+
+\textcolor{red}{Propale de Fabien pour complèter la réponse ici.
+We have added a discussion (où?) on the fact that our approach reifies the levels using "generic" models and rely on the process of construction to implement the levels of the challenge.
+We hope, it makes it easier to see that it is a contribution to multi-level domain.}
+
+%\todo{Answering this is going to be tricky!}
 \end{response}
@@
 It is not clear, how federation works in this case.}
-We have improved the description and we hope this is much clearer. TODO.
+We use federation only for the tooling part of our answer.
+In practice, we only use the reification of models through VirtualModel to emulate multi-level as shown in figure 4.
+We have improved the description and we hope this is much clearer.
 \end{response}
@@
 The text mentions the (linguistic) instantiation mechanism of the FML Language… how does it work exactly?}
-We hope we have improved the explanation. \texttt{InstanceInstance} has been replaced with \texttt{ObjectInstance}. TODO.
+We have improved the explanation in sections 2 and 3.
+Note that \texttt{InstanceInstance} has been replaced with \texttt{ObjectInstance}.
 \end{response}
@@
 I strongly suggest following the UML guidelines here and use rolenames (similarly to multiplicity).}
-TODO.
+\textcolor{red}{
+We have improved figure 6. The notation is close to UML, but is not UML. We hope it is understandable enough.}
 \end{response}
+
+% Fabien ok pour moi
 
@@
 What are the contexts of the constraints (to which concept are they attached to)?}
-TODO.
+MultiInstance(s) are introduced in our answer metamodel (figure 6).
+As such, their semantic is defined at the ontological level, i.e.
+in the metamodel.
+It is an ad hoc mechanism introduced to fulfill requirements.
+
+{\color{red}hum, j'ai l'impression qu'il manque un peu de détail pour répondre sur la partie multi-instance.
+Il me semble qu'en plus d'être ad hoc, le mécanisme est manuel (mais je me trompe peut être).
+Si c'est vrai, il faudrait le dire.}
+
+%On ne gère pas de multi-instances dans la sémantique FML, on gère ca seulement au niveau ontologique (FlexoConcept Instance et MultiInstance, figure 6).
+C'est un mécanisme ad-hoc utilisé pour le traitement des requirements.
+
+The constraints are defined at FlexoConcept level, and are designed of invariant being evaluated by the FML interpreter, which may raise constraint violation at run-time.
+An example of constraint is presented for requirement P9, in TaskType flexo concept (p 8).
+
 \end{response}
@@
 Is this a convention, or are such methods configured somehow to replace the usual constructors?
-According to the text, “same pattern applies” in several cases, but these patterns could be also included, because it is not completely clear to what part of the behavior “same” refers.
-In same cases, the diagram uses “/type”, while in the code it is “type” – why?
+According to the text, "same pattern applies" in several cases, but these patterns could be also included, because it is not completely clear to what part of the behavior "same" refers.
+In same cases, the diagram uses "/type", while in the code it is "type" -- why?
 The specialization of Actors is encoded by FML, wouldn’t be easier to use the built-in specialization instead?}
-TODO.
+\textcolor{red}{
+- newProcess est un vrai behaviour défini dans ProcessType (pattern factory) et permet de spécifier l'instantiation d'un Process à partir d'un ProcessType (c'est la partie réalisation de l'instantiation ontologique)
+- difficile de répondre sur le "same pattern applies"
+- /type raffine type défini à l'étage plus haut
+- pas compris la question sur la spécialisation des Actors
+}
+
 \end{response}
@@
 9.) is defined.}
-TODO.
+We have improved the description of the relationship between FML and our metamodels (use of colors in many figures). 
+The tooling uses model federation, but the description of our answer relies only on VirtualModel and FlexoConcept of FML (figure 2).
+The tooling is suggested in figure (new) 9.
+The tooling is not detailed, but the VirtualModel ProcessEditor presented in figure (new) 9 defines a lot of interactions, where texts of (old) figure 9 are found.
+
+The business logic of the diagram tooling in encoded in FML, and more particulary for this editor in ProcessEditor VirtualModel presented figure 9.
+Structure of diagramming, interactions and presentation layer (eg texts) are defined in various FlexoConcept defined in VirtualModel ProcessEditor.
+
 \end{response}
@@
 In the case of ontological instantiation, the term “dual concept nature” and “instance nature” is mentioned… what do these terms mean?}
-TODO.
+On a rectifié et ré-écrit ces bouts de texte
+TODO (Salvador).
+
+\textcolor{red}{Conceptualization means that a concept is created to fulfill or meet a requirement. }
+
 \end{response}
@@
 
-\begin{response}{Minor issues/questions:} 
+\begin{response}{Minor issues/questions: 
 \begin{itemize}
     \item Are there any tools supporting model federation besides Openflexo? 
-    \item Section 2. “In contrast to approaches that compose metamodels into a single large metamodel \item grouping all needed entities,…” - I do not see, why this is relevant here
+    \item Section 2. “In contrast to approaches that compose metamodels into a single large metamodel grouping all needed entities,…” - I do not see, why this is relevant here
     \item Figure 5 – concept Sequencing – the roles “in” and “out” are not represented by lines… why?
@@
 
-\end{itemize}
-TODO. \todo{The idea is to group all the answers to the minor questions here.}
+\end{itemize}}
+We do not know equivalent tools as far as we know.
+
+In section 2 we wanted to compare our federation approach to other that integrate all metamodels together.
+We advocate reifying models (with virtual models) helps structuring, organizing, federating models.
+
+Figure 5. We explain that we ignore these link for the sake of readability.
+
+ProgrammingLanguages is modelled as an enumeration of names (Java, Cobol, etc) from FML.
+
+You're right, Coding is actually a CodingTask, as shown in figure 7.
+
+%\todo{The idea is to group all the answers to the minor questions here.}
 \end{response}
@@
 Finally, in terms of the pattern, why should anyone care - why would should they use the pattern?}
-TODO.
+You are right. We intensively use the Type/Instance pattern. 
+We made this choice because it was easier to implement with Openflexo. 
+We make it more explicit in the text.
 \end{response}
@@
 It would be appropriate (perhaps at the end of the article) to revisit these features and to clearly demonstrate that their approach supports them and that they have concrete utility in some way in the context of the challenge.}
-
 We have added a paragraph in the Conclusions section indicating how Openflexo/FML permitted us to overcome the mentioned limitations of the strict modeling paradigm and how this was used in order to build our solution to the Multi-Level Process challenge.
@@
 There is a reified relationships between a type and its instance and therefore conditions can be defined that establish both static and dynamic semantics.
-I am confused regaring why this is not a key feature of the semantics for the challenge as it is implemented by the authors.
+I am confused regarding why this is not a key feature of the semantics for the challenge as it is implemented by the authors.
 Constraints are provided very sparingly compared to implementation of behaviour in FML.
@@
 It would be more abstract than the current 'method' style semantics.}
-TODO.
+\textcolor{red}{
+We designed our answer with the Type/Instance pattern. We think, we could have implemented it with another approach such as Clabject.
+We had to define the expected semantic of the challenge.
+To do that, FML offers constraints, but we can also define our own constraint at the ontological level (metamodel).
+We have here a trade-off.
+Reuse the constraint from FML which is efficient but may be too restricted, and define our own constraints, which is more adaptable, but has to be implemented.
+Due to the size of the problem and the expected freedom, we choose the second way.
+Le language FML permet de définir des comportements reifiant la sémantique ontologique (notamment du pattern type/instance, par exemple le comportement newProcess() dans ProcessType).
+Le language FML permet de définir également des contraintes sur les concepts.
+Les deux approches présentent chacune leurs avantages et leurs inconvénients.
+La première approche guide l'utilisateur dans une conception de modèle conforme, alors que la seconde permet plus de flexibilité (mais aussi de faire n'importe quoi).
+Tout est donc une question de choix, la question est plutot d'ordre méthodologique.
+Nous, on a plutot choisi de faire plus un outil "guidé", d'où l'utilisation de beaucoup de code impératif.
+On utilise tout de même les contraintes.
+On a essayé aussi de faire des choix multiples et variés pour montrer l'expressivité du langage FML.}
+
 \end{response}
@@
 
+\textcolor{red}{
 We tried to illustrate subsections of section 6 by giving examples, by referencing figures and sections (subsections 6.2, 6.3, 6.4, 6.6, 6.7 and 6.11 have been updated).
-
-Concerning the fact that "MLM is the key to the ability of Openflexo to create tooling", the reviewer probably read "MLM" instead of "FML" (Federation Modeling Language).
+Concerning the fact that "MLM is the key to the ability of Openflexo to create tooling", the reviewer probably read "MLM" instead of "FML" (Federation Modeling Language).}
 \end{response}
@@
 }
-\todo[inline]{While it is true that model slots and other features of Openflexo/FML are not used in order to build our solution,  we believe including them in our description help readers understand the framework.
+\textcolor{red}{While it is true that model slots and other features of Openflexo/FML are not used in order to build our solution,  we believe including them in our description help readers understand the framework.
 We therefore choose to keep the description as is.}

### 2021-10-14 6-modelpresentation.tex
Updates from Overleaf

@@
 
+Note that in all following figures, blue is used to represent \FML virtual models and flexo concepts (VirtualModel and FlexoConcept in figure \ref{fig:mm}) , see figures \ref{fig:BPMNSubsetExample}, \ref{fig:MultilevelArchitecture}, \ref{fig:LinguisticAndOntologicInstantiation}, \ref{fig:BaseMetamodel}, \ref{fig:AcmeFullArchitecture},  \ref{fig:ToolingArchitecture}, while brown is used for instances of FlexoConcept, see figures \ref{fig:BPMNSubsetExample}, \ref{fig:MultilevelArchitecture}, \ref{fig:LinguisticAndOntologicInstantiation}, \ref{fig:AcmeFullArchitecture},  \ref{fig:ToolingArchitecture}.
+% 3, 4, 5, 6, 7, 8, 9
 % Expliquer que le use case XSure ne nécessitait pas l'extension de metamodèle requise pour le usecase Acme Software Developement Process
@@
  \centering
+     \includegraphics[width=\textwidth]{Figures/ScreenshotFMLEditor.png}
+     \caption{Screenshot of the FML metamodel editor}
+    \label{fig:ScreenshotFMLEditor}
+\end{figure*}
+
+\begin{figure*}
+ \centering
      \includegraphics[width=\textwidth]{Figures/ScreenshotEnactedProcessEditor.png}

### 2021-10-14 responseLetter.tex
Updates from Overleaf

@@
 
-\textcolor{red}{We introduced at the end of the introduction a description of our approach.
-In a word, we use FML (linguistic level) as a tool to build models and meta-models that allows multi-level modelling emulation.
+\textcolor{orange}{We fully agree with the reviewer on the fact that it is the flexibility of the model federation approach what enables us to provide a solution to the multi-level process challenge while using a tool not specifically designed for multi-level modelling.
+Of course, this comes together with a number of trade-offs which we discuss in the paper (e.g., see the the end of the introduction section).}
+
+\textcolor{red}{We have included at the end of the introduction section a more detailed description of our approach.
+In a word, we use FML (linguistic level) as a tool to build models and meta-models permitting us to emulate multi-level modelling.
 The evolution of models (multi-levels) is encoded as a hierarchy of models and meta-models.
-We do multi-level modelling on demands, and propose an organization guided by models evolution.
-}
+In this sense, we do on-demand multi-level modelling, and propose an organization guided by the evolution of models.}
 
@@
 It is not clear, how federation works in this case.}
-We use federation only for the tooling part of our answer.
-In practice, we only use the reification of models through VirtualModel to emulate multi-level as shown in figure 4.
+
+\textcolor{orange}{We have refactored the Technology section in order to make it, hopefully, easier to follow.
+Apart from improvements in the text, we have notably improved Figure 2 so that now a lot more of details are visible and contributed a new figure (Figure 3) showing how the FML concepts are used in a classical modeling scenario.
+Note however that we use model federation only for the tooling part of our answer to the challenge.
+In practice, we rely solely in the reification of models through the VirtualModel concept in order to emulate multi-levels as shown in Figure 4.}
+
+%We use federation only for the tooling part of our answer.
+In practice, we only use %the reification of models through VirtualModel to emulate multi-level as shown in %figure 4.
 We have improved the description and we hope this is much clearer.
@@
 The text mentions the (linguistic) instantiation mechanism of the FML Language… how does it work exactly?}
-We have improved the explanation in sections 2 and 3.
-Note that \texttt{InstanceInstance} has been replaced with \texttt{ObjectInstance}.
+
+\textcolor{orange}{As indicated in the answer to the previous reviewer's comment we have refactored Section 2 in order to provide more details and make explanations easier to follow.
+We have done the same with }
+
+We have improved the explanation in sections 2 and 3. Note that \texttt{InstanceInstance} has been replaced with \texttt{ObjectInstance}.
 \end{response}

### 2021-10-14 responseLetter.tex
synchro

@@
 
-\textcolor{orange}{As indicated in the answer to the previous reviewer's comment we have refactored Section 2 in order to provide more details and make explanations easier to follow.
-We have done the same with }
+\textcolor{orange}{As indicated in the answer to the previous comment we have refactored Section 2 in order to provide more details and make explanations easier to follow.
+We have done the same with Section 3.
+Note that \texttt{InstanceInstance} has been replaced with \texttt{ObjectInstance}} \todo{Probably it would be nice to add here some concrete answers instead of directing the reviewer to read two sections.}
 
-We have improved the explanation in sections 2 and 3. Note that \texttt{InstanceInstance} has been replaced with \texttt{ObjectInstance}.
+%We have improved the explanation in sections 2 and 3. Note that %\texttt{InstanceInstance} has been replaced with \texttt{ObjectInstance}.
 \end{response}
@@
 I strongly suggest following the UML guidelines here and use rolenames (similarly to multiplicity).}
-\textcolor{red}{
-We have improved figure 6. The notation is close to UML, but is not UML. We hope it is understandable enough.}
+
+\textcolor{orange}{Thanks for pointing this out.
+Initially we decided against using rolenames for the sake of readability (e.g., by avoiding clutter near the association ends).
+However, we have realized now that the alternative we used hampered readability even more.
+In that sense, we use now rolenames (an exception is made for the \emph{Gateway} associations with \emph{TaskType} as explained in the caption of Figure 6)}
+
+%\textcolor{red}{
+%We have improved figure 6. The notation is close to UML, but is not UML. We hope it %is understandable enough.}
 \end{response}
@@
 The specialization of Actors is encoded by FML, wouldn’t be easier to use the built-in specialization instead?}
+
+{\color{teal}The behavior \texttt{newProcess} is defined in the \texttt{ProcessType} concept following a \emph{factory} design pattern.
+Rather than being created directly a process instance is therefore created from its type.
+This realizes the ontological instantiation.
+}
 \textcolor{red}{
-- newProcess est un vrai behaviour défini dans ProcessType (pattern factory) et permet de spécifier l'instantiation d'un Process à partir d'un ProcessType (c'est la partie réalisation de l'instantiation ontologique)
-- difficile de répondre sur le "same pattern applies"
-- /type raffine type défini à l'étage plus haut
-- pas compris la question sur la spécialisation des Actors
+%- newProcess est un vrai behaviour défini dans ProcessType (pattern factory) et permet de spécifier l'instantiation d'un Process à partir d'un ProcessType (c'est la partie réalisation de l'instantiation ontologique)
+- difficile de répondre sur le "same pattern applies" /type raffine type défini à l'étage plus haut (fd ne comprends pas la question)
+- pas compris la question sur la spécialisation des Actors (fd ne comprends pas la question)
 }
@@
 \begin{itemize}
-    \item Are there any tools supporting model federation besides Openflexo? 
-    \item Section 2. “In contrast to approaches that compose metamodels into a single large metamodel grouping all needed entities,…” - I do not see, why this is relevant here
-    \item Figure 5 – concept Sequencing – the roles “in” and “out” are not represented by lines… why?
-    \item What does the term “sub-concept” mean exactly? Please explain it before the first use
-    \item ProgrammingLanguage concept is an enumeration… how is it modelled?
-    \item “The Acme metamodel also defines a specialized task type Coding” – isn’t it CodingTask?
-    \item “The value of the property producedArtifactTypes” – is it property, or role?
+    \item Are there any tools supporting model federation besides Openflexo? \textcolor{orange}{To the best of our knowledge there are not other model federation tools comparable to Openflexo/FML.}
+    \item Section 2.
+“In contrast to approaches that compose metamodels into a single large metamodel grouping all needed entities,…” - I do not see, why this is relevant here \textcolor{orange}{Our intention was to compare the federation approach to approaches that integrate all metamodels together.
+We argue that reifying models (with virtual models) helps structuring, organizing, federating models.}
+    \item Figure 5 – concept Sequencing – the roles “in” and “out” are not represented by lines… why? \textcolor{orange}{As explained in a previous answer, we do this for the sake of readability.}
+    \item What does the term “sub-concept” mean exactly? Please explain it before the first use \todo{TODO}
+    \item ProgrammingLanguage concept is an enumeration… how is it modelled? \textcolor{orange}{ProgrammingLanguages is modelled as an enumeration of names (Java, Cobol, etc) from FML. }
+    \item “The Acme metamodel also defines a specialized task type Coding” – isn’t it CodingTask? \textcolor{orange}{Yes, this is correct. Coding is actually a CodingTask, as shown in figure 7.}
+    \item “The value of the property producedArtifactTypes” – is it property, or role? \todo{TODO}
+\end{itemize}}
 
-\end{itemize}}
-We do not know equivalent tools as far as we know.
+%We do not know equivalent tools as far as we know.
 
-In section 2 we wanted to compare our federation approach to other that integrate all metamodels together.
-We advocate reifying models (with virtual models) helps structuring, organizing, federating models.
+%In section 2 we wanted to compare our federation approach to other that integrate all %metamodels together.
+We advocate reifying models (with virtual models) helps %structuring, organizing, federating models.
 
-Figure 5. We explain that we ignore these link for the sake of readability.
+%Figure 5. We explain that we ignore these link for the sake of readability.
 
-ProgrammingLanguages is modelled as an enumeration of names (Java, Cobol, etc) from FML.
+%ProgrammingLanguages is modelled as an enumeration of names (Java, Cobol, etc) from %FML.
 
-You're right, Coding is actually a CodingTask, as shown in figure 7.
+%You're right, Coding is actually a CodingTask, as shown in figure 7.
 
-%\todo{The idea is to group all the answers to the minor questions here.}
 \end{response}
@@
 Finally, in terms of the pattern, why should anyone care - why would should they use the pattern?}
-You are right. We intensively use the Type/Instance pattern. 
-We made this choice because it was easier to implement with Openflexo. 
-We make it more explicit in the text.
+
+\textcolor{orange}{The reviewer is absolutely right.
+We decided to use the type-object pattern as it was easier to integrate as an ad-hoc mechanism in our solution to the challenge than other multi-level enabling mechanisms.
+In that sense we would like to remark that this, together with the \emph{VirtualModel} concept of the Openflexo/FML framework sufficed to meet all the requirements of the challenge.
+Note however that we intend to explore the integration of other mechanisms (such as clabjets or potency) as future work.}
+
+%You are right. We intensively use the Type/Instance pattern. 
+%We made this choice because it was easier to implement with Openflexo. 
+%We make it more explicit in the text.
 \end{response}
@@
 It would be appropriate (perhaps at the end of the article) to revisit these features and to clearly demonstrate that their approach supports them and that they have concrete utility in some way in the context of the challenge.}
+
 We have added a paragraph in the Conclusions section indicating how Openflexo/FML permitted us to overcome the mentioned limitations of the strict modeling paradigm and how this was used in order to build our solution to the Multi-Level Process challenge.
@@
 
-\textcolor{red}{
-We tried to illustrate subsections of section 6 by giving examples, by referencing figures and sections (subsections 6.2, 6.3, 6.4, 6.6, 6.7 and 6.11 have been updated).
-Concerning the fact that "MLM is the key to the ability of Openflexo to create tooling", the reviewer probably read "MLM" instead of "FML" (Federation Modeling Language).}
+\textcolor{orange}{We have refactored the subsections of Section 6 in order to include concrete examples and references to the description of our solution whenever it was possible and/or relevant.
+Concretely, subsections 6.2, 6.3, 6.4, 6.6, 6.7 and 6.11 have been updated.}
+
+\todo[inline]{Here we have to explain how openflexo helps to create tooling!}
+
+%textcolor{red}{
+%We tried to illustrate subsections of section 6 by giving examples, by referencing %figures and sections (subsections 6.2, 6.3, 6.4, 6.6, 6.7 and 6.11 have been %updated).
+Concerning the fact that "MLM is the key to the ability of Openflexo to %create tooling", the reviewer probably read "MLM" instead of "FML" (Federation %Modeling Language).}
 \end{response}
@@
 }
-\textcolor{red}{While it is true that model slots and other features of Openflexo/FML are not used in order to build our solution,  we believe including them in our description help readers understand the framework.
+
+\textcolor{orange}{While it is true that model slots and other features of Openflexo/FML are not used in order to build our solution,  we believe including them in our description help readers understand the framework.
 We therefore choose to keep the description as is.}
@@
 (a) In fig 5: why are the in/out gateway features shown as properties and not as associations?
+\textcolor{orange}{This is done for the sake of readability in order to avoid cluttering the diagram with (possibly uninteresting) associations.}
 

### 2021-10-14 10-conclusions.tex
synchro

@@
 
-The features of \FML enable a great modelling flexibility which, coupled with an appropriate methodology, allow us to achieve
+The features of \FML enable a great modeling flexibility which, coupled with an appropriate methodology, allow us to achieve
 the multi/level capabilities without dedicated tooling in order to solve problems which require it. We use virtual models to "implement" levels on demand.

### 2021-10-14 3-introduction.tex
synchro

@@
 This strict approach shows its limitations when modeling complex domains requiring more than one level of specialization, \eg to adapt the model to application sub domains.
-Indeed, the two-level approach fails to acknowledge and support: 1) the existence of different forms of classification and/or instantiation (\eg ontological vs linguistic); and 2) the duality type-object for model elements.
+Indeed, the two-level approach fails to acknowledge and support: 1) the existence of different forms of classification and/or instantiation (\eg ontological vs linguistics); and 2) the duality type-object for model elements.
 Therefore, while it may still be used for the aforementioned complex modeling tasks, the strict approach entails the introduction of accidental complexity in both the modeling process and the resulting modeling artefacts.
@@
 
-Our modeling approach is based on a language (\FML) which provides the (linguistic) concepts that are used to define  the multi-levels models and meta-models that are needed.
+Our modeling approach is based on a language (\FML) which provides the (linguistics) concepts that are used to define  the multi-levels models and meta-models that are needed.
 \FML allows to define models as first class entities (which we call VirtualModel) and to link them together (by extension).
@@
 
-Our answer to the challenge is therefore not an answer using a classical multilevel approach that would allow an efficiency comparison with another classical multilevel answer, but an approach that proposes another organization of multilevels through the translation of needs into levels over time and the possibility to build an evolving hierarchy of adapted models.
-This is made possible by (1) the reification of the notion of model and the possibility to interconnect them (specialization) and (2) the freedom of meta-modeling (low model conformity, flexible meta-modeling).
+Our answer to the challenge is therefore not an answer using a classical multi-level approach that would allow an efficiency comparison with another classical multi-level answer, but an approach that proposes another organization of multi-levels through the translation of needs into levels over time and the possibility to build an evolving hierarchy of adapted models.
+This is made possible by (1) the reification of the notion of a model and (2) the possibility to interconnect models (specialization).% and (2) the freedom of meta-modeling (low model conformity, flexible meta-modeling).
 

### 2021-10-14 4-technology.tex
Updates from Overleaf
synchro

@@
 So to provide an adaptable framework, we have developed a flexible modeling language to provide modeling features without strict modeling levels.
-In a federation, the targeted modelling approaches can be unknown at the beginning and also can integrate several modeling levels.
+In a federation, the targeted modeling approaches can be unknown at the beginning and also can integrate several modeling levels.
 
 We decided to use this approach since it offers the possibility to
-link and to navigate among levels. Before we describe the overall
+link and to navigate among levels. Before we describe the 
 architecture in next section, here are some key concepts implemented
@@
 
-This framework relies on the architecture of Figure~\ref{fig:mf}. A federation
+This framework relies on the architecture of figure~\ref{fig:mf}. A federation
 gathers a set of conceptual models, named \emph{virtual models} and a
@@
 \emph{technological space} and uses the language of its specific
-paradigm while a virtual model is built using the Flexo Modeling
+paradigm while a virtual model is built using the Federation Modeling
 Language (\FML). Each federated model is an autonomous
@@
 
-A UML like representation of a simplified metamodel of \FML is provided in Figure~\ref{fig:mm}.
+A UML like representation of a simplified metamodel of \FML is provided in figure~\ref{fig:mm}.
 \FML is language designed to define models as virtual models.
@@
 \FML is designed to define not only the structure of virtual models but also the collection of actions that can be performed on them.
-These actions are called \emph{behaviors} and rely on behavioural primitives called \emph{EditionAction}.
+These actions are called \emph{behaviors} and rely on behavioral primitives called \emph{EditionAction}.
 These \emph{behaviors} can either be called or triggered by events.
@@
 
-\noteSylvain{A la fin de la rédaction, s'arranger pour que la figure 3 (fig:BPMNSubsetExample) et le listing soient cote a cote, et aussi que le listing tienne sur une seule page}
+%\noteSylvain{A la fin de la rédaction, s'arranger pour que la figure 3 (fig:BPMNSubsetExample) et le listing soient cote a cote, et aussi que le listing tienne sur une seule page}
 

### 2021-10-14 6-modelpresentation.tex
put VM and FC in textsf

--- 2021-10-06 6-modelpresentation.tex
@@
 
+Note that in all following figures, blue is used to represent \FML virtual models and flexo concepts (\textsf{VirtualModel} and \textsf{FlexoConcept} in figure \ref{fig:mm}) , see figures \ref{fig:BPMNSubsetExample}, \ref{fig:MultilevelArchitecture}, \ref{fig:LinguisticAndOntologicInstantiation}, \ref{fig:BaseMetamodel}, \ref{fig:AcmeFullArchitecture},  \ref{fig:ToolingArchitecture}, while brown is used for instances of FlexoConcept, see figures \ref{fig:BPMNSubsetExample}, \ref{fig:MultilevelArchitecture}, \ref{fig:LinguisticAndOntologicInstantiation}, \ref{fig:AcmeFullArchitecture},  \ref{fig:ToolingArchitecture}.
+% 3, 4, 5, 6, 7, 8, 9
 % Expliquer que le use case XSure ne nécessitait pas l'extension de metamodèle requise pour le usecase Acme Software Developement Process

### 2021-10-14 responseLetter.tex
mainly typo style

@@
 \begin{response}{We very much welcome a "conventional" response to the challenge but would like to see the respective reviewer questions addressed in a revised version which is explicit as possible about the advantages and the disadvantages of the emulation.
-You may find the references listed below useful to keep explanations concise by referencing related work (note that reviewer 3’s “Type-Instance pattern” is also known as the “Type Object” pattern).
+You may find the references listed below useful to keep explanations concise by referencing related work (note that reviewer 3’s "Type-Instance pattern” is also known as the “Type Object” pattern).
 You should also respond to other concerns reviewers raise, either by making changes to the submission or in the form of an authors' response.}
@@
 The Figure \ref{fig:BPMNSubsetExample} (above) shows how to realize a small subset of BPMN with Openflexo/FML (assuming a metamodel of the full specification of BPMN exists).
-Concretely, we show  how a VirtualModel (SimplifiedBPMN) can filter an existing model (BPMN) and be instantiated.
+Concretely, we show  how a virtual model (SimplifiedBPMN) can filter an existing model (BPMN) and be instantiated.
 Note that connection with external sources is not required for our solution to the challenge and is only represented in order to illustrate the capabilities of Openflexo/FML.
@@
 
-\textcolor{orange}{The only two concepts from Figured 2 represented elsewhere in the paper are the \emph{VirtualModel} and \emph{FlexoConcept} elements, which are represented in blue in figures 3, 5, 6, 7 and 9.
+\textcolor{orange}{The only two concepts from Figured 2 represented elsewhere in the paper are the \textsf{VirtualModel} and \textsf{FlexoConcept} elements, which are represented in blue in figures 3, 5, 6, 7 and 9.
 Then, we use the brown color to represent the instantiation of these concepts.
@@
 Note however that we use model federation only for the tooling part of our answer to the challenge.
-In practice, we rely solely in the reification of models through the VirtualModel concept in order to emulate multi-levels as shown in Figure 4.}
+In practice, we rely solely in the reification of models through the \textsf{VirtualModel} concept in order to emulate multi-levels as shown in Figure 4.}
 
@@
 We have improved the description of the relationship between FML and our metamodels (use of colors in many figures).
-The tooling uses model federation, but the description of our answer relies only on VirtualModel and FlexoConcept of FML (figure 2).
+The tooling uses model federation, but the description of our answer relies only on \textsf{VirtualModel} and \textsf{FlexoConcept} of FML (figure 2).
 The tooling is suggested in figure (new) 9.
@@
 
-The business logic of the diagram tooling in encoded in FML, and more particulary for this editor in ProcessEditor VirtualModel presented figure 9.
-Structure of diagramming, interactions and presentation layer (eg texts) are defined in various FlexoConcept defined in VirtualModel ProcessEditor.
+The business logic of the diagram tooling in encoded in FML, and more particulary for this editor in \texttt{ProcessEditor} virtual model presented figure 9.
+Structure of diagramming, interactions and presentation layer (eg texts) are defined in various concepts defined in the virtual model \texttt{ProcessEditor}.
 
@@
 We decided to use the type-object pattern as it was easier to integrate as an ad-hoc mechanism in our solution to the challenge than other multi-level enabling mechanisms.
-In that sense we would like to remark that this, together with the \emph{VirtualModel} concept of the Openflexo/FML framework sufficed to meet all the requirements of the challenge.
+In that sense we would like to remark that this, together with the \emph{virtual model} notion of the Openflexo/FML framework sufficed to meet all the requirements of the challenge.
 Note however that we intend to explore the integration of other mechanisms (such as clabjets or potency) as future work.}

### 2021-10-14 10-conclusions.tex
typo a FML, UML + trailing space

@@
 In that sense, our solution seamlessly integrates the type-object pattern which allows us to dynamically create different \emph{instances} of model elements (e.g., \emph{TaskType}, among others) and use them for typing other model elements.
-This, together with the FML core support for (level agnostic) classical linguistic instantiation, permitted us to merge linguistic and ontological instantiation in order to provide a satisfactory solution to the Multi-Level Process challenge.
+This, together with the \FML core support for (level agnostic) classical linguistic instantiation, permitted us to merge linguistic and ontological instantiation in order to provide a satisfactory solution to the Multi-Level Process challenge.
 

### 2021-10-14 4-technology.tex
typo a FML, UML + trailing space

@@
 
-A UML like representation of a simplified metamodel of \FML is provided in figure~\ref{fig:mm}.
+A \UML like representation of a simplified metamodel of \FML is provided in figure~\ref{fig:mm}.
 \FML is language designed to define models as virtual models.
@@
 Openflexo\footnote{\url{https://github.com/openflexo-team}}, is
-developed as an open source initiative. This tool offers a \FML
+developed as an open source initiative. This tool offers an \FML
 execution engine with an interactive virtual model design environment.

### 2021-10-14 6-modelpresentation.tex
typo a FML, UML + trailing space

--- 2021-10-06 6-modelpresentation.tex
@@
 
+Note that in all following figures, blue is used to represent \FML virtual models and flexo concepts (\textsf{VirtualModel} and \textsf{FlexoConcept} in figure \ref{fig:mm}) , see figures \ref{fig:BPMNSubsetExample}, \ref{fig:MultilevelArchitecture}, \ref{fig:LinguisticAndOntologicInstantiation}, \ref{fig:BaseMetamodel}, \ref{fig:AcmeFullArchitecture},  \ref{fig:ToolingArchitecture}, while brown is used for instances of \textsf{FlexoConcept}, see figures \ref{fig:BPMNSubsetExample}, \ref{fig:MultilevelArchitecture}, \ref{fig:LinguisticAndOntologicInstantiation}, \ref{fig:AcmeFullArchitecture},  \ref{fig:ToolingArchitecture}.
+% 3, 4, 5, 6, 7, 8, 9
 % Expliquer que le use case XSure ne nécessitait pas l'extension de metamodèle requise pour le usecase Acme Software Developement Process

### 2021-10-14 responseLetter.tex
Updates from Overleaf

@@
 We are grateful to the reviewers for their interest, effort, and suggestions.
-We detail in the following pages how we have addressed the reviewers' concerns; we are confident that the paper is now improved, and hope that you will agree.
-We have organized the responses to the comments on a per-reviewer basis, focusing, as suggested, on:
+The  following pages detail how we have addressed the reviewers' concerns.
+We are confident that the paper is now improved, and hope that you will agree.
+The responses to the comments are organized on a per-reviewer basis, focusing, as suggested, on:
 
@@
 \item Improving the presentation.
-\item Provide a more explicit description of how the multi-level modelling emulation is achieved.
+\item Providing a more explicit description of how the multi-level modelling emulation is achieved.
 \item Discussing the general advantages and limitations of the multi-level emulation, including a clarification of what the implications would be of supporting more "levels" than the challenge required.
@@
 \bigskip
-Sylvain, Joel, Jean-Christophe, Antoine, Fabien and Salvador
+Sylvain, Joël, Jean-Christophe, Antoine, Fabien and Salvador
 
@@
 
-\begin{response}{the reviewers have made suggestions for improving the presentation but also raised a number of issues requiring clarification.
+\begin{response}{The reviewers have made suggestions for improving the presentation but also raised a number of issues requiring clarification.
 An overarching theme is the request for a more explicit description of how the multi-level modelling emulation is achieved and what its general advantages and limitations are, including a clarification of what the implications would be of supporting more "levels" than the challenge required.}
@@
 We have improved the presentation of both, Openflexo/FML and our solution to the challenge so that: 1) the techniques we use to \emph{achieve} multi-level capabilities are more explicit; and, 2) the advantages and disadvantages of using them in order to fulfil the challenge requirements become clearer.
-Notably, we have added many details in the final paragraphs of the introduction section, improved the Section 2, devoted to the description of OpenFlexo, refactored figures 2, 4, 5, 6, 7 and 9 and, added figures 3 and 10.
+Notably, we have added many details in the final paragraphs of the introduction section, improved the Section 2, devoted to the description of Openflexo, refactored figures 2, 4, 5, 6, 7 and 9 and, added figures 3 and 10.
 \end{response}
@@
 \begin{response}{We very much welcome a "conventional" response to the challenge but would like to see the respective reviewer questions addressed in a revised version which is explicit as possible about the advantages and the disadvantages of the emulation.
-You may find the references listed below useful to keep explanations concise by referencing related work (note that reviewer 3’s "Type-Instance pattern” is also known as the “Type Object” pattern).
+You may find the references listed below useful to keep explanations concise by referencing related work (note that reviewer 3's "Type-Instance pattern" is also known as the "Type Object" pattern).
 You should also respond to other concerns reviewers raise, either by making changes to the submission or in the form of an authors' response.}
@@
 Additionally, the related work section have been extended in order to enhance the comparison between our approach and previous solutions to the challenge.}
+% fd ok
 
@@
 The Figure \ref{fig:BPMNSubsetExample} (above) shows how to realize a small subset of BPMN with Openflexo/FML (assuming a metamodel of the full specification of BPMN exists).
-Concretely, we show  how a virtual model (SimplifiedBPMN) can filter an existing model (BPMN) and be instantiated.
+Concretely, we show  how a virtual model (\texttt{SimplifiedBPMN}) can filter an existing model (BPMN) and be instantiated.
 Note that connection with external sources is not required for our solution to the challenge and is only represented in order to illustrate the capabilities of Openflexo/FML.
@@
 The previous version of the figure depicting the main concepts of the FML language (Figure 2) was too succinct.
-We have fully refactored it so that now important concepts such as \emph{Type} are shown.
+We have fully refactored it so that now important concepts such as \textsf{Type} are shown.
 We have correspondingly improved the text explaining the figure too.
 We hope this changes make understanding the core concepts of the FML language easier.}
+%fd ok, j'ajouterais bien aussi :
+{\color{teal} Furthermore, to make it easier to distinguish the concepts of FML, the figure 2 adopts a slightly different graphical syntax while the caption insists on the fact that it is a representation.
+Indeed, FML is implemented by an interpreter that understands these concepts.
+For example, defining a behavior consists in providing its code using the FML language syntax (like in a programming language).}
 
@@
 
-\textcolor{orange}{The only two concepts from Figured 2 represented elsewhere in the paper are the \textsf{VirtualModel} and \textsf{FlexoConcept} elements, which are represented in blue in figures 3, 5, 6, 7 and 9.
+\textcolor{orange}{The only two concepts from Figure 2 represented elsewhere in the paper are the \textsf{VirtualModel} and \textsf{FlexoConcept} elements, which are represented in blue in figures 3, 5, 6, 7 and 9.
 Then, we use the brown color to represent the instantiation of these concepts.
@@
 Regarding the notation, it is indeed similar to UML but it has different semantics as explained in the paper.}
+%fd ok
 
@@
 \textcolor{orange}{Thanks for pointing out this inconsistency. We have completed Figure 6 (now Figure 7) so that it contains all the necessary elements for Figure 7 (now Figure 8).}
+%fd ok
 
@@
 
-\textcolor{orange}{Our manuscript includes now a new figure (Figure 10) that shos a screen capture of the FML metamodel editor.}
+\textcolor{orange}{Our manuscript now includes a new figure (Figure 10) that shows a screen capture of the FML metamodel editor.}
+%fd ok
 %Figure 10 now shows a screen capture of the FML metamodel editor.
@@
 The tooling is suggested in figure (new) 9.
-The tooling is not detailed, but the VirtualModel ProcessEditor presented in figure (new) 9 defines a lot of interactions, where texts of (old) figure 9 are found.
+The tooling is not detailed, but the virtual model \texttt{ProcessEditor} presented in figure (new) 9 defines a lot of interactions, where texts of (old) figure 9 are found.
 
-The business logic of the diagram tooling in encoded in FML, and more particulary for this editor in \texttt{ProcessEditor} virtual model presented figure 9.
-Structure of diagramming, interactions and presentation layer (eg texts) are defined in various concepts defined in the virtual model \texttt{ProcessEditor}.
+The business logic of the diagram tooling is encoded in FML, and more specifically for this editor in the \texttt{ProcessEditor} virtual model presented in figure 9.
+Structure of diagramming, interactions and presentation layer (\emph{e.g.} texts) are defined in various concepts defined in the virtual model \texttt{ProcessEditor}.
 

### 2021-10-14 responseLetter.tex
synchro

@@
 
-\textcolor{orange}{We fully agree with the reviewer on the fact that it is the flexibility of the model federation approach what enables us to provide a solution to the multi-level process challenge while using a tool not specifically designed for multi-level modelling.
-Of course, this comes together with a number of trade-offs which we discuss in the paper (e.g., see the the end of the introduction section).}
-
+\textcolor{orange}{We fully agree with the reviewer on the fact that it is the flexibility of the model federation approach that enables us to provide a solution to the multi-level process challenge while using a tool not specifically designed for multi-level modelling.
+Of course, this comes together with a number of trade-offs which we discuss in the paper (\emph{e.g.}, see the the end of the introduction section).}
+%fd ok
 \textcolor{red}{We have included at the end of the introduction section a more detailed description of our approach.
@@
 
-\textcolor{red}{Propale de Fabien pour complèter la réponse ici.
+\textcolor{teal}{Propale de Fabien pour complèter la réponse ici.
 We have added a discussion (où?) on the fact that our approach reifies the levels using "generic" models and rely on the process of construction to implement the levels of the challenge.
@@
 In practice, we rely solely in the reification of models through the \textsf{VirtualModel} concept in order to emulate multi-levels as shown in Figure 4.}
-
+%fd ok
 %We use federation only for the tooling part of our answer.
@@
 
+{\color{red} FD: je pense aussi qu'il faut être plus précis dans cette réponse.
+J'aurais bien dit que l'instantiation linguistique consiste juste à utiliser le langage FML et respecter sa syntaxe et ses contraintes de typage et que pour l'ontologique, c'est uniquement par le process ou l'on s'impose de faire des factory depuis le type (Sylvain ?)}
+
 %We have improved the explanation in sections 2 and 3. Note that %\texttt{InstanceInstance} has been replaced with \texttt{ObjectInstance}.
@@
 However, we have realized now that the alternative we used hampered readability even more.
-In that sense, we use now rolenames (an exception is made for the \emph{Gateway} associations with \emph{TaskType} as explained in the caption of Figure 6)}
+In that sense, we use now rolenames (an exception is made for the \texttt{Gateway} associations with \texttt{TaskType} as explained in the caption of Figure 6)}
+%fd ok
 
@@
 
-{\color{red}hum, j'ai l'impression qu'il manque un peu de détail pour répondre sur la partie multi-instance.
+{\color{red}fd: hum, j'ai l'impression qu'il manque un peu de détail pour répondre sur la partie multi-instance.
 Il me semble qu'en plus d'être ad hoc, le mécanisme est manuel (mais je me trompe peut être).
@@
 The constraints are defined at FlexoConcept level, and are designed of invariant being evaluated by the FML interpreter, which may raise constraint violation at run-time.
-An example of constraint is presented for requirement P9, in TaskType flexo concept (p 8).
+An example of constraint is presented for requirement P9, in \texttt{TaskType} concept (p 8).
 
@@
 Structure of diagramming, interactions and presentation layer (\emph{e.g.} texts) are defined in various concepts defined in the virtual model \texttt{ProcessEditor}.
-
+%fd ok
 \end{response}
@@
 TODO.
+
+{\color{teal}hum, je crois qu'il n'a pas bien compris mais c'est difficile de le dire comme cela.
+je pense que l'on pourrait indiquer que chaque niveau est représenté par un virtual model et donc ajouter un niveau se ferait en ajoutant un virtual model.
+En revanche, il y a peu de support automatisé et nous reposons sur une approche manuel pour gérer correctement ces niveaux.
+La limite est probablement plus dans la faible aide (me semble t il) Et pour le dernier point, comme notre outil repose sur de l'interpretation il est possible de travailler au deux niveaux à la fois.
+Par contre, il faut une certaine discipline pour ne pas tout casser (SYLVAIN ?)}
 \end{response}
@@
     \item Section 2.
-“In contrast to approaches that compose metamodels into a single large metamodel grouping all needed entities,…” - I do not see, why this is relevant here \textcolor{orange}{Our intention was to compare the federation approach to approaches that integrate all metamodels together.
+"In contrast to approaches that compose metamodels into a single large metamodel grouping all needed entities,…" - I do not see, why this is relevant here \textcolor{orange}{Our intention was to compare the federation approach to approaches that integrate all metamodels together.
 We argue that reifying models (with virtual models) helps structuring, organizing, federating models.}
-    \item Figure 5 – concept Sequencing – the roles “in” and “out” are not represented by lines… why? \textcolor{orange}{As explained in a previous answer, we do this for the sake of readability.}
-    \item What does the term “sub-concept” mean exactly? Please explain it before the first use \todo{TODO}
-    \item ProgrammingLanguage concept is an enumeration… how is it modelled? \textcolor{orange}{ProgrammingLanguages is modelled as an enumeration of names (Java, Cobol, etc) from FML. }
-    \item “The Acme metamodel also defines a specialized task type Coding” – isn’t it CodingTask? \textcolor{orange}{Yes, this is correct. Coding is actually a CodingTask, as shown in figure 7.}
-    \item “The value of the property producedArtifactTypes” – is it property, or role? \todo{TODO}
+    \item Figure 5 - concept Sequencing - the roles "in" and "out" are not represented by lines… why? \textcolor{orange}{As explained in a previous answer, we do this for the sake of readability.}
+    \item What does the term "sub-concept" mean exactly? Please explain it before the first use \todo{TODO}
+    \item ProgrammingLanguage concept is an enumeration… how is it modelled?
+\textcolor{orange}{ProgrammingLanguages is modelled as an enumeration of names (Java, Cobol, etc) using the notion of enumeration of the FML language.
+}
+    \item "The Acme metamodel also defines a specialized task type Coding" - isn’t it CodingTask? \textcolor{orange}{Yes, this is correct. Coding is actually a CodingTask, as shown in figure 7.}
+    \item "The value of the property producedArtifactTypes" - is it property, or role?
+\todo{TODO, Sylvain, il me semble que c'est la seule occurrence de property, on pourrait basculer sur role il me semble}
 \end{itemize}}
@@
 Note however that we intend to explore the integration of other mechanisms (such as clabjets or potency) as future work.}
+%fd ok
 
@@
 Notably, we mention how the type/object pattern allowed us to use model element instances as types in order to achieve ontological instantiation.
+%fd ok
 \end{response}
@@
 It would be more abstract than the current 'method' style semantics.}
-\textcolor{red}{
+{\color{red}
 We designed our answer with the Type/Instance pattern. We think, we could have implemented it with another approach such as Clabject.
@@
 We have here a trade-off.
-Reuse the constraint from FML which is efficient but may be too restricted, and define our own constraints, which is more adaptable, but has to be implemented.
-Due to the size of the problem and the expected freedom, we choose the second way.
-Le language FML permet de définir des comportements reifiant la sémantique ontologique (notamment du pattern type/instance, par exemple le comportement newProcess() dans ProcessType).
+Reusing the constraint from FML is efficient but may be too restrictive for the user.
+Defining our own constraints is more adaptable, but it has to be implemented and the produced model may be incorrect.
+Due to the size of the problem and the expected freedom, we chose the second way.
+To make it easier to build correct models, we propose several built-in behavior in charge of the various operation and notably the instantiation operations.
+This result in a "constrained" tool rather than a "restrictive" metamodel.
+
+Our implementation of the challenge was also an opportunity to showcase the expressivity of FML.
+So we tried, in our solution, to reuse several mechanism of FML.
+For example, sometime we use behavior ({\color{teal} un exemple}) and sometimes we use constraints ({\color{teal} un exemple}).}
+
+%fd, j'ai essayé de traduire et complèter, j'espère ne pas avoir trahi
+
+%Le language FML permet de définir des comportements reifiant la sémantique ontologique (notamment du pattern type/instance, par exemple le comportement newProcess() dans ProcessType).
 Le language FML permet de définir également des contraintes sur les concepts.
@@
 On utilise tout de même les contraintes.
-On a essayé aussi de faire des choix multiples et variés pour montrer l'expressivité du langage FML.}
+On a essayé aussi de faire des choix multiples et variés pour montrer l'expressivité du langage FML.
 
@@
 \textcolor{orange}{While it is true that model slots and other features of Openflexo/FML are not used in order to build our solution,  we believe including them in our description help readers understand the framework.
-We therefore choose to keep the description as is.}
+We therefore chose to keep the description as is.
+While explicitly expressing that they are only marginally used in our solution (only for the tooling).}
 \end{response}
@@
 (b) On p. 7 'specialization' -> 'specializations'.}
-TODO.
+Done.
 \end{response}

### 2021-10-15 responseLetter.tex
sync overleaf->github

@@
 % fd ok
+% sg ok
 
@@
 % Fabien OK pour moi
+% sg ok
 
@@
 For example, defining a behavior consists in providing its code using the FML language syntax (like in a programming language).}
+% sg d'accord avec cette proposition
 
@@
 %fd ok
+%sg ok
 
@@
 %fd ok
+% sg ok
 \textcolor{red}{We have included at the end of the introduction section a more detailed description of our approach.
@@
 We hope, it makes it easier to see that it is a contribution to multi-level domain.}
+
+% sg ok : je pense que c'est dans la section 3
 
@@
 %fd ok
+%sg ok
 %We use federation only for the tooling part of our answer.
@@
 
+{\color{red} Sylvain: (Nouvelle proposition de réponse) As indicated in the answer to the previous comment we have refactored Section 2 and 3 in order to provide more details and make explanations easier to follow.
+Linguistic instantiation in this context refers to the use of FML language for our solution.
+This offers syntax and semantics, as well as type checking and various constraints checking.
+We tried to show how the use of FML helps in meeting requirements of the challenge.
+On the other side, we exploit the ontologic instantiation though an extensive use of the "Type Object" pattern in two versions (with single and multiple cardinalities), controlling the ontologic instantiation from the type, using a classical factory pattern.
+Additionaly, we agree that the choice of \texttt{InstanceInstance} was not a good decision, we have replaced with \texttt{ObjectInstance}.}
+
 %We have improved the explanation in sections 2 and 3. Note that %\texttt{InstanceInstance} has been replaced with \texttt{ObjectInstance}.

### 2021-10-15 4-technology.tex
relecture + typos Fabien

@@
 A \UML like representation of a simplified metamodel of \FML is provided in figure~\ref{fig:mm}.
-\FML is language designed to define models as virtual models.
+\FML is a language designed to define models as virtual models.
 A \emph{virtual model} is composed of a set of \emph{concepts}, while itself being a \emph{concept}.
@@
 \FML is designed to define not only the structure of virtual models but also the collection of actions that can be performed on them.
-These actions are called \emph{behaviors} and rely on behavioral primitives called \emph{EditionAction}.
+These actions are called \emph{behaviors} and rely on behavioral primitives called \textsf{EditionAction}.
 These \emph{behaviors} can either be called or triggered by events.
@@
 
-To illustrate this, figure \ref{fig:BPMNSubsetExample} shows how to realize a small subset of BPMN with Openflexo/FML.
+To illustrate this, figure~\ref{fig:BPMNSubsetExample} shows how to realize a small subset of BPMN with Openflexo/FML.
 Concretely, we have designed a \textit{virtual model} (\texttt{SimplifiedBPMN}) with 3 \textit{concepts} representing a process and its contained tasks.
@@
 The following listing shows the \FML code of this \textit{virtual model}.
+{\color{red}Notice that the composition relation of figure~\ref{fig:BPMNSubsetExample} between \texttt{BPMNProcess} and \texttt{BPMNTask} is realized in the FML code by concept containment.
+Whenever an instance of \texttt{BPMNTask} is created, it must be in a container (instance of \texttt{BPMNProcess}) and the two instances get linked together.}\noteFabien{A valider}
 

### 2021-10-15 5-analysis.tex
typo a FML, UML + trailing space
relecture + typos Fabien

@@
 
-During the next step of our approach, we try to take into account the reusability of the relationships by identifying the semantics of them.
-The goal is the identification of concepts with their behavior.
+During the next step of our approach, we try to take into account the reusability of the relationships by identifying their semantics. The goal is the identification of concepts with their behavior.
 
@@
 Like in a lot of modeling approaches, these  steps could also be achieved in any order and iteratively.
-But our goal during the challenge's analysis remains to produce a \FML virtual model architecture.
+But our goal during the challenge's analysis remains to produce an \FML virtual model architecture.
 %for our federation.

### 2021-10-15 6-modelpresentation.tex
relecture + typos Fabien

--- 2021-10-14 6-modelpresentation.tex
@@
 
-Note that in all following figures, blue is used to represent \FML virtual models and flexo concepts (VirtualModel and FlexoConcept in figure \ref{fig:mm}) , see figures \ref{fig:BPMNSubsetExample}, \ref{fig:MultilevelArchitecture}, \ref{fig:LinguisticAndOntologicInstantiation}, \ref{fig:BaseMetamodel}, \ref{fig:AcmeFullArchitecture},  \ref{fig:ToolingArchitecture}, while brown is used for instances of FlexoConcept, see figures \ref{fig:BPMNSubsetExample}, \ref{fig:MultilevelArchitecture}, \ref{fig:LinguisticAndOntologicInstantiation}, \ref{fig:AcmeFullArchitecture},  \ref{fig:ToolingArchitecture}.
+Note that in all following figures, blue is used to represent \FML virtual models and concepts (\textsf{VirtualModel} and \textsf{FlexoConcept} in figure \ref{fig:mm}), see figures \ref{fig:BPMNSubsetExample}, \ref{fig:MultilevelArchitecture}, \ref{fig:LinguisticAndOntologicInstantiation}, \ref{fig:BaseMetamodel}, \ref{fig:AcmeFullArchitecture},  \ref{fig:ToolingArchitecture}, while brown is used for instances of \textsf{FlexoConcept}, see figures \ref{fig:BPMNSubsetExample}, \ref{fig:MultilevelArchitecture}, \ref{fig:LinguisticAndOntologicInstantiation}, \ref{fig:AcmeFullArchitecture},  \ref{fig:ToolingArchitecture}.
 % 3, 4, 5, 6, 7, 8, 9
@@
 
-The \texttt{ActorType} concept is a sub-concept of \texttt{Type} and the \texttt{allowedActorTypes} relation to \texttt{ActorType} defined in \texttt{TaskType} (with 0..* cardinality) captures \textbf{P5} requirement.
+The \texttt{ActorType} concept inherits from the \texttt{Type} concept and the \texttt{allowedActorTypes} relation to \texttt{ActorType} defined in \texttt{TaskType} (with 0..* cardinality) captures \textbf{P5} requirement.
 Requirement \textbf{P6} is symmetrically satisfied with \texttt{allowedActors} relation to \texttt{Actor} also defined in \texttt{TaskType} (with 0..* cardinality).
-The same modeling pattern applies to \texttt{ArtefactType} defined as a sub-concept of \texttt{Type}, and both relations \texttt{usedArtifactTypes} and \texttt{producedArtifactTypes} defined in \texttt{TaskType} (\textbf{P7}).
+The same modeling pattern applies to \texttt{ArtefactType} also extending the concept \texttt{Type}, and both relations \texttt{usedArtifactTypes} and \texttt{producedArtifactTypes} defined in \texttt{TaskType} (\textbf{P7}).
 \texttt{TaskType} additionally exposes an \texttt{expectedDuration} role (expressed in number of days), satisfying \textbf{P8}.
@@
 When the drag and drop interactor is applied for a new item, the tool allows choosing the concept type to be instantiated.
-For example, when a \texttt{TaskType} is created, the user must choose a sub-concept of \texttt{TaskType} -- in Acme use case it can be \texttt{SETaskTask} or \texttt{CodingTaskType} or default value \texttt{TaskType}).
+For example, when a \texttt{TaskType} is created, the user must choose a concept inheriting from \texttt{TaskType} -- in Acme use case it can be \texttt{SETaskTask} or \texttt{CodingTaskType} or default value \texttt{TaskType}).
 

### 2021-10-15 responseLetter.tex

@@
 \textcolor{orange}{In the answer to the previous editor's comment we describe the improvements of our revision, which includes an emphasis in discussing trade-offs related to the emulation of multi-level modeling by using the Openflexo/FML and the type-object pattern.
-Additionally, the related work section have been extended in order to enhance the comparison between our approach and previous solutions to the challenge.}
+Additionally, the related work section has been extended in order to enhance the comparison between our approach and previous solutions to the challenge.}
 % fd ok
 % sg ok
+% jc ok
 
@@
 % sg ok
+% jc ok
 
@@
 \end{figure}
+
+\todo{SG: Est ce possible de rajouter ce listing ou ca n'avance à rien ?}
 
@@
 We have fully refactored it so that now important concepts such as \textsf{Type} are shown.
-We have correspondingly improved the text explaining the figure too.
+We also have correspondingly improved the text explaining the figure.
 We hope this changes make understanding the core concepts of the FML language easier.}
@@
 % sg d'accord avec cette proposition
+% jc ok
 
@@
 %sg ok
+% jc ok
 
@@
 %fd ok
+% jc ok
 %Figure 10 now shows a screen capture of the FML metamodel editor.
@@
 In this sense, we do on-demand multi-level modelling, and propose an organization guided by the evolution of models.}
+% jc ok
 
-\textcolor{teal}{Propale de Fabien pour complèter la réponse ici.
-We have added a discussion (où?) on the fact that our approach reifies the levels using "generic" models and rely on the process of construction to implement the levels of the challenge.
+\textcolor{teal}{Propale de Fabien pour compléter la réponse ici.
+We have added a discussion (où?) on the fact that our approach reifies the levels using "generic" models and relies on the process of construction to implement the levels of the challenge.
 We hope, it makes it easier to see that it is a contribution to multi-level domain.}
@@
 % sg ok : je pense que c'est dans la section 3
+% jc euh… je ne vois pas dans la section 3.
 
@@
 %sg ok
+% jc ok
 %We use federation only for the tooling part of our answer.
@@
 We tried to show how the use of FML helps in meeting requirements of the challenge.
-On the other side, we exploit the ontologic instantiation though an extensive use of the "Type Object" pattern in two versions (with single and multiple cardinalities), controlling the ontologic instantiation from the type, using a classical factory pattern.
-Additionaly, we agree that the choice of \texttt{InstanceInstance} was not a good decision, we have replaced with \texttt{ObjectInstance}.}
+On the other side, we exploit the ontologic instantiation through an extensive use of the "Type Object" pattern in two versions (with single and multiple cardinalities), controlling the ontologic instantiation from the type, using a classical factory pattern.
+Additionaly, we agree that the choice of \texttt{InstanceInstance} was not a good decision, we have replaced it with \texttt{ObjectInstance}.}
+% jc ok (avec la nouvelle proposition de Sylvain)
 
@@
 \begin{response}{Overall, the solutions for the requirements are correctly presented in this section, but there are some spots not clear enough.
-E.
-g.
+E.g.
 regarding ontological instantiation, the solution uses two kinds: single and multiple.
@@
 
+{\color{red}Sylvain: (Nouvelle proposition de réponse) MultiInstance(s) are introduced in our answer metamodel (figure 6).
+As such, their semantic is defined at the ontological level, i.e.
+in the metamodel.
+It is an ad hoc mechanism introduced to fulfill requirements.
+This means that corresponding semantics is not handled by FML but in the designed virtual models, with specific behaviours implementing underlying semantics.}
+
 %On ne gère pas de multi-instances dans la sémantique FML, on gère ca seulement au niveau ontologique (FlexoConcept Instance et MultiInstance, figure 6).
@@
 
+% jc ok
 \end{response}
@@
 This realizes the ontological instantiation.
-}
+There is no notion "usual" or "default" constructors in FML.
+Whenever a user creates a new instance, it has to chose one constructor.
+Here, we impose methodologically to the user to create a process instance through the chosen process type.
+There is no enforcement in our solution.
+SYLVAIN ?
+
+Concerning the use of "/type" and "type", the former one is a way to stress that the relation is derived, i.e.
+it is a refinement of the "type" relation defined in another level.
+In the FML code, we use "type" due to the fact there is no notion of level.
+(It is explained in the lasta paragraph of 4.1 on page 8.)}
 \textcolor{red}{
 %- newProcess est un vrai behaviour défini dans ProcessType (pattern factory) et permet de spécifier l'instantiation d'un Process à partir d'un ProcessType (c'est la partie réalisation de l'instantiation ontologique)
-- difficile de répondre sur le "same pattern applies" /type raffine type défini à l'étage plus haut (fd ne comprends pas la question)
-- pas compris la question sur la spécialisation des Actors (fd ne comprends pas la question)
+\begin{itemize}
+\item difficile de répondre sur le "same pattern applies"  (fd finalement, je pense qu'il ne comprend pas bien notre usage de "same pattern" car il a l'impression qu'il y a plusieurs pattern => peut etre faudrait-il les nommer (specialization, factory, ...)) ; (jc) j'ai l'impression que la question est un peu à côté de la plaque
+\item pas compris la question sur la spécialisation des Actors (fd ne comprend pas la question, jc non plus)
+\end{itemize}
 }
@@
 %fd ok
+%sg ok
+%jc ok
 \end{response}
@@
 In the case of ontological instantiation, the term “dual concept nature” and “instance nature” is mentioned… what do these terms mean?}
-On a rectifié et ré-écrit ces bouts de texte
-TODO (Salvador).
 
-\textcolor{red}{Conceptualization means that a concept is created to fulfill or meet a requirement. }
+%On a rectifié et ré-écrit ces bouts de texte
+%TODO (Salvador).
+
+\textcolor{orange}{Conceptualization means that a concept is created to fulfill or meet a requirement.
+As for the mention to a "dual concept nature", we have rewritten that part as it was indeed confusing.
+Now it reads: the fulfillment of a requirement is obtained by the relation between an instance and its type definition instance.}
+%jc ok
 
@@
 Another problem may arise if we would use the framework as a prototyping tool and change the metamodel and the models at the same time.}
-TODO.
 
@@
 Par contre, il faut une certaine discipline pour ne pas tout casser (SYLVAIN ?)}
+
+{\color{red}Sylvain: (Nouvelle proposition de réponse) Indeed, our solution relies on a ad-hoc architecture where multi-level is emulated.
+However, our solution presents reified layers though virtual model inheritance hierarchy.
+Although there is no automatized support, the proposed methodology allows to provide an arbitrary number of layers and some extension capabilities and reuse support (e.g.
+adding a level would mean creating a new virtual model extending an existing one).
+This is made possible by the fact that FML allows one to work on different levels at a time.
+The limitations of this approach resides in a good comprehension of the architecture and a strong expertise to maintain global consistency.}
+%jc ok (j'ai probablement loupé la première réponse)
 \end{response}
@@
     \item Figure 5 - concept Sequencing - the roles "in" and "out" are not represented by lines… why? \textcolor{orange}{As explained in a previous answer, we do this for the sake of readability.}
-    \item What does the term "sub-concept" mean exactly? Please explain it before the first use \todo{TODO}
+    \item What does the term "sub-concept" mean exactly?
+Please explain it before the first use \textcolor{orange}{Indeed, this was unclear.
+This was related to concept inheritance.
+We have removed its use.}
     \item ProgrammingLanguage concept is an enumeration… how is it modelled?
@@
     \item "The value of the property producedArtifactTypes" - is it property, or role?
-\todo{TODO, Sylvain, il me semble que c'est la seule occurrence de property, on pourrait basculer sur role il me semble}
+\textcolor{orange}{Indeed, this was unclear.
+Figure 2 has been completed to show all kind of properties (\texttt{FlexoRole} is a specialization of \texttt{FlexoProperty}).
+The two terms "property" and "role" are equivalent in the context of \texttt{FlexoRole} use.
+}
+    % fd: TODO, Syl-vain, il me semble que c’est la seule occurrence de property,on pourrait basculer sur role il me semble
 \end{itemize}}
@@
 %fd ok
+%sg ok
 \end{response}
@@
 So we tried, in our solution, to reuse several mechanism of FML.
-For example, sometime we use behavior ({\color{teal} un exemple}) and sometimes we use constraints ({\color{teal} un exemple}).}
+For example, sometime we use behavior ({\color{teal} P17, P18 and P19}) and sometimes we use constraints ({\color{teal} P9 and P17}).}
 

### 2021-10-18 responseLetter.tex
test same pattern ;)

@@
 
-\todo{SG: Est ce possible de rajouter ce listing ou ca n'avance à rien ?}
+\todo{SG: Est ce possible de rajouter ce listing ou ca n'avance à rien ?. SM: If you want to add it, it should be ok, but remember then to update the answer text.}
 
@@
 % jc ok
+% Ok avec l'ajout de Fabien
 
@@
 % jc euh… je ne vois pas dans la section 3.
+% sm: ok.. if we find where it is ;)
 
@@
 % jc ok (avec la nouvelle proposition de Sylvain)
+% sm: ok
 
@@
 % jc ok
+% sm ok
 \end{response}
@@
 9.) is defined.}
+
 We have improved the description of the relationship between FML and our metamodels (use of colors in many figures).
@@
 %jc ok
+%sm ok
 \end{response}
@@
 %jc ok (j'ai probablement loupé la première réponse)
+%sm ok.. but we could even ignore this comment as it does not really request action.
 \end{response}
@@
 Due to the size of the problem and the expected freedom, we chose the second way.
-To make it easier to build correct models, we propose several built-in behavior in charge of the various operation and notably the instantiation operations.
+To make it easier to build correct models, we propose several built-in behavior in charge of the various operations and notably the instantiation operations.
 This result in a "constrained" tool rather than a "restrictive" metamodel.
@@
 %fd, j'ai essayé de traduire et complèter, j'espère ne pas avoir trahi
+%sm ok
 
@@
 
-\todo[inline]{Here we have to explain how openflexo helps to create tooling!}
+\todo[inline]{Here we have to explain how openflexo helps to create tooling.
+Just a line or two should suffice.
+Or maybe I have misanderstood the question...
+Is the reviewer really wrong and thought we were talking about MLM?
+Then just copy answer below from: Concerning...}
 

### 2021-10-18 responseLetter.tex
synchro

@@
 
-\textcolor{orange}{In the answer to the previous editor's comment we describe the improvements of our revision, which includes an emphasis in discussing trade-offs related to the emulation of multi-level modeling by using the Openflexo/FML and the type-object pattern.
-Additionally, the related work section has been extended in order to enhance the comparison between our approach and previous solutions to the challenge.}
+%\textcolor{orange}{
+In the answer to the previous editor's comment we describe the improvements of our revision, which includes an emphasis in discussing trade-offs related to the emulation of multi-level modeling by using the Openflexo/FML and the type-object pattern.
+Additionally, the related work section has been extended in order to enhance the comparison between our approach and previous solutions to the challenge.%}
 % fd ok
@@
 
-\textcolor{orange}{We agree with the reviewer.
+%\textcolor{orange}{
+We agree with the reviewer.
 The previous version of the figure depicting the main concepts of the FML language (Figure 2) was too succinct.
@@
 We also have correspondingly improved the text explaining the figure.
-We hope this changes make understanding the core concepts of the FML language easier.}
+We hope this changes make understanding the core concepts of the FML language easier.
+%}
 %fd ok, j'ajouterais bien aussi :
-{\color{teal} Furthermore, to make it easier to distinguish the concepts of FML, the figure 2 adopts a slightly different graphical syntax while the caption insists on the fact that it is a representation.
+%{\color{teal} 
+Furthermore, to make it easier to distinguish the concepts of FML, the figure 2 adopts a slightly different graphical syntax while the caption insists on the fact that it is a representation.
 Indeed, FML is implemented by an interpreter that understands these concepts.
-For example, defining a behavior consists in providing its code using the FML language syntax (like in a programming language).}
+For example, defining a behavior consists in providing its code using the FML language syntax (like in a programming language).
+%}
 % sg d'accord avec cette proposition
@@
 
-\textcolor{orange}{The only two concepts from Figure 2 represented elsewhere in the paper are the \textsf{VirtualModel} and \textsf{FlexoConcept} elements, which are represented in blue in figures 3, 5, 6, 7 and 9.
+%\textcolor{orange}{
+The only two concepts from Figure 2 represented elsewhere in the paper are the \textsf{VirtualModel} and \textsf{FlexoConcept} elements, which are represented in blue in figures 3, 5, 6, 7 and 9.
 Then, we use the brown color to represent the instantiation of these concepts.
 We hope this makes the connection between the Figure 2 and the other representations in the paper clearer.
-Regarding the notation, it is indeed similar to UML but it has different semantics as explained in the paper.}
+Regarding the notation, it is indeed similar to UML but it has different semantics as explained in the paper.
+%}
 %fd ok
@@
 
-\textcolor{orange}{Thanks for pointing out this inconsistency. We have completed Figure 6 (now Figure 7) so that it contains all the necessary elements for Figure 7 (now Figure 8).}
+%\textcolor{orange}{
+Thanks for pointing out this inconsistency. We have completed Figure 6 (now Figure 7) so that it contains all the necessary elements for Figure 7 (now Figure 8).
+%}
 %fd ok
@@
 
-\textcolor{orange}{Our manuscript now includes a new figure (Figure 10) that shows a screen capture of the FML metamodel editor.}
+%\textcolor{orange}{
+Our manuscript now includes a new figure (Figure 10) that shows a screen capture of the FML metamodel editor.%}
 %fd ok
@@
 
-\textcolor{orange}{We fully agree with the reviewer on the fact that it is the flexibility of the model federation approach that enables us to provide a solution to the multi-level process challenge while using a tool not specifically designed for multi-level modelling.
-Of course, this comes together with a number of trade-offs which we discuss in the paper (\emph{e.g.}, see the the end of the introduction section).}
+%\textcolor{orange}{
+We fully agree with the reviewer on the fact that it is the flexibility of the model federation approach that enables us to provide a solution to the multi-level process challenge while using a tool not specifically designed for multi-level modelling.
+Of course, this comes together with a number of trade-offs which we discuss in the paper (\emph{e.g.}, see the the end of the introduction section).%}
 %fd ok
 % sg ok
-\textcolor{red}{We have included at the end of the introduction section a more detailed description of our approach.
+%\textcolor{red}{
+We have included at the end of the introduction section a more detailed description of our approach.
 In a word, we use FML (linguistic level) as a tool to build models and meta-models permitting us to emulate multi-level modelling.
 The evolution of models (multi-levels) is encoded as a hierarchy of models and meta-models.
-In this sense, we do on-demand multi-level modelling, and propose an organization guided by the evolution of models.}
+In this sense, we do on-demand multi-level modelling, and propose an organization guided by the evolution of models.%}
 % jc ok
@@
 
-\textcolor{orange}{We have refactored the Technology section in order to make it, hopefully, easier to follow.
+%\textcolor{orange}{
+We have refactored the Technology section in order to make it, hopefully, easier to follow.
 Apart from improvements in the text, we have notably improved Figure 2 so that now a lot more of details are visible and contributed a new figure (Figure 3) showing how the FML concepts are used in a classical modeling scenario.
 Note however that we use model federation only for the tooling part of our answer to the challenge.
-In practice, we rely solely in the reification of models through the \textsf{VirtualModel} concept in order to emulate multi-levels as shown in Figure 4.}
+In practice, we rely solely in the reification of models through the \textsf{VirtualModel} concept in order to emulate multi-levels as shown in Figure 4.%}
 %fd ok
@@
 
-\textcolor{orange}{As indicated in the answer to the previous comment we have refactored Section 2 in order to provide more details and make explanations easier to follow.
+%\textcolor{orange}{As indicated in the answer to the previous comment we have refactored Section 2 in order to provide more details and make explanations easier to follow.
 We have done the same with Section 3.
@@
 
-{\color{red} FD: je pense aussi qu'il faut être plus précis dans cette réponse.
+%{\color{red} FD: je pense aussi qu'il faut être plus précis dans cette réponse.
 J'aurais bien dit que l'instantiation linguistique consiste juste à utiliser le langage FML et respecter sa syntaxe et ses contraintes de typage et que pour l'ontologique, c'est uniquement par le process ou l'on s'impose de faire des factory depuis le type (Sylvain ?)}
@@
 
-\textcolor{orange}{Thanks for pointing this out.
+%\textcolor{orange}{
+Thanks for pointing this out.
 Initially we decided against using rolenames for the sake of readability (e.g., by avoiding clutter near the association ends).
 However, we have realized now that the alternative we used hampered readability even more.
-In that sense, we use now rolenames (an exception is made for the \texttt{Gateway} associations with \texttt{TaskType} as explained in the caption of Figure 6)}
+In that sense, we use now rolenames (an exception is made for the \texttt{Gateway} associations with \texttt{TaskType} as explained in the caption of Figure 6).
+%}
 %fd ok

### 2021-10-18 3-introduction.tex
replace typo VirtalModel
synchro

@@
 
-This strict approach shows its limitations when modeling complex domains requiring more than one level of specialization, \eg to adapt the model to application sub domains.
+This strict approach shows its limitations when modeling complex domains requiring more than one level of specialization, \eg to adapt the model to application subdomains.
 Indeed, the two-level approach fails to acknowledge and support: 1) the existence of different forms of classification and/or instantiation (\eg ontological vs linguistics); and 2) the duality type-object for model elements.
@@
 
-Our modeling approach is based on a language (\FML) which provides the (linguistics) concepts that are used to define  the multi-levels models and meta-models that are needed.
-\FML allows to define models as first class entities (which we call VirtualModel) and to link them together (by extension).
+Our modeling approach is based on a language (\FML) which provides the (linguistics) concepts that are used to define the multi-level models and meta-models that are needed.
+\FML allows defining models as first class entities (which we call \emph{virtual models}) and to link them together (by extension).
 In the \mpc case study, we exploit this possibility to:
@@
 
-Contrary to a classical multi-level approach, in which the notion of multi-level is integrated in the language and must therefore be respected, our approach allows to build multi-levels on demand, and with the appropriate form.
+Contrary to a classical multi-level approach, in which the notion of multi-level is integrated in the language and must therefore be respected, our approach allows building multi-level on demand, and with the appropriate form.
 Without multilevel-specific operators/concepts, the "custom" construction must be done explicitly during the model development and evolution process.

### 2021-10-18 4-technology.tex
synchro

@@
 The following listing shows the \FML code of this \textit{virtual model}.
-{\color{red}Notice that the composition relation of figure~\ref{fig:BPMNSubsetExample} between \texttt{BPMNProcess} and \texttt{BPMNTask} is realized in the FML code by concept containment.
-Whenever an instance of \texttt{BPMNTask} is created, it must be in a container (instance of \texttt{BPMNProcess}) and the two instances get linked together.}\noteFabien{A valider}
+Note that the composition relation of figure~\ref{fig:BPMNSubsetExample} between \texttt{BPMNProcess} and \texttt{BPMNTask} is realized in the FML code by concept containment.
+Whenever an instance of \texttt{BPMNTask} is created, it must be in a container (instance of \texttt{BPMNProcess}) and the two instances get linked together.
+%\noteFabien{A valider}
 

### 2021-10-18 responseLetter.tex
sync overleaf->github

@@
 What are the contexts of the constraints (to which concept are they attached to)?}
-MultiInstance(s) are introduced in our answer metamodel (figure 6).
+%MultiInstance(s) are introduced in our answer metamodel (figure 6).
 As such, their semantic is defined at the ontological level, i.e.
@@
 
-{\color{red}fd: hum, j'ai l'impression qu'il manque un peu de détail pour répondre sur la partie multi-instance.
+%{\color{red}fd: hum, j'ai l'impression qu'il manque un peu de détail pour répondre sur la partie multi-instance.
 Il me semble qu'en plus d'être ad hoc, le mécanisme est manuel (mais je me trompe peut être).

### 2021-10-18 responseLetter.tex
sync response overleaf->github

@@
 
-\todo{SG: Est ce possible de rajouter ce listing ou ca n'avance à rien ?. SM: If you want to add it, it should be ok, but remember then to update the answer text.}
+%non (jc)
+%\todo{SG: Est ce possible de rajouter ce listing ou ca n'avance à rien ?. SM: If you want to add it, it should be ok, but remember then to update the answer text.}
 
@@
 \begin{response}{There are some minor typos throughout the manuscript (eg also regarding the reference Jeusfeld2019; choose -> chose; axis -> axes; aN FML virtual model...; etc.)}
+
 Done.
@@
 In a word, we use FML (linguistic level) as a tool to build models and meta-models permitting us to emulate multi-level modelling.
+We use virtual models to implement the levels.
 The evolution of models (multi-levels) is encoded as a hierarchy of models and meta-models.
@@
 
-\textcolor{teal}{Propale de Fabien pour compléter la réponse ici.
+%\textcolor{teal}{Propale de Fabien pour compléter la réponse ici.
 We have added a discussion (où?) on the fact that our approach reifies the levels using "generic" models and relies on the process of construction to implement the levels of the challenge.
@@
 
-{\color{red} Sylvain: (Nouvelle proposition de réponse) As indicated in the answer to the previous comment we have refactored Section 2 and 3 in order to provide more details and make explanations easier to follow.
+%{\color{red} Sylvain: (Nouvelle proposition de réponse) 
+As indicated in the answer to the previous comment we have refactored Section 2 and 3 in order to provide more details and make explanations easier to follow.
 Linguistic instantiation in this context refers to the use of FML language for our solution.
@@
 On the other side, we exploit the ontologic instantiation through an extensive use of the "Type Object" pattern in two versions (with single and multiple cardinalities), controlling the ontologic instantiation from the type, using a classical factory pattern.
-Additionaly, we agree that the choice of \texttt{InstanceInstance} was not a good decision, we have replaced it with \texttt{ObjectInstance}.}
+Additionaly, we agree that the choice of \texttt{InstanceInstance} was not a good decision, we have replaced it with \texttt{ObjectInstance}.%}
 % jc ok (avec la nouvelle proposition de Sylvain)
@@
 
-{\color{red}Sylvain: (Nouvelle proposition de réponse) MultiInstance(s) are introduced in our answer metamodel (figure 6).
+%{\color{red}Sylvain: (Nouvelle proposition de réponse) 
+MultiInstance(s) are introduced in our answer metamodel (figure 6).
 As such, their semantic is defined at the ontological level, i.e.
@@
 It is an ad hoc mechanism introduced to fulfill requirements.
-This means that corresponding semantics is not handled by FML but in the designed virtual models, with specific behaviours implementing underlying semantics.}
+This means that corresponding semantics is not handled by FML but in the designed virtual models, with specific behaviours implementing underlying semantics.%}
 
@@
 This realizes the ontological instantiation.
-There is no notion "usual" or "default" constructors in FML.
+There is no notion of "usual" or "default" constructors in FML.
 Whenever a user creates a new instance, it has to chose one constructor.
@@
 
+As for the use of the formule \emph{smae pattern}, we have rewritten the concerned sentences so that they now explicitely name the us
+
 Concerning the use of "/type" and "type", the former one is a way to stress that the relation is derived, i.e.
 it is a refinement of the "type" relation defined in another level.
-In the FML code, we use "type" due to the fact there is no notion of level.
-(It is explained in the lasta paragraph of 4.1 on page 8.)}
+In the FML code, we use "type" due to the fact there is no notion of level (this is explained in the last paragraph of 4.1 on page 8.)}
 \textcolor{red}{

### 2021-10-18 responseLetter.tex
changes

@@
 
-As for the use of the formule \emph{smae pattern}, we have rewritten the concerned sentences so that they now explicitely name the us
+As for the use of the formula \emph{same pattern}, we have rewritten the concerned sentences so that they either explicitly name the used pattern or do not refers to patterns at all.
 

### 2021-10-18 6-modelpresentation.tex
fix model presentation in order to remove same pattern expression

@@
 
-This scheme relies on \FML dynamic binding mechanism to delegate to task types the responsibility of instances creation.
-An instance of \texttt{Process} references its unique type \texttt{ProcessType} by the specialized \texttt{/type} role.
-Each instance of \texttt{TaskType} is ontologically instantiated with a \texttt{Task} (\textbf{P11}), using the same pattern where \texttt{TaskType} has the responsibility to manage the ontological instantiation.
-A \texttt{Task} references its unique \texttt{TaskType}, and defines a \texttt{begin date} and an \texttt{end date} basic role (\textbf{P12}).
+This scheme relies on \FML dynamic binding mechanism to delegate to task types
+the responsibility of instances creation. An instance of \texttt{Process}
+references its unique type \texttt{ProcessType} by the specialized
+\texttt{/type} role. Each instance of \texttt{TaskType} is ontologically
+instantiated with a \texttt{Task} (\textbf{P11}), using the same factory
+pattern where \texttt{TaskType} has the responsibility to manage the
+ontological instantiation. A \texttt{Task} references its unique
+\texttt{TaskType}, and defines a \texttt{begin date} and an \texttt{end date}
+basic role (\textbf{P12}).
 
-The same pattern applies for the used and produced artifacts by roles \texttt{usedArtifacts}, \texttt{producedArtifacts} and \texttt{performingActors} defined in \texttt{Task} (\textbf{P13}).
-An instance of \texttt{Artifact} specializes \texttt{MultiInstance} and references a set of \texttt{ArtifactType} through the specialized \texttt{/types} role (\textbf{P14} and \textbf{P16}).
-Likewise, \texttt{Actor} specializes \texttt{MultiInstance} and references a set of \texttt{ActorType} through the specialized \texttt{/types} relation (\textbf{P15}).
+The used and produced artifacts are implemented by roles
+\texttt{usedArtifacts}, \texttt{producedArtifacts} and
+\texttt{performingActors} defined in \texttt{Task} (\textbf{P13}). An instance
+of \texttt{Artifact} specializes \texttt{MultiInstance} and references a set of
+\texttt{ArtifactType} through the specialized \texttt{/types} role
+(\textbf{P14} and \textbf{P16}). Likewise, \texttt{Actor} specializes
+\texttt{MultiInstance} and references a set of \texttt{ActorType} through the
+specialized \texttt{/types} relation (\textbf{P15}).
 
 The semantics is unclear relatively to the instantiation policy for artifacts.
-In the requirements, nothing is explicitly stated about the link between the \emph{artifact type} of a \emph{task type} and the types of the artifacts of a corresponding task.
-Here, we assume that task execution implies that for each used and produced \emph{artifact type} defined in a related \emph{task type}, it exists at least one artifact of the right type.
-The following excerpt of \FML code shows a partial implementation of this.
-The same pattern applies to produced artifacts.
+In the requirements, nothing is explicitly stated about the link between the
+\emph{artifact type} of a \emph{task type} and the types of the artifacts of a
+corresponding task.  Here, we assume that task execution implies that for each
+used and produced \emph{artifact type} defined in a related \emph{task type},
+it exists at least one artifact of the right type. The following excerpt of
+\FML code shows a partial implementation of this. We proceed likewise for
+produced artifacts.
 

### 2021-10-18 responseLetter.tex

--- 2021-10-18 responseLetter.tex
@@
 
-{\color{teal}The behavior \texttt{newProcess} is defined in the \texttt{ProcessType} concept following a \emph{factory} design pattern.
+%{\color{teal}
+The behavior \texttt{newProcess} is defined in the \texttt{ProcessType} concept following a \emph{factory} design pattern.
 Rather than being created directly a process instance is therefore created from its type.
@@
 
-As for the use of the formule \emph{smae pattern}, we have rewritten the concerned sentences so that they now explicitely name the us
+As for the use of the formula \emph{same pattern}, we have rewritten the concerned sentences so that they either explicitly name the used pattern or do not refer to patterns at all.
 
@@
 it is a refinement of the "type" relation defined in another level.
-In the FML code, we use "type" due to the fact there is no notion of level (this is explained in the last paragraph of 4.1 on page 8.)}
-\textcolor{red}{
+In the FML code, we use "type" due to the fact there is no notion of level (this is explained in the last paragraph of 4.1 on page 8.)
+
+The specialization of Actors follows the same approach as the other specializations, therefore we do not understand the question.%}
+%\textcolor{red}{
 %- newProcess est un vrai behaviour défini dans ProcessType (pattern factory) et permet de spécifier l'instantiation d'un Process à partir d'un ProcessType (c'est la partie réalisation de l'instantiation ontologique)
-\begin{itemize}
-\item difficile de répondre sur le "same pattern applies"  (fd finalement, je pense qu'il ne comprend pas bien notre usage de "same pattern" car il a l'impression qu'il y a plusieurs pattern => peut etre faudrait-il les nommer (specialization, factory, ...)) ; (jc) j'ai l'impression que la question est un peu à côté de la plaque
-\item pas compris la question sur la spécialisation des Actors (fd ne comprend pas la question, jc non plus)
-\end{itemize}
-}
+%\begin{itemize}
+%\item difficile de répondre sur le "same pattern applies"  (fd finalement, je pense qu'il ne comprend pas bien notre usage de "same pattern" car il a l'impression qu'il y a plusieurs pattern => peut etre faudrait-il les nommer (specialization, factory, ...)) ; (jc) j'ai l'impression que la question est un peu à côté de la plaque
+%\item pas compris la question sur la spécialisation des Actors (fd ne comprend pas la question, jc non plus)
+%\end{itemize}
+%}
 
@@
 
-\textcolor{orange}{Conceptualization means that a concept is created to fulfill or meet a requirement.
+%\textcolor{orange}{
+Conceptualization means that a concept is created to fulfill or meet a requirement.
 As for the mention to a "dual concept nature", we have rewritten that part as it was indeed confusing.
-Now it reads: the fulfillment of a requirement is obtained by the relation between an instance and its type definition instance.}
+Now it reads: the fulfillment of a requirement is obtained by the relation between an instance and its type definition instance.%}
 %jc ok
@@
 
-{\color{teal}hum, je crois qu'il n'a pas bien compris mais c'est difficile de le dire comme cela.
+%{\color{teal}hum, je crois qu'il n'a pas bien compris mais c'est difficile de le dire comme cela.
 je pense que l'on pourrait indiquer que chaque niveau est représenté par un virtual model et donc ajouter un niveau se ferait en ajoutant un virtual model.
@@
 
-{\color{red}Sylvain: (Nouvelle proposition de réponse) Indeed, our solution relies on a ad-hoc architecture where multi-level is emulated.
+%{\color{red}Sylvain: (Nouvelle proposition de réponse)
+Indeed, our solution relies on a ad-hoc architecture where multi-level is emulated.
 However, our solution presents reified layers though virtual model inheritance hierarchy.
@@
 This is made possible by the fact that FML allows one to work on different levels at a time.
-The limitations of this approach resides in a good comprehension of the architecture and a strong expertise to maintain global consistency.}
+The limitations of this approach resides in a good comprehension of the architecture and a strong expertise to maintain global consistency.%}
 %jc ok (j'ai probablement loupé la première réponse)
@@
 \begin{itemize}
-    \item Are there any tools supporting model federation besides Openflexo? \textcolor{orange}{To the best of our knowledge there are not other model federation tools comparable to Openflexo/FML.}
+    \item Are there any tools supporting model federation besides Openflexo? \textcolor{blue}{To the best of our knowledge there are not other model federation tools comparable to Openflexo/FML.}
     \item Section 2.
-"In contrast to approaches that compose metamodels into a single large metamodel grouping all needed entities,…" - I do not see, why this is relevant here \textcolor{orange}{Our intention was to compare the federation approach to approaches that integrate all metamodels together.
+"In contrast to approaches that compose metamodels into a single large metamodel grouping all needed entities,…" - I do not see, why this is relevant here \textcolor{blue}{Our intention was to compare the federation approach to approaches that integrate all metamodels together.
 We argue that reifying models (with virtual models) helps structuring, organizing, federating models.}
-    \item Figure 5 - concept Sequencing - the roles "in" and "out" are not represented by lines… why? \textcolor{orange}{As explained in a previous answer, we do this for the sake of readability.}
+    \item Figure 5 - concept Sequencing - the roles "in" and "out" are not represented by lines… why? \textcolor{blue}{As explained in a previous answer, we do this for the sake of readability.}
     \item What does the term "sub-concept" mean exactly?
-Please explain it before the first use \textcolor{orange}{Indeed, this was unclear.
+Please explain it before the first use \textcolor{blue}{Indeed, this was unclear.
 This was related to concept inheritance.
@@
     \item ProgrammingLanguage concept is an enumeration… how is it modelled?
-\textcolor{orange}{ProgrammingLanguages is modelled as an enumeration of names (Java, Cobol, etc) using the notion of enumeration of the FML language.
+\textcolor{blue}{ProgrammingLanguages is modelled as an enumeration of names (Java, Cobol, etc.) using the notion of enumeration of the FML language.
 }
-    \item "The Acme metamodel also defines a specialized task type Coding" - isn’t it CodingTask? \textcolor{orange}{Yes, this is correct. Coding is actually a CodingTask, as shown in figure 7.}
+    \item "The Acme metamodel also defines a specialized task type Coding" - isn’t it CodingTask? \textcolor{blue}{Yes, this is correct. Coding is actually a CodingTask, as shown in figure 7.}
     \item "The value of the property producedArtifactTypes" - is it property, or role?
-\textcolor{orange}{Indeed, this was unclear.
+\textcolor{blue}{Indeed, this was unclear.
 Figure 2 has been completed to show all kind of properties (\texttt{FlexoRole} is a specialization of \texttt{FlexoProperty}).
-The two terms "property" and "role" are equivalent in the context of \texttt{FlexoRole} use.
-}
+The two terms "property" and "role" are equivalent in the context of \texttt{FlexoRole} use.}
     % fd: TODO, Syl-vain, il me semble que c’est la seule occurrence de property,on pourrait basculer sur role il me semble
@@
 
-\textcolor{orange}{The reviewer is absolutely right.
+%\textcolor{orange}{
+The reviewer is absolutely right.
 We decided to use the type-object pattern as it was easier to integrate as an ad-hoc mechanism in our solution to the challenge than other multi-level enabling mechanisms.
 In that sense we would like to remark that this, together with the \emph{virtual model} notion of the Openflexo/FML framework sufficed to meet all the requirements of the challenge.
-Note however that we intend to explore the integration of other mechanisms (such as clabjets or potency) as future work.}
+Note however that we intend to explore the integration of other mechanf ”usual” or ”default” constructors in FML.
+Whenever auser creates a new instance, it has to chose one constructor.
+Here, we imposemethodologically to the user to crisms (such as clabjets or potency) as future work.%}
 %fd ok
@@
 It would be more abstract than the current 'method' style semantics.}
-{\color{red}
+
+%{\color{red}
 We designed our answer with the Type/Instance pattern. We think, we could have implemented it with another approach such as Clabject.
@@
 Our implementation of the challenge was also an opportunity to showcase the expressivity of FML.
-So we tried, in our solution, to reuse several mechanism of FML.
-For example, sometime we use behavior ({\color{teal} P17, P18 and P19}) and sometimes we use constraints ({\color{teal} P9 and P17}).}
+So we tried, in our solution, to reuse several mechanisms of FML.
+For example, sometimes we use behaviors (P17, P18 and P19) and sometimes we use constraints (P9 and P17).%}
 
@@
 
-\textcolor{orange}{We have refactored the subsections of Section 6 in order to include concrete examples and references to the description of our solution whenever it was possible and/or relevant.
-Concretely, subsections 6.2, 6.3, 6.4, 6.6, 6.7 and 6.11 have been updated.}
+%\textcolor{orange}{
+We have refactored the subsections of Section 6 in order to include concrete examples and references to the description of our solution whenever it was possible and/or relevant.
+Concretely, subsections 6.2, 6.3, 6.4, 6.6, 6.7 and 6.11 have been updated.%}
 
-\todo[inline]{Here we have to explain how openflexo helps to create tooling.
+%\todo[inline]{Here we have to explain how openflexo helps to create tooling.
 Just a line or two should suffice.
@@
 Then just copy answer below from: Concerning...}
+
+Concerning the fact that "MLM is the key to the ability of Openflexo to create tooling", the reviewer probably read "MLM" instead of "FML" (Federation Modeling Language).
+The tooling is built over the federation approach but it is not detailed because we think it is out of scope.
 
@@
 
-\textcolor{orange}{While it is true that model slots and other features of Openflexo/FML are not used in order to build our solution,  we believe including them in our description help readers understand the framework.
+%\textcolor{orange}{
+While it is true that model slots and other features of Openflexo/FML are not used in order to build our solution,  we believe including them in our description help readers understand the framework.
 We therefore chose to keep the description as is.
-While explicitly expressing that they are only marginally used in our solution (only for the tooling).}
+While explicitly expressing that they are only marginally used in our solution (only for the tooling).%}
 \end{response}
@@
 (a) In fig 5: why are the in/out gateway features shown as properties and not as associations?
-\textcolor{orange}{This is done for the sake of readability in order to avoid cluttering the diagram with (possibly uninteresting) associations.}
 
-(b) On p. 7 'specialization' -> 'specializations'.}
-Done.
+\textcolor{blue}{This is done for the sake of readability in order to avoid cluttering the diagram with (possibly uninteresting) associations.}
+
+(b) On p. 7 'specialization' -> 'specializations'. \textcolor{blue}{Done.}} 
 \end{response}

### 2021-10-18 10-conclusions.tex
Antidote done

@@
 The features of \FML enable a great modeling flexibility which, coupled with an appropriate methodology, allow us to achieve 
-the multi/level capabilities without dedicated tooling in order to solve problems which require it. We use virtual models to "implement" levels on demand.
+the multi-level capabilities without dedicated tooling, in order to solve problems which require it. We use virtual models to "implement" levels on demand.
 %\noteSM{Say that the flexibility of our approach coupled with a methodology allows us to achieve multi/level capabilities without specific tooling in order to solve problems which require it.}

### 2021-10-18 3-introduction.tex
Antidote done

@@
 Indeed, the two-level approach fails to acknowledge and support: 1) the existence of different forms of classification and/or instantiation (\eg ontological vs linguistics); and 2) the duality type-object for model elements.
-Therefore, while it may still be used for the aforementioned complex modeling tasks, the strict approach entails the introduction of accidental complexity in both the modeling process and the resulting modeling artefacts.
+Therefore, while it may still be used for the aforementioned complex modeling tasks, the strict approach entails the introduction of accidental complexity in both the modeling process and the resulting modeling artifacts.
 

### 2021-10-18 5-analysis.tex
Antidote done

@@
 
-Based on this approach, we organize our set of models following the architecture of the aforementioned Figure \ref{fig:MultilevelArchitecture}.
-The resulting multi-level architecture is organized in two virtual models, one for the core concepts metamodel containing the definition of Processes and Tasks, and one, the Acme metamodel, extending the previous model to specialize the Acme definitions.
+Based on this approach, we organize our set of models following the architecture of the aforementioned figure \ref{fig:MultilevelArchitecture}.
+The resulting multi-level architecture is organized in two virtual models, one defining the core concepts metamodel including the definition of Processes and Tasks, and one, the Acme metamodel, extending the previous model to specialize the Acme definitions.
 The \FML language defines multiple inheritance concept between virtual models as illustrated in the section~\ref{sec:AcmeSoftwareDevelopmentProcess}.

### 2021-10-18 6-modelpresentation.tex
Antidote done

@@
 
-The designed multilevel architecture shown in Figure~\ref{fig:MultilevelArchitecture} captures the two use cases described in the Process Challenge.
+The designed multilevel architecture shown in figure~\ref{fig:MultilevelArchitecture} captures the two use cases described in the Process Challenge.
 
@@
 
-The Figure~\ref{fig:AcmeFullArchitecture} highlights all conceptual levels of the architecture.
+The figure~\ref{fig:AcmeFullArchitecture} highlights all conceptual levels of the architecture.
 This figure is only partial at instance level and not all instances are represented.
@@
 
-The requirement \textbf{S8} is simply expressed by the definition of the \inst{Testing} instance of \texttt{SETaskType}, \inst{Tester} instance of \texttt{ActorType}, and \inst{Test report} instance of \texttt{ArtifactType}, as showed in the figure \ref{fig:AcmeSoftwareDevelopmentProcess}.
+The requirement \textbf{S8} is simply expressed by the definition of the \inst{Testing} instance of \texttt{SETaskType}, \inst{Tester} instance of \texttt{ActorType}, and \inst{Test report} instance of \texttt{ArtifactType}, as shown in the figure \ref{fig:AcmeSoftwareDevelopmentProcess}.
 
@@
 
-We took advantage of model federation and the availability of diagraming features though the \textit{Diagramming Technology Adapter} to implement two interactive graphical tools built on top of the conceptual levels detailed in previous sections.
+We took advantage of model federation and the availability of diagraming features though the \textit{Diagraming Technology Adapter} to implement two interactive graphical tools built on top of the conceptual levels detailed in previous sections.
 The figure \ref{fig:ToolingArchitecture} shows the overview of the tooling architecture and is detailed in the next section.
@@
 All tasks are created from their \texttt{TaskType} definition, and required assignments apply (if for example \inst{Coding} \texttt{TaskType} defines COBOL as output programming language, the related task is automatically assigned to \inst{Ann Smith}).
-Each task has a status as well as a set of performing actors, a begin and end date, and used and produced artifacts.
+Each task has a status as well as a set of performing actors, a begin and end dates, and used and produced artifacts.
 
-A demonstrative video for the proposed tooling is available on our website\footnote{\url{https://research.openflexo.org/MLMChallenge.html}}, as well as download and installation instructions.
+A demonstrative video for the developed tooling is available on our website\footnote{\url{https://research.openflexo.org/MLMChallenge.html}}, as well as download and installation instructions.

### 2021-10-18 8-discussion.tex
Antidote done

@@
 
-In this section, we discuss our multilevel model solution with regard to the required aspects
+In this section, we discuss our multi-level model solution with regard to the required aspects
 mentioned in the challenge.
@@
 
-  The Openflexo approach is level-agnostic: ``levels'' have no specific nature
+  The Openflexo approach is level agnostic: ``levels'' have no specific nature
   and there are no numbered levels. In our solution, concepts of a given level
@@
 %  ACME metamodel and the ACME model.
-Figure~\ref{fig:MultilevelArchitecture} shows the multilevel architecture of
+Figure~\ref{fig:MultilevelArchitecture} shows the multi-level architecture of
 our solution.
@@
   linked each other in a transparent way, using inheritance or instantiation.
-  For instance in the figure~\ref{fig:AcmeFullArchitecture}, the
+  For instance, in the figure~\ref{fig:AcmeFullArchitecture}, the
   \texttt{Developer} concept from \texttt{Acme} metamodel specializes the
@@
   users. Thus, the integrity of the contents relies on the users when they
-  build the metamodels and the models. The section~\ref{sec:model} describes
+  build the metamodels and the models. Section~\ref{sec:model} describes
   how the requirements are captured and how behaviors are implemented with

### 2021-10-18 9-relatedworks.tex
typo a FML, UML + trailing space
Antidote done

@@
 The first level describes generic processes, the second level software engineering processes, and the third level, software engineering enacted processes.
-Additionally, the authors use~\emph{linguistic extensions}~\citep{metadepth}, a mechanism to  linguistically extend ontological instance models at any level, in order to introduce artefact types and tasks duration in the second level.
+Additionally, the authors use~\emph{linguistic extensions}~\citep{metadepth}, a mechanism to  linguistically extend ontological instance models at any level, in order to introduce artifact types and tasks duration in the second level.
 Our solution is similar to theirs w.r.t.
 the grouping and organization of concepts (e.g., a model for general process, concepts, an extension for software engineering process) but it replaces the use of clabjets and potency with the use of the type-object pattern~\citep{typeObject}.
-The so called \emph{linguistic extensions} are natively supported in Openflexo/FML.
+The so-called \emph{linguistic extensions} are natively supported in Openflexo/\FML.
 
@@
 \cite{multiecore2019} use MultEcore \citep{multecore2016} to provide a solution to the challenge.
-MultEcore uses, (un)pluggable linguistic levels (e.g, there is not a \emph{fixed} core metamodel defining the concept of clabjet), an extension of the two-level cascading technique \citep{atkinson2005concepts} (this is achieved with model transformations to transform instance models into instantiable models) and potency.
+MultEcore uses (un)pluggable linguistic levels (\eg, there is not a \emph{fixed} core metamodel defining the concept of clabjet), an extension of the two-level cascading technique \citep{atkinson2005concepts} (this is achieved with model transformations to transform instance models into instantiable models) and potency.
 Regarding the challenge, their solution uses 4 levels.
@@
 The third level refines the aforementioned models to adapt them to the ACME and Xsure processes.
-Finally, process instances lie in level four.
+Finally, process instances lie in level 4.
 Additionally, the authors use an orthogonal linguistic hierarchy to support alternative names for every model element.
@@
 This sub-module contains in turn a sub-module representing the coding process type whereas concrete coding processes are represented in subsequent sub-modules.
-There are two main differences between the DeepTelos and Openflexo/\FML: 1) we use the type-object pattern instead of the powertype pattern; 2) DeepTelos reify their support for multi-level modeling while ours remains ad-hoc.
+There are two main differences between the DeepTelos and Openflexo/\FML: 1) we use the type-object pattern instead of the powertype pattern; 2) DeepTelos reifies their support for multi-level modeling while our remains ad hoc.
 
@@
 Finally, \cite{dmla2019} use the Dynamic Multi-Layer Algebra (DMLA) \parencite{dmla2017} in order to solve the process challenge.
-DMLA is a level-blind modeling framework which is fully customizable (\eg different types of instantiation may be implemented) and includes support for deep characterization (instantiation may refer to any other element disregarding hierarchy levels and a sort of potency in the form of fluid metamodeling at the entity level exists).
+DMLA is a level blind modeling framework which is fully customizable (\eg different types of instantiation may be implemented) and includes support for deep characterization (instantiation may refer to any other element disregarding hierarchy levels and a sort of potency in the form of fluid metamodeling at the entity level exists).
 Their proposed solution separates the process challenge in two separate domains, namely, the task definition domain and the process definition domain.

### 2021-10-18 responseLetter.tex
sync overleaf->github

@@
 \item Improving the presentation.
-\item Providing a more explicit description of how the multi-level modelling emulation is achieved.
+\item Providing a more explicit description of how the multi-level modeling emulation is achieved.
 \item Discussing the general advantages and limitations of the multi-level emulation, including a clarification of what the implications would be of supporting more "levels" than the challenge required.
@@
 
-We have improved the presentation of both, Openflexo/FML and our solution to the challenge so that: 1) the techniques we use to \emph{achieve} multi-level capabilities are more explicit; and, 2) the advantages and disadvantages of using them in order to fulfil the challenge requirements become clearer.
+We have improved the presentation of both, Openflexo/FML and our solution to the challenge so that: 1) the techniques we use to \emph{achieve} multi-level capabilities are more explicit; and, 2) the advantages and disadvantages of using them in order to fulfill the challenge requirements become clearer.
 Notably, we have added many details in the final paragraphs of the introduction section, improved the Section 2, devoted to the description of Openflexo, refactored figures 2, 4, 5, 6, 7 and 9 and, added figures 3 and 10.
@@
 %\textcolor{orange}{
-In the answer to the previous editor's comment we describe the improvements of our revision, which includes an emphasis in discussing trade-offs related to the emulation of multi-level modeling by using the Openflexo/FML and the type-object pattern.
+In the answer to the previous editor's comment, we describe the improvements of our revision, which includes an emphasis in discussing trade-offs related to the emulation of multi-level modeling by using the Openflexo/FML and the type-object pattern.
 Additionally, the related work section has been extended in order to enhance the comparison between our approach and previous solutions to the challenge.%}
@@
 Note that connection with external sources is not required for our solution to the challenge and is only represented in order to illustrate the capabilities of Openflexo/FML.
-Therefore, in the article we choose not to give this level of details and focus on the left part of the figure.
+Therefore, in the article we choose not to give this level of detail and focus on the left part of the figure.
 % Fabien OK pour moi
@@
 We also have correspondingly improved the text explaining the figure.
-We hope this changes make understanding the core concepts of the FML language easier.
+We hope these changes make understanding the core concepts of the FML language easier.
 %}
@@
 %\textcolor{orange}{
-We fully agree with the reviewer on the fact that it is the flexibility of the model federation approach that enables us to provide a solution to the multi-level process challenge while using a tool not specifically designed for multi-level modelling.
-Of course, this comes together with a number of trade-offs which we discuss in the paper (\emph{e.g.}, see the the end of the introduction section).%}
+We fully agree with the reviewer on the fact that it is the flexibility of the model federation approach that enables us to provide a solution to the multi-level process challenge while using a tool not specifically designed for multi-level modeling.
+Of course, this comes together with a number of trade-offs which we discuss in the paper (\emph{e.g.}, see the end of the introduction section).%}
 %fd ok
@@
 We have included at the end of the introduction section a more detailed description of our approach.
-In a word, we use FML (linguistic level) as a tool to build models and meta-models permitting us to emulate multi-level modelling.
+In a word, we use FML (linguistic level) as a tool to build models and meta-models permitting us to emulate multi-level modeling.
 We use virtual models to implement the levels.
 The evolution of models (multi-levels) is encoded as a hierarchy of models and meta-models.
-In this sense, we do on-demand multi-level modelling, and propose an organization guided by the evolution of models.%}
+In this sense, we do on demand multi-level modeling, and propose an organization guided by the evolution of models.%}
 % jc ok
@@
 We have refactored the Technology section in order to make it, hopefully, easier to follow.
-Apart from improvements in the text, we have notably improved Figure 2 so that now a lot more of details are visible and contributed a new figure (Figure 3) showing how the FML concepts are used in a classical modeling scenario.
-Note however that we use model federation only for the tooling part of our answer to the challenge.
+Apart from improvements in the text, we have notably improved Figure 2 so that now a lot more details are visible and contributed a new figure (Figure 3) showing how the FML concepts are used in a classical modeling scenario.
+Note, however, that we use model federation only for the tooling part of our answer to the challenge.
 In practice, we rely solely in the reification of models through the \textsf{VirtualModel} concept in order to emulate multi-levels as shown in Figure 4.%}
@@
 %{\color{red} Sylvain: (Nouvelle proposition de réponse) 
-As indicated in the answer to the previous comment we have refactored Section 2 and 3 in order to provide more details and make explanations easier to follow.
+As indicated in the answer to the previous comment, we have refactored Section 2 and 3 in order to provide more details and make explanations easier to follow.
 Linguistic instantiation in this context refers to the use of FML language for our solution.
@@
 We tried to show how the use of FML helps in meeting requirements of the challenge.
-On the other side, we exploit the ontologic instantiation through an extensive use of the "Type Object" pattern in two versions (with single and multiple cardinalities), controlling the ontologic instantiation from the type, using a classical factory pattern.
-Additionaly, we agree that the choice of \texttt{InstanceInstance} was not a good decision, we have replaced it with \texttt{ObjectInstance}.%}
+On the other side, we exploit the ontological instantiation through an extensive use of the "Type Object" pattern in two versions (with single and multiple cardinalities), controlling the ontological instantiation from the type, using a classical factory pattern.
+Additionally, we agree that the choice of \texttt{InstanceInstance} was not a good decision, we have replaced it with \texttt{ObjectInstance}.%}
 % jc ok (avec la nouvelle proposition de Sylvain)
@@
 Thanks for pointing this out.
-Initially we decided against using rolenames for the sake of readability (e.g., by avoiding clutter near the association ends).
+Initially we decided against using role names for the sake of readability (e.g., by avoiding clutter near the association ends).
 However, we have realized now that the alternative we used hampered readability even more.
-In that sense, we use now rolenames (an exception is made for the \texttt{Gateway} associations with \texttt{TaskType} as explained in the caption of Figure 6).
+In that sense, we now use role names (an exception is made for the \texttt{Gateway} associations with \texttt{TaskType} as explained in the caption of Figure 6).
 %}
@@
 It is an ad hoc mechanism introduced to fulfill requirements.
-This means that corresponding semantics is not handled by FML but in the designed virtual models, with specific behaviours implementing underlying semantics.%}
+This means that corresponding semantics is not handled by FML but in the designed virtual models, with specific behaviors implementing underlying semantics.%}
 
@@
 
-The constraints are defined at FlexoConcept level, and are designed of invariant being evaluated by the FML interpreter, which may raise constraint violation at run-time.
+The constraints are defined at FlexoConcept level, and are designed with invariant being evaluated by the FML interpreter, which may raise constraint violation at run-time.
 An example of constraint is presented for requirement P9, in \texttt{TaskType} concept (p 8).
@@
 There is no notion of "usual" or "default" constructors in FML.
-Whenever a user creates a new instance, it has to chose one constructor.
+Whenever a user creates a new instance, it has to choose one constructor.
 Here, we impose methodologically to the user to create a process instance through the chosen process type.
 There is no enforcement in our solution.
-SYLVAIN ?
 
-As for the use of the formula \emph{same pattern}, we have rewritten the concerned sentences so that they either explicitly name the used pattern or do not refer to patterns at all.
+As for the use of the formula \emph{same pattern}, we have rewritten the concerned sentences so that they either explicitly name the used pattern, or do not refer to patterns at all.
 
@@
 it is a refinement of the "type" relation defined in another level.
-In the FML code, we use "type" due to the fact there is no notion of level (this is explained in the last paragraph of 4.1 on page 8.)
+In the FML code, we use "type" due to the fact there is no notion of levels (this is explained in the last paragraph of 4.1 on page 8.)
 
@@
 The business logic of the diagram tooling is encoded in FML, and more specifically for this editor in the \texttt{ProcessEditor} virtual model presented in figure 9.
-Structure of diagramming, interactions and presentation layer (\emph{e.g.} texts) are defined in various concepts defined in the virtual model \texttt{ProcessEditor}.
+Structure of diagraming, interactions and presentation layer (\emph{e.g.} texts) are defined in various concepts defined in the virtual model \texttt{ProcessEditor}.
 %fd ok
@@
 Conceptualization means that a concept is created to fulfill or meet a requirement.
-As for the mention to a "dual concept nature", we have rewritten that part as it was indeed confusing.
+As for the mention of a "dual concept nature", we have rewritten that part as it was indeed confusing.
 Now it reads: the fulfillment of a requirement is obtained by the relation between an instance and its type definition instance.%}
@@
 %{\color{red}Sylvain: (Nouvelle proposition de réponse)
-Indeed, our solution relies on a ad-hoc architecture where multi-level is emulated.
-However, our solution presents reified layers though virtual model inheritance hierarchy.
-Although there is no automatized support, the proposed methodology allows to provide an arbitrary number of layers and some extension capabilities and reuse support (e.g.
-adding a level would mean creating a new virtual model extending an existing one).
+Indeed, our solution relies on a ad hoc architecture where multi-level is emulated.
+However, our solution presents reified layers through virtual model inheritance hierarchy.
+Although there is no automatized support, the proposed methodology allows providing an arbitrary number of layers and some extension capabilities and reuse support (e.g., adding a level would mean creating a new virtual model extending an existing one).
 This is made possible by the fact that FML allows one to work on different levels at a time.
@@
 The reviewer is absolutely right.
-We decided to use the type-object pattern as it was easier to integrate as an ad-hoc mechanism in our solution to the challenge than other multi-level enabling mechanisms.
-In that sense we would like to remark that this, together with the \emph{virtual model} notion of the Openflexo/FML framework sufficed to meet all the requirements of the challenge.
-Note however that we intend to explore the integration of other mechanf ”usual” or ”default” constructors in FML.
-Whenever auser creates a new instance, it has to chose one constructor.
-Here, we imposemethodologically to the user to crisms (such as clabjets or potency) as future work.%}
+We decided to use the type-object pattern as it was easier to integrate as an ad hoc mechanism in our solution to the challenge than other multi-level enabling mechanisms.
+In that sense we would like to remark that this, together with the \emph{virtual model} notion of the Openflexo/FML framework, is sufficient to meet all the requirements of the challenge.
+Note, however, that we intend to explore the integration of other mechanisms (such as clabjets or potency) as future work.%}
 %fd ok

### 2022-02-23 1-main.tex
Polishing related work.
sync (subtitle)
pdf metadata ✔

@@
 
+\hypersetup{
+  pdfauthor={Sylvain Guérin, Joel Champeau, Jean-Christophe Bach, Antoine
+  Beugnard, Fabien Dagnat, Salvador Martinez},
+  pdftitle={Multi-level modeling with Openflexo/\FML -- A contribution to the \mlpc},
+  pdfkeywords={Multi-Level Modeling Challenge, Federation Modeling Language,
+  Openflexo},
+  %abstract
+  pdfsubject={Model federation is a multi-model \emph{management} approach based on the use of virtual models and loosely coupled links.
+The models in a federation remain autonomous and represented in their original technological spaces whereas virtual models and links (which are not level bounded) serve as control components used to present different views to the users and  maintain synchronization.
+In this paper we tackle the \emph{MULTI process modeling challenge}, which consists in providing a solution to the problem of specifying and enacting processes.
+Solutions must fulfill a number of requirements for a process representation defined at an abstract process-definition level and at various more concrete domain-specific levels, resulting in a multi-level hierarchy of related models.
+We present a solution based on model federation and discuss the advantages and limitations of using this approach for multi-level modeling.
+Concretely, we use virtual models and more precisely the Federation Modeling Language ({FML}) that serves to describe them as the main building block in order to solve the process modeling challenge whereas the federation aspect is used as a means to provide editing tooling for the resulting process language.
+Our solution is fully implemented with the Openflexo framework.
+It fulfills all the challenge requirements.
+}
+}
+
 \begin{document}
@@
     \title{Multi-level modeling with Openflexo/\FML}
-    \subtitle{A Contribution to the \mlpc}
+    \subtitle{A contribution to the \mlpc} % A contribution to the Multi-Level Process Challenge
 
@@
     \abstract{\input{2-abstract}}
-    \keywords{Multilevel Modeling Challenge \and Federation Modeling Language
+    \keywords{Multi-Level Modeling Challenge \and Federation Modeling Language
     \and Openflexo}

### 2022-02-23 3-introduction.tex
style guidelines: quotation marks -> \enquote

@@
 
-Contrary to a classical multi-level approach, in which the notion of multi-level is integrated in the language and must therefore be respected, our approach allows building multi-level on demand, and with the appropriate form.
-Without multilevel-specific operators/concepts, the "custom" construction must be done explicitly during the model development and evolution process.
+Contrary to a classical multi-level approach, in which the notion of
+multi-level is integrated in the language and must therefore be respected, our
+approach allows building multi-level on demand, and with the appropriate form.
+Without multilevel-specific operators/concepts, the \enquote{custom}
+construction must be done explicitly during the model development and evolution
+process.
 

### 2022-02-23 6-modelpresentation.tex
style guidelines: quotation marks -> \enquote

@@
 Figure \ref{fig:ToolingArchitecture} details the architecture of these two tools.
+
 The left of the figure presents the \textit{ProcessType graphical editor}.
-\texttt{ProcessTypeEditor} is modeled as a virtual model declaring two model slots (represented with bold circles).
-The first model slot references an instance of Acme metamodel, while the second model slot references an instance of diagram, conforms to the \textit{ProcessType diagram specification}\footnote{a diagram "metamodel" which defines and specify structure and graphical representations for edited items}.
-When executed, this tool manages a graphical view for an instance of Acme metamodel (the \textit{Acme model}) and a specific \texttt{ProcessType} instance.
-This tool allows representing and editing a \texttt{ProcessType}.
-The highly reflective nature of \FML and its tooling should be noted.
-When the drag and drop interactor is applied for a new item, the tool allows choosing the concept type to be instantiated.
-For example, when a \texttt{TaskType} is created, the user must choose a concept inheriting from \texttt{TaskType} -- in Acme use case it can be \texttt{SETaskTask} or \texttt{CodingTaskType} or default value \texttt{TaskType}).
+\texttt{ProcessTypeEditor} is modeled as a virtual model declaring two model
+slots (represented with bold circles). The first model slot references an
+instance of Acme metamodel, while the second model slot references an instance
+of diagram, conforms to the \textit{ProcessType diagram
+specification}\footnote{a diagram \enquote{metamodel} which defines and specify
+structure and graphical representations for edited items}. When executed, this
+tool manages a graphical view for an instance of Acme metamodel (the
+\textit{Acme model}) and a specific \texttt{ProcessType} instance. This tool
+allows representing and editing a \texttt{ProcessType}. The highly reflective
+nature of \FML and its tooling should be noted. When the drag and drop
+interactor is applied for a new item, the tool allows choosing the concept type
+to be instantiated. For example, when a \texttt{TaskType} is created, the user
+must choose a concept inheriting from \texttt{TaskType} -- in Acme use case it
+can be \texttt{SETaskTask} or \texttt{CodingTaskType} or default value
+\texttt{TaskType}).
 

### 2022-02-23 10-conclusions.tex
style guidelines: quotation marks -> \enquote
style guidelines: * [sS]ection -> Sect. * [fF]igure -> Fig. * [tT]able -> Tab. * few s/ \ref/~\ref/

@@
 including models and metamodels, thanks to the flexibility provided by the
-metametamodeling infrastructure offered by Openflexo.
-This infrastructure helped us to overcome the limitations of the strict modeling paradigm mentioned in Section \ref{sec:introduction}, this is, the lack of support for different forms of classifications and for the duality type/object.
-In that sense, our solution seamlessly integrates the type-object pattern which allows us to dynamically create different \emph{instances} of model elements (e.g., \emph{TaskType}, among others) and use them for typing other model elements.
+metametamodeling infrastructure offered by Openflexo. This infrastructure
+helped us to overcome the limitations of the strict modeling paradigm mentioned
+in Sect.~\ref{sec:introduction}, this is, the lack of support for different
+forms of classifications and for the duality type/object. In that sense, our
+solution seamlessly integrates the type-object pattern which allows us to
+dynamically create different \emph{instances} of model elements (\eg, \emph{TaskType}, among others) and use them for typing other model elements.
 This, together with the \FML core support for (level agnostic) classical linguistic instantiation, permitted us to merge linguistic and ontological instantiation in order to provide a satisfactory solution to the Multi-Level Process challenge.
@@
 The features of \FML enable a great modeling flexibility which, coupled with an appropriate methodology, allow us to achieve 
-the multi-level capabilities without dedicated tooling, in order to solve problems which require it. We use virtual models to "implement" levels on demand.
+the multi-level capabilities without dedicated tooling, in order to solve
+problems which require it. We use virtual models to \enquote{implement} levels on demand.
+
 %\noteSM{Say that the flexibility of our approach coupled with a methodology allows us to achieve multi/level capabilities without specific tooling in order to solve problems which require it.}

### 2022-02-23 3-introduction.tex
style guidelines: * [sS]ection -> Sect. * [fF]igure -> Fig. * [tT]able -> Tab. * few s/ \ref/~\ref/

@@
 
-The rest of the paper is organized as follows.
-Section~\ref{sec:technology} presents our approach and the technology it relies on.
-Section~\ref{sec:analysis} is the analysis of the problem.
-We detail our model in section~\ref{sec:model} and show how it satisfies the challenge requirements in section~\ref{sec:requirements}.
-We assess the modeling solution in section~\ref{sec:discussion}.
-Section~\ref{sec:relatedwork} presents the related work before the conclusions.%~\ref{sec:conclusions}.
+The rest of the paper is organized as follows. Sect.~\ref{sec:technology}
+presents our approach and the technology it relies on. Sect.~\ref{sec:analysis}
+is the analysis of the problem. We detail our model in Sect.~\ref{sec:model}
+and show how it satisfies the challenge requirements in
+Sect.~\ref{sec:requirements}. We assess the modeling solution in
+Sect.~\ref{sec:discussion}. Sect.~\ref{sec:relatedwork} presents the related work before the conclusions.%~\ref{sec:conclusions}.
 

### 2022-02-23 4-technology.tex
style guidelines: * [sS]ection -> Sect. * [fF]igure -> Fig. * [tT]able -> Tab. * few s/ \ref/~\ref/

@@
 
-This framework relies on the architecture of figure~\ref{fig:mf}. A federation
+This framework relies on the architecture of Fig.~\ref{fig:mf}. A federation
 gathers a set of conceptual models, named \emph{virtual models} and a
@@
 
-A \UML like representation of a simplified metamodel of \FML is provided in figure~\ref{fig:mm}.
+A \UML like representation of a simplified metamodel of \FML is provided in Fig.~\ref{fig:mm}.
 \FML is a language designed to define models as virtual models.
@@
 
-To illustrate this, figure~\ref{fig:BPMNSubsetExample} shows how to realize a small subset of BPMN with Openflexo/FML.
+To illustrate this, Fig.~\ref{fig:BPMNSubsetExample} shows how to realize a small subset of BPMN with Openflexo/FML.
 Concretely, we have designed a \textit{virtual model} (\texttt{SimplifiedBPMN}) with 3 \textit{concepts} representing a process and its contained tasks.
@@
 The following listing shows the \FML code of this \textit{virtual model}.
-Note that the composition relation of figure~\ref{fig:BPMNSubsetExample} between \texttt{BPMNProcess} and \texttt{BPMNTask} is realized in the FML code by concept containment.
+Note that the composition relation of Fig.~\ref{fig:BPMNSubsetExample} between \texttt{BPMNProcess} and \texttt{BPMNTask} is realized in the FML code by concept containment.
 Whenever an instance of \texttt{BPMNTask} is created, it must be in a container (instance of \texttt{BPMNProcess}) and the two instances get linked together.

### 2022-02-23 5-analysis.tex
style guidelines: * [sS]ection -> Sect. * [fF]igure -> Fig. * [tT]able -> Tab. * few s/ \ref/~\ref/

@@
 %for our federation.
-As shown in the figure~\ref{fig:MultilevelArchitecture}, we defined two virtual models (Base and Acme processes abstractions) instantiated by two virtual model instances (XSure and Acme processes).
+As shown in the Fig.~\ref{fig:MultilevelArchitecture}, we defined two virtual models (Base and Acme processes abstractions) instantiated by two virtual model instances (XSure and Acme processes).
 As the figure shows the virtual models play the role of metamodels, in the sense of concept definition with a level-agnostic approach.
@@
 
-Our analysis of the use case leads to identify two main modeling axes (as presented in the figure \ref{fig:LinguisticAndOntologicInstantiation}):
+Our analysis of the use case leads to identify two main modeling axes (as presented in the Fig.~\ref{fig:LinguisticAndOntologicInstantiation}):
 \begin{itemize}
-    \item The horizontal axis is characterized as the ontological instantiation axis, in the sense that the domain type definition (\ie \texttt{ProcessType}) is referenced by an instance definition (\ie \texttt{Process}).
-In our base metamodel, each domain type definition is referenced by its instance definition, as developed in the section \ref{sec:ProcessEnactment}
+    \item The horizontal axis is characterized as the ontological instantiation
+      axis, in the sense that the domain type definition (\ie
+      \texttt{ProcessType}) is referenced by an instance definition (\ie
+      \texttt{Process}). In our base metamodel, each domain type definition is
+      referenced by its instance definition, as developed in the Sect.~\ref{sec:ProcessEnactment}
 
-\item The vertical axis is viewed as the linguistic instantiation.
-The Metamodel (a virtual model) can be instantiated as a model thanks to the instantiation mechanism of \FML.
-This mechanism is similar to the classical object/instance mechanism of the object languages but without any constraint on the referenced virtual model, \ie metamodel.
-The set of resulting instances comes from several virtual models of any abstraction level as detailed in the section~\ref{sec:AcmeSoftwareDevelopmentProcess}
+\item The vertical axis is viewed as the linguistic instantiation. The
+  Metamodel (a virtual model) can be instantiated as a model thanks to the
+  instantiation mechanism of \FML. This mechanism is similar to the classical
+  object/instance mechanism of the object languages but without any constraint
+  on the referenced virtual model, \ie metamodel. The set of resulting
+  instances comes from several virtual models of any abstraction level as
+  detailed in the Sect.~\ref{sec:AcmeSoftwareDevelopmentProcess}
 
@@
 
-Based on this approach, we organize our set of models following the architecture of the aforementioned figure \ref{fig:MultilevelArchitecture}.
-The resulting multi-level architecture is organized in two virtual models, one defining the core concepts metamodel including the definition of Processes and Tasks, and one, the Acme metamodel, extending the previous model to specialize the Acme definitions.
-The \FML language defines multiple inheritance concept between virtual models as illustrated in the section~\ref{sec:AcmeSoftwareDevelopmentProcess}.
+Based on this approach, we organize our set of models following the
+architecture of the aforementioned Fig.~\ref{fig:MultilevelArchitecture}. The
+resulting multi-level architecture is organized in two virtual models, one
+defining the core concepts metamodel including the definition of Processes and
+Tasks, and one, the Acme metamodel, extending the previous model to specialize
+the Acme definitions. The \FML language defines multiple inheritance concept
+between virtual models as illustrated in the Sect.~\ref{sec:AcmeSoftwareDevelopmentProcess}.
 

### 2022-02-23 6-modelpresentation.tex
style guidelines: * [sS]ection -> Sect. * [fF]igure -> Fig. * [tT]able -> Tab. * few s/ \ref/~\ref/

@@
 
-The designed multilevel architecture shown in figure~\ref{fig:MultilevelArchitecture} captures the two use cases described in the Process Challenge.
+The designed multilevel architecture shown in Fig.~\ref{fig:MultilevelArchitecture} captures the two use cases described in the Process Challenge.
 
-Note that in all following figures, blue is used to represent \FML virtual models and concepts (\textsf{VirtualModel} and \textsf{FlexoConcept} in figure \ref{fig:mm}), see figures \ref{fig:BPMNSubsetExample}, \ref{fig:MultilevelArchitecture}, \ref{fig:LinguisticAndOntologicInstantiation}, \ref{fig:BaseMetamodel}, \ref{fig:AcmeFullArchitecture},  \ref{fig:ToolingArchitecture}, while brown is used for instances of \textsf{FlexoConcept}, see figures \ref{fig:BPMNSubsetExample}, \ref{fig:MultilevelArchitecture}, \ref{fig:LinguisticAndOntologicInstantiation}, \ref{fig:AcmeFullArchitecture},  \ref{fig:ToolingArchitecture}.
+Note that in all following figures, blue is used to represent \FML virtual models and concepts (\textsf{VirtualModel} and \textsf{FlexoConcept} in Fig.~\ref{fig:mm}), see figures \ref{fig:BPMNSubsetExample}, \ref{fig:MultilevelArchitecture}, \ref{fig:LinguisticAndOntologicInstantiation}, \ref{fig:BaseMetamodel}, \ref{fig:AcmeFullArchitecture},  \ref{fig:ToolingArchitecture}, while brown is used for instances of \textsf{FlexoConcept}, see figures \ref{fig:BPMNSubsetExample}, \ref{fig:MultilevelArchitecture}, \ref{fig:LinguisticAndOntologicInstantiation}, \ref{fig:AcmeFullArchitecture},  \ref{fig:ToolingArchitecture}.
 % 3, 4, 5, 6, 7, 8, 9
@@
 
-In this section, we present the base metamodel presented in Figure~\ref{fig:BaseMetamodel} with the \textit{XSure} insurance domain use case, whose partial description was provided in the challenge description.
-The requirements \textbf{P1} to \textbf{P19} of XSure insurance domain are straightforwardly implemented by instantiating the \emph{XSure model}, as an instance of base metamodel (the left side of figure~\ref{fig:MultilevelArchitecture}).
+In this section, we present the base metamodel presented in Fig.~\ref{fig:BaseMetamodel} with the \textit{XSure} insurance domain use case, whose partial description was provided in the challenge description.
+The requirements \textbf{P1} to \textbf{P19} of XSure insurance domain are straightforwardly implemented by instantiating the \emph{XSure model}, as an instance of base metamodel (the left side of Fig.~\ref{fig:MultilevelArchitecture}).
 
-Figure~\ref{fig:BaseMetamodel} represents this base metamodel with a \UML-like formalism, well adapted to represent \FML concepts and their instances.
+Fig.~\ref{fig:BaseMetamodel} represents this base metamodel with a \UML-like formalism, well adapted to represent \FML concepts and their instances.
 A \textit{Concept} of \FML is represented by a \UML class where roles of basic types are attributes.
@@
 
-Our proposition relies on ontological instantiation as presented in figure \ref{fig:LinguisticAndOntologicInstantiation}, with a common root concept \texttt{ModelingElement}.
+Our proposition relies on ontological instantiation as presented in Fig.~\ref{fig:LinguisticAndOntologicInstantiation}, with a common root concept \texttt{ModelingElement}.
 Two types of ontological instantiation are needed to meet the requirements of the challenge.
@@
 
-We first present the process definition part of the base metamodel, located on the left of Figure \ref{fig:BaseMetamodel}.
+We first present the process definition part of the base metamodel, located on the left of Fig.~\ref{fig:BaseMetamodel}.
 
@@
 \label{sec:ProcessEnactment}
-We now present the process enactment part of base metamodel, located right of figure~\ref{fig:BaseMetamodel}.
+We now present the process enactment part of base metamodel, located right of Fig.~\ref{fig:BaseMetamodel}.
 All concepts defined in this subsection are either specializations of the \texttt{Instance} concept (if they have exactly one type) or the \texttt{MultiInstance} concept (when they have several types).
@@
 The Base metamodel described previously is too generic to capture all domain-specific aspects of this use case.
-We chose to complete the architectural hierarchy with a specific virtual model, specific to the Acme software development process metamodel, shown in figure~\ref{fig:AcmeFullArchitecture}.
+We chose to complete the architectural hierarchy with a specific virtual model, specific to the Acme software development process metamodel, shown in Fig.~\ref{fig:AcmeFullArchitecture}.
 The \textit{Acme} metamodel specializes the Base metamodel, and \textit{Acme model} is an instance of the \textit{Acme} metamodel.
@@
 
-The figure~\ref{fig:AcmeFullArchitecture} highlights all conceptual levels of the architecture.
+The Fig.~\ref{fig:AcmeFullArchitecture} highlights all conceptual levels of the architecture.
 This figure is only partial at instance level and not all instances are represented.
@@
 
-The \textit{Software Development Process} for Acme company is presented in the figure~\ref{fig:AcmeSoftwareDevelopmentProcess}.
+The \textit{Software Development Process} for Acme company is presented in the Fig.~\ref{fig:AcmeSoftwareDevelopmentProcess}.
 It is a screen capture of the tool developed for the challenge and described in the next section.
@@
 
-The requirement \textbf{S8} is simply expressed by the definition of the \inst{Testing} instance of \texttt{SETaskType}, \inst{Tester} instance of \texttt{ActorType}, and \inst{Test report} instance of \texttt{ArtifactType}, as shown in the figure \ref{fig:AcmeSoftwareDevelopmentProcess}.
+The requirement \textbf{S8} is simply expressed by the definition of the \inst{Testing} instance of \texttt{SETaskType}, \inst{Tester} instance of \texttt{ActorType}, and \inst{Test report} instance of \texttt{ArtifactType}, as shown in the Fig.~\ref{fig:AcmeSoftwareDevelopmentProcess}.
 
@@
 This must be completed with an association between an artifact produced and the task that validates it.
-This is modeled by the reference to \inst{Test Cases} as a produced artifact type in \inst{Test Case design} and the reference to \inst{Test Cases} as a used artifact in \inst{Test Design review}, as shown in figure~\ref{fig:AcmeSoftwareDevelopmentProcess}.
+This is modeled by the reference to \inst{Test Cases} as a produced artifact type in \inst{Test Case design} and the reference to \inst{Test Cases} as a used artifact in \inst{Test Design review}, as shown in Fig.~\ref{fig:AcmeSoftwareDevelopmentProcess}.
 
@@
 We took advantage of model federation and the availability of diagraming features though the \textit{Diagraming Technology Adapter} to implement two interactive graphical tools built on top of the conceptual levels detailed in previous sections.
-The figure \ref{fig:ToolingArchitecture} shows the overview of the tooling architecture and is detailed in the next section.
+The Fig.~\ref{fig:ToolingArchitecture} shows the overview of the tooling architecture and is detailed in the next section.
 A first tool offers a graphical edition of a Process type.
@@
 
-The Base metamodel is completed with some behaviors implementing an execution semantics for the executed processes.
-All tasks -- instantiated from \texttt{TaskType} for a given enacted process -- manage a status.
-This status is either \texttt{Not startable} (when not assigned to a performing actor or when required input artifacts are not available), \texttt{Startable}, \texttt{Started}, \texttt{Completable} (when all output artifacts are ready) or \texttt{Completed}.
-A \texttt{Task} also manages a set of performing actors, a begin and end date, some used and produced artifacts.
-The \texttt{Gateway} business logic has also been implemented through the implementation of an abstract \texttt{execute(Process)} behavior.
-The management of artifacts whose semantics follows rules defined in section~\ref{sec:ProcessEnactment} (\textbf{P13}) is also implemented.
+The Base metamodel is completed with some behaviors implementing an execution
+semantics for the executed processes. All tasks -- instantiated from
+\texttt{TaskType} for a given enacted process -- manage a status. This status
+is either \texttt{Not startable} (when not assigned to a performing actor or
+when required input artifacts are not available), \texttt{Startable},
+\texttt{Started}, \texttt{Completable} (when all output artifacts are ready) or
+\texttt{Completed}. A \texttt{Task} also manages a set of performing actors, a
+begin and end date, some used and produced artifacts. The \texttt{Gateway}
+business logic has also been implemented through the implementation of an
+abstract \texttt{execute(Process)} behavior. The management of artifacts whose
+semantics follows rules defined in Sect.~\ref{sec:ProcessEnactment} (\textbf{P13}) is also implemented.
 
-Figure \ref{fig:ToolingArchitecture} details the architecture of these two tools.
+Fig.~\ref{fig:ToolingArchitecture} details the architecture of these two tools.
 
@@
 
-The other tool, called \textit{Enacted process graphical editor} and shown on figure~\ref{fig:ScreenshotEnactedProcessEditor}, provides process edition and tasks assignations through a graphical visualization displaying the process being executed.
-This tool is represented on the right side of the figure~\ref{fig:ToolingArchitecture}.
+The other tool, called \textit{Enacted process graphical editor} and shown on Fig.~\ref{fig:ScreenshotEnactedProcessEditor}, provides process edition and tasks assignations through a graphical visualization displaying the process being executed.
+This tool is represented on the right side of the Fig.~\ref{fig:ToolingArchitecture}.
 It follows the same architecture pattern of the \textit{ProcessType graphical editor}, with two model slots referencing both the model and a diagram.

### 2022-02-23 7-requirementssatisfaction.tex
style guidelines: * [sS]ection -> Sect. * [fF]igure -> Fig. * [tT]able -> Tab. * few s/ \ref/~\ref/

--- 2021-10-07 7-requirementssatisfaction.tex
@@
 
-In the previous section~\ref{sec:model}, we demonstrated that our solution satisfies all the challenge requirements.
+In the previous Sect.~\ref{sec:model}, we demonstrated that our solution satisfies all the challenge requirements.
 In this section, we summarize the different techniques used to satisfy them, without repeating all the justifications.
@@
 
-Table~\ref{tab:RequirementsSatisfactionBaseMetamodel} summarizes the requirement coverage for the base metamodel illustrated by XSure insurance use case while table~\ref{tab:RequirementsSatisfactionAcme} shows requirements satisfaction for Acme software engineering process.
+Tab.~\ref{tab:RequirementsSatisfactionBaseMetamodel} summarizes the requirement coverage for the base metamodel illustrated by XSure insurance use case while Tab.~\ref{tab:RequirementsSatisfactionAcme} shows requirements satisfaction for Acme software engineering process.
 Syntactic conformance is more precisely classified into six subcategories:
@@
 
-    \item \emph{Ontological instantiation} (\textit{O.Inst.}): fulfillment of a requirement is obtained by the relation between an instance and its type definition instance (for example the notion of \textit{developer} of \textbf{S3} is implemented by both a concept in \textit{Acme metamodel} and an instance in \textit{Acme model}).
+    \item \emph{Ontological instantiation} (\textit{O.Inst.}): fulfillment of a requirement is obtained by the relation between an instance and its type definition instance (for example the notion of \textit{developer} of \textbf{S3} is implemented by both a concept in \textit{Acme metamodel} and an instance in \textit{Acme model}, see bottom left of Fig.~\ref{fig:AcmeFullArchitecture}).
 
@@
 
-\noteSylvain{updated}
+%\noteSylvain{updated}
 
-%Table~\ref{tab:RequirementsSatisfactionBaseMetamodel}
-%Table~\ref{tab:RequirementsSatisfactionAcme}
+%Tab.~\ref{tab:RequirementsSatisfactionBaseMetamodel}
+%Tab.~\ref{tab:RequirementsSatisfactionAcme}
 

### 2022-02-23 8-discussion.tex
style guidelines: quotation marks -> \enquote
style guidelines: * [sS]ection -> Sect. * [fF]igure -> Fig. * [tT]able -> Tab. * few s/ \ref/~\ref/

@@
   Our solution uses the basic modeling constructs depicted by the \FML core
-  metamodel (Figure~\ref{fig:mm}) and described in
-  the section~\ref{sec:technology}. Models are \emph{virtual models}. They are
+  metamodel (Fig.~\ref{fig:mm}) and described in
+  the Sect.~\ref{sec:technology}. Models are \emph{virtual models}. They are
   composed of \emph{concepts}, being themselves concepts. Concepts may have
@@
   instances of diagram from our diagraming TA). The use of slots has been
-  described in the section~\ref{sec:model} and is illustrated by bold circles in
-  the Figure~\ref{fig:ToolingArchitecture}.
+  described in the Sect.~\ref{sec:model} and is illustrated by bold circles in
+  the Fig.~\ref{fig:ToolingArchitecture}.
 
@@
 %\ref{fig:BPMNSubsetExample}
-%  Figure~\ref{fig:AcmeFullArchitecture} shows three levels: the metamodel, the
+%  Fig.~\ref{fig:AcmeFullArchitecture} shows three levels: the metamodel, the
 %  ACME metamodel and the ACME model.
-Figure~\ref{fig:MultilevelArchitecture} shows the multi-level architecture of
+Fig.~\ref{fig:MultilevelArchitecture} shows the multi-level architecture of
 our solution.
@@
   linked each other in a transparent way, using inheritance or instantiation.
-  For instance, in the figure~\ref{fig:AcmeFullArchitecture}, the
+  For instance, in the Fig.~\ref{fig:AcmeFullArchitecture}, the
   \texttt{Developer} concept from \texttt{Acme} metamodel specializes the
@@
   users. Thus, the integrity of the contents relies on the users when they
-  build the metamodels and the models. Section~\ref{sec:model} describes
+  build the metamodels and the models. Sect.~\ref{sec:model} describes
   how the requirements are captured and how behaviors are implemented with
@@
   \inst{Design Task} instance into a concept whose instances could have been
-  "using agile methodology".
+  \enquote{using agile methodology}.
 

### 2022-02-23 9-relatedworks.tex
style guidelines: * [sS]ection -> Sect. * [fF]igure -> Fig. * [tT]able -> Tab. * few s/ \ref/~\ref/

@@
 
-A first description of process modeling as a multi-level modeling problem was proposed by~\citet{lara2018refactoring} in the context of a catalogue of refactoring for multi-level models (in a simpler form with fewer constraints and requirements w.r.t.
-the challenge version).
-A solution is provided with MetaDepth~\parencite{metadepth}, which supports modeling with any number of levels, dual ontological/linguistic typing and deep characterization through potency~\citep{potency}.
-For their solution to the multi-level process problem they use three levels.
-The first level describes generic processes, the second level software engineering processes, and the third level, software engineering enacted processes.
-Additionally, the authors use~\emph{linguistic extensions}~\citep{metadepth}, a mechanism to  linguistically extend ontological instance models at any level, in order to introduce artifact types and tasks duration in the second level.
-Our solution is similar to theirs w.r.t.
-the grouping and organization of concepts (e.g., a model for general process, concepts, an extension for software engineering process) but it replaces the use of clabjets and potency with the use of the type-object pattern~\citep{typeObject}.
+A first description of process modeling as a multi-level modeling problem was
+proposed by~\citet{lara2018refactoring} in the context of a catalogue of
+refactoring for multi-level models (in a simpler form with fewer constraints
+and requirements w.r.t. the challenge version). A solution is provided with
+MetaDepth~\parencite{metadepth}, which supports modeling with any number of
+levels, dual ontological/linguistic typing and deep characterization through
+potency~\citep{potency}. For their solution to the multi-level process problem
+they use three levels. The first level describes generic processes, the second
+level software engineering processes, and the third level, software engineering
+enacted processes. Additionally, the authors use~\emph{linguistic
+extensions}~\citep{metadepth}, a mechanism to  linguistically extend
+ontological instance models at any level, in order to introduce artifact types
+and tasks duration in the second level. Our solution is similar to theirs
+w.r.t.
+the grouping and organization of concepts (\eg, a model for general process, concepts, an extension for software engineering process) but it replaces the use of clabjets and potency with the use of the type-object pattern~\citep{typeObject}.
 The so-called \emph{linguistic extensions} are natively supported in Openflexo/\FML.
@@
 
-More similar to us,~\citet{deeptelos2019} uses DeepTelos \parencite{deeptelos2016}, an extension of the Telos language \parencite{telos1990} in order to solve the process challenge.
-As with Openflexo/\FML, they do not use explicit (numbered) levels nor potency.
-However, unlike Openflexo, DeepTelos integrates a multi-level modeling specific construct similar to the PowerType \parencite{atkinson2001essence} pattern they call \emph{most general instances} which the authors use to simulate potency.
-Models in DeepTelos are organized in (tree-like) module hierarchies, where submodules can see all the concepts defined in parent modules.
-The authors use this module system in order to \emph{organize} their solution for the Multi-level process challenge.
-Concretely they created a hierarchy of modules such as the top module contains base concepts and formulas required for multi-level modeling (e.g., the support for \emph{most general instances}), and a sub-module contains process definitions fulfilling requirements P1 to P19.
+More similar to us,~\citet{deeptelos2019} uses DeepTelos
+\parencite{deeptelos2016}, an extension of the Telos language
+\parencite{telos1990} in order to solve the process challenge. As with
+Openflexo/\FML, they do not use explicit (numbered) levels nor potency.
+However, unlike Openflexo, DeepTelos integrates a multi-level modeling specific
+construct similar to the PowerType \parencite{atkinson2001essence} pattern they
+call \emph{most general instances} which the authors use to simulate potency.
+Models in DeepTelos are organized in (tree-like) module hierarchies, where
+submodules can see all the concepts defined in parent modules. The authors use
+this module system in order to \emph{organize} their solution for the
+Multi-level process challenge. Concretely they created a hierarchy of modules
+such as the top module contains base concepts and formulas required for
+multi-level modeling (\eg, the support for \emph{most general instances}), and a sub-module contains process definitions fulfilling requirements P1 to P19.
 This sub-module contains in turn a sub-module representing the coding process type whereas concrete coding processes are represented in subsequent sub-modules.

### 2022-02-23 6-modelpresentation.tex
s/ \ref/~\ref

@@
 
-Note that in all following figures, blue is used to represent \FML virtual models and concepts (\textsf{VirtualModel} and \textsf{FlexoConcept} in Fig.~\ref{fig:mm}), see figures \ref{fig:BPMNSubsetExample}, \ref{fig:MultilevelArchitecture}, \ref{fig:LinguisticAndOntologicInstantiation}, \ref{fig:BaseMetamodel}, \ref{fig:AcmeFullArchitecture},  \ref{fig:ToolingArchitecture}, while brown is used for instances of \textsf{FlexoConcept}, see figures \ref{fig:BPMNSubsetExample}, \ref{fig:MultilevelArchitecture}, \ref{fig:LinguisticAndOntologicInstantiation}, \ref{fig:AcmeFullArchitecture},  \ref{fig:ToolingArchitecture}.
+Note that in all following figures, blue is used to represent \FML virtual models and concepts (\textsf{VirtualModel} and \textsf{FlexoConcept} in Fig.~\ref{fig:mm}), see figures~\ref{fig:BPMNSubsetExample},~\ref{fig:MultilevelArchitecture},~\ref{fig:LinguisticAndOntologicInstantiation},~\ref{fig:BaseMetamodel},~\ref{fig:AcmeFullArchitecture}, ~\ref{fig:ToolingArchitecture}, while brown is used for instances of \textsf{FlexoConcept}, see figures~\ref{fig:BPMNSubsetExample},~\ref{fig:MultilevelArchitecture},~\ref{fig:LinguisticAndOntologicInstantiation},~\ref{fig:AcmeFullArchitecture},~\ref{fig:ToolingArchitecture}.
 % 3, 4, 5, 6, 7, 8, 9

### 2022-02-23 6-modelpresentation.tex
style guidelines: figures -> Fig.

--- 2022-02-23 6-modelpresentation.tex
@@
 
-Note that in all following figures, blue is used to represent \FML virtual models and concepts (\textsf{VirtualModel} and \textsf{FlexoConcept} in Fig.~\ref{fig:mm}), see figures \ref{fig:BPMNSubsetExample}, \ref{fig:MultilevelArchitecture}, \ref{fig:LinguisticAndOntologicInstantiation}, \ref{fig:BaseMetamodel}, \ref{fig:AcmeFullArchitecture},  \ref{fig:ToolingArchitecture}, while brown is used for instances of \textsf{FlexoConcept}, see figures \ref{fig:BPMNSubsetExample}, \ref{fig:MultilevelArchitecture}, \ref{fig:LinguisticAndOntologicInstantiation}, \ref{fig:AcmeFullArchitecture},  \ref{fig:ToolingArchitecture}.
+Note that in all following figures, blue is used to represent \FML virtual
+models and concepts (\textsf{VirtualModel} and \textsf{FlexoConcept} in
+Fig.~\ref{fig:mm}), see
+Fig.~\ref{fig:BPMNSubsetExample},~\ref{fig:MultilevelArchitecture},~\ref{fig:LinguisticAndOntologicInstantiation},~\ref{fig:BaseMetamodel},~\ref{fig:AcmeFullArchitecture},
+~\ref{fig:ToolingArchitecture}, while brown is used for instances of
+\textsf{FlexoConcept}, see Fig.~\ref{fig:BPMNSubsetExample},~\ref{fig:MultilevelArchitecture},~\ref{fig:LinguisticAndOntologicInstantiation},~\ref{fig:AcmeFullArchitecture},~\ref{fig:ToolingArchitecture}.
 % 3, 4, 5, 6, 7, 8, 9

### 2022-02-23 6-modelpresentation.tex
proper use of -, --, --- and $-$

@@
 The Base metamodel is completed with some behaviors implementing an execution
-semantics for the executed processes. All tasks -- instantiated from
-\texttt{TaskType} for a given enacted process -- manage a status. This status
+semantics for the executed processes. All tasks --- instantiated from
+\texttt{TaskType} for a given enacted process --- manage a status. This status
 is either \texttt{Not startable} (when not assigned to a performing actor or
@@
 to be instantiated. For example, when a \texttt{TaskType} is created, the user
-must choose a concept inheriting from \texttt{TaskType} -- in Acme use case it
+must choose a concept inheriting from \texttt{TaskType} (in Acme use case it
 can be \texttt{SETaskTask} or \texttt{CodingTaskType} or default value

### 2022-02-24 1-main.tex
bookmarks

@@
 
-\newcommand{\mpc}{MULTI process challenge\xspace}%keep it for the moment
-\newcommand{\mlpc}{Multi-Level Process Challenge\xspace}
-
 \usepackage{multirow}
@@
 \usepackage{hyperref}
-\usepackage{tikz-uml}
+%\usepackage{tikz-uml}
 %\usepackage{float}
+\usepackage{algorithmic,algorithm}
+
+\newcommand{\mpc}{MULTI process challenge\xspace}%keep it for the moment
+\newcommand{\mlpc}{Multi-Level Process Challenge\xspace}
 
 \newcommand{\inst}[1]{\textsf{'#1'}}
-
-%\usepackage{amsmath,mathtools}%amssymb,
-\usepackage{algorithmic,algorithm}
 
@@
   pdftitle={Multi-level modeling with Openflexo/\FML -- A contribution to the \mlpc},
+  bookmarks=true,
+  unicode=true,
+  pdfpagelabels=true,
   pdfkeywords={Multi-Level Modeling Challenge, Federation Modeling Language,

### 2022-02-24 4-technology.tex
checklist: initialisms (\BPMN)

@@
     \includegraphics[width=\columnwidth]{Figures/BPMNSubsetExample.pdf}
-    \caption{An example of BPMN subset showing both the model and an instance of that model}
+    \caption{An example of \BPMN subset showing both the model and an instance of that model}
     \label{fig:BPMNSubsetExample}
@@
 
-To illustrate this, Fig.~\ref{fig:BPMNSubsetExample} shows how to realize a small subset of BPMN with Openflexo/FML.
+To illustrate this, Fig.~\ref{fig:BPMNSubsetExample} shows how to realize a
+small subset of \BPMN with Openflexo/FML.
 Concretely, we have designed a \textit{virtual model} (\texttt{SimplifiedBPMN}) with 3 \textit{concepts} representing a process and its contained tasks.

### 2022-02-24 9-relatedworks.tex
citations and references: replace \citet, resp. \citep, by \textcite, resp. \parencite

@@
 
-A plethora of multi-level modeling approaches and tools\footnote{\url{https://homepages.ecs.vuw.ac.nz/Groups/MultiLevelModeling/MultiTools}} with different foundations have appeared in recent years~\citep{somogyi2021playground}.
+A plethora of multi-level modeling approaches and tools\footnote{\url{https://homepages.ecs.vuw.ac.nz/Groups/MultiLevelModeling/MultiTools}} with different foundations have appeared in recent years~\parencite{somogyi2021playground}.
 Comparing them lies out of the scope of this paper.
@@
 A first description of process modeling as a multi-level modeling problem was
-proposed by~\citet{lara2018refactoring} in the context of a catalogue of
+proposed by~\textcite{lara2018refactoring} in the context of a catalogue of
 refactoring for multi-level models (in a simpler form with fewer constraints
@@
 levels, dual ontological/linguistic typing and deep characterization through
-potency~\citep{potency}. For their solution to the multi-level process problem
+potency~\parencite{potency}. For their solution to the multi-level process problem
 they use three levels. The first level describes generic processes, the second
@@
 enacted processes. Additionally, the authors use~\emph{linguistic
-extensions}~\citep{metadepth}, a mechanism to  linguistically extend
+extensions}~\parencite{metadepth}, a mechanism to  linguistically extend
 ontological instance models at any level, in order to introduce artifact types
@@
 w.r.t.
-the grouping and organization of concepts (\eg, a model for general process, concepts, an extension for software engineering process) but it replaces the use of clabjets and potency with the use of the type-object pattern~\citep{typeObject}.
+the grouping and organization of concepts (\eg, a model for general process, concepts, an extension for software engineering process) but it replaces the use of clabjets and potency with the use of the type-object pattern~\parencite{typeObject}.
 The so-called \emph{linguistic extensions} are natively supported in Openflexo/\FML.
 
-%\textcolor{red}{Our solution to the Multi-Level process challenge relies on the use of %the type/object pattern~\citep{typeObject} for achieving ontological instantiation and %the core capacity of Openflexo/FML to reference models disregarding abstraction levels.
+%\textcolor{red}{Our solution to the Multi-Level process challenge relies on the use of %the type/object pattern~\parencite{typeObject} for achieving ontological instantiation and %the core capacity of Openflexo/FML to reference models disregarding abstraction levels.
 %Concretely, we use a metamodel to group the required concepts for describing universal %processes and an extending metamodel that adapts it to the domain of software %engineering processes.
@@
 
-\cite{multiecore2019} use MultEcore \citep{multecore2016} to provide a solution to the challenge.
-MultEcore uses (un)pluggable linguistic levels (\eg, there is not a \emph{fixed} core metamodel defining the concept of clabjet), an extension of the two-level cascading technique \citep{atkinson2005concepts} (this is achieved with model transformations to transform instance models into instantiable models) and potency.
+\textcite{multiecore2019} use MultEcore \parencite{multecore2016} to provide a solution to the challenge.
+MultEcore uses (un)pluggable linguistic levels (\eg, there is not a \emph{fixed} core metamodel defining the concept of clabjet), an extension of the two-level cascading technique \parencite{atkinson2005concepts} (this is achieved with model transformations to transform instance models into instantiable models) and potency.
 Regarding the challenge, their solution uses 4 levels.
@@
 
-More similar to us,~\citet{deeptelos2019} uses DeepTelos
+More similar to us,~\textcite{deeptelos2019} uses DeepTelos
 \parencite{deeptelos2016}, an extension of the Telos language
@@
 
-Finally, \cite{dmla2019} use the Dynamic Multi-Layer Algebra (DMLA) \parencite{dmla2017} in order to solve the process challenge.
+Finally, \textcite{dmla2019} use the Dynamic Multi-Layer Algebra (DMLA) \parencite{dmla2017} in order to solve the process challenge.
 DMLA is a level blind modeling framework which is fully customizable (\eg different types of instantiation may be implemented) and includes support for deep characterization (instantiation may refer to any other element disregarding hierarchy levels and a sort of potency in the form of fluid metamodeling at the entity level exists).

### 2022-02-24 1-main.tex
comment macros for commenting

@@
 
-\usepackage{todonotes}
+% \usepackage{todonotes}
 
-\newcommand{\noteSM}[1]{\todo[color=red!40, author=\textbf{Salvador}, inline, caption={}]{#1}}
-\newcommand{\noteJC}[1]{\todo[color=pink!40, author=\textbf{JC}, inline, caption={}]{#1}}
-\newcommand{\noteSylvain}[1]{\todo[color=green!40, author=\textbf{Sylvain}, inline, caption={}]{#1}}
-\newcommand{\noteAntoine}[1]{\todo[color=yellow!40, author=\textbf{Antoine}, inline, caption={}]{#1}}
-\newcommand{\noteFabien}[1]{\todo[color=orange!40, author=\textbf{Fabien}, inline, caption={}]{#1}}
-\newcommand{\noteJoel}[1]{\todo[color=blue!20, author=\textbf{Joel}, inline, caption={}]{#1}}
+% \newcommand{\noteSM}[1]{\todo[color=red!40, author=\textbf{Salvador}, inline, caption={}]{#1}}
+% \newcommand{\noteJC}[1]{\todo[color=pink!40, author=\textbf{JC}, inline, caption={}]{#1}}
+% \newcommand{\noteSylvain}[1]{\todo[color=green!40, author=\textbf{Sylvain}, inline, caption={}]{#1}}
+% \newcommand{\noteAntoine}[1]{\todo[color=yellow!40, author=\textbf{Antoine}, inline, caption={}]{#1}}
+% \newcommand{\noteFabien}[1]{\todo[color=orange!40, author=\textbf{Fabien}, inline, caption={}]{#1}}
+% \newcommand{\noteJoel}[1]{\todo[color=blue!20, author=\textbf{Joel}, inline, caption={}]{#1}}
 

### 2022-02-24 1-main.tex
comment useless packages
comment useless font

@@
 
+\usepackage{bookmark}
 \usepackage{multirow}
 \usepackage{listings}
-\usepackage{xcolor}
-\usepackage{hyperref}
+%\usepackage{xcolor}
+%\usepackage{hyperref}
 %\usepackage{tikz-uml}
 %\usepackage{float}
-\usepackage{algorithmic,algorithm}
+%\usepackage{algorithmic,algorithm}
 
@@
 
-\DeclareTextFontCommand{\mytexttt}{\ttfamily\hyphenchar\font=45\relax}
+%\DeclareTextFontCommand{\mytexttt}{\ttfamily\hyphenchar\font=45\relax}
 
@@
   pdftitle={Multi-level modeling with Openflexo/\FML -- A contribution to the \mlpc},
-  bookmarks=true,
   unicode=true,
-  pdfpagelabels=true,
   pdfkeywords={Multi-Level Modeling Challenge, Federation Modeling Language,
@@
 \label{sec:introduction}
+\pdfbookmark[section]{Introduction}{sec:introduction}
 \input{3-introduction.tex}
@@
 \label{sec:technology}
+\pdfbookmark[section]{Technology}{sec:technology}
 \input{4-technology.tex}
@@
 \label{sec:analysis}
+\pdfbookmark[section]{Analysis}{sec:analysis}
 \input{5-analysis.tex}
@@
 \label{sec:model}
+\pdfbookmark[section]{Model presentation}{sec:model}
 \input{6-modelpresentation.tex}
@@
 \label{sec:requirements}
+\pdfbookmark[section]{Satisfaction of requirements}{sec:requirements}
 \input{7-requirementssatisfaction.tex}
@@
 \label{sec:discussion}
+\pdfbookmark[section]{Assessment of the modeling solution}{sec:discussion}
 \input{8-discussion.tex}
@@
 \label{sec:relatedwork}
+\pdfbookmark[section]{Related work}{sec:relatedwork}
 \input{9-relatedworks.tex}
@@
 \label{sec:conclusions}
+\pdfbookmark[section]{Conclusions}{sec:conclusions}
 \input{10-conclusions.tex}

### 2022-02-25 9-relatedworks.tex
complete biblio.bib + fix redundant citation

@@
 levels, dual ontological/linguistic typing and deep characterization through
-potency~\parencite{potency}. For their solution to the multi-level process problem
+potency~\parencite{atkinson2001essence}. For their solution to the multi-level process problem
 they use three levels. The first level describes generic processes, the second

### 2022-03-01 6-modelpresentation.tex
fix spaces

@@
 Fig.~\ref{fig:mm}), see
-Fig.~\ref{fig:BPMNSubsetExample},~\ref{fig:MultilevelArchitecture},~\ref{fig:LinguisticAndOntologicInstantiation},~\ref{fig:BaseMetamodel},~\ref{fig:AcmeFullArchitecture},
-~\ref{fig:ToolingArchitecture}, while brown is used for instances of
+Fig.~\ref{fig:BPMNSubsetExample},~\ref{fig:MultilevelArchitecture},~\ref{fig:LinguisticAndOntologicInstantiation},~\ref{fig:BaseMetamodel},~\ref{fig:AcmeFullArchitecture},~\ref{fig:ToolingArchitecture}, while brown is used for instances of
 \textsf{FlexoConcept}, see Fig.~\ref{fig:BPMNSubsetExample},~\ref{fig:MultilevelArchitecture},~\ref{fig:LinguisticAndOntologicInstantiation},~\ref{fig:AcmeFullArchitecture},~\ref{fig:ToolingArchitecture}.
@@
     % \includegraphics[width=1.0 \columnwidth]{Figures/Metamodel.pdf}
-    \includegraphics[width=1.0 \textwidth]{Figures/Metamodel.pdf}
+    \includegraphics[width=1.0 \textwidth]{Figures/Metamodel-1.5.pdf}
      \caption{Process management base metamodel (\texttt{Gateway}'s associations with \texttt{TaskType} are represented with attributes)}
@@
  \centering
-     \includegraphics[width=1.0 \textwidth]{Figures/AcmeFullArchitecture.pdf}
+     \includegraphics[width=1.0 \textwidth]{Figures/AcmeFullArchitecture-1.5.pdf}
      \caption{Acme software development process architecture (not all instances are shown for the sake of readability)}
@@
  \centering
-     \includegraphics[width=0.9 \textwidth]{Figures/ToolingArchitecture.pdf}
+     \includegraphics[width=0.9 \textwidth]{Figures/ToolingArchitecture-1.5.pdf}
      \caption{Tooling architecture}

### 2022-03-09 2-abstract.tex
Updates from Overleaf

--- 2021-09-27 2-abstract.tex
@@
 We present a solution based on model federation and discuss the advantages and limitations of using this approach for multi-level modeling.
-Concretely, we use virtual models and more precisely the Flexo Modeling Language ({FML}) that serves to describe them as the main building block in order to solve the process modeling challenge whereas the federation aspect is used as a means to provide editing tooling for the resulting process language.
-Our solution is fully implemented with the Openflexo framework.
-It fulfills all the challenge requirements.
+Concretely, we use virtual models and more precisely the Federation Modeling Language ({FML}) that serves to describe them as the main building block in order to solve the process modeling challenge whereas the federation feature {\color{red} Do you have a better word than aspect here?} is used as a means to provide editing tools {\color{red} I would rather use "tools" instead of "tooling".
+Tooling is the process of adding/making tools and you are already using "provide" before the "editing tooling"} for the resulting process language.
+Our solution fulfills all the challenge requirements and is fully implemented with the Openflexo framework.
+{\color{red} For the list of keywords, nobody will search your article using them :), use generic search keywords that point to your article.
+}
 

### 2022-03-09 4-technology.tex
convert drawio pdf to 1.5
fix spaces
Updates from Overleaf

@@
   federation} is a way to assemble models using some kind of
-low-coupling links. It has been studied to answer the \OMG' RFP
+low-coupling links. It was initially  formulated to respond to the \OMG's RFP
 Semantic Modeling for Information Federation~\parencite{simf}. In
@@
     \centering
-    \includegraphics[width=1.0 \textwidth]{Figures/FMLMetaModel.pdf}
+    \includegraphics[width=1.0 \textwidth]{Figures/FMLMetaModel-1.5.pdf}
     \caption{Representation of the main concepts of {FML} language}
@@
 Language (\FML). Each federated model is an autonomous
-element that may evolve with its own tooling. The virtual models
+element that may evolve with its own tooling {\color{red} tooling is used in place of tool throughout the article.
+I suggest clarifying it and then editing it all along the article}.
+The virtual models
 serve as control elements binding the federated models together.
-In this paper, we mainly use conceptual models and more precisely \FML its domain specific modeling language. The only use of the federation aspect is on the tooling made for the process language.
+In this paper, we mainly use conceptual models and more precisely \FML, its domain specific modeling language. The only use of the federation aspect is on the tooling made for the process language.
 
-A \UML like representation of a simplified metamodel of \FML is provided in Fig.~\ref{fig:mm}.
-\FML is a language designed to define models as virtual models.
+{\color{red} This paragraph needs a parallel in naming convention between Figure 2 and the text.
+FlexoConcept, FlexoBehavior, etc.} A \UML like representation of a simplified metamodel of \FML is provided in Fig.~\ref{fig:mm}.
+\FML is a language designed to define virtual models.
 A \emph{virtual model} is composed of a set of \emph{concepts}, while itself being a \emph{concept}.
@@
     \centering
-    \includegraphics[width=\columnwidth]{Figures/BPMNSubsetExample.pdf}
+    \includegraphics[width=\columnwidth]{Figures/BPMNSubsetExample-1.5.pdf}
     \caption{An example of \BPMN subset showing both the model and an instance of that model}

### 2022-03-09 5-analysis.tex

@@
 
-In a first step, the federation approach is mainly based on modeling relationships between several models, independently of their abstraction level and the model architecture.
+In a first step, the federation approach is mainly based on modeling relationships between several models, independent of their abstraction levels and the modeling architecture.
 
@@
     \centering
-    \includegraphics[width=0.9 \columnwidth]{Figures/MultilevelArchitecture.pdf}
+    \includegraphics[width=0.9 \columnwidth]{Figures/MultilevelArchitecture-1.5.pdf}
     \caption{Multilevel architecture of our solution}
@@
     \centering
-    \includegraphics[width=1.0 \columnwidth]{Figures/Instantiation.pdf}
+    \includegraphics[width=1.0 \columnwidth]{Figures/Instantiation-1.5.pdf}
     \caption{Problem  domain linguistic and ontologic instantiation}
@@
 
-Based on this approach, we organize our set of models following the
+{\color{red} I do not know what is the legend of figure 5.
+However, if all four modeled concepts are virtual models, their graphical representation should be the same.
+I find the same problem with figure 3.
+As a matter of fact all models from Figure 2, 3 and 5 use various representations and one wonders when it is UML, when it is a conceptual model?}Based on this approach, we organize our set of models following the
 architecture of the aforementioned Fig.~\ref{fig:MultilevelArchitecture}. The
@@
 In our approach this virtual model instance level can not be specialized or extended but all the other virtual models could be extended by any concepts, as a new federated model.
-Also, the instance level can be instantiated from any virtual model, which represents any metamodel level.
+Furthermore, the instance level can be instantiated from any virtual model, which represents any metamodel level.
 

### 2022-03-10 3-introduction.tex
proper use of -, --, --- and $-$

@@
 In order to tackle this problem, the \emph{multi-level} modeling paradigm has been introduced.
-Contrary to the strict approach, multi-level modeling advocates the use of a flexible number of levels as well as more flexible relations between them.
-This paradigm is gaining traction lately within the modeling community as evidenced by the contribution of many new multi-level modeling approaches and tools such as Melanee~\parencite{melanee}, MetaDepth~\parencite{metadepth}, MultEcore~\parencite{multecore2016}, DeepTelos~\parencite{deeptelos2016} or DMLA~\parencite{dmla2017}.
-Taking advantage of this vibrant state of affairs, and in order to foster discussion and enable comparison between competing approaches, a multi-level modeling challenge has been created.
+Contrary to the strict approach, multi-level modeling advocates the use of a flexible number of levels as well as more flexible relations between them{\color{red} The article of ICSE by me and Fabien can be cited here}.
+This paradigm is gaining traction within the modeling community as evidenced by the contribution of many new multi-level modeling approaches and tools such as Melanee~\parencite{melanee}, MetaDepth~\parencite{metadepth}, MultEcore~\parencite{multecore2016}, DeepTelos~\parencite{deeptelos2016} or DMLA~\parencite{dmla2017}.
+Taking advantage of this vibrant state of affairs, and in order to foster discussion and enable comparison between competing approaches, a multi-level modeling challenge has been created{\color{red} Who creating this challenge, any reference?}.
 This paper is a response to the latest multi-level modeling challenge
@@
 Solutions to the \emph{Multi-level Process Challenge} must fulfill a number of requirements for a process representation defined at an abstract process-definition level and at various more concrete domain-specific levels, resulting in a multi-level hierarchy of related models.
-We provide here a solution based on model federation~\parencite{Golra2016-federation}.
+In this article, we respond to this challenge using a solution based on model federation~\parencite{Golra2016-federation}.
 
 Model federation is a multi-model management approach based on the use of virtual models and loosely coupled links.
-The models in a federation remain autonomous and represented in their original technological spaces whereas virtual models (also called conceptual models) and links serve as control components used to present different views to the users and maintain synchronization.
+The models in a federation remain autonomous and are represented in their original technological spaces whereas virtual models (also called conceptual models) and links that serve as control components used to present different views to the users and maintain synchronization.
 Our solution is based on the model federation infrastructure.
-Concretely, we use Openflexo and its internal Federation Modeling Language (\FML), a language to create, link and manage virtual models, in order to solve the MULTI process challenge while the federation aspect is used solely so as to provide tooling for the resulting process language.
+Concretely, we use Openflexo and its internal Federation Modeling Language (\FML), a language to create, link and manage virtual models, in order to solve the MULTI process challenge while the federation aspect{\color{red} I dont like "aspect", it is a vague term} is used solely so as to provide tooling for the resulting process language.
 Our solution, which is fully implemented and executable, meets all the requirements.
@@
 Our modeling approach is based on a language (\FML) which provides the (linguistics) concepts that are used to define the multi-level models and meta-models that are needed.
-\FML allows defining models as first class entities (which we call \emph{virtual models}) and to link them together (by extension).
+\FML allows defining models as first class entities (which we call \emph{virtual models}) and linking them together (by extension).
 In the \mpc case study, we exploit this possibility to:
@@
     modeling requirements of the case study,
-    \item Align this hierarchy of models with requirements evolution ($P_1-P_{19}$ then $S_1-S_{13}$).
+    \item Align this hierarchy of models with requirements evolution
+      ($P_1$--$P_{19}$ then $S_1$--$S_{13}$).
 \end{itemize}
@@
 multi-level is integrated in the language and must therefore be respected, our
-approach allows building multi-level on demand, and with the appropriate form.
+approach allows building multi-level{\color{red} multi-level "something" or multiple levels} on demand, and with the appropriate form.
 Without multilevel-specific operators/concepts, the \enquote{custom}
@@
 
-Our answer to the challenge is therefore not an answer using a classical multi-level approach that would allow an efficiency comparison with another classical multi-level answer, but an approach that proposes another organization of multi-levels through the translation of needs into levels over time and the possibility to build an evolving hierarchy of adapted models.
+Our solution to the challenge is therefore not a response using a classical multi-level approach that would allow an efficiency comparison with another classical multi-level solutions, but an approach that proposes another organization of multi-level hierarchy through the translation of needs into levels over time and the possibility to build an evolving hierarchy of adapted models.
 This is made possible by (1) the reification of the notion of a model and (2) the possibility to interconnect models (specialization).% and (2) the freedom of meta-modeling (low model conformity, flexible meta-modeling).

### 2022-03-10 1-main.tex
rephrasing: s/aspect/feature

@@
 % \newcommand{\noteJoel}[1]{\todo[color=blue!20, author=\textbf{Joel}, inline, caption={}]{#1}}
+ \newcommand{\noteFahad}[1]{\todo[color=purple!20, author=\textbf{Fahad}, inline, caption={}]{#1}}
 
@@
 We present a solution based on model federation and discuss the advantages and limitations of using this approach for multi-level modeling.
-Concretely, we use virtual models and more precisely the Federation Modeling Language ({FML}) that serves to describe them as the main building block in order to solve the process modeling challenge whereas the federation aspect is used as a means to provide editing tooling for the resulting process language.
+Concretely, we use virtual models and more precisely the Federation Modeling Language ({FML}) that serves to describe them as the main building block in order to solve the process modeling challenge whereas the federation feature is used as a means to provide editing tooling for the resulting process language.
 Our solution is fully implemented with the Openflexo framework.
@@
     \abstract{\input{2-abstract}}
-    \keywords{Multi-Level Modeling Challenge \and Federation Modeling Language
-    \and Openflexo}
+    \keywords{Multi-Level Modeling \and System modeling languages \and Abstraction, modeling and modularity \and Reusability }
+    %ACM
+    % Software and its engineering/Software notations and tools/System description languages/System modeling languages
+    % Software and its engineering/Software organization and properties/Software system structures/Abstraction, modeling and modularity
+    % Software and its engineering/Software creation and management/Software development techniques/Reusability
     \bibliography{biblio.bib}

### 2022-03-10 2-abstract.tex
rephrasing: s/aspect/feature

@@
 We present a solution based on model federation and discuss the advantages and limitations of using this approach for multi-level modeling.
-Concretely, we use virtual models and more precisely the Federation Modeling Language ({FML}) that serves to describe them as the main building block in order to solve the process modeling challenge whereas the federation feature {\color{red} Do you have a better word than aspect here?} is used as a means to provide editing tools {\color{red} I would rather use "tools" instead of "tooling".
+Concretely, we use virtual models and more precisely the Federation Modeling Language ({FML}) that serves to describe them as the main building block in order to solve the process modeling challenge whereas the federation feature is used as a means to provide editing tools {\color{red} I would rather use "tools" instead of "tooling".
 Tooling is the process of adding/making tools and you are already using "provide" before the "editing tooling"} for the resulting process language.

### 2022-03-10 3-introduction.tex
rephrasing: s/aspect/feature

@@
 Our solution is based on the model federation infrastructure.
-Concretely, we use Openflexo and its internal Federation Modeling Language (\FML), a language to create, link and manage virtual models, in order to solve the MULTI process challenge while the federation aspect{\color{red} I dont like "aspect", it is a vague term} is used solely so as to provide tooling for the resulting process language.
+Concretely, we use Openflexo and its internal Federation Modeling Language (\FML), a language to create, link and manage virtual models, in order to solve the MULTI process challenge while the federation feature is used solely so as to provide tooling for the resulting process language.
 Our solution, which is fully implemented and executable, meets all the requirements.

### 2022-03-10 4-technology.tex
rephrasing: s/aspect/feature

@@
 serve as control elements binding the federated models together.
-In this paper, we mainly use conceptual models and more precisely \FML, its domain specific modeling language. The only use of the federation aspect is on the tooling made for the process language.
+In this paper, we mainly use conceptual models and more precisely \FML, its
+domain specific modeling language. The only use of the federation feature is on the tooling made for the process language.
 
@@
 
-A parallel to object-oriented approach can be useful to understand \FML\footnote{Some aspects of \FML do not exist in object oriented approach.}.
+A parallel to object-oriented approach can be useful to understand \FML\footnote{Some features of \FML do not exist in object oriented approach.}.
 A concept corresponds to a class, its properties to the attributes of the class and its behaviors to the methods of the class.

### 2022-03-11 3-introduction.tex
cite the challenge

@@
 This paradigm is gaining traction within the modeling community as evidenced by the contribution of many new multi-level modeling approaches and tools such as Melanee~\parencite{melanee}, MetaDepth~\parencite{metadepth}, MultEcore~\parencite{multecore2016}, DeepTelos~\parencite{deeptelos2016} or DMLA~\parencite{dmla2017}.
-Taking advantage of this vibrant state of affairs, and in order to foster discussion and enable comparison between competing approaches, a multi-level modeling challenge has been created{\color{red} Who creating this challenge, any reference?}.
-This paper is a response to the latest multi-level modeling challenge
+Taking advantage of this vibrant state of affairs, and in order to foster discussion and enable comparison between competing approaches, a multi-level modeling challenge has been created for the MULTI workshop.
+This paper is a response to the latest multi-level modeling challenge~\parencite{multiProcessChallenge-emisaj}
+%\footnote{It can be found here: \url{http://purl.org/emisajchallenge}. It is an updated version of the challenge used in~\textcite{multiProcessChallenge2019}}
 which consists in providing a solution to the problem of specifying and enacting processes.

### 2022-03-11 3-introduction.tex
fix multi-level / multiple levels problem

@@
 multi-level is integrated in the language and must therefore be respected, our
-approach allows building multi-level{\color{red} multi-level "something" or multiple levels} on demand, and with the appropriate form.
+approach allows building multiple levels on demand, and with the appropriate form.
 Without multilevel-specific operators/concepts, the \enquote{custom}

### 2022-03-11 5-analysis.tex
remove comments concerning the figures: we discussed the representation during the writing of the article and we agreed on the current version. Reviewers did not seem to be lost and the article has been accepted, therefore I would let the figures unchanged

@@
 
-{\color{red} I do not know what is the legend of figure 5.
-However, if all four modeled concepts are virtual models, their graphical representation should be the same.
-I find the same problem with figure 3.
-As a matter of fact all models from Figure 2, 3 and 5 use various representations and one wonders when it is UML, when it is a conceptual model?}Based on this approach, we organize our set of models following the
+%{\color{red} I do not know what is the legend of figure 5.  However, if all
+%  four modeled concepts are virtual models, their graphical representation
+%  should be the same. I find the same problem with figure 3. As a matter of
+%  fact all models from Figure 2, 3 and 5 use various representations and one
+%  wonders when it is UML, when it is a conceptual model?}{\color{pink} (JC) we
+%  discussed the reprenstation during the writing and agreed on the current
+%version. Reviewers did not seem to be lost, therefore I would let the figures
+%unchanged}
+
+Based on this approach, we organize our set of models following the
 architecture of the aforementioned Fig.~\ref{fig:MultilevelArchitecture}. The

### 2022-03-11 3-introduction.tex
cite LIMM paper (icse2011-limm) in the list of multi-level modeling approaches and tools

@@
 
-In order to tackle this problem, the \emph{multi-level} modeling paradigm has been introduced.
-Contrary to the strict approach, multi-level modeling advocates the use of a flexible number of levels as well as more flexible relations between them{\color{red} The article of ICSE by me and Fabien can be cited here}.
-This paradigm is gaining traction within the modeling community as evidenced by the contribution of many new multi-level modeling approaches and tools such as Melanee~\parencite{melanee}, MetaDepth~\parencite{metadepth}, MultEcore~\parencite{multecore2016}, DeepTelos~\parencite{deeptelos2016} or DMLA~\parencite{dmla2017}.
-Taking advantage of this vibrant state of affairs, and in order to foster discussion and enable comparison between competing approaches, a multi-level modeling challenge has been created for the MULTI workshop.
-This paper is a response to the latest multi-level modeling challenge~\parencite{multiProcessChallenge-emisaj}
+In order to tackle this problem, the \emph{multi-level} modeling paradigm has
+been introduced. Contrary to the strict approach, multi-level modeling
+advocates the use of a flexible number of levels as well as more flexible
+relations between them. %{\color{red} The article of ICSE by me and Fabien can be
+%cited here} {(JC) \color{pink} Effectivement. intégré dans la liste}
+This paradigm is gaining traction within the modeling community as evidenced by
+the contribution of many new multi-level modeling approaches and tools such as
+Melanee~\parencite{melanee}, LIMM~\parencite{icse2011-limm},
+MetaDepth~\parencite{metadepth}, MultEcore~\parencite{multecore2016},
+DeepTelos~\parencite{deeptelos2016} or DMLA~\parencite{dmla2017}. Taking
+advantage of this vibrant state of affairs, and in order to foster discussion
+and enable comparison between competing approaches, a multi-level modeling
+challenge has been created for the MULTI workshop. This paper is a response to
+the latest multi-level modeling challenge~\parencite{multiProcessChallenge-emisaj}
 %\footnote{It can be found here: \url{http://purl.org/emisajchallenge}. It is an updated version of the challenge used in~\textcite{multiProcessChallenge2019}}
-which consists in providing a solution to the problem of specifying and enacting processes.
-Solutions to the \emph{Multi-level Process Challenge} must fulfill a number of requirements for a process representation defined at an abstract process-definition level and at various more concrete domain-specific levels, resulting in a multi-level hierarchy of related models.
-In this article, we respond to this challenge using a solution based on model federation~\parencite{Golra2016-federation}.
+which consists in providing a solution to the problem of specifying and
+enacting processes. Solutions to the \emph{Multi-level Process Challenge} must
+fulfill a number of requirements for a process representation defined at an
+abstract process-definition level and at various more concrete domain-specific
+levels, resulting in a multi-level hierarchy of related models. In this
+article, we respond to this challenge using a solution based on model
+federation~\parencite{Golra2016-federation}.
 

### 2022-03-15 1-main.tex
Synchro après réunion du 15 mars

--- 2022-02-24 1-main.tex
@@
     \abstract{\input{2-abstract}}
-    \keywords{Multi-Level Modeling Challenge \and Federation Modeling Language
-    \and Openflexo}
+    \keywords{Model federation \and Multi-Level Modeling \and System modeling languages \and Abstraction, modeling and modularity \and Reusability }
+    %ACM
+    % Software and its engineering/Software notations and tools/System description languages/System modeling languages
+    % Software and its engineering/Software organization and properties/Software system structures/Abstraction, modeling and modularity
+    % Software and its engineering/Software creation and management/Software development techniques/Reusability
     \bibliography{biblio.bib}

### 2022-03-15 10-conclusions.tex
Synchro après réunion du 15 mars

@@
 dynamically create different \emph{instances} of model elements (\eg, \emph{TaskType}, among others) and use them for typing other model elements.
-This, together with the \FML core support for (level agnostic) classical linguistic instantiation, permitted us to merge linguistic and ontological instantiation in order to provide a satisfactory solution to the Multi-Level Process challenge.
+This, together with the \FML core support for (level-agnostic) classical linguistic instantiation, permitted us to merge linguistic and ontological instantiation in order to provide a satisfactory solution to the Multi-Level Process challenge.
 
-The features of \FML enable a great modeling flexibility which, coupled with an appropriate methodology, allow us to achieve 
+The features of \FML enable modeling flexibility which, coupled with an appropriate methodology, allow us to achieve 
 the multi-level capabilities without dedicated tooling, in order to solve

### 2022-03-15 6-modelpresentation.tex
Synchro après réunion du 15 mars

@@
 The satisfaction of requirements is explained throughout this presentation.
-
 %\todo[inline]{phrase d'intro pour expliquer que les req étaient couverts au fur et à mesure (côté systématique)}
@@
 The designed multilevel architecture shown in Fig.~\ref{fig:MultilevelArchitecture} captures the two use cases described in the Process Challenge.
-
 Note that in all following figures, blue is used to represent \FML virtual
@@
 Fig.~\ref{fig:BPMNSubsetExample},~\ref{fig:MultilevelArchitecture},~\ref{fig:LinguisticAndOntologicInstantiation},~\ref{fig:BaseMetamodel},~\ref{fig:AcmeFullArchitecture},~\ref{fig:ToolingArchitecture}, while brown is used for instances of
-\textsf{FlexoConcept}, see Fig.~\ref{fig:BPMNSubsetExample},~\ref{fig:MultilevelArchitecture},~\ref{fig:LinguisticAndOntologicInstantiation},~\ref{fig:AcmeFullArchitecture},~\ref{fig:ToolingArchitecture}.
+\textsf{FlexoConcept}, see Fig.~\ref{fig:BPMNSubsetExample},~\ref{fig:MultilevelArchitecture},~\ref{fig:LinguisticAndOntologicInstantiation},~\ref{fig:AcmeFullArchitecture},~\ref{fig:ToolingArchitecture}.{\color{red} color scheming is good, but fails when document printed in gray scale.
+It might be worthwhile to consider patterns as well.}
 % 3, 4, 5, 6, 7, 8, 9
@@
 Two types of ontological instantiation are needed to meet the requirements of the challenge.
-One provides that some instances conform to only one type (\texttt{Process} and \texttt{Task} do respectively conform to \texttt{ProcessType} and \texttt{TaskType}).
+One provides that some instances conform to only one type (\texttt{Process} and \texttt{Task} do respectively conform to \texttt{ProcessType} and \texttt{TaskType}) {\color{red}Rephrase the previous sentence.
+"One provides ..."  where do you describe the second?}.
 While some entities define their conformity to several types (\texttt{Actor} and \texttt{Artifact} do respectively conform to several \texttt{ActorType} and \texttt{ArtifactType}).
@@
 
-We first present the process definition part of the base metamodel, located on the left of Fig.~\ref{fig:BaseMetamodel}.
+We first{\color{red} Use "first" when you have second third coming up later} present the process definition part of the base metamodel, located on the left of Fig.~\ref{fig:BaseMetamodel}.
 
@@
 
-The \texttt{ActorType} concept inherits from the \texttt{Type} concept and the \texttt{allowedActorTypes} relation to \texttt{ActorType} defined in \texttt{TaskType} (with 0..* cardinality) captures \textbf{P5} requirement.
+The \texttt{ActorType} concept inherits from the \texttt{Type} concept and the \texttt{allowedActorTypes} relation to \texttt{ActorType} defined in \texttt{TaskType} (with 0..* cardinality) captures the requirement \textbf{P5}.
 Requirement \textbf{P6} is symmetrically satisfied with \texttt{allowedActors} relation to \texttt{Actor} also defined in \texttt{TaskType} (with 0..* cardinality).
@@
 The \texttt{Actor} concept defines a boolean attribute called \texttt{isSenior}, while \texttt{TaskType} defines an additional \texttt{isCritical} boolean attribute, indicating that some instances are flagged as critical and must be performed by senior actors.
-To fulfill \textbf{P9} requirement, a supplementary constraint is required for \texttt{TaskType} and is captured through the following invariant expressed in the \FML language:
+To fulfill \textbf{P9}, a supplementary constraint is required for \texttt{TaskType} and is captured through the following invariant expressed in the \FML language:
 
@@
 
-This invariant should be completed with additional constraints defined in the \texttt{Task} concept, which apply to performing actors assigned to enact tasks.
+This invariant should be completed with additional constraints defined in the \texttt{Task} concept, which applies to performing actors assigned to enact tasks.
 
@@
 \label{sec:ProcessEnactment}
-We now present the process enactment part of base metamodel, located right of Fig.~\ref{fig:BaseMetamodel}.
+We now present the process enactment part of base metamodel, located at the right of Fig.~\ref{fig:BaseMetamodel}.
 All concepts defined in this subsection are either specializations of the \texttt{Instance} concept (if they have exactly one type) or the \texttt{MultiInstance} concept (when they have several types).
@@
 
-The semantics is unclear relatively to the instantiation policy for artifacts.
+The semantics is relatively unclear to the instantiation policy for artifacts.
 In the requirements, nothing is explicitly stated about the link between the
@@
 
-Each concept inherits from \texttt{ModelingElement}, which defines a \texttt{lastUpdated} attribute with \texttt{Date} type, and thus satisfies \textbf{P19} requirement.
+Each concept inherits from \texttt{ModelingElement}, which defines a \texttt{lastUpdated} attribute with \texttt{Date} type, and thus satisfies the requirement \textbf{P19}.
 
@@
 
-The challenge describes in a second part a Software engineering process for a fictional Acme company.
+The challenge describes a software engineering process for a fictional Acme company in the second part.
 The Base metamodel described previously is too generic to capture all domain-specific aspects of this use case.
@@
 This metamodel is completed with the \texttt{ProgrammingLanguage} concept, defined as an enumeration (\texttt{Java}, \texttt{C}, \texttt{COBOL}).
-All the instances required to capture the challenge use case are defined in the final Acme model (lower part of the figure).
+All the instances required to capture the use case of the challenge are defined in the final Acme model (lower part of the figure {\color{red} missing reference}).
 Instances are represented with rounded boxes and linguistic instantiation is represented using dashed connectors.
@@
 
-The capture of \textbf{S3} requirement is performed while defining \inst{Coding} as an instance of concept \texttt{CodingTaskType} and defining a relation \texttt{languages} to \texttt{ProgrammingLanguage} with 1..* cardinality.
+The requirement \textbf{S3} is captured by defining \inst{Coding} as an instance of concept \texttt{CodingTaskType} and defining a relation \texttt{languages} to \texttt{ProgrammingLanguage} with 1..* cardinality.
 The \texttt{newTask(String)} behavior overrides the generic implementation by instantiating a \texttt{CodingTask} concept, specializing \texttt{Task}, as shown in the following excerpt of \FML code:
@@
 
-\textbf{S5} is more ambiguous as a task type \texttt{CodingTask} defines one or more programming languages but produces code which is expressed in one language only.
+\textbf{S5} is more ambiguous as a task type \texttt{CodingTask} defines one or more programming languages but produces code which is expressed in only one language.
 This requirement is captured through the redefinition of the behavior \texttt{declaresRequiredProducedArtifacts} where programming language should also match.
@@
 
-\textbf{S13} requirement has been previously partially fulfilled.
+The requirement \textbf{S13} has been previously partially fulfilled.
 This must be completed with an association between an artifact produced and the task that validates it.
@@
 Process enactment is operated from a \texttt{ProcessType}, and must be defined using a name identifying the newly instantiated process.
-All tasks are created from their \texttt{TaskType} definition, and required assignments apply (if for example \inst{Coding} \texttt{TaskType} defines COBOL as output programming language, the related task is automatically assigned to \inst{Ann Smith}).
+All tasks are created from their \texttt{TaskType} definitions, and required assignments apply (if for example \inst{Coding} \texttt{TaskType} defines COBOL as output programming language, the related task is automatically assigned to \inst{Ann Smith}).
 Each task has a status as well as a set of performing actors, a begin and end dates, and used and produced artifacts.

### 2022-03-15 7-requirementssatisfaction.tex
Synchro après réunion du 15 mars

@@
 
-    \item \emph{Linguistic instantiation} (\textit{L.Inst.}): a requirement is satisfied by the instantiation of one or more concepts (for example for \textbf{S1}: \inst{Requirement analysis} is defined as an instance of \texttt{SETaskType}).
+    \item \emph{Linguistic instantiation} (\textit{L.Inst.}): the requirements are satisfied by the instantiation of one or more concepts (for example for \textbf{S1}: \inst{Requirement analysis} is defined as an instance of \texttt{SETaskType}).
 

### 2022-03-15 8-discussion.tex
Synchro après réunion du 15 mars

@@
   more or fewer levels depending on the variations of the use case. However,
-  the number of levels is related to the problem, our solution uses 3 levels.
+  the number of levels is related to the problem.  For example, our solution uses 3 levels for this multi-level process challenge.
 
@@
   relationships are not an issue. Model elements of different levels can be
-  linked each other in a transparent way, using inheritance or instantiation.
+  linked to each other in a transparent way, using inheritance or instantiation.
   For instance, in the Fig.~\ref{fig:AcmeFullArchitecture}, the
@@
   As for cross-level relationships, cross-level constraints do not have a
-  specific nature. They are like any constraint a user can define by adding a
+  specific nature. They are like any constraint that a user can define by adding a
   behavior to a model element. Therefore, if a constraint is mandatory when
   establishing a relationship between elements from different levels, the user
-  has to create it manually. This can be a limit of our approach: the cost of
+  has to create it manually. This can be a limit of our approach; the cost of
   the flexibility of our tooling is a limited number of automated behaviors.
@@
   solution easily. The challenge itself can be seen as a validation of the
-  extensibility ability of our solution: it was decomposed into two steps
+  extensibility of our solution: it was decomposed into two steps
   (Xsure process, then ACME process) which can be seen as a simulation of the

### 2022-03-15 9-relatedworks.tex
Synchro après réunion du 15 mars

@@
 A plethora of multi-level modeling approaches and tools\footnote{\url{https://homepages.ecs.vuw.ac.nz/Groups/MultiLevelModeling/MultiTools}} with different foundations have appeared in recent years~\parencite{somogyi2021playground}.
-Comparing them lies out of the scope of this paper.
+A complete comparison of all these approaches is out of the scope of this paper.
 In this sense, we limit the following discussion to the presentation and comparison of previous solutions to the MULTI Process Challenge.
@@
 multi-level modeling (\eg, the support for \emph{most general instances}), and a sub-module contains process definitions fulfilling requirements P1 to P19.
-This sub-module contains in turn a sub-module representing the coding process type whereas concrete coding processes are represented in subsequent sub-modules.
+This sub-module in turn contains another sub-module representing the coding process type whereas concrete coding processes are represented in subsequent sub-modules.
 There are two main differences between the DeepTelos and Openflexo/\FML: 1) we use the type-object pattern instead of the powertype pattern; 2) DeepTelos reifies their support for multi-level modeling while our remains ad hoc.

### 2022-03-16 1-main.tex
clarification between tool and tooling. Tooling was often replaced by tool(s) or toolset

--- 2022-03-10 1-main.tex
@@
 % \newcommand{\noteSM}[1]{\todo[color=red!40, author=\textbf{Salvador}, inline, caption={}]{#1}}
-% \newcommand{\noteJC}[1]{\todo[color=pink!40, author=\textbf{JC}, inline, caption={}]{#1}}
+ \newcommand{\noteJC}[1]{\todo[color=pink!40, author=\textbf{JC}, inline, caption={}]{#1}}
 % \newcommand{\noteSylvain}[1]{\todo[color=green!40, author=\textbf{Sylvain}, inline, caption={}]{#1}}
@@
 We present a solution based on model federation and discuss the advantages and limitations of using this approach for multi-level modeling.
-Concretely, we use virtual models and more precisely the Federation Modeling Language ({FML}) that serves to describe them as the main building block in order to solve the process modeling challenge whereas the federation feature is used as a means to provide editing tooling for the resulting process language.
-Our solution is fully implemented with the Openflexo framework.
-It fulfills all the challenge requirements.
-}
+Concretely, we use virtual models and more precisely the Federation Modeling Language ({FML}) that serves to describe them as the main building block in order to solve the process modeling challenge whereas the federation feature is used as a means to provide editing tools for the resulting process language.
+Our solution fulfills all the challenge requirements and is fully implemented with the Openflexo framework.}
 }

### 2022-03-16 2-abstract.tex
clarification between tool and tooling. Tooling was often replaced by tool(s) or toolset

--- 2021-10-13 2-abstract.tex
@@
-Model federation is a multi-model \emph{management} approach based on the use of virtual models and loosely coupled links.
-The models in a federation remain autonomous and represented in their original technological spaces whereas virtual models and links (which are not level bounded) serve as control components used to present different views to the users and  maintain synchronization.
-In this paper we tackle the \emph{MULTI process modeling challenge}, which consists in providing a solution to the problem of specifying and enacting processes.
-Solutions must fulfill a number of requirements for a process representation defined at an abstract process-definition level and at various more concrete domain-specific levels, resulting in a multi-level hierarchy of related models.
-We present a solution based on model federation and discuss the advantages and limitations of using this approach for multi-level modeling.
-Concretely, we use virtual models and more precisely the Federation Modeling Language ({FML}) that serves to describe them as the main building block in order to solve the process modeling challenge whereas the federation aspect is used as a means to provide editing tooling for the resulting process language.
-Our solution is fully implemented with the Openflexo framework.
-It fulfills all the challenge requirements.
+Model federation is a multi-model \emph{management} approach based on the use
+of virtual models and loosely coupled links. The models in a federation remain
+autonomous and represented in their original technological spaces whereas
+virtual models and links (which are not level bounded) serve as control
+components used to present different views to the users and maintain
+synchronization. In this paper we tackle the \emph{MULTI process modeling
+challenge}, which consists in providing a solution to the problem of specifying
+and enacting processes. Solutions must fulfill a number of requirements for a
+process representation defined at an abstract process-definition level and at
+various more concrete domain-specific levels, resulting in a multi-level
+hierarchy of related models. We present a solution based on model federation
+and discuss the advantages and limitations of using this approach for
+multi-level modeling. Concretely, we use virtual models and more precisely the
+Federation Modeling Language ({FML}) that serves to describe them as the main
+building block in order to solve the process modeling challenge whereas the
+federation feature is used as a means to provide editing tools for the
+resulting process language. Our solution fulfills all the challenge
+requirements and is fully implemented with the Openflexo framework.
 

### 2022-03-16 3-introduction.tex
clarification between tool and tooling. Tooling was often replaced by tool(s) or toolset

@@
-Model-driven engineering (\MDE) has traditionally adopted a strict hierarchical two-level approach, where metamodels reside in a certain meta-level, and models are created one meta-level below by using types from the metamodel.
-Notable examples of this approach are the widespread Eclipse Modeling Framework (\EMF)~\parencite{emf} and the Object Management Group Meta-Object Facility (\MOF)~\parencite{omg2013mof}.
+Model-driven engineering (\MDE) has traditionally adopted a strict hierarchical
+two-level approach, where metamodels reside in a certain meta-level, and models
+are created one meta-level below by using types from the metamodel. Notable
+examples of this approach are the widespread Eclipse Modeling Framework
+(\EMF)~\parencite{emf} and the Object Management Group Meta-Object Facility
+(\MOF)~\parencite{omg2013mof}.
 
-This strict approach shows its limitations when modeling complex domains requiring more than one level of specialization, \eg to adapt the model to application subdomains.
-Indeed, the two-level approach fails to acknowledge and support: 1) the existence of different forms of classification and/or instantiation (\eg ontological vs linguistics); and 2) the duality type-object for model elements.
-Therefore, while it may still be used for the aforementioned complex modeling tasks, the strict approach entails the introduction of accidental complexity in both the modeling process and the resulting modeling artifacts.
+This strict approach shows its limitations when modeling complex domains
+requiring more than one level of specialization, \eg to adapt the model to
+application subdomains. Indeed, the two-level approach fails to acknowledge and
+support: 1) the existence of different forms of classification and/or
+instantiation (\eg ontological vs linguistics); and 2) the duality type-object
+for model elements. Therefore, while it may still be used for the
+aforementioned complex modeling tasks, the strict approach entails the
+introduction of accidental complexity in both the modeling process and the
+resulting modeling artifacts.
 
@@
 
-Model federation is a multi-model management approach based on the use of virtual models and loosely coupled links.
-The models in a federation remain autonomous and are represented in their original technological spaces whereas virtual models (also called conceptual models) and links that serve as control components used to present different views to the users and maintain synchronization.
-Our solution is based on the model federation infrastructure.
-Concretely, we use Openflexo and its internal Federation Modeling Language (\FML), a language to create, link and manage virtual models, in order to solve the MULTI process challenge while the federation feature is used solely so as to provide tooling for the resulting process language.
-Our solution, which is fully implemented and executable, meets all the requirements.
-%\noteFabien{Le paragraphe qui précède me semble redire des choses de la section 2, comment fait-on ?}
+Model federation is a multi-model management approach based on the use of
+virtual models and loosely coupled links. The models in a federation remain
+autonomous and are represented in their original technological spaces whereas
+virtual models (also called conceptual models) and links that serve as control
+components used to present different views to the users and maintain
+synchronization. Our solution is based on the model federation infrastructure.
+Concretely, we use Openflexo and its internal Federation Modeling Language
+(\FML), a language to create, link and manage virtual models, in order to solve
+the MULTI process challenge while the federation feature is used solely so as
+to provide tools for the resulting process language. Our solution, which is
+fully implemented and executable, meets all the requirements.
+%\noteFabien{Le paragraphe qui précède me semble redire des choses de la
+%section 2, comment fait-on ?}
 
-Our modeling approach is based on a language (\FML) which provides the (linguistics) concepts that are used to define the multi-level models and meta-models that are needed.
-\FML allows defining models as first class entities (which we call \emph{virtual models}) and linking them together (by extension).
-In the \mpc case study, we exploit this possibility to:
+Our modeling approach is based on a language (\FML) which provides the
+(linguistics) concepts that are used to define the multi-level models and
+meta-models that are needed. \FML allows defining models as first class
+entities (which we call \emph{virtual models}) and linking them together (by
+extension). In the \mpc case study, we exploit this possibility to:
 \begin{itemize}
@@
 Sect.~\ref{sec:requirements}. We assess the modeling solution in
-Sect.~\ref{sec:discussion}. Sect.~\ref{sec:relatedwork} presents the related work before the conclusions.%~\ref{sec:conclusions}.
+Sect.~\ref{sec:discussion}. Sect.~\ref{sec:relatedwork} presents the related
+work before the conclusions.%~\ref{sec:conclusions}.
 

### 2022-03-16 4-technology.tex
clarification between tool and tooling. Tooling was often replaced by tool(s) or toolset

@@
 This framework relies on the architecture of Fig.~\ref{fig:mf}. A federation
-gathers a set of conceptual models, named \emph{virtual models} and a
-set of \emph{federated models}. Each federated model pertains to a
-\emph{technological space} and uses the language of its specific
-paradigm while a virtual model is built using the Federation Modeling
-Language (\FML). Each federated model is an autonomous
-element that may evolve with its own tooling {\color{red} tooling is used in place of tool throughout the article.
-I suggest clarifying it and then editing it all along the article}.
-The virtual models
-serve as control elements binding the federated models together.
-In this paper, we mainly use conceptual models and more precisely \FML, its
-domain specific modeling language. The only use of the federation feature is on the tooling made for the process language.
+gathers a set of conceptual models, named \emph{virtual models} and a set of
+\emph{federated models}. Each federated model pertains to a \emph{technological
+space} and uses the language of its specific paradigm while a virtual model is
+built using the Federation Modeling Language (\FML). Each federated model is an
+autonomous element that may evolve with its own tools. The virtual models serve
+as control elements binding the federated models together.  In this paper, we
+mainly use conceptual models and more precisely \FML, its domain specific
+modeling language. The only use of the federation feature is on the toolset for
+the process language.
 

### 2022-03-16 6-modelpresentation.tex
clarification between tool and tooling. Tooling was often replaced by tool(s) or toolset

--- 2022-03-01 6-modelpresentation.tex
@@
 
-The presentation of our solution to the \mlpc follows a systematic methodology where modeling choices are introduced step by step, as the requirements are stated.
-The satisfaction of requirements is explained throughout this presentation.
+The presentation of our solution to the \mlpc follows a systematic methodology
+where modeling choices are introduced step by step, as the requirements are
+stated. The satisfaction of requirements is explained throughout this
+presentation.
 
@@
 
-The designed multilevel architecture shown in Fig.~\ref{fig:MultilevelArchitecture} captures the two use cases described in the Process Challenge.
+The designed multilevel architecture shown in
+Fig.~\ref{fig:MultilevelArchitecture} captures the two use cases described in
+the Process Challenge.
 
@@
 Fig.~\ref{fig:mm}), see
-Fig.~\ref{fig:BPMNSubsetExample},~\ref{fig:MultilevelArchitecture},~\ref{fig:LinguisticAndOntologicInstantiation},~\ref{fig:BaseMetamodel},~\ref{fig:AcmeFullArchitecture},~\ref{fig:ToolingArchitecture}, while brown is used for instances of
-\textsf{FlexoConcept}, see Fig.~\ref{fig:BPMNSubsetExample},~\ref{fig:MultilevelArchitecture},~\ref{fig:LinguisticAndOntologicInstantiation},~\ref{fig:AcmeFullArchitecture},~\ref{fig:ToolingArchitecture}.
-% 3, 4, 5, 6, 7, 8, 9
-% Expliquer que le use case XSure ne nécessitait pas l'extension de metamodèle requise pour le usecase Acme Software Developement Process
+Fig.~\ref{fig:BPMNSubsetExample},~\ref{fig:MultilevelArchitecture},~\ref{fig:LinguisticAndOntologicInstantiation},~\ref{fig:BaseMetamodel},~\ref{fig:AcmeFullArchitecture},~\ref{fig:ToolingArchitecture},
+while brown is used for instances of \textsf{FlexoConcept}, see
+Fig.~\ref{fig:BPMNSubsetExample},~\ref{fig:MultilevelArchitecture},~\ref{fig:LinguisticAndOntologicInstantiation},~\ref{fig:AcmeFullArchitecture},~\ref{fig:ToolingArchitecture}.
+% 3, 4, 5, 6, 7, 8, 9 Expliquer que le use case XSure ne nécessitait pas
+% l'extension de metamodèle requise pour le usecase Acme Software Developement
+% Process
 
@@
 
-In this section, we present the base metamodel presented in Fig.~\ref{fig:BaseMetamodel} with the \textit{XSure} insurance domain use case, whose partial description was provided in the challenge description.
-The requirements \textbf{P1} to \textbf{P19} of XSure insurance domain are straightforwardly implemented by instantiating the \emph{XSure model}, as an instance of base metamodel (the left side of Fig.~\ref{fig:MultilevelArchitecture}).
+In this section, we present the base metamodel presented in
+Fig.~\ref{fig:BaseMetamodel} with the \textit{XSure} insurance domain use case,
+whose partial description was provided in the challenge description. The
+requirements \textbf{P1} to \textbf{P19} of XSure insurance domain are
+straightforwardly implemented by instantiating the \emph{XSure model}, as an
+instance of base metamodel (the left side of
+Fig.~\ref{fig:MultilevelArchitecture}).
 
-Fig.~\ref{fig:BaseMetamodel} represents this base metamodel with a \UML-like formalism, well adapted to represent \FML concepts and their instances.
-A \textit{Concept} of \FML is represented by a \UML class where roles of basic types are attributes.
-The roles whose types are concepts are represented by an arrow from their concept to their type, with their name on the arrow.
-The cardinality follows the \UML practice.
-For example, the role \texttt{parentActorTypes} of the concept \texttt{ActorType} has the type \texttt{ActorType} and the * cardinality.
+Fig.~\ref{fig:BaseMetamodel} represents this base metamodel with a \UML-like
+formalism, well adapted to represent \FML concepts and their instances. A
+\textit{Concept} of \FML is represented by a \UML class where roles of basic
+types are attributes. The roles whose types are concepts are represented by an
+arrow from their concept to their type, with their name on the arrow. The
+cardinality follows the \UML practice.  For example, the role
+\texttt{parentActorTypes} of the concept \texttt{ActorType} has the type
+\texttt{ActorType} and the * cardinality.
 
-Our proposition relies on ontological instantiation as presented in Fig.~\ref{fig:LinguisticAndOntologicInstantiation}, with a common root concept \texttt{ModelingElement}.
-Two types of ontological instantiation are needed to meet the requirements of the challenge.
-One provides that some instances conform to only one type (\texttt{Process} and \texttt{Task} do respectively conform to \texttt{ProcessType} and \texttt{TaskType}).
-While some entities define their conformity to several types (\texttt{Actor} and \texttt{Artifact} do respectively conform to several \texttt{ActorType} and \texttt{ArtifactType}).
-These instantiations are respectively expressed using the relations \texttt{type} and \texttt{types} between \texttt{Type}, \texttt{Instance} and \texttt{MultiInstance} concepts.
-In the diagram, the realization of these relations are indicated by a derived relation, \texttt{/type} or \texttt{/types}, as for example between \texttt{Process} and \texttt{ProcessType} or between \texttt{Actor} and \texttt{ActorType}.
+Our proposition relies on ontological instantiation as presented in
+Fig.~\ref{fig:LinguisticAndOntologicInstantiation}, with a common root concept
+\texttt{ModelingElement}. Two types of ontological instantiation are needed to
+meet the requirements of the challenge. One provides that some instances
+conform to only one type (\texttt{Process} and \texttt{Task} do respectively
+conform to \texttt{ProcessType} and \texttt{TaskType}). While some entities
+define their conformity to several types (\texttt{Actor} and \texttt{Artifact}
+do respectively conform to several \texttt{ActorType} and
+\texttt{ArtifactType}). These instantiations are respectively expressed using
+the relations \texttt{type} and \texttt{types} between \texttt{Type},
+\texttt{Instance} and \texttt{MultiInstance} concepts. In the diagram, the
+realization of these relations are indicated by a derived relation,
+\texttt{/type} or \texttt{/types}, as for example between \texttt{Process} and
+\texttt{ProcessType} or between \texttt{Actor} and \texttt{ActorType}.
 
@@
 
-We first present the process definition part of the base metamodel, located on the left of Fig.~\ref{fig:BaseMetamodel}.
+We first present the process definition part of the base metamodel, located on
+the left of Fig.~\ref{fig:BaseMetamodel}.
 
@@
 
-\texttt{ProcessType} is a specialization of the \texttt{Type} concept, and references a collection of \texttt{TaskType} through the composition relation \texttt{taskTypes} with (0..*) cardinality (\textbf{P1}).
-A \texttt{TaskType} is embedded in a \texttt{ProcessType} and inherits from its context.
-To illustrate this in \textit{XSure} insurance domain use case, \textit{XSure model} defines \inst{Claim Handling}, instance of \texttt{ProcessType}, and \inst{Receive Claim}, \inst{Assess Claim} and \inst{Pay premium}, instances of \texttt{TaskType}.
+\texttt{ProcessType} is a specialization of the \texttt{Type} concept, and
+references a collection of \texttt{TaskType} through the composition relation
+\texttt{taskTypes} with (0..*) cardinality (\textbf{P1}). A \texttt{TaskType}
+is embedded in a \texttt{ProcessType} and inherits from its context. To
+illustrate this in \textit{XSure} insurance domain use case, \textit{XSure
+model} defines \inst{Claim Handling}, instance of \texttt{ProcessType}, and
+\inst{Receive Claim}, \inst{Assess Claim} and \inst{Pay premium}, instances of
+\texttt{TaskType}.
 
-\texttt{ProcessType} also references a collection of gateways, reified with the \texttt{Gateway} concept hierarchy.
-\texttt{Gateway} is a specialization of \texttt{Type} and is specialized by \texttt{Sequencing}, \texttt{AndSplit}, \texttt{AndJoin}, \texttt{OrSplit} and \texttt{OrJoin} concepts (\textbf{P2}).
-Depending on its type and following its underlying operational semantics, a gateway defines one or more inputs and one or more outputs.
-A \texttt{ProcessType} additionally exposes a unique initial \texttt{TaskType} and a collection of final \texttt{TaskType} with both roles \texttt{initialTaskType} (single cardinality) and \texttt{finalTaskType} (cardinality 0..*) (\textbf{P3}).
-\texttt{TaskType} exposes a creator role as a reference to an \texttt{Actor} concept (\textbf{P4}), which is at the same conceptual level.
+\texttt{ProcessType} also references a collection of gateways, reified with the
+\texttt{Gateway} concept hierarchy. \texttt{Gateway} is a specialization of
+\texttt{Type} and is specialized by \texttt{Sequencing}, \texttt{AndSplit},
+\texttt{AndJoin}, \texttt{OrSplit} and \texttt{OrJoin} concepts (\textbf{P2}).
+Depending on its type and following its underlying operational semantics, a
+gateway defines one or more inputs and one or more outputs. A
+\texttt{ProcessType} additionally exposes a unique initial \texttt{TaskType}
+and a collection of final \texttt{TaskType} with both roles
+\texttt{initialTaskType} (single cardinality) and \texttt{finalTaskType}
+(cardinality 0..*) (\textbf{P3}). \texttt{TaskType} exposes a creator role as a
+reference to an \texttt{Actor} concept (\textbf{P4}), which is at the same
+conceptual level.
 
-The following listing shows an excerpt of the \FML code modeling some core concepts of the process modeling base metamodel.
+The following listing shows an excerpt of the \FML code modeling some core
+concepts of the process modeling base metamodel.
 
@@
 
-The \texttt{ActorType} concept inherits from the \texttt{Type} concept and the \texttt{allowedActorTypes} relation to \texttt{ActorType} defined in \texttt{TaskType} (with 0..* cardinality) captures \textbf{P5} requirement.
-Requirement \textbf{P6} is symmetrically satisfied with \texttt{allowedActors} relation to \texttt{Actor} also defined in \texttt{TaskType} (with 0..* cardinality).
-The same modeling pattern applies to \texttt{ArtefactType} also extending the concept \texttt{Type}, and both relations \texttt{usedArtifactTypes} and \texttt{producedArtifactTypes} defined in \texttt{TaskType} (\textbf{P7}).
-\texttt{TaskType} additionally exposes an \texttt{expectedDuration} role (expressed in number of days), satisfying \textbf{P8}.
+The \texttt{ActorType} concept inherits from the \texttt{Type} concept and the
+\texttt{allowedActorTypes} relation to \texttt{ActorType} defined in
+\texttt{TaskType} (with 0..* cardinality) captures \textbf{P5} requirement.
+Requirement \textbf{P6} is symmetrically satisfied with \texttt{allowedActors}
+relation to \texttt{Actor} also defined in \texttt{TaskType} (with 0..*
+cardinality). The same modeling pattern applies to \texttt{ArtefactType} also
+extending the concept \texttt{Type}, and both relations
+\texttt{usedArtifactTypes} and \texttt{producedArtifactTypes} defined in
+\texttt{TaskType} (\textbf{P7}). \texttt{TaskType} additionally exposes an
+\texttt{expectedDuration} role (expressed in number of days), satisfying
+\textbf{P8}.
 
-The \texttt{Actor} concept defines a boolean attribute called \texttt{isSenior}, while \texttt{TaskType} defines an additional \texttt{isCritical} boolean attribute, indicating that some instances are flagged as critical and must be performed by senior actors.
-To fulfill \textbf{P9} requirement, a supplementary constraint is required for \texttt{TaskType} and is captured through the following invariant expressed in the \FML language:
+The \texttt{Actor} concept defines a boolean attribute called
+\texttt{isSenior}, while \texttt{TaskType} defines an additional
+\texttt{isCritical} boolean attribute, indicating that some instances are
+flagged as critical and must be performed by senior actors. To fulfill
+\textbf{P9} requirement, a supplementary constraint is required for
+\texttt{TaskType} and is captured through the following invariant expressed in
+the \FML language:
 
@@
 
-This invariant should be completed with additional constraints defined in the \texttt{Task} concept, which apply to performing actors assigned to enact tasks.
+This invariant should be completed with additional constraints defined in the
+\texttt{Task} concept, which apply to performing actors assigned to enact
+tasks.
 
@@
 \label{sec:ProcessEnactment}
-We now present the process enactment part of base metamodel, located right of Fig.~\ref{fig:BaseMetamodel}.
-All concepts defined in this subsection are either specializations of the \texttt{Instance} concept (if they have exactly one type) or the \texttt{MultiInstance} concept (when they have several types).
+We now present the process enactment part of base metamodel, located right of
+Fig.~\ref{fig:BaseMetamodel}. All concepts defined in this subsection are
+either specializations of the \texttt{Instance} concept (if they have exactly
+one type) or the \texttt{MultiInstance} concept (when they have several types).
 
-\texttt{Process} represents an enacted \texttt{ProcessType}, as defined in previous subsection (\textbf{P10}).
-\FML defines behavioral features called behavior.
-\texttt{ProcessType} defines the behavior \texttt{newProcess(String)}, taking the name of the process to enact as argument:
+\texttt{Process} represents an enacted \texttt{ProcessType}, as defined in
+previous subsection (\textbf{P10}). \FML defines behavioral features called
+behavior. \texttt{ProcessType} defines the behavior
+\texttt{newProcess(String)}, taking the name of the process to enact as
+argument:
 
@@
 
-Authorization for an actor to perform a task (\textbf{P17}) is captured either by the role \texttt{allowedActors} or the role \texttt{allowedActorTypes} defined in \texttt{TaskType}.
-This mechanism is completed by the behaviors \texttt{isAuthorizedActor}, \texttt{isValidActor} and \texttt{isValidActorType} defined in \texttt{TaskType}:
+Authorization for an actor to perform a task (\textbf{P17}) is captured either
+by the role \texttt{allowedActors} or the role \texttt{allowedActorTypes}
+defined in \texttt{TaskType}. This mechanism is completed by the behaviors
+\texttt{isAuthorizedActor}, \texttt{isValidActor} and \texttt{isValidActorType}
+defined in \texttt{TaskType}:
 
@@
 
-The \texttt{Task} concept delegates this authorization to its task type, as shown in the following \FML code:
+The \texttt{Task} concept delegates this authorization to its task type, as
+shown in the following \FML code:
 
@@
 
-Enforcing those constraints is finally performed by the definition of this invariant in the \texttt{Task} concept:
+Enforcing those constraints is finally performed by the definition of this
+invariant in the \texttt{Task} concept:
 
@@
 
-The default behavior states that all actors and actor types are valid for all task types.
-This modeling scheme offers many extension points, by the redefinition of some behaviors in the inherited concepts (although none were required in the context of \emph{XSure} use case).
+The default behavior states that all actors and actor types are valid for all
+task types. This modeling scheme offers many extension points, by the
+redefinition of some behaviors in the inherited concepts (although none were
+required in the context of \emph{XSure} use case).
 
-Actor types specialization is captured by the \texttt{parentActorTypes} relation defined in \texttt{ActorType} (\textbf{P18}).
-This is completed by both the definition of the \texttt{hasActorType(ActorType)} behavior in \texttt{Actor} and the recursive behavior \texttt{isOrSpecializes(ActorType)} in \texttt{ActorType}:
+Actor types specialization is captured by the \texttt{parentActorTypes}
+relation defined in \texttt{ActorType} (\textbf{P18}). This is completed by
+both the definition of the \texttt{hasActorType(ActorType)} behavior in
+\texttt{Actor} and the recursive behavior \texttt{isOrSpecializes(ActorType)}
+in \texttt{ActorType}:
 
@@
 
-Each concept inherits from \texttt{ModelingElement}, which defines a \texttt{lastUpdated} attribute with \texttt{Date} type, and thus satisfies \textbf{P19} requirement.
+Each concept inherits from \texttt{ModelingElement}, which defines a
+\texttt{lastUpdated} attribute with \texttt{Date} type, and thus satisfies
+\textbf{P19} requirement.
 
@@
 
-The challenge describes in a second part a Software engineering process for a fictional Acme company.
-The Base metamodel described previously is too generic to capture all domain-specific aspects of this use case.
-We chose to complete the architectural hierarchy with a specific virtual model, specific to the Acme software development process metamodel, shown in Fig.~\ref{fig:AcmeFullArchitecture}.
-The \textit{Acme} metamodel specializes the Base metamodel, and \textit{Acme model} is an instance of the \textit{Acme} metamodel.
-The metamodel inheritance is implemented by virtual model's inheritance.
-Any instance of the Acme model is either an instance of a concept defined in the Base metamodel, or a concept defined in the specialized \textit{Acme} metamodel.
+The challenge describes in a second part a Software engineering process for a
+fictional Acme company. The Base metamodel described previously is too generic
+to capture all domain-specific aspects of this use case. We chose to complete
+the architectural hierarchy with a specific virtual model, specific to the Acme
+software development process metamodel, shown in
+Fig.~\ref{fig:AcmeFullArchitecture}.  The \textit{Acme} metamodel specializes
+the Base metamodel, and \textit{Acme model} is an instance of the \textit{Acme}
+metamodel. The metamodel inheritance is implemented by virtual model's
+inheritance. Any instance of the Acme model is either an instance of a concept
+defined in the Base metamodel, or a concept defined in the specialized
+\textit{Acme} metamodel.
 % FML follows a classical object-oriented semantics. ???
@@
 
-The Fig.~\ref{fig:AcmeFullArchitecture} highlights all conceptual levels of the architecture.
-This figure is only partial at instance level and not all instances are represented.
-It also contains an instance of enacted a \textit{Software Engineering Process} at the bottom.
-The Acme metamodel specializes some concepts of the Base metamodel, for example, \texttt{SETaskType} extends \texttt{TaskType} and \texttt{SEArtifactType} extends \texttt{ArtifactType}.
-These concepts are further specialized  \texttt{CodingTaskType} extends \texttt{SETaskType} and \texttt{CodeArtifactType} extends \texttt{SEArtifactType}.
-The Acme metamodel also defines a specialized task type \texttt{Coding} extending \texttt{Task}, and provides the \texttt{Developer} concept extending \texttt{ActorType}.
-This metamodel is completed with the \texttt{ProgrammingLanguage} concept, defined as an enumeration (\texttt{Java}, \texttt{C}, \texttt{COBOL}).
-All the instances required to capture the challenge use case are defined in the final Acme model (lower part of the figure).
-Instances are represented with rounded boxes and linguistic instantiation is represented using dashed connectors.
+The Fig.~\ref{fig:AcmeFullArchitecture} highlights all conceptual levels of the
+architecture. This figure is only partial at instance level and not all
+instances are represented. It also contains an instance of enacted a
+\textit{Software Engineering Process} at the bottom. The Acme metamodel
+specializes some concepts of the Base metamodel, for example,
+\texttt{SETaskType} extends \texttt{TaskType} and \texttt{SEArtifactType}
+extends \texttt{ArtifactType}. These concepts are further specialized
+\texttt{CodingTaskType} extends \texttt{SETaskType} and
+\texttt{CodeArtifactType} extends \texttt{SEArtifactType}.  The Acme metamodel
+also defines a specialized task type \texttt{Coding} extending \texttt{Task},
+and provides the \texttt{Developer} concept extending \texttt{ActorType}. This
+metamodel is completed with the \texttt{ProgrammingLanguage} concept, defined
+as an enumeration (\texttt{Java}, \texttt{C}, \texttt{COBOL}). All the
+instances required to capture the challenge use case are defined in the final
+Acme model (lower part of the figure). Instances are represented with rounded
+boxes and linguistic instantiation is represented using dashed connectors.
 
-The \textit{Software Development Process} for Acme company is presented in the Fig.~\ref{fig:AcmeSoftwareDevelopmentProcess}.
-It is a screen capture of the tool developed for the challenge and described in the next section.
+The \textit{Software Development Process} for Acme company is presented in the
+Fig.~\ref{fig:AcmeSoftwareDevelopmentProcess}. It is a screen capture of the
+tool developed for the challenge and described in the next section.
 
@@
 
-\inst{Requirements analysis} is an instance of \texttt{SETaskType}, \inst{Analyst} is an instance of \texttt{ActorType} and \inst{Requirement specifications} is an instance of \texttt{SEArtifactType}.
-The value of the property \texttt{producedArtifactTypes} of \inst{Requirements analysis} is \texttt{\{}\inst{Requirement specifications}\texttt{\}} and is \texttt{\{}\inst{Analyst}\texttt{\}} for \texttt{allowedActorTypes} (\textbf{S1}).
-Similarly, \inst{Test case design} is an instance of \texttt{SETaskType}, with a value of \texttt{true} for its property \texttt{isCritical}, and bound to \inst{Analyst} by the \texttt{allowedActorTypes} relation.
-\inst{Test case design} produces \inst{Test cases} instance of \texttt{SETaskType} (\textbf{S2}).
-The latter is used in \inst{Test design review} (\texttt{producedArtifactTypes} relation) (\textbf{S1} and \textbf{S13}, satisfied with \textbf{P9}).
+\inst{Requirements analysis} is an instance of \texttt{SETaskType},
+\inst{Analyst} is an instance of \texttt{ActorType} and \inst{Requirement
+specifications} is an instance of \texttt{SEArtifactType}. The value of the
+property \texttt{producedArtifactTypes} of \inst{Requirements analysis} is
+\texttt{\{}\inst{Requirement specifications}\texttt{\}} and is
+\texttt{\{}\inst{Analyst}\texttt{\}} for \texttt{allowedActorTypes}
+(\textbf{S1}). Similarly, \inst{Test case design} is an instance of
+\texttt{SETaskType}, with a value of \texttt{true} for its property
+\texttt{isCritical}, and bound to \inst{Analyst} by the
+\texttt{allowedActorTypes} relation. \inst{Test case design} produces
+\inst{Test cases} instance of \texttt{SETaskType} (\textbf{S2}). The latter is
+used in \inst{Test design review} (\texttt{producedArtifactTypes} relation)
+(\textbf{S1} and \textbf{S13}, satisfied with \textbf{P9}).
 
-The capture of \textbf{S3} requirement is performed while defining \inst{Coding} as an instance of concept \texttt{CodingTaskType} and defining a relation \texttt{languages} to \texttt{ProgrammingLanguage} with 1..* cardinality.
-The \texttt{newTask(String)} behavior overrides the generic implementation by instantiating a \texttt{CodingTask} concept, specializing \texttt{Task}, as shown in the following excerpt of \FML code:
+The capture of \textbf{S3} requirement is performed while defining
+\inst{Coding} as an instance of concept \texttt{CodingTaskType} and defining a
+relation \texttt{languages} to \texttt{ProgrammingLanguage} with 1..*
+cardinality.  The \texttt{newTask(String)} behavior overrides the generic
+implementation by instantiating a \texttt{CodingTask} concept, specializing
+\texttt{Task}, as shown in the following excerpt of \FML code:
 
@@
 
-The \texttt{CodeArtifact} concept extends \texttt{SEArtifact}, and defines a relation \texttt{languages} to \texttt{ProgrammingLanguage} with 1..* cardinality (\textbf{S4}).
+The \texttt{CodeArtifact} concept extends \texttt{SEArtifact}, and defines a
+relation \texttt{languages} to \texttt{ProgrammingLanguage} with 1..*
+cardinality (\textbf{S4}).
 
-\textbf{S5} is more ambiguous as a task type \texttt{CodingTask} defines one or more programming languages but produces code which is expressed in one language only.
-This requirement is captured through the redefinition of the behavior \texttt{declaresRequiredProducedArtifacts} where programming language should also match.
-\textbf{S6} is guaranteed though the \texttt{languages} relation defined in \texttt{CodeArtifact} and the following invariant declared in \texttt{CodeArtifact}:
+\textbf{S5} is more ambiguous as a task type \texttt{CodingTask} defines one or
+more programming languages but produces code which is expressed in one language
+only. This requirement is captured through the redefinition of the behavior
+\texttt{declaresRequiredProducedArtifacts} where programming language should
+also match.  \textbf{S6} is guaranteed though the \texttt{languages} relation
+defined in \texttt{CodeArtifact} and the following invariant declared in
+\texttt{CodeArtifact}:
 \begin{lstlisting}
@@
 
-The requirement \textbf{S8} is simply expressed by the definition of the \inst{Testing} instance of \texttt{SETaskType}, \inst{Tester} instance of \texttt{ActorType}, and \inst{Test report} instance of \texttt{ArtifactType}, as shown in the Fig.~\ref{fig:AcmeSoftwareDevelopmentProcess}.
+The requirement \textbf{S8} is simply expressed by the definition of the
+\inst{Testing} instance of \texttt{SETaskType}, \inst{Tester} instance of
+\texttt{ActorType}, and \inst{Test report} instance of \texttt{ArtifactType},
+as shown in the Fig.~\ref{fig:AcmeSoftwareDevelopmentProcess}.
 
-A critical task must additionally produce artifacts that must be associated with a validation task.
-This is modeled with a supplementary relation \texttt{validatedWith} defined in \texttt{Artifact} and referencing \texttt{Artifact} validating it (\textbf{S9}).
+A critical task must additionally produce artifacts that must be associated
+with a validation task. This is modeled with a supplementary relation
+\texttt{validatedWith} defined in \texttt{Artifact} and referencing
+\texttt{Artifact} validating it (\textbf{S9}).
 
-All software engineering artifacts defined in the context of Acme Software Engineering Process are  instances of \texttt{SEArtifact}.
-This concept defines two attributes: \texttt{responsible} (an \texttt{Actor} instance), and \texttt{versionNumber} (an integer).
-It thus fulfills \textbf{S10}.
-\inst{Bob Brown} is declared as an instance of \texttt{Actor}, and references \inst{Analyst} and \inst{Tester} (\texttt{ActorType} instances).
-He is also referenced by all instances of \texttt{SETaskType} as the creator for related task types (\textbf{S11}).
+All software engineering artifacts defined in the context of Acme Software
+Engineering Process are  instances of \texttt{SEArtifact}. This concept defines
+two attributes: \texttt{responsible} (an \texttt{Actor} instance), and
+\texttt{versionNumber} (an integer). It thus fulfills \textbf{S10}. \inst{Bob
+Brown} is declared as an instance of \texttt{Actor}, and references
+\inst{Analyst} and \inst{Tester} (\texttt{ActorType} instances). He is also
+referenced by all instances of \texttt{SETaskType} as the creator for related
+task types (\textbf{S11}).
 
-For \textbf{S12}, we assume that all tasks may define an expected duration, which might be checked during process execution.
-This is modeled by the \texttt{expectedDuration} attribute in \texttt{TaskType}.
-Alternatively, this could be modeled through the definition of a \texttt{SETesting} concept, as a specialization of \texttt{SETaskType}, and the instantiation of \inst{Testing} as an instance of \texttt{SETesting}.
-The business logic expressed by \textbf{S12} requirement should then be redefined in \texttt{SETesting}.
+For \textbf{S12}, we assume that all tasks may define an expected duration,
+which might be checked during process execution. This is modeled by the
+\texttt{expectedDuration} attribute in \texttt{TaskType}. Alternatively, this
+could be modeled through the definition of a \texttt{SETesting} concept, as a
+specialization of \texttt{SETaskType}, and the instantiation of \inst{Testing}
+as an instance of \texttt{SETesting}. The business logic expressed by
+\textbf{S12} requirement should then be redefined in \texttt{SETesting}.
 
-\textbf{S13} requirement has been previously partially fulfilled.
-This must be completed with an association between an artifact produced and the task that validates it.
-This is modeled by the reference to \inst{Test Cases} as a produced artifact type in \inst{Test Case design} and the reference to \inst{Test Cases} as a used artifact in \inst{Test Design review}, as shown in Fig.~\ref{fig:AcmeSoftwareDevelopmentProcess}.
+\textbf{S13} requirement has been previously partially fulfilled. This must be
+completed with an association between an artifact produced and the task that
+validates it. This is modeled by the reference to \inst{Test Cases} as a
+produced artifact type in \inst{Test Case design} and the reference to
+\inst{Test Cases} as a used artifact in \inst{Test Design review}, as shown in
+Fig.~\ref{fig:AcmeSoftwareDevelopmentProcess}.
 
-\subsection{Openflexo tooling}
+\subsection{Openflexo toolset}
 \label{subsec:tooling}
@@
      \includegraphics[width=0.9 \textwidth]{Figures/ToolingArchitecture-1.5.pdf}
-     \caption{Tooling architecture}
+     \caption{Tool architecture}
     \label{fig:ToolingArchitecture}
@@
 
-Our solution is fully implemented within the Openflexo tool. Both use cases have been modeled in the interactive design environment and can be executed by the \FML execution engine.
+Our solution is fully implemented within the Openflexo tool. Both use cases
+have been modeled in the interactive design environment and can be executed by
+the \FML execution engine.
 
-We took advantage of model federation and the availability of diagraming features though the \textit{Diagraming Technology Adapter} to implement two interactive graphical tools built on top of the conceptual levels detailed in previous sections.
-The Fig.~\ref{fig:ToolingArchitecture} shows the overview of the tooling architecture and is detailed in the next section.
-A first tool offers a graphical edition of a Process type.
-The second one offers an enactment feature (the instantiation of a process from its process type), the ability within an enacted process to assign tasks to actors, and a graphical visualization of this process execution.
+We took advantage of model federation and the availability of diagraming
+features though the \textit{Diagraming Technology Adapter} to implement two
+interactive graphical tools built on top of the conceptual levels detailed in
+previous sections. The Fig.~\ref{fig:ToolingArchitecture} shows the overview of
+the tool architecture and is detailed in the next section. A first tool offers
+a graphical edition of a Process type. The second one offers an enactment
+feature (the instantiation of a process from its process type), the ability
+within an enacted process to assign tasks to actors, and a graphical
+visualization of this process execution.
 
@@
 abstract \texttt{execute(Process)} behavior. The management of artifacts whose
-semantics follows rules defined in Sect.~\ref{sec:ProcessEnactment} (\textbf{P13}) is also implemented.
+semantics follows rules defined in Sect.~\ref{sec:ProcessEnactment}
+(\textbf{P13}) is also implemented.
 
@@
 allows representing and editing a \texttt{ProcessType}. The highly reflective
-nature of \FML and its tooling should be noted. When the drag and drop
+nature of \FML and its tools should be noted. When the drag and drop
 interactor is applied for a new item, the tool allows choosing the concept type
@@
 
-The other tool, called \textit{Enacted process graphical editor} and shown on Fig.~\ref{fig:ScreenshotEnactedProcessEditor}, provides process edition and tasks assignations through a graphical visualization displaying the process being executed.
-This tool is represented on the right side of the Fig.~\ref{fig:ToolingArchitecture}.
-It follows the same architecture pattern of the \textit{ProcessType graphical editor}, with two model slots referencing both the model and a diagram.
-Process enactment is operated from a \texttt{ProcessType}, and must be defined using a name identifying the newly instantiated process.
-All tasks are created from their \texttt{TaskType} definition, and required assignments apply (if for example \inst{Coding} \texttt{TaskType} defines COBOL as output programming language, the related task is automatically assigned to \inst{Ann Smith}).
-Each task has a status as well as a set of performing actors, a begin and end dates, and used and produced artifacts.
+The other tool, called \textit{Enacted process graphical editor} and shown on
+Fig.~\ref{fig:ScreenshotEnactedProcessEditor}, provides process edition and
+tasks assignations through a graphical visualization displaying the process
+being executed. This tool is represented on the right side of the
+Fig.~\ref{fig:ToolingArchitecture}. It follows the same architecture pattern of
+the \textit{ProcessType graphical editor}, with two model slots referencing
+both the model and a diagram. Process enactment is operated from a
+\texttt{ProcessType}, and must be defined using a name identifying the newly
+instantiated process. All tasks are created from their \texttt{TaskType}
+definition, and required assignments apply (if for example \inst{Coding}
+\texttt{TaskType} defines COBOL as output programming language, the related
+task is automatically assigned to \inst{Ann Smith}). Each task has a status as
+well as a set of performing actors, a begin and end dates, and used and
+produced artifacts.
 
-A demonstrative video for the developed tooling is available on our website\footnote{\url{https://research.openflexo.org/MLMChallenge.html}}, as well as download and installation instructions.
+A demonstrative video for the developed tool is available on our
+website\footnote{\url{https://research.openflexo.org/MLMChallenge.html}}, as
+well as download and installation instructions.

### 2022-03-16 7-requirementssatisfaction.tex
clarification between tool and tooling. Tooling was often replaced by tool(s) or toolset

--- 2022-02-23 7-requirementssatisfaction.tex
@@
 
-In the previous Sect.~\ref{sec:model}, we demonstrated that our solution satisfies all the challenge requirements.
-In this section, we summarize the different techniques used to satisfy them, without repeating all the justifications.
-We classify these techniques into three categories:
+In the previous Sect.~\ref{sec:model}, we demonstrated that our solution
+satisfies all the challenge requirements. In this section, we summarize the
+different techniques used to satisfy them, without repeating all the
+justifications. We classify these techniques into three categories:
 \begin{itemize}
-    \item \emph{syntactic conformance}: the requirements are structurally satisfied when the model conforms to its metamodel,
-    \item \emph{constraints checking}: the requirements are fulfilled when all instances satisfy both syntactic conformance and the evaluation of invariants expressed for related concepts,
-    \item \emph{tooling}: the requirements are enforced by the implementation of the associated tooling.
+    \item \emph{syntactic conformance}: the requirements are structurally
+      satisfied when the model conforms to its metamodel,
+    \item \emph{constraints checking}: the requirements are fulfilled when all
+      instances satisfy both syntactic conformance and the evaluation of
+      invariants expressed for related concepts,
+    \item \emph{tooling}: the requirements are enforced by the implementation
+      of the associated tool.
 \end{itemize}
 
-Tab.~\ref{tab:RequirementsSatisfactionBaseMetamodel} summarizes the requirement coverage for the base metamodel illustrated by XSure insurance use case while Tab.~\ref{tab:RequirementsSatisfactionAcme} shows requirements satisfaction for Acme software engineering process.
-Syntactic conformance is more precisely classified into six subcategories:
+Tab.~\ref{tab:RequirementsSatisfactionBaseMetamodel} summarizes the requirement
+coverage for the base metamodel illustrated by XSure insurance use case while
+Tab.~\ref{tab:RequirementsSatisfactionAcme} shows requirements satisfaction for
+Acme software engineering process. Syntactic conformance is more precisely
+classified into six subcategories:
 
 \begin{itemize}
-    \item \emph{Conceptualization} (\textit{Conc.}): a concept is created to fulfill or meet the requirement (for example for \textbf{P1}: \textit{process type} is conceptualized in a concept \texttt{ProcessType}).
+    \item \emph{Conceptualization} (\textit{Conc.}): a concept is created to
+      fulfill or meet the requirement (for example for \textbf{P1}:
+      \textit{process type} is conceptualized in a concept
+    \texttt{ProcessType}).
 
-    \item \emph{Specialization} (\textit{Spec.}): a concept is specialized in the sense of object-oriented modeling (for example for \textbf{P2}: \texttt{Gateway} defines an abstract behavior \texttt{execute()} and \texttt{Sequencing}, \texttt{AndSplit}, \texttt{AndJoin}, \texttt{OrSplit} and \texttt{OrJoin} are defined as inherited concepts implementing specific business logic).
+    \item \emph{Specialization} (\textit{Spec.}): a concept is specialized in
+      the sense of object-oriented modeling (for example for \textbf{P2}:
+      \texttt{Gateway} defines an abstract behavior \texttt{execute()} and
+    \texttt{Sequencing}, \texttt{AndSplit}, \texttt{AndJoin}, \texttt{OrSplit}
+  and \texttt{OrJoin} are defined as inherited concepts implementing specific
+business logic).
 
-    \item \emph{Composition} (\textit{Comp.}): some concepts are in relation, or specific attributes were added to some concepts (for example for \textbf{P3}: a \textit{process type} has one \textit{initial task type}).
+    \item \emph{Composition} (\textit{Comp.}): some concepts are in relation,
+      or specific attributes were added to some concepts (for example for
+      \textbf{P3}: a \textit{process type} has one \textit{initial task type}).
 
-    \item \emph{Linguistic instantiation} (\textit{L.Inst.}): a requirement is satisfied by the instantiation of one or more concepts (for example for \textbf{S1}: \inst{Requirement analysis} is defined as an instance of \texttt{SETaskType}).
+    \item \emph{Linguistic instantiation} (\textit{L.Inst.}): a requirement is
+      satisfied by the instantiation of one or more concepts (for example for
+      \textbf{S1}: \inst{Requirement analysis} is defined as an instance of
+    \texttt{SETaskType}).
 
-    \item \emph{Ontological instantiation} (\textit{O.Inst.}): fulfillment of a requirement is obtained by the relation between an instance and its type definition instance (for example the notion of \textit{developer} of \textbf{S3} is implemented by both a concept in \textit{Acme metamodel} and an instance in \textit{Acme model}, see bottom left of Fig.~\ref{fig:AcmeFullArchitecture}).
+    \item \emph{Ontological instantiation} (\textit{O.Inst.}): fulfillment of a
+      requirement is obtained by the relation between an instance and its type
+      definition instance (for example the notion of \textit{developer} of
+      \textbf{S3} is implemented by both a concept in \textit{Acme metamodel}
+    and an instance in \textit{Acme model}, see bottom left of
+  Fig.~\ref{fig:AcmeFullArchitecture}).
 
-    \item \emph{Behavioral modeling} (\textit{B.Mod.}): a requirement is satisfied by the definition of one or more behaviors, which may be combined with constraints checking and associated tooling (for example \textbf{P17}).
+    \item \emph{Behavioral modeling} (\textit{B.Mod.}): a requirement is
+  satisfied by the definition of one or more behaviors, which may be combined
+  with constraints checking and associated tools (for example \textbf{P17}).
   \end{itemize}

### 2022-03-16 8-discussion.tex
clarification between tool and tooling. Tooling was often replaced by tool(s) or toolset

--- 2022-02-23 8-discussion.tex
@@
 
-%\todo[inline]{on pourrait aussi répondre ici aux 3 questions posées
-%explicitement (copiées dans l'intro pour le moment), a priori non, le chef
-%« trouve que c'est chiant »}
-
-%\noteJC{pour le moment, seulement réponse aux différents points, on ne voit pas
-%de discussion sur les choix, sur les potentiels compromis/déficiencies}
-
-In this section, we discuss our multi-level model solution with regard to the required aspects
-mentioned in the challenge.
+In this section, we discuss our multi-level model solution with regard to the
+required aspects mentioned in the challenge.
 
@@
   %Explain the basic modeling constructs used in the solution.
-
-%  \noteJC{pas clair pour moi ce qui est entendu par « basic modeling constructs », ce qui est dans la partie 2 ?}
 
@@
   \texttt{Developer} concept from \texttt{Acme} metamodel specializes the
-  \texttt{ActorType} of the base metamodel. The \texttt{Software Engineering Process} of the
-  \texttt{Acme} model is an instance of \texttt{ProcessType} in the metamodel.
+  \texttt{ActorType} of the base metamodel. The \texttt{Software Engineering
+  Process} of the \texttt{Acme} model is an instance of \texttt{ProcessType} in
+  the metamodel.
 %  \textcolor{red}{Phrase faite a relire sur Developer\dots}
@@
   has to create it manually. This can be a limit of our approach: the cost of
-  the flexibility of our tooling is a limited number of automated behaviors.
+  the flexibility of our toolset is a limited number of automated behaviors.
 
@@
 We use the Openflexo infrastructure to build models and metamodels together,
-and to provide dedicated tooling to edit a Process type, to enact a process and
+and to provide dedicated tools to edit a Process type, to enact a process and
 to execute it. Being able to provide a solution and its associated tools
@@
 
-Our tooling includes mechanisms to verify the models. First, syntactical
+Our toolset includes mechanisms to verify the models. First, syntactical
 consistency is realized by cardinality checks and by typing. Second, the

### 2022-03-16 6-modelpresentation.tex
typos, missing reference, rephrasing

@@
 the Process Challenge.
-
 Note that in all following figures, blue is used to represent \FML virtual
@@
 Fig.~\ref{fig:LinguisticAndOntologicInstantiation}, with a common root concept
-\texttt{ModelingElement}. Two types of ontological instantiation are needed to
+\texttt{ModelingElement}. Two types of ontological instantiations are needed to
 meet the requirements of the challenge. One provides that some instances
@@
 
-We first present the process definition part of the base metamodel, located on
-the left of Fig.~\ref{fig:BaseMetamodel}.
+In this subsection, we present the process definition part of the base
+metamodel, located on the left of Fig.~\ref{fig:BaseMetamodel}.
 
@@
 \label{sec:ProcessEnactment}
-We now present the process enactment part of base metamodel, located right of
-Fig.~\ref{fig:BaseMetamodel}. All concepts defined in this subsection are
-either specializations of the \texttt{Instance} concept (if they have exactly
-one type) or the \texttt{MultiInstance} concept (when they have several types).
+In this subsection, we present the process enactment part of base metamodel,
+located right of Fig.~\ref{fig:BaseMetamodel}. All concepts defined in this
+subsection are either specializations of the \texttt{Instance} concept (if they
+  have exactly one type) or the \texttt{MultiInstance} concept (when they have
+  several types).
 
@@
 instances required to capture the challenge use case are defined in the final
-Acme model (lower part of the figure). Instances are represented with rounded
-boxes and linguistic instantiation is represented using dashed connectors.
+Acme model (lower part of the Fig.~\ref{fig:AcmeFullArchitecture}). Instances
+are represented with rounded boxes and linguistic instantiation is represented
+using dashed connectors.
 

### 2022-03-16 1-main.tex
trix to fix cosmetics in a footnote

@@
+\RequirePackage[hyphens]{url}
 \documentclass[american]{emisa}
@@
 \usepackage{listings}
+%\usepackage[hyphens]{url}
 %\usepackage{xcolor}
@@
   unicode=true,
+  breaklinks=true,
   pdfkeywords={Multi-Level Modeling Challenge, Federation Modeling Language,

### 2022-03-17 8-discussion.tex

@@
   As for cross-level relationships, cross-level constraints do not have a
-  specific nature. They are like any constraint a user can define by adding a
+  specific nature. They are like any constraint that a user can define by adding a
   behavior to a model element. Therefore, if a constraint is mandatory when

### 2022-03-17 4-technology.tex
parallel in naming convention between text and fig.2 => explicit mention of FlexoThings

@@
 
-{\color{red} This paragraph needs a parallel in naming convention between Figure 2 and the text.
-FlexoConcept, FlexoBehavior, etc.} A \UML like representation of a simplified metamodel of \FML is provided in Fig.~\ref{fig:mm}.
-\FML is a language designed to define virtual models.
-A \emph{virtual model} is composed of a set of \emph{concepts}, while itself being a \emph{concept}.
-Hence, \emph{virtual models} are structuring units forming architectures while \emph{concepts} are the core entities.
-A concept has a set of \emph{properties} and \emph{behaviors}.
-\emph{Properties} can be either simple variables, roles (pointers to a modeling element in a \emph{virtual model} or an external technology-specific model), or properties bound to complex control graphs.
-\FML is designed to define not only the structure of virtual models but also the collection of actions that can be performed on them.
-These actions are called \emph{behaviors} and rely on behavioral primitives called \textsf{EditionAction}.
-These \emph{behaviors} can either be called or triggered by events.
-The reactive behaviors are useful when a federated model evolution needs to trigger a computation.
-Note that we do not exploit this possibility in the challenge.
+A \UML like representation of a simplified metamodel of \FML is provided in
+Fig.~\ref{fig:mm}. \FML is a language designed to define virtual models. A
+\emph{virtual model} is composed of a set of \emph{concepts} (named
+\textsf{FlexoConcept} in Fig.~\ref{fig:mm}), while itself being a
+\emph{concept}. Hence, \emph{virtual models} are structuring units forming
+architectures while \emph{concepts} are the core entities. A concept has a set
+of \emph{properties} (\textsf{FlexoProperty}) and \emph{behaviors}
+(\textsf{FlexoBehaviour}). \emph{Properties} can be either simple variables,
+roles (pointers to a modeling element in a \emph{virtual model} or an external
+technology-specific model), or properties bound to complex control graphs. \FML
+is designed to define not only the structure of virtual models but also the
+collection of actions that can be performed on them. These actions are called
+\emph{behaviors} and rely on behavioral primitives called
+\textsf{EditionAction}. These \emph{behaviors} can either be called or
+triggered by events. The reactive behaviors are useful when a federated model
+evolution needs to trigger a computation. Note that we do not exploit this
+possibility in the challenge.
 
-A parallel to object-oriented approach can be useful to understand \FML\footnote{Some features of \FML do not exist in object oriented approach.}.
-A concept corresponds to a class, its properties to the attributes of the class and its behaviors to the methods of the class.
-These properties have types defining the kind of value the role will point at runtime.
-Whenever an element external to the federation space is used, one needs to use a \emph{model slot}.
-A model slot is a mediation entity in charge of giving access to external elements using a \emph{technology adapter}\footnote{It is a reusable library that defines connections between the \FML execution engine and a particular technological space.}.
+A parallel to object-oriented approach can be useful to understand
+\FML\footnote{Some features of \FML do not exist in object oriented approach.}.
+A concept corresponds to a class, its properties to the attributes of the class
+and its behaviors to the methods of the class. These properties have types
+defining the kind of value the role will point at runtime. Whenever an element
+external to the federation space is used, one needs to use a \emph{model slot}.
+A model slot is a mediation entity in charge of giving access to external
+elements using a \emph{technology adapter}\footnote{It is a reusable library
+that defines connections between the \FML execution engine and a particular
+technological space.}.
 
@@
 To illustrate this, Fig.~\ref{fig:BPMNSubsetExample} shows how to realize a
-small subset of \BPMN with Openflexo/FML.
-Concretely, we have designed a \textit{virtual model} (\texttt{SimplifiedBPMN}) with 3 \textit{concepts} representing a process and its contained tasks.
-An instance of this \textit{virtual model} is presented in the bottom of the figure, showing instances with their \textsf{instanceOf} relationships with their model.
+small subset of \BPMN with Openflexo/FML. Concretely, we have designed a
+\textit{virtual model} (\texttt{SimplifiedBPMN}) with 3 \textit{concepts}
+representing a process and its contained tasks. An instance of this
+\textit{virtual model} is presented in the bottom of the figure, showing
+instances with their \textsf{instanceOf} relationships with their model.
 %
-The following listing shows the \FML code of this \textit{virtual model}.
-Note that the composition relation of Fig.~\ref{fig:BPMNSubsetExample} between \texttt{BPMNProcess} and \texttt{BPMNTask} is realized in the FML code by concept containment.
-Whenever an instance of \texttt{BPMNTask} is created, it must be in a container (instance of \texttt{BPMNProcess}) and the two instances get linked together.
-%\noteFabien{A valider}
+The following listing shows the \FML code of this \textit{virtual model}. Note
+that the composition relation of Fig.~\ref{fig:BPMNSubsetExample} between
+\texttt{BPMNProcess} and \texttt{BPMNTask} is realized in the FML code by
+concept containment. Whenever an instance of \texttt{BPMNTask} is created, it
+must be in a container (instance of \texttt{BPMNProcess}) and the two instances
+get linked together. %\noteFabien{A valider}
 
@@
 
-Finally, tools have their own model. We have taken advantage of Openflexo features to build, in addition to the models  required by the challenge, a drawing tool that makes our
-solution (partially) executable.
+Finally, tools have their own model. We have taken advantage of Openflexo
+features to build, in addition to the models  required by the challenge, a
+drawing tool that makes our solution (partially) executable.

### 2022-03-18 6-modelpresentation.tex
try to rephrase a paragraph

@@
 \texttt{ModelingElement}. Two types of ontological instantiations are needed to
-meet the requirements of the challenge. One provides that some instances
-conform to only one type (\texttt{Process} and \texttt{Task} do respectively
-conform to \texttt{ProcessType} and \texttt{TaskType}). While some entities
-define their conformity to several types (\texttt{Actor} and \texttt{Artifact}
-do respectively conform to several \texttt{ActorType} and
-\texttt{ArtifactType}). These instantiations are respectively expressed using
+meet the requirements of the challenge. 
+The first one is characterized by the fact that instances conform to only one
+type (\texttt{Process} and \texttt{Task} do respectively conform to
+\texttt{ProcessType} and \texttt{TaskType}). While the second type of
+ontological instantiations allows some entities to define their conformity to
+several types (\texttt{Actor} and \texttt{Artifact} do respectively conform to
+several \texttt{ActorType} and \texttt{ArtifactType}). 
+%One provides that some instances
+%conform to only one type (\texttt{Process} and \texttt{Task} do respectively
+%conform to \texttt{ProcessType} and \texttt{TaskType}). While some entities
+%define their conformity to several types (\texttt{Actor} and \texttt{Artifact}
+%do respectively conform to several \texttt{ActorType} and
+%\texttt{ArtifactType}). 
+These instantiations are respectively expressed using
 the relations \texttt{type} and \texttt{types} between \texttt{Type},
@@
 the \FML language:
-
 \begin{lstlisting}

### 2022-03-25 4-technology.tex
retrait d'une incise

@@
 work together. As an illustration of this feature, we developed a
-free-modeling editor -- freeing oneself from the bonds of
-model/metamodel conformity -- that is presented
-in~\parencite{models2016-freemodel}. Another notable feature of
+free-modeling editor presented
+in~\parencite{models2016-freemodel} that frees the designer from the bonds of
+model/metamodel conformity. Another notable feature of
 this approach is the strong decoupling among tools that remain usable

### 2022-03-25 4-technology.tex
fix space between footnote number and punctuation

--- 2022-03-17 4-technology.tex
@@
 A parallel to object-oriented approach can be useful to understand
-\FML\footnote{Some features of \FML do not exist in object oriented approach.}.
-A concept corresponds to a class, its properties to the attributes of the class
-and its behaviors to the methods of the class. These properties have types
-defining the kind of value the role will point at runtime. Whenever an element
-external to the federation space is used, one needs to use a \emph{model slot}.
-A model slot is a mediation entity in charge of giving access to external
-elements using a \emph{technology adapter}\footnote{It is a reusable library
-that defines connections between the \FML execution engine and a particular
-technological space.}.
+\FML\footnote{Some features of \FML do not exist in object oriented
+approach.}\!.  A concept corresponds to a class, its properties to the
+attributes of the class and its behaviors to the methods of the class. These
+properties have types defining the kind of value the role will point at
+runtime. Whenever an element external to the federation space is used, one
+needs to use a \emph{model slot}.  A model slot is a mediation entity in charge
+of giving access to external elements using a \emph{technology
+adapter}\footnote{It is a reusable library that defines connections between the
+\FML execution engine and a particular technological space.}.
 

### 2022-03-25 6-modelpresentation.tex
fix metamodel: cardinality of creator (between Actor and TaskType) was * and should be 1
format all requirements with new formatting

@@
     % \includegraphics[width=1.0 \columnwidth]{Figures/Metamodel.pdf}
-    \includegraphics[width=1.0 \textwidth]{Figures/Metamodel-1.5.pdf}
+    %\includegraphics[width=1.0 \textwidth]{Figures/Metamodel-1.5.pdf}
+    %fix creator cardinality (between Actor and TaskType)
+    \includegraphics[width=1.0 \textwidth]{Figures/Metamodel-17.pdf}
      \caption{Process management base metamodel (\texttt{Gateway}'s associations with \texttt{TaskType} are represented with attributes)}
@@
 whose partial description was provided in the challenge description. The
-requirements \textbf{P1} to \textbf{P19} of XSure insurance domain are
+requirements \reqp{1} to \reqp{19} of XSure insurance domain are
 straightforwardly implemented by instantiating the \emph{XSure model}, as an
@@
 references a collection of \texttt{TaskType} through the composition relation
-\texttt{taskTypes} with (0..*) cardinality (\textbf{P1}). A \texttt{TaskType}
+\texttt{taskTypes} with (0..*) cardinality (\reqp{1}). A \texttt{TaskType}
 is embedded in a \texttt{ProcessType} and inherits from its context. To
@@
 \texttt{Type} and is specialized by \texttt{Sequencing}, \texttt{AndSplit},
-\texttt{AndJoin}, \texttt{OrSplit} and \texttt{OrJoin} concepts (\textbf{P2}).
+\texttt{AndJoin}, \texttt{OrSplit} and \texttt{OrJoin} concepts (\reqp{2}).
 Depending on its type and following its underlying operational semantics, a
@@
 \texttt{initialTaskType} (single cardinality) and \texttt{finalTaskType}
-(cardinality 0..*) (\textbf{P3}). \texttt{TaskType} exposes a creator role as a
-reference to an \texttt{Actor} concept (\textbf{P4}), which is at the same
+(cardinality 0..*) (\reqp{3}). \texttt{TaskType} exposes a creator role as a
+reference to an \texttt{Actor} concept (\reqp{4}), which is at the same
 conceptual level.
@@
 \texttt{allowedActorTypes} relation to \texttt{ActorType} defined in
-\texttt{TaskType} (with 0..* cardinality) captures \textbf{P5} requirement.
-Requirement \textbf{P6} is symmetrically satisfied with \texttt{allowedActors}
+\texttt{TaskType} (with 0..* cardinality) captures \reqp{5} requirement.
+Requirement \reqp{6} is symmetrically satisfied with \texttt{allowedActors}
 relation to \texttt{Actor} also defined in \texttt{TaskType} (with 0..*
@@
 \texttt{usedArtifactTypes} and \texttt{producedArtifactTypes} defined in
-\texttt{TaskType} (\textbf{P7}). \texttt{TaskType} additionally exposes an
+\texttt{TaskType} (\reqp{7}). \texttt{TaskType} additionally exposes an
 \texttt{expectedDuration} role (expressed in number of days), satisfying
-\textbf{P8}.
+\reqp{8}.
 
@@
 flagged as critical and must be performed by senior actors. To fulfill
-\textbf{P9} requirement, a supplementary constraint is required for
+\reqp{9} requirement, a supplementary constraint is required for
 \texttt{TaskType} and is captured through the following invariant expressed in
@@
 \texttt{Process} represents an enacted \texttt{ProcessType}, as defined in
-previous subsection (\textbf{P10}). \FML defines behavioral features called
+previous subsection (\reqp{10}). \FML defines behavioral features called
 behavior. \texttt{ProcessType} defines the behavior
@@
 \texttt{/type} role. Each instance of \texttt{TaskType} is ontologically
-instantiated with a \texttt{Task} (\textbf{P11}), using the same factory
+instantiated with a \texttt{Task} (\reqp{11}), using the same factory
 pattern where \texttt{TaskType} has the responsibility to manage the
@@
 \texttt{TaskType}, and defines a \texttt{begin date} and an \texttt{end date}
-basic role (\textbf{P12}).
+basic role (\reqp{12}).
 
@@
 \texttt{usedArtifacts}, \texttt{producedArtifacts} and
-\texttt{performingActors} defined in \texttt{Task} (\textbf{P13}). An instance
+\texttt{performingActors} defined in \texttt{Task} (\reqp{13}). An instance
 of \texttt{Artifact} specializes \texttt{MultiInstance} and references a set of
 \texttt{ArtifactType} through the specialized \texttt{/types} role
-(\textbf{P14} and \textbf{P16}). Likewise, \texttt{Actor} specializes
+(\reqp{14} and \reqp{16}). Likewise, \texttt{Actor} specializes
 \texttt{MultiInstance} and references a set of \texttt{ActorType} through the
-specialized \texttt{/types} relation (\textbf{P15}).
+specialized \texttt{/types} relation (\reqp{15}).
 
@@
 
-Authorization for an actor to perform a task (\textbf{P17}) is captured either
+Authorization for an actor to perform a task (\reqp{17}) is captured either
 by the role \texttt{allowedActors} or the role \texttt{allowedActorTypes}
@@
 Actor types specialization is captured by the \texttt{parentActorTypes}
-relation defined in \texttt{ActorType} (\textbf{P18}). This is completed by
+relation defined in \texttt{ActorType} (\reqp{18}). This is completed by
 both the definition of the \texttt{hasActorType(ActorType)} behavior in
@@
 \texttt{lastUpdated} attribute with \texttt{Date} type, and thus satisfies
-\textbf{P19} requirement.
+\reqp{19} requirement.
 
@@
 \texttt{\{}\inst{Analyst}\texttt{\}} for \texttt{allowedActorTypes}
-(\textbf{S1}). Similarly, \inst{Test case design} is an instance of
+(\reqs{1}). Similarly, \inst{Test case design} is an instance of
 \texttt{SETaskType}, with a value of \texttt{true} for its property
@@
 \texttt{allowedActorTypes} relation. \inst{Test case design} produces
-\inst{Test cases} instance of \texttt{SETaskType} (\textbf{S2}). The latter is
+\inst{Test cases} instance of \texttt{SETaskType} (\reqs{2}). The latter is
 used in \inst{Test design review} (\texttt{producedArtifactTypes} relation)
-(\textbf{S1} and \textbf{S13}, satisfied with \textbf{P9}).
+(\reqs{1} and \reqs{13}, satisfied with \reqp{9}).
 
-The capture of \textbf{S3} requirement is performed while defining
+The capture of \reqs{3} requirement is performed while defining
 \inst{Coding} as an instance of concept \texttt{CodingTaskType} and defining a
@@
 relation \texttt{languages} to \texttt{ProgrammingLanguage} with 1..*
-cardinality (\textbf{S4}).
+cardinality (\reqs{4}).
 
-\textbf{S5} is more ambiguous as a task type \texttt{CodingTask} defines one or
+\reqs{5} is more ambiguous as a task type \texttt{CodingTask} defines one or
 more programming languages but produces code which is expressed in one language
@@
 \texttt{declaresRequiredProducedArtifacts} where programming language should
-also match.  \textbf{S6} is guaranteed though the \texttt{languages} relation
+also match.  \reqs{6} is guaranteed though the \texttt{languages} relation
 defined in \texttt{CodeArtifact} and the following invariant declared in
@@
 
-\inst{Ann Smith} is the only one allowed to perform coding in COBOL (\textbf{S7}). This is implemented through the redefinition of \texttt{newTask(String)} in \texttt{CodingTaskType}:
+\inst{Ann Smith} is the only one allowed to perform coding in COBOL (\reqs{7}). This is implemented through the redefinition of \texttt{newTask(String)} in \texttt{CodingTaskType}:
 
@@
 
-The requirement \textbf{S8} is simply expressed by the definition of the
+The requirement \reqs{8} is simply expressed by the definition of the
 \inst{Testing} instance of \texttt{SETaskType}, \inst{Tester} instance of
@@
 \texttt{validatedWith} defined in \texttt{Artifact} and referencing
-\texttt{Artifact} validating it (\textbf{S9}).
+\texttt{Artifact} validating it (\reqs{9}).
 
@@
 two attributes: \texttt{responsible} (an \texttt{Actor} instance), and
-\texttt{versionNumber} (an integer). It thus fulfills \textbf{S10}. \inst{Bob
+\texttt{versionNumber} (an integer). It thus fulfills \reqs{10}. \inst{Bob
 Brown} is declared as an instance of \texttt{Actor}, and references
@@
 referenced by all instances of \texttt{SETaskType} as the creator for related
-task types (\textbf{S11}).
+task types (\reqs{11}).
 
-For \textbf{S12}, we assume that all tasks may define an expected duration,
+For \reqs{12}, we assume that all tasks may define an expected duration,
 which might be checked during process execution. This is modeled by the
@@
 as an instance of \texttt{SETesting}. The business logic expressed by
-\textbf{S12} requirement should then be redefined in \texttt{SETesting}.
+\reqs{12} requirement should then be redefined in \texttt{SETesting}.
 
-\textbf{S13} requirement has been previously partially fulfilled. This must be
+\reqs{13} requirement has been previously partially fulfilled. This must be
 completed with an association between an artifact produced and the task that
@@
 semantics follows rules defined in Sect.~\ref{sec:ProcessEnactment}
-(\textbf{P13}) is also implemented.
+(\reqp{13}) is also implemented.
 

### 2022-03-25 7-requirementssatisfaction.tex
format all requirements with new formatting

@@
     \item \emph{Conceptualization} (\textit{Conc.}): a concept is created to
-      fulfill or meet the requirement (for example for \textbf{P1}:
+      fulfill or meet the requirement (for example for \reqp{1}:
       \textit{process type} is conceptualized in a concept
@@
     \item \emph{Specialization} (\textit{Spec.}): a concept is specialized in
-      the sense of object-oriented modeling (for example for \textbf{P2}:
+      the sense of object-oriented modeling (for example for \reqp{2}:
       \texttt{Gateway} defines an abstract behavior \texttt{execute()} and
@@
       or specific attributes were added to some concepts (for example for
-      \textbf{P3}: a \textit{process type} has one \textit{initial task type}).
+      \reqp{3}: a \textit{process type} has one \textit{initial task type}).
 
@@
       satisfied by the instantiation of one or more concepts (for example for
-      \textbf{S1}: \inst{Requirement analysis} is defined as an instance of
+      \reqs{1}: \inst{Requirement analysis} is defined as an instance of
     \texttt{SETaskType}).
@@
       definition instance (for example the notion of \textit{developer} of
-      \textbf{S3} is implemented by both a concept in \textit{Acme metamodel}
+      \reqs{3} is implemented by both a concept in \textit{Acme metamodel}
     and an instance in \textit{Acme model}, see bottom left of
@@
   satisfied by the definition of one or more behaviors, which may be combined
-  with constraints checking and associated tools (for example \textbf{P17}).
+  with constraints checking and associated tools (for example \reqp{17}).
   \end{itemize}
@@
   \hline
-    \textbf{P1}  & \checkmark &            & \checkmark &            &            &            &            & \\
-    \textbf{P2}  & \checkmark & \checkmark & \checkmark &            &            &            &            & \\
-    \textbf{P3}  &            &            & \checkmark &            &            &            &            & \\
-    \textbf{P4}  & \checkmark &            &            &            & \checkmark &            &            & \\
-    \textbf{P5}  & \checkmark &            & \checkmark &            &            &            &            & \\
-    \textbf{P6}  & \checkmark &            & \checkmark &            &  &            &            & \\
-    \textbf{P7}  & \checkmark &            & \checkmark &            &            &            &            & \\
-    \textbf{P8}  &            &            & \checkmark &            &            &            &            & \\
-    \textbf{P9}  &            &            & \checkmark &            &            &            & \checkmark & \\
-    \textbf{P10} & \checkmark &            &            &            & \checkmark &            &            & \\
-    \textbf{P11} & \checkmark &            & \checkmark &            & \checkmark &            &            & \\
-    \textbf{P12} &            &            & \checkmark &            &            &            &            & \\
-    \textbf{P13} & \checkmark &            & \checkmark &            & \checkmark &            &            & \\
-    \textbf{P14} & \checkmark &            &            &            & \checkmark &            &            & \\
-    \textbf{P15} & \checkmark &            &            &            & \checkmark &            &            & \\
-    \textbf{P16} & \checkmark &            &            &            & \checkmark &            &            & \\
-    \textbf{P17} & \checkmark           &            &            &            &            & \checkmark & \checkmark & \checkmark \\
-    \textbf{P18} & \checkmark           &            &            &            &            & \checkmark &            & \\
-    \textbf{P19} &            & \checkmark & \checkmark &            &            & \checkmark &            & \\
+    \reqp{1}  & \checkmark &            & \checkmark &            &            &            &            & \\
+    \reqp{2}  & \checkmark & \checkmark & \checkmark &            &            &            &            & \\
+    \reqp{3}  &            &            & \checkmark &            &            &            &            & \\
+    \reqp{4}  & \checkmark &            &            &            & \checkmark &            &            & \\
+    \reqp{5}  & \checkmark &            & \checkmark &            &            &            &            & \\
+    \reqp{6}  & \checkmark &            & \checkmark &            &  &            &            & \\
+    \reqp{7}  & \checkmark &            & \checkmark &            &            &            &            & \\
+    \reqp{8}  &            &            & \checkmark &            &            &            &            & \\
+    \reqp{9}  &            &            & \checkmark &            &            &            & \checkmark & \\
+    \reqp{10} & \checkmark &            &            &            & \checkmark &            &            & \\
+    \reqp{11} & \checkmark &            & \checkmark &            & \checkmark &            &            & \\
+    \reqp{12} &            &            & \checkmark &            &            &            &            & \\
+    \reqp{13} & \checkmark &            & \checkmark &            & \checkmark &            &            & \\
+    \reqp{14} & \checkmark &            &            &            & \checkmark &            &            & \\
+    \reqp{15} & \checkmark &            &            &            & \checkmark &            &            & \\
+    \reqp{16} & \checkmark &            &            &            & \checkmark &            &            & \\
+    \reqp{17} & \checkmark           &            &            &            &            & \checkmark & \checkmark & \checkmark \\
+    \reqp{18} & \checkmark           &            &            &            &            & \checkmark &            & \\
+    \reqp{19} &            & \checkmark & \checkmark &            &            & \checkmark &            & \\
   \hline
@@
   \hline
-    \textbf{S1}  &            &            &            & \checkmark &            &            &            & \\
-    \textbf{S2}  &            &            &            & \checkmark &            &            &            & \\
-    \textbf{S3}  & \checkmark & \checkmark & \checkmark & \checkmark &            &            &            & \checkmark \\
-    \textbf{S4}  & \checkmark & \checkmark & \checkmark & \checkmark &            &            &            & \\
-    \textbf{S5}  &            & \checkmark &            & \checkmark &            & \checkmark &            & \\
-    \textbf{S6}  &            & \checkmark &            & \checkmark &            & \checkmark & \checkmark & \\
-    \textbf{S7}  &            &            &            & \checkmark &            & \checkmark &            & \\
-    \textbf{S8}  &            &            &            & \checkmark &            &            &            & \\
-    \textbf{S9}  &            &            &            &            &            & \checkmark & \checkmark & \\
-    \textbf{S10} & \checkmark & \checkmark & \checkmark &            &            &            &            & \\
-    \textbf{S11} &            &            &            & \checkmark &            &            &            & \checkmark \\
-    \textbf{S12} &            &            &            & \checkmark &            &            &            & \checkmark \\
-    \textbf{S13} &            &            &            & \checkmark &            &            &            & \\
+    \reqs{1}  &            &            &            & \checkmark &            &            &            & \\
+    \reqs{2}  &            &            &            & \checkmark &            &            &            & \\
+    \reqs{3}  & \checkmark & \checkmark & \checkmark & \checkmark &            &            &            & \checkmark \\
+    \reqs{4}  & \checkmark & \checkmark & \checkmark & \checkmark &            &            &            & \\
+    \reqs{5}  &            & \checkmark &            & \checkmark &            & \checkmark &            & \\
+    \reqs{6}  &            & \checkmark &            & \checkmark &            & \checkmark & \checkmark & \\
+    \reqs{7}  &            &            &            & \checkmark &            & \checkmark &            & \\
+    \reqs{8}  &            &            &            & \checkmark &            &            &            & \\
+    \reqs{9}  &            &            &            &            &            & \checkmark & \checkmark & \\
+    \reqs{10} & \checkmark & \checkmark & \checkmark &            &            &            &            & \\
+    \reqs{11} &            &            &            & \checkmark &            &            &            & \checkmark \\
+    \reqs{12} &            &            &            & \checkmark &            &            &            & \checkmark \\
+    \reqs{13} &            &            &            & \checkmark &            &            &            & \\
   \hline

### 2022-03-25 3-introduction.tex
change requirements format + macros
typo

@@
 autonomous and are represented in their original technological spaces whereas
-virtual models (also called conceptual models) and links that serve as control
+virtual models (also called conceptual models) and links serve as control
 components used to present different views to the users and maintain

### 2022-03-25 4-technology.tex
typo
shorter code

--- 2022-03-25 4-technology.tex
@@
 after federations are made.
-So to provide an adaptable framework, we have developed a flexible modeling language to provide modeling features without strict modeling levels.
+So to provide an adaptable framework, we have developed a flexible modeling language offering modeling features without strict modeling levels.
 In a federation, the targeted modeling approaches can be unknown at the beginning and also can integrate several modeling levels.
@@
       execute() {
-        log("Execute task "+name);
         ...
@@
       // BPMNTask concept
-      ...
-
     }

### 2022-03-25 4-technology.tex
experiment on placement

--- 2022-03-25 4-technology.tex
@@
     \centering
-    \includegraphics[width=1.0 \textwidth]{Figures/FMLMetaModel-1.5.pdf}
+    \includegraphics[width=0.95 \textwidth]{Figures/FMLMetaModel-1.5.pdf}
     \caption{Representation of the main concepts of {FML} language}
@@
 
-A parallel to object-oriented approach can be useful to understand
+A parallel to object-oriented approach is useful to understand
 \FML\footnote{Some features of \FML do not exist in object oriented
@@
 are connected to external elements through model slot instances.
-
-% Sylvain : relire et vérifier
-
-\begin{figure}[t]
-    \centering
-    \includegraphics[width=\columnwidth]{Figures/BPMNSubsetExample-1.5.pdf}
-    \caption{An example of \BPMN subset showing both the model and an instance of that model}
-    \label{fig:BPMNSubsetExample}
-\end{figure}
-
+%
 To illustrate this, Fig.~\ref{fig:BPMNSubsetExample} shows how to realize a
@@
 
-%\noteSylvain{A la fin de la rédaction, s'arranger pour que la figure 3 (fig:BPMNSubsetExample) et le listing soient cote a cote, et aussi que le listing tienne sur une seule page}
+\begin{figure}[t]
+    \centering
+    \includegraphics[width=\columnwidth]{Figures/BPMNSubsetExample-1.5.pdf}
+    \caption{An example of \BPMN subset showing both the model and an instance of that model}
+    \label{fig:BPMNSubsetExample}
+\end{figure}
 

### 2022-03-25 6-modelpresentation.tex
rewording for placement

@@
 
-This invariant should be completed with additional constraints defined in the
-\texttt{Task} concept, which apply to performing actors assigned to enact
+This invariant should be extended with additional constraints defined in the
+\texttt{Task} concept, which apply to actors assigned to enact
 tasks.
@@
 
-The used and produced artifacts are implemented by roles
+A \texttt{Task} defines the roles
 \texttt{usedArtifacts}, \texttt{producedArtifacts} and
-\texttt{performingActors} defined in \texttt{Task} (\reqp{13}). An instance
+\texttt{performingActors}  (\reqp{13}). An instance
 of \texttt{Artifact} specializes \texttt{MultiInstance} and references a set of
@@
 The \texttt{Task} concept delegates this authorization to its task type, as
-shown in the following \FML code:
+shown below.
 
@@
 
-Enforcing those constraints is finally performed by the definition of this
-invariant in the \texttt{Task} concept:
+Enforcing those constraints is finally performed by the following
+invariant of the \texttt{Task} concept:
 

### 2022-03-26 6-modelpresentation.tex
cosmetics

@@
 The Base metamodel is completed with some behaviors implementing an execution
-semantics for the executed processes. All tasks --- instantiated from
-\texttt{TaskType} for a given enacted process --- manage a status. This status
-is either \texttt{Not startable} (when not assigned to a performing actor or
-when required input artifacts are not available), \texttt{Startable},
+semantics for the executed processes. All tasks manage a status and are
+instantiated from \texttt{TaskType} for a given enacted process. This status is
+either \texttt{Not startable} (when not assigned to a performing actor or when
+required input artifacts are not available), \texttt{Startable},
 \texttt{Started}, \texttt{Completable} (when all output artifacts are ready) or
@@
 abstract \texttt{execute(Process)} behavior. The management of artifacts whose
-semantics follows rules defined in Sect.~\ref{sec:ProcessEnactment}
-(\reqp{13}) is also implemented.
+semantics follows rules defined in Sect.~\ref{sec:ProcessEnactment} (\reqp{13})
+is also implemented.
 
 Fig.~\ref{fig:ToolingArchitecture} details the architecture of these two tools.
-
 The left of the figure presents the \textit{ProcessType graphical editor}.
@@
 allows representing and editing a \texttt{ProcessType}. The highly reflective
-nature of \FML and its tools should be noted. When the drag and drop
-interactor is applied for a new item, the tool allows choosing the concept type
-to be instantiated. For example, when a \texttt{TaskType} is created, the user
-must choose a concept inheriting from \texttt{TaskType} (in Acme use case it
-can be \texttt{SETaskTask} or \texttt{CodingTaskType} or default value
+nature of \FML and its tools should be noted. When the drag and drop interactor
+is applied for a new item, the tool allows choosing the concept type to be
+instantiated. For example, when a \texttt{TaskType} is created, the user must
+choose a concept inheriting from \texttt{TaskType} (in Acme use case it can be
+  \texttt{SETaskTask} or \texttt{CodingTaskType} or default value
 \texttt{TaskType}).

### 2022-03-26 1-main.tex
change requirements format + macros
rm useless package (tikz-uml)
fix keywords in metadata

@@
 %\usepackage{hyperref}
-%\usepackage{tikz-uml}
 %\usepackage{float}
@@
 \newcommand{\inst}[1]{\textsf{'#1'}}
+
+\newcommand{\req}[2]{\textsf{#1}$_{#2}$}
+\newcommand{\reqp}[1]{\req{P}{#1}}
+\newcommand{\reqs}[1]{\req{S}{#1}}
+%\newcommand{\reqp}[1]{\textsf{P}$_{#1}$}
+%\newcommand{\reqs}[1]{\textsf{S}$_{#1}$}
 
@@
   breaklinks=true,
-  pdfkeywords={Multi-Level Modeling Challenge, Federation Modeling Language,
-  Openflexo},
+  pdfkeywords={Model federation, Multi-Level Modeling, System modeling
+  languages, Abstraction, modeling and modularity, Reusability, Openflexo},
   %abstract
@@
     \abstract{\input{2-abstract}}
-    \keywords{Multi-Level Modeling \and System modeling languages \and Abstraction, modeling and modularity \and Reusability }
+    \keywords{Model federation \and Multi-Level Modeling \and System modeling languages \and Abstraction, modeling and modularity \and Reusability }
     %ACM

### 2022-03-31 7-requirementssatisfaction.tex
clarification

@@
     \item \emph{syntactic conformance}: the requirements are structurally
-      satisfied when the model conforms to its metamodel,
+      satisfied when the model conforms to its metamodel
+
     \item \emph{constraints checking}: the requirements are fulfilled when all
       instances satisfy both syntactic conformance and the evaluation of
-      invariants expressed for related concepts,
-    \item \emph{tooling}: the requirements are enforced by the implementation
-      of the associated tool.
+      invariants expressed for related concepts
+
+    \item \emph{process tools}: the requirements are enforced by the
+      implementation of the process tools illustrated by
+      Fig.~\ref{fig:ToolingArchitecture}. It can be done by using behaviors
+      defined in the models or by using the process tools manually.
+
 \end{itemize}
@@
    \hline
-    & \multicolumn{6}{c|}{Syntactic conformance} & Constraints & \multirow{2}{*}{Tooling}\\
+    & \multicolumn{6}{c|}{Syntactic conformance} & Constraints & \multirow{2}{*}{Process tools}\\
    \cline{1-7}

### 2022-04-06 conclusions.tex
renaming of .tex files

--- 2022-02-23 10-conclusions.tex
@@
 
-We fulfill all the challenge requirements and we propose a tooling that
+We fulfill all the challenge requirements and we propose a tool that
 demonstrates the usability of our solution. We developed an ad hoc solution,
@@
 solution seamlessly integrates the type-object pattern which allows us to
-dynamically create different \emph{instances} of model elements (\eg, \emph{TaskType}, among others) and use them for typing other model elements.
-This, together with the \FML core support for (level agnostic) classical linguistic instantiation, permitted us to merge linguistic and ontological instantiation in order to provide a satisfactory solution to the Multi-Level Process challenge.
+dynamically create different \emph{instances} of model elements (\eg,
+\emph{TaskType}, among others) and use them for typing other model elements.
+This, together with the \FML core support for (level agnostic) classical
+linguistic instantiation, permitted us to merge linguistic and ontological
+instantiation in order to provide a satisfactory solution to the Multi-Level
+Process challenge.
 
-The features of \FML enable a great modeling flexibility which, coupled with an appropriate methodology, allow us to achieve 
-the multi-level capabilities without dedicated tooling, in order to solve
-problems which require it. We use virtual models to \enquote{implement} levels on demand.
+The features of \FML enable a great modeling flexibility which, coupled with an
+appropriate methodology, allow us to achieve the multi-level capabilities
+without dedicated tool, in order to solve problems which require it. We use
+virtual models to \enquote{implement} levels on demand.
 

### 2022-04-06 discussion.tex
renaming of .tex files

--- 2022-03-17 8-discussion.tex
@@
   \FML. For instance, \texttt{isAuthorizedActor} behavior is needed to enforce
-  \textbf{P17} requirement, leading to writing \FML code in \texttt{TaskType}
+  \reqp{17} requirement, leading to writing \FML code in \texttt{TaskType}
   and \texttt{Task}. Note that in our approach, metamodels are built in an
@@
 users add behaviors when they build the metamodels and the models. For
-instance, \textbf{S6} requirement is constrained by an invariant implemented in
+instance, \reqs{6} requirement is constrained by an invariant implemented in
 \texttt{CodeArtifact}. It ensures that \texttt{artifactType} has the right type

### 2022-04-06 relatedworks.tex
renaming of .tex files

--- 2022-02-25 9-relatedworks.tex
@@
 
-A plethora of multi-level modeling approaches and tools\footnote{\url{https://homepages.ecs.vuw.ac.nz/Groups/MultiLevelModeling/MultiTools}} with different foundations have appeared in recent years~\parencite{somogyi2021playground}.
-Comparing them lies out of the scope of this paper.
-In this sense, we limit the following discussion to the presentation and comparison of previous solutions to the MULTI Process Challenge.
+A plethora of multi-level modeling approaches and
+tools\footnote{\raggedright\url{https://homepages.ecs.vuw.ac.nz/Groups/MultiLevelModeling/MultiTools}}
+with different foundations have appeared in recent
+years~\parencite{somogyi2021playground}. Comparing them lies out of the scope
+of this paper. In this sense, we limit the following discussion to the
+presentation and comparison of previous solutions to the MULTI Process
+Challenge.
 
@@
 levels, dual ontological/linguistic typing and deep characterization through
-potency~\parencite{atkinson2001essence}. For their solution to the multi-level process problem
-they use three levels. The first level describes generic processes, the second
-level software engineering processes, and the third level, software engineering
-enacted processes. Additionally, the authors use~\emph{linguistic
-extensions}~\parencite{metadepth}, a mechanism to  linguistically extend
-ontological instance models at any level, in order to introduce artifact types
-and tasks duration in the second level. Our solution is similar to theirs
-w.r.t.
-the grouping and organization of concepts (\eg, a model for general process, concepts, an extension for software engineering process) but it replaces the use of clabjets and potency with the use of the type-object pattern~\parencite{typeObject}.
-The so-called \emph{linguistic extensions} are natively supported in Openflexo/\FML.
+potency~\parencite{atkinson2001essence}. For their solution to the multi-level
+process problem they use three levels. The first level describes generic
+processes, the second level software engineering processes, and the third
+level, software engineering enacted processes. Additionally, the authors
+use~\emph{linguistic extensions}~\parencite{metadepth}, a mechanism to
+linguistically extend ontological instance models at any level, in order to
+introduce artifact types and tasks duration in the second level. Our solution
+is similar to theirs w.r.t. the grouping and organization of concepts (\eg, a
+model for general process, concepts, an extension for software engineering
+process) but it replaces the use of clabjets and potency with the use of the
+type-object pattern~\parencite{typeObject}. The so-called \emph{linguistic
+extensions} are natively supported in Openflexo/\FML.
 
@@
 
-\textcite{multiecore2019} use MultEcore \parencite{multecore2016} to provide a solution to the challenge.
-MultEcore uses (un)pluggable linguistic levels (\eg, there is not a \emph{fixed} core metamodel defining the concept of clabjet), an extension of the two-level cascading technique \parencite{atkinson2005concepts} (this is achieved with model transformations to transform instance models into instantiable models) and potency.
-Regarding the challenge, their solution uses 4 levels.
-The first level represents generic processes and constitutes the root level for two ontological hierarchies, one for the software engineering domain and the other for the insurance domain.
-Each hierarchy contains three other ontological levels.
-The second level refines generic process to software engineering and insurance processes.
-The third level refines the aforementioned models to adapt them to the ACME and Xsure processes.
-Finally, process instances lie in level 4.
-Additionally, the authors use an orthogonal linguistic hierarchy to support alternative names for every model element.
-With respect to our solution, the authors use more models, as we require less refinement steps.
-However accidental complexity is introduced by our approach as we need additional constructs and constraints for the typing relations which are otherwise built-in in MultEcore.
+\textcite{multiecore2019} use MultEcore \parencite{multecore2016} to provide a
+solution to the challenge. MultEcore uses (un)pluggable linguistic levels (\eg,
+there is not a \emph{fixed} core metamodel defining the concept of clabjet), an
+extension of the two-level cascading technique \parencite{atkinson2005concepts}
+(this is achieved with model transformations to transform instance models into
+instantiable models) and potency. Regarding the challenge, their solution uses
+4 levels. The first level represents generic processes and constitutes the
+root level for two ontological hierarchies, one for the software engineering
+domain and the other for the insurance domain. Each hierarchy contains three
+other ontological levels. The second level refines generic process to software
+engineering and insurance processes. The third level refines the aforementioned
+models to adapt them to the ACME and Xsure processes. Finally, process
+instances lie in level 4. Additionally, the authors use an orthogonal
+linguistic hierarchy to support alternative names for every model element.
+With respect to our solution, the authors use more models, as we require less
+refinement steps. However accidental complexity is introduced by our approach
+as we need additional constructs and constraints for the typing relations which
+are otherwise built-in in MultEcore.
 
@@
 such as the top module contains base concepts and formulas required for
-multi-level modeling (\eg, the support for \emph{most general instances}), and a sub-module contains process definitions fulfilling requirements P1 to P19.
-This sub-module contains in turn a sub-module representing the coding process type whereas concrete coding processes are represented in subsequent sub-modules.
-There are two main differences between the DeepTelos and Openflexo/\FML: 1) we use the type-object pattern instead of the powertype pattern; 2) DeepTelos reifies their support for multi-level modeling while our remains ad hoc.
+multi-level modeling (\eg, the support for \emph{most general instances}), and
+a sub-module contains process definitions fulfilling requirements P1 to P19.
+This sub-module contains in turn a sub-module representing the coding process
+type whereas concrete coding processes are represented in subsequent
+sub-modules. There are two main differences between the DeepTelos and
+Openflexo/\FML: 1) we use the type-object pattern instead of the powertype
+pattern; 2) DeepTelos reifies their support for multi-level modeling while our
+remains ad hoc.
 
@@
 
-Finally, \textcite{dmla2019} use the Dynamic Multi-Layer Algebra (DMLA) \parencite{dmla2017} in order to solve the process challenge.
-DMLA is a level blind modeling framework which is fully customizable (\eg different types of instantiation may be implemented) and includes support for deep characterization (instantiation may refer to any other element disregarding hierarchy levels and a sort of potency in the form of fluid metamodeling at the entity level exists).
-Their proposed solution separates the process challenge in two separate domains, namely, the task definition domain and the process definition domain.
-Wrappers are used in order to reuse tasks in processes.
-Their solution does not use inheritance, as it is not supported by the framework.
-It does not use explicitly separated models either.
+Finally, \textcite{dmla2019} use the Dynamic Multi-Layer Algebra (DMLA)
+\parencite{dmla2017} in order to solve the process challenge. DMLA is a level
+blind modeling framework which is fully customizable (\eg different types of
+instantiation may be implemented) and includes support for deep
+characterization (instantiation may refer to any other element disregarding
+hierarchy levels and a sort of potency in the form of fluid metamodeling at the
+entity level exists). Their proposed solution separates the process challenge
+in two separate domains, namely, the task definition domain and the process
+definition domain. Wrappers are used in order to reuse tasks in processes.
+Their solution does not use inheritance, as it is not supported by the
+framework. It does not use explicitly separated models either.
 
